### 2022-11-14 src/tex/ms.tex
[showyourwork] first commit

% Define document class
\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}

% Begin!
\begin{document}

% Title
\title{An open source scientific article}

% Author list
\author{@jiayindong}

% Abstract with filler text
\begin{abstract}
    Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
    Donec vehicula augue eu neque, morbi tristique senectus et netus et.
    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
    Phasellus eu tellus sit amet tortor gravida placerat.
    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
    Praesent eget sem vel leo ultrices bibendum.
    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
    Curabitur auctor semper nulla donec varius orci eget risus.
    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
    Duis eget orci sit amet orci dignissim rutrum.
\end{abstract}

% Main body with filler text
\section{Introduction}
\label{sec:intro}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
Donec vehicula augue eu neque, morbi tristique senectus et netus et.
Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
Phasellus eu tellus sit amet tortor gravida placerat.
Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
Praesent eget sem vel leo ultrices bibendum.
Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
Curabitur auctor semper nulla donec varius orci eget risus.
Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
Duis eget orci sit amet orci dignissim rutrum.

Nam dui ligula, fringilla a, euismod sodales, sollici- tudin vel, wisi.
Morbi auctor lorem non justo, nam lacus libero, pretium at, lobortis vitae.
Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna.
Morbi ac orci et nisl hendrerit mollis, suspendisse ut massa, cras nec ante.
Pellentesque a nulla cum sociis natoque penatibus et magnis dis parturient.
Aliquam tincidunt urna, nulla ullamcorper vestibulum turpis.
Pellentesque cursus luctus mauris \citep{Luger2021}.

\bibliography{bib}

\end{document}

### 2022-11-14 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 % Title
-\title{An open source scientific article}
+\title{True Stellar Obliquity Distribution of Exoplanets}
 
-% Author list
-\author{@jiayindong}
+\correspondingauthor{Jiayin Dong}
+\email{jdong@flatironinstitute.org}
+
+\newcommand{\FlatironCCA}{Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Avenue, New York, NY 10010, USA}
+
+\author[0000-0002-3610-6953]{Jiayin Dong}
+\altaffiliation{Flatiron Research Fellow}
+\affiliation{\FlatironCCA}
+
+\author[0000-0002-9328-5652]{Dan Foreman-Mackey}
+\affiliation{\FlatironCCA}
 
@@
 \begin{abstract}
-    Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
-    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
-    Donec vehicula augue eu neque, morbi tristique senectus et netus et.
-    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
-    Phasellus eu tellus sit amet tortor gravida placerat.
-    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
-    Praesent eget sem vel leo ultrices bibendum.
-    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
-    Curabitur auctor semper nulla donec varius orci eget risus.
-    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
-    Duis eget orci sit amet orci dignissim rutrum.
+We build a hierarchical Bayesian framework to constrain exoplanets' true stellar obliquity ($\psi$) distribution.
+Using the sky-projected stellar obliquities ($\lambda$)  of a sample of 162 exoplanets, we infer their $\cos{\psi}$ distribution using a four-parameter Beta mixture model under the assumption of a uniform cosine of the stellar inclination ($\cos{i_\star}$) bounded between $-1$ and $1$.
+We find a significant pile-up of $\cos{\psi}$ near 1, suggesting that the majority of exoplanetary systems are well aligned.
+For misaligned systems, we find $\cos{\psi}$ is uniformly distributed between $-0.75$ to $0.5$ and have a slight but insignificant preference to $\cos{\psi} = -0.25$, corresponding to $\psi = 104.5\degr$.
+A pile-up at $\cos{\psi} = 0$ is not found.
 \end{abstract}
@@
 
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
-Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
-Donec vehicula augue eu neque, morbi tristique senectus et netus et.
-Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
-Phasellus eu tellus sit amet tortor gravida placerat.
-Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
-Praesent eget sem vel leo ultrices bibendum.
-Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
-Curabitur auctor semper nulla donec varius orci eget risus.
-Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
-Duis eget orci sit amet orci dignissim rutrum.
-
-Nam dui ligula, fringilla a, euismod sodales, sollici- tudin vel, wisi.
-Morbi auctor lorem non justo, nam lacus libero, pretium at, lobortis vitae.
-Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna.
-Morbi ac orci et nisl hendrerit mollis, suspendisse ut massa, cras nec ante.
-Pellentesque a nulla cum sociis natoque penatibus et magnis dis parturient.
-Aliquam tincidunt urna, nulla ullamcorper vestibulum turpis.
-Pellentesque cursus luctus mauris \citep{Luger2021}.
-
 \bibliography{bib}

### 2022-11-15 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \usepackage{showyourwork}
+
+\received{}
+\revised{}
+\accepted{}
+
+\submitjournal{AAS Journals}
 
@@
 
+Stellar obliquity $\psi$ is described as the angle between a planet's orbital axis $\hat{i}_{\rm orb}$ and its host star's spin axis $\hat{i}_{\star}$.
+If both $\hat{i}_{\rm orb}$ and $\hat{i}_{\star}$ are random unit vectors with uniform distribution on a three-dimensional sphere, $\cos{\psi}$ will be uniformly distributed between $-1$ and 1.
+
+\begin{figure}[ht!]
+    \script{hbm_istar.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/hbm_istar.pdf}
+        \caption{Plot showing a bunch of random numbers.}
+        \label{fig:hbm_istar}
+    \end{centering}
+\end{figure}
+
+\begin{figure}
+    \script{hbm_all.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/hbm_all.pdf}
+        \caption{Plot showing a bunch of random numbers.}
+        \label{fig:hbm_all}
+    \end{centering}
+\end{figure}
+
 \bibliography{bib}

### 2022-11-15 src/tex/ms.tex
[showyourwork] overleaf sync

@@
         \includegraphics[width=\linewidth]{figures/hbm_istar.pdf}
-        \caption{Plot showing a bunch of random numbers.}
+        \caption{Hierarchical Bayesian model to infer \emph{true} stellar obliquity for planets with $\hat{i}_{\star}$ inferred from their rotation periods.}
         \label{fig:hbm_istar}
@@
 
-\begin{figure}
+\begin{figure*}
     \script{hbm_all.py}
@@
         \includegraphics[width=\linewidth]{figures/hbm_all.pdf}
-        \caption{Plot showing a bunch of random numbers.}
+        \caption{Hierarchical Bayesian model to infer \emph{true} stellar obliquity for all planets with sky-projected stellar obliquities measured.}
         \label{fig:hbm_all}
     \end{centering}
-\end{figure}
+\end{figure*}
 

### 2022-11-15 src/scripts/hbm_all.py
add hbm graphs

@@ -1,0 +1,118 @@
+import paths
+
+import sys
+import subprocess
+
+# implement pip as a subprocess:
+subprocess.check_call([sys.executable, '-m', 'pip', 'install',
+'daft'])
+import daft
+
+from matplotlib import rc
+rc('font', **{'family':'sans-serif'})
+rc('text', usetex=True)
+
+pgm = daft.PGM(observed_style="inner")
+
+# Hierarchical parameters.
+pgm.add_node("w", r"$w$", 1.5, 2.7)
+pgm.add_node("alpha", r"$\alpha$", 1.5, 2.15)
+pgm.add_node("beta", r"$\beta$", 1.5, 1.6)
+
+# Latent variable.
+pgm.add_node("lam", r"$\lambda$", 3, 4)
+pgm.add_node("obslam", r"$\lambda$", 3, 3.2, observed=True)
+
+pgm.add_node("coslam", r"$\cos{\lambda}$", 4, 3.2)
+pgm.add_node("errlam", r"$\sigma_\lambda$", 2.2, 3.2, fixed=True)
+
+
+pgm.add_node("cosi", r"$\cos{i}$", 5, 4)
+pgm.add_node("sini", r"$\sin{i}$", 5, 3.2)
+
+pgm.add_node("Rstar", r"$R_{\star}$", 5.8, 4)
+pgm.add_node("Prot", r"$P_{\rm rot}$", 6.6, 4)
+
+pgm.add_node("v", r"$v$", 6.2, 3.2)
+
+pgm.add_node("cospsi", r"$\cos{\psi}$", 4.5, 1.9)
+
+pgm.add_node("vsini", r"$v\sin{i}$", 6.2, 2.4)
+pgm.add_node("obsvsini", r"$v\sin{i}$", 6.2, 1.6, observed=True)
+pgm.add_node("errvsini", r"$\sigma_{v\sin{i}}$", 7, 1.6, fixed=True)
+
+# Add in the edges.
+pgm.add_edge("lam", "coslam")
+pgm.add_edge("lam", "obslam")
+pgm.add_edge("errlam", "obslam")
+
+pgm.add_edge("cosi", "sini")
+
+pgm.add_edge("Rstar", "v")
+pgm.add_edge("Prot", "v")
+
+
+pgm.add_edge("sini", "cospsi")
+pgm.add_edge("coslam", "cospsi")
+
[…]

### 2022-11-15 src/scripts/hbm_istar.py

@@ -1,0 +1,76 @@
+import paths
+
+import sys
+import subprocess
+
+# implement pip as a subprocess:
+subprocess.check_call([sys.executable, '-m', 'pip', 'install',
+'daft'])
+import daft
+
+from matplotlib import rc
+rc('font', **{'family':'sans-serif'})
+rc('text', usetex=True)
+
+pgm = daft.PGM(observed_style="inner")
+
+# Hierarchical parameters.
+pgm.add_node("w", r"$w$", 1.5, 2.7)
+pgm.add_node("alpha", r"$\alpha$", 1.5, 2.15)
+pgm.add_node("beta", r"$\beta$", 1.5, 1.6)
+
+# Latent and observed variables.
+pgm.add_node("lam", r"$\lambda$", 3, 4)
+pgm.add_node("obslam", r"$\lambda$", 3, 3.2, observed=True)
+
+pgm.add_node("coslam", r"$\cos{\lambda}$", 4, 3.2)
+pgm.add_node("errlam", r"$\sigma_\lambda$", 2.2, 3.2, fixed=True)
+
+
+pgm.add_node("cosi", r"$\cos{i}$", 5, 4)
+pgm.add_node("sini", r"$\sin{i}$", 5, 3.2)
+
+pgm.add_node("Rstar", r"$R_{\star}$", 5.8, 4)
+pgm.add_node("Prot", r"$P_{\rm rot}$", 6.6, 4)
+
+pgm.add_node("v", r"$v$", 6.2, 3.2)
+
+pgm.add_node("cospsi", r"$\cos{\psi}$", 4.5, 1.9)
+
+pgm.add_node("vsini", r"$v\sin{i}$", 6.2, 2.4)
+pgm.add_node("obsvsini", r"$v\sin{i}$", 6.2, 1.4, observed=True)
+pgm.add_node("errvsini", r"$\sigma_{v\sin{i}}$", 7, 1.4, fixed=True)
+
+# Add in the edges.
+pgm.add_edge("lam", "coslam")
+pgm.add_edge("lam", "obslam")
+pgm.add_edge("errlam", "obslam")
+
+pgm.add_edge("cosi", "sini")
+
+pgm.add_edge("Rstar", "v")
+pgm.add_edge("Prot", "v")
+
+
+pgm.add_edge("sini", "cospsi")
+pgm.add_edge("coslam", "cospsi")
+
+pgm.add_edge("sini", "vsini")
[…]

### 2022-11-15 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 We build a hierarchical Bayesian framework to constrain exoplanets' true stellar obliquity ($\psi$) distribution.
-Using the sky-projected stellar obliquities ($\lambda$)  of a sample of 162 exoplanets, we infer their $\cos{\psi}$ distribution using a four-parameter Beta mixture model under the assumption of a uniform cosine of the stellar inclination ($\cos{i_\star}$) bounded between $-1$ and $1$.
+Using the sky-projected stellar obliquities ($\lambda$)  of a sample of 162 exoplanets, we infer their $\cos{\psi}$ distribution using a four-parameter Beta mixture model with the assumption of a uniform cosine of stellar inclination ($\cos{i_\star}$) distribution for each planet.
 We find a significant pile-up of $\cos{\psi}$ near 1, suggesting that the majority of exoplanetary systems are well aligned.
-For misaligned systems, we find $\cos{\psi}$ is uniformly distributed between $-0.75$ to $0.5$ and have a slight but insignificant preference to $\cos{\psi} = -0.25$, corresponding to $\psi = 104.5\degr$.
-A pile-up at $\cos{\psi} = 0$ is not found.
+For misaligned systems, we find $\cos{\psi}$ is uniformly distributed between $-0.75$ to $0.5$ with no significant pipe-up near $\cos{\psi} = 0$.
 \end{abstract}

### 2022-11-15 src/scripts/hbm_all.py
update hbm graphs

@@ -12,7 +12,7 @@ from matplotlib import rc
 rc('font', **{'family':'sans-serif'})
 rc('text', usetex=True)
 
-pgm = daft.PGM(observed_style="inner")
+pgm = daft.PGM(observed_style="outer", dpi=150, node_unit=1.1, node_ec='k')
 
 # Hierarchical parameters.
 pgm.add_node("w", r"$w$", 1.5, 2.7)
@@ -27,8 +27,8 @@ pgm.add_node("coslam", r"$\cos{\lambda}$", 4, 3.2)
 pgm.add_node("errlam", r"$\sigma_\lambda$", 2.2, 3.2, fixed=True)
 
 
-pgm.add_node("cosi", r"$\cos{i}$", 5, 4)
-pgm.add_node("sini", r"$\sin{i}$", 5, 3.2)
+pgm.add_node("cosi", r"$\cos{i_{\star}}$", 5, 4)
+pgm.add_node("sini", r"$\sin{i_{\star}}$", 5, 3.2)
 
 pgm.add_node("Rstar", r"$R_{\star}$", 5.8, 4)
 pgm.add_node("Prot", r"$P_{\rm rot}$", 6.6, 4)
@@ -37,9 +37,9 @@ pgm.add_node("v", r"$v$", 6.2, 3.2)
 
 pgm.add_node("cospsi", r"$\cos{\psi}$", 4.5, 1.9)
 
-pgm.add_node("vsini", r"$v\sin{i}$", 6.2, 2.4)
-pgm.add_node("obsvsini", r"$v\sin{i}$", 6.2, 1.6, observed=True)
-pgm.add_node("errvsini", r"$\sigma_{v\sin{i}}$", 7, 1.6, fixed=True)
+pgm.add_node("vsini", r"$v\sin{i_{\star}}$", 6.2, 2.4)
+pgm.add_node("obsvsini", r"$v\sin{i_{\star}}$", 6.2, 1.6, observed=True)
+pgm.add_node("errvsini", r"$\sigma_{v\sin{i_{\star}}}$", 7, 1.6, fixed=True)
 
 # Add in the edges.
 pgm.add_edge("lam", "coslam")
@@ -87,18 +87,18 @@ pgm.add_edge("lam_", "obslam_")
 pgm.add_edge("errlam_", "obslam_")
 
 
-pgm.add_node("sini_", r"$\sin{i}$", -2, 3.2)
[…]

### 2022-11-15 src/scripts/hbm_istar.py

@@ -12,14 +12,14 @@ from matplotlib import rc
 rc('font', **{'family':'sans-serif'})
 rc('text', usetex=True)
 
-pgm = daft.PGM(observed_style="inner")
+pgm = daft.PGM(observed_style="outer", dpi=150, node_unit=1.1, node_ec='k')
 
 # Hierarchical parameters.
 pgm.add_node("w", r"$w$", 1.5, 2.7)
 pgm.add_node("alpha", r"$\alpha$", 1.5, 2.15)
 pgm.add_node("beta", r"$\beta$", 1.5, 1.6)
 
-# Latent and observed variables.
+# Latent variable.
 pgm.add_node("lam", r"$\lambda$", 3, 4)
 pgm.add_node("obslam", r"$\lambda$", 3, 3.2, observed=True)
 
@@ -27,8 +27,8 @@ pgm.add_node("coslam", r"$\cos{\lambda}$", 4, 3.2)
 pgm.add_node("errlam", r"$\sigma_\lambda$", 2.2, 3.2, fixed=True)
 
 
-pgm.add_node("cosi", r"$\cos{i}$", 5, 4)
-pgm.add_node("sini", r"$\sin{i}$", 5, 3.2)
+pgm.add_node("cosi", r"$\cos{i_{\star}}$", 5, 4)
+pgm.add_node("sini", r"$\sin{i_{\star}}$", 5, 3.2)
 
 pgm.add_node("Rstar", r"$R_{\star}$", 5.8, 4)
 pgm.add_node("Prot", r"$P_{\rm rot}$", 6.6, 4)
@@ -37,9 +37,9 @@ pgm.add_node("v", r"$v$", 6.2, 3.2)
 
 pgm.add_node("cospsi", r"$\cos{\psi}$", 4.5, 1.9)
 
-pgm.add_node("vsini", r"$v\sin{i}$", 6.2, 2.4)
-pgm.add_node("obsvsini", r"$v\sin{i}$", 6.2, 1.4, observed=True)
-pgm.add_node("errvsini", r"$\sigma_{v\sin{i}}$", 7, 1.4, fixed=True)
+pgm.add_node("vsini", r"$v\sin{i_{\star}}$", 6.2, 2.4)
+pgm.add_node("obsvsini", r"$v\sin{i_{\star}}$", 6.2, 1.4, observed=True)
[…]

### 2022-11-15 src/scripts/hbm_all.py
fix latex issue

@@ -1,3 +1,7 @@
+import os
+from pathlib import Path
+os.environ["PATH"] += os.pathsep + str(Path.home() / "bin")
+
 import paths
 
 import sys

### 2022-11-18 src/tex/ms.tex
[showyourwork] overleaf sync

--- 2022-11-15 src/tex/ms.tex
@@
 
-\submitjournal{AAS Journals}
+\submitjournal{the AAS Journals}
 
@@
 \begin{abstract}
-We build a hierarchical Bayesian framework to constrain exoplanets' true stellar obliquity ($\psi$) distribution.
-Using the sky-projected stellar obliquities ($\lambda$)  of a sample of 162 exoplanets, we infer their $\cos{\psi}$ distribution using a four-parameter Beta mixture model under the assumption of a uniform cosine of the stellar inclination ($\cos{i_\star}$) bounded between $-1$ and $1$.
-We find a significant pile-up of $\cos{\psi}$ near 1, suggesting that the majority of exoplanetary systems are well aligned.
-For misaligned systems, we find $\cos{\psi}$ is uniformly distributed between $-0.75$ to $0.5$ and have a slight but insignificant preference to $\cos{\psi} = -0.25$, corresponding to $\psi = 104.5\degr$.
-A pile-up at $\cos{\psi} = 0$ is not found.
+We build a hierarchical Bayesian framework to constrain true stellar obliquity ($\psi$) distribution in exoplanetary systems.
+Using the sky-projected stellar obliquities ($\lambda$) of a sample of 162 exoplanets, we infer their $\cos{\psi}$ distribution using a four-parameter Beta mixture model with the assumption of a uniform cosine of stellar inclination ($i_\star$) distribution for each planet.
+We find $\cos{\psi}$ has a significant pile-up near 1, suggesting the majority of exoplanetary systems are well aligned.
+For misaligned systems, $\cos{\psi}$ is nearly uniformly distributed between $-0.75$ to $0.75$ with no significant pipe-up near $\cos{\psi} = 0$, suggesting the stellar spin axis is randomly distributed around the planet's orbital axis.
+Our Bayesian framework could reproduce the excess of perpendicular planets observed in systems with constrained $i_\star$.
+However, we find the sample of misaligned systems is limited by small number statistics and likely biased because misaligned systems are often found in hot-star systems, whereas the gravity-darkening method, a technique commonly used to constrain the inclinations of hot stars ($T_{\rm eff} > 7000$K), is only sensitive to detect perpendicular planets.
 \end{abstract}
@@
 Stellar obliquity $\psi$ is described as the angle between a planet's orbital axis $\hat{i}_{\rm orb}$ and its host star's spin axis $\hat{i}_{\star}$.
-If both $\hat{i}_{\rm orb}$ and $\hat{i}_{\star}$ are random unit vectors with uniform distribution on a three-dimensional sphere, $\cos{\psi}$ will be uniformly distributed between $-1$ and 1.
+If both $\hat{i}_{\rm orb}$ and $\hat{i}_{\star}$ are independent, random unit vectors with uniform distribution on a three-dimensional sphere, $\cos{\psi}$ will be uniformly distributed between $-1$ and $1$.
+In reality, the sky-projected stellar obliquity $\lambda$ can only be measured in transiting exoplanet systems, limiting $\hat{i}_{\rm orb}$ close to $\sim 90\degr$.
+
+Stellar obliquity $\psi$ can be calculated from $\hat{i}_{\rm orb}$ and $\hat{i}_{\star}$ as the following:
+\begin{equation}
+    \cos{\psi} = \sin{i_\star}\sin{i_{\rm orb}}\cos{\lambda} + \cos{i_\star}\cos{i_{\rm orb}},
+\end{equation}
+where $i_{\rm orb}$ and $i_\star$ are the projections of $\hat{i}_{\rm orb}$ and $\hat{i}_{\star}$ along the observer's line of sight, and $\lambda$ is the sky-projected stellar obliquity.
+When $i_{\rm orb} = 90\degr$, $\cos{\psi} = \sin{i_\star}\cos{\lambda}$.
+
+In most exoplanetary systems, only the sky-projected obliquity $\lambda$ is constrained.
+We could then infer $\psi$ given the measured $\lambda$ and assuming a randomly distributed $i_\star$ (uniformly distributed in $\cos{i_\star}$).
+In some systems, $i_\star$ can be constrained via photometric rotational modulation, gravitational darkening or asteroseismology.
+In such cases, an informative $i_\star$ prior can be used.
+See Figure\ref{hbm} as an illustration.
+
+Using a sample of stars with $i_\star$, Albrecht+21 found the majority of planets are aligned.
+For misaligned planets, their stellar obliquities are preferentially distributed between 80--125$\degr$.
+Furthermore, they found the $\cos{\psi}$ distribution of misaligned systems is not uniformly distributed but clustered near 0.
+
+In this work, we build a Bayesian framework to evaluate the distribution of $\cos{\psi}$ and extend the analysis to the full sample exoplanets that have sky-projected stellar obliquity measurements.
+
+\section{Stellar Obliquity Distribution} \label{sec:hbm}
+We model the whole population stellar obliquity distribution and also the sub-population with $i_\star$ constrained. See Figure\ref{hbm} as an illustration.
 
 \begin{figure}[ht!]
-    \script{hbm_istar.py}
+    \script{hbm.py}
     \begin{centering}
-        \includegraphics[width=\linewidth]{figures/hbm_istar.pdf}
-        \caption{Plot showing a bunch of random numbers.}
-        \label{fig:hbm_istar}
+        \includegraphics[width=\linewidth]{figures/hbm.pdf}
+        \caption{Hierarchical Bayesian model to infer \emph{true} stellar obliquity for planets with $i_{\star}$ inferred from their rotation periods.}
+        \label{fig:hbm}
     \end{centering}
@@
 
-\begin{figure}
-    \script{hbm_all.py}
+We use a four-parameter Beta distribution. The framework is shown in Figure~\ref{fig:hbm}.
+
+\begin{itemize}
+    \item Uniform $\cos{i}$
+    \item Uniform $\lambda$
+    \item Observed Normal $\lambda$
+    \item Observed Normal $R_\star$
+    \item Observed Normal $P_{\rm rot}$
+\end{itemize}
+
+We find the $i_\star$ sample, there is a pile-up of planets near $-0.25$ whereas, for all planet sample, the distribution is mostly flat. See Figure 2.
+
+What makes the difference?
+
+\begin{figure*}[ht!]
+    \script{dist.py}
     \begin{centering}
-        \includegraphics[width=\linewidth]{figures/hbm_all.pdf}
-        \caption{Plot showing a bunch of random numbers.}
-        \label{fig:hbm_all}
+        \includegraphics[width=\linewidth]{figures/dist.pdf}
+        \caption{$\cos{\psi}$ distributions for the sample with $i_\star$ measured and all planets.}
+        \label{fig:dist}
     \end{centering}
-\end{figure}
+\end{figure*}
+
+\section{Hypothesis} \label{sec:tests}
+
+\subsection{Stellar Teff $>$ 7000K?}
+All planets with host star effective temperatures greater than 7000 K are near polar.
+Note that all these stars are expected to be fast rotators that limit the Rossiter-McLaughlin effect measurement.
+Doppler Tomography can be applied to these stars, but the majority of the measurements are conducted by gravitational darkening that is biased toward polar planets.
+
+We removed 7/21 planets with host star effective temperatures greater than 7000 K and redo the analysis. Interestingly, we find the results do not change. See Figure~\ref{fig:teteff_cutff}.
+
+\begin{figure*}[ht!]
+    \script{teff_cut.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/teff_cut.pdf}
+        \caption{$\cos{\psi}$ distributions for the $i_\star$ sample including/not including $>$7000 K stars.}
+        \label{fig:teff_cut}
+    \end{centering}
+\end{figure*}
+
+\subsection{Small vsini}
+Albrecht et al.
+21 that underestimating $v\sin{i_\star}$, overestimating $R_\star$, or underestimating $P_{\rm rot}$ could all lead to biased estimation on $\cos{\psi}$.
+We performed another test to evaluate the sample with random $i_\star$ prior (i.e., uniform on $\cos{i}$).
+Interestingly, we find a very similar distribution to the $\cos{\psi}$ distributions inferred with informative $i_\star$ priors.
+See Figure~\ref{fig:rand_inc}.
+
+\begin{figure*}[ht!]
+    \script{rand_inc.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/rand_inc.pdf}
+        \caption{$\cos{\psi}$ distributions for the $i_\star$ sample with/without informative $i_\star$ priors.}
+        \label{fig:rand_inc}
+    \end{centering}
+\end{figure*}
+
+\subsection{Statistical fluke}
+Could it just be a statistical fluke?
+We perform the test by randomly drawing 20 planets out of the 40 misaligned systems.
+We find in half of the time, we could not reproduce the distribution but result in a pile-up of $\cos{\psi}$.
+See Figure~\ref{fig:bootstrap}.
+
+\begin{figure*}[ht!]
+    \script{bootstrap.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/bootstrap.pdf}
+        \caption{$\cos{\psi}$ distributions for the $i_\star$ sample with/without informative $i_\star$ priors.}
+        \label{fig:bootstrap}
+    \end{centering}
+\end{figure*}
+
+\section{Discussion}
+No strong clustering?
 

### 2022-11-18 src/tex/ms.tex
[showyourwork] overleaf sync

@@
         \includegraphics[width=\linewidth]{figures/hbm.pdf}
-        \caption{Hierarchical Bayesian model to infer \emph{true} stellar obliquity for planets with $i_{\star}$ inferred from their rotation periods.}
+        \caption{Hierarchical Bayesian framework to infer stellar obliquity distribution described by hyperparameters $\theta$ from individual system's stellar obliquity $\psi_n$.
+$\psi_n$ is calculated by the system's sky-projected stellar obliquity $\lambda_n$ and stellar inclination $i_{\star, n}$.}
         \label{fig:hbm}
@@
 
-\begin{figure*}[ht!]
-    \script{dist.py}
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/dist.pdf}
-        \caption{$\cos{\psi}$ distributions for the sample with $i_\star$ measured and all planets.}
-        \label{fig:dist}
-    \end{centering}
-\end{figure*}
+% \begin{figure*}[ht!]
+%     \script{dist.py}
+%     \begin{centering}
+%         \includegraphics[width=\linewidth]{figures/dist.pdf}
+%         \caption{$\cos{\psi}$ distributions for the sample with $i_\star$ measured and all planets.}
+%         \label{fig:dist}
+%     \end{centering}
+% \end{figure*}
 
@@
 
-\begin{figure*}[ht!]
-    \script{teff_cut.py}
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/teff_cut.pdf}
-        \caption{$\cos{\psi}$ distributions for the $i_\star$ sample including/not including $>$7000 K stars.}
-        \label{fig:teff_cut}
-    \end{centering}
-\end{figure*}
+% \begin{figure*}[ht!]
+%     \script{teff_cut.py}
+%     \begin{centering}
+%         \includegraphics[width=\linewidth]{figures/teff_cut.pdf}
+%         \caption{$\cos{\psi}$ distributions for the $i_\star$ sample including/not including $>$7000 K stars.}
+%         \label{fig:teff_cut}
+%     \end{centering}
+% \end{figure*}
 
@@
 
-\begin{figure*}[ht!]
-    \script{rand_inc.py}
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/rand_inc.pdf}
-        \caption{$\cos{\psi}$ distributions for the $i_\star$ sample with/without informative $i_\star$ priors.}
-        \label{fig:rand_inc}
-    \end{centering}
-\end{figure*}
+% \begin{figure*}[ht!]
+%     \script{rand_inc.py}
+%     \begin{centering}
+%         \includegraphics[width=\linewidth]{figures/rand_inc.pdf}
+%         \caption{$\cos{\psi}$ distributions for the $i_\star$ sample with/without informative $i_\star$ priors.}
+%         \label{fig:rand_inc}
+%     \end{centering}
+% \end{figure*}
 
@@
 
-\begin{figure*}[ht!]
-    \script{bootstrap.py}
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/bootstrap.pdf}
-        \caption{$\cos{\psi}$ distributions for the $i_\star$ sample with/without informative $i_\star$ priors.}
-        \label{fig:bootstrap}
-    \end{centering}
-\end{figure*}
+% \begin{figure*}[ht!]
+%     \script{bootstrap.py}
+%     \begin{centering}
+%         \includegraphics[width=\linewidth]{figures/bootstrap.pdf}
+%         \caption{$\cos{\psi}$ distributions for the $i_\star$ sample with/without informative $i_\star$ priors.}
+%         \label{fig:bootstrap}
+%     \end{centering}
+% \end{figure*}
 

### 2022-11-18 src/scripts/hbm.py
new hbm graph

@@ -1,0 +1,62 @@
+import os
+from pathlib import Path
+os.environ["PATH"] += os.pathsep + str(Path.home() / "bin")
+
+import paths
+
+import sys
+import subprocess
+
+# implement pip as a subprocess:
+subprocess.check_call([sys.executable, '-m', 'pip', 'install',
+'daft'])
+import daft
+
+from matplotlib import rc
+rc('font', **{'family':'sans-serif'})
+rc('text', usetex=True)
+
+pgm = daft.PGM(observed_style="outer",node_unit=1, node_ec='k',dpi=150,line_width=0.8)
+               
+# Hierarchical parameters.
+pgm.add_node("beta", r"$\beta$", 2, 2.3)
+
+# Latent variable.
+pgm.add_node("psi", r"$\psi_n$", 2, 1.5)
+
+pgm.add_node("lambda", r"$\lambda_n$", 1., 1.5)
+
+pgm.add_node("istar", r"$i_{\star,n}$", 3., 1.5)
+
+pgm.add_node("pstar", r"$\theta_{\star,n}$", 3.6, 1.5)
+
+
+pgm.add_edge("beta", "psi")
+
+pgm.add_edge("lambda", "psi")
+
+pgm.add_edge("istar", "psi")
+
+
+pgm.add_node("obs_lambda", r"$\hat{\lambda}_n$", 1., 0.7, observed=True)
+
+pgm.add_node("obs_vsini", r"$\hat{v}\sin{i}_{\star, n}$", 3., 0.7, observed=True, aspect=1.3, fontsize=8)
+
+pgm.add_edge("lambda", "obs_lambda")
+
+pgm.add_edge("pstar", "obs_vsini", plot_params=dict(ls=(0, (1, 2)),head_width=0,head_length=0))
+pgm.add_edge("pstar", "obs_vsini", plot_params={'ls':''})
+
+pgm.add_edge("istar", "obs_vsini", plot_params=dict(ls=(0, (1, 2)),head_width=0,head_length=0))
+pgm.add_edge("istar", "obs_vsini", plot_params={'ls':''})
+
+pgm.add_edge("lambda", "obs_vsini", plot_params=dict(ls=(0, (1, 2)),head_width=0,head_length=0))
+pgm.add_edge("lambda", "obs_vsini", plot_params={'ls':''})
+
+# And a plate.
[…]

### 2022-11-18 src/scripts/hbm_all.py

@@ -1,122 +1,0 @@
-import os
-from pathlib import Path
-os.environ["PATH"] += os.pathsep + str(Path.home() / "bin")
-
-import paths
-
-import sys
-import subprocess
-
-# implement pip as a subprocess:
-subprocess.check_call([sys.executable, '-m', 'pip', 'install',
-'daft'])
-import daft
-
-from matplotlib import rc
-rc('font', **{'family':'sans-serif'})
-rc('text', usetex=True)
-
-pgm = daft.PGM(observed_style="outer", dpi=150, node_unit=1.1, node_ec='k')
-
-# Hierarchical parameters.
-pgm.add_node("w", r"$w$", 1.5, 2.7)
-pgm.add_node("alpha", r"$\alpha$", 1.5, 2.15)
-pgm.add_node("beta", r"$\beta$", 1.5, 1.6)
-
-# Latent variable.
-pgm.add_node("lam", r"$\lambda$", 3, 4)
-pgm.add_node("obslam", r"$\lambda$", 3, 3.2, observed=True)
-
-pgm.add_node("coslam", r"$\cos{\lambda}$", 4, 3.2)
-pgm.add_node("errlam", r"$\sigma_\lambda$", 2.2, 3.2, fixed=True)
-
-
-pgm.add_node("cosi", r"$\cos{i_{\star}}$", 5, 4)
-pgm.add_node("sini", r"$\sin{i_{\star}}$", 5, 3.2)
-
-pgm.add_node("Rstar", r"$R_{\star}$", 5.8, 4)
-pgm.add_node("Prot", r"$P_{\rm rot}$", 6.6, 4)
-
-pgm.add_node("v", r"$v$", 6.2, 3.2)
-
-pgm.add_node("cospsi", r"$\cos{\psi}$", 4.5, 1.9)
-
-pgm.add_node("vsini", r"$v\sin{i_{\star}}$", 6.2, 2.4)
-pgm.add_node("obsvsini", r"$v\sin{i_{\star}}$", 6.2, 1.6, observed=True)
-pgm.add_node("errvsini", r"$\sigma_{v\sin{i_{\star}}}$", 7, 1.6, fixed=True)
-
-# Add in the edges.
-pgm.add_edge("lam", "coslam")
-pgm.add_edge("lam", "obslam")
-pgm.add_edge("errlam", "obslam")
-
[…]

### 2022-11-18 src/scripts/hbm_istar.py

@@ -1,80 +1,0 @@
-import os
-from pathlib import Path
-os.environ["PATH"] += os.pathsep + str(Path.home() / "bin")
-
-import paths
-
-import sys
-import subprocess
-
-# implement pip as a subprocess:
-subprocess.check_call([sys.executable, '-m', 'pip', 'install',
-'daft'])
-import daft
-
-from matplotlib import rc
-rc('font', **{'family':'sans-serif'})
-rc('text', usetex=True)
-
-pgm = daft.PGM(observed_style="outer", dpi=150, node_unit=1.1, node_ec='k')
-
-# Hierarchical parameters.
-pgm.add_node("w", r"$w$", 1.5, 2.7)
-pgm.add_node("alpha", r"$\alpha$", 1.5, 2.15)
-pgm.add_node("beta", r"$\beta$", 1.5, 1.6)
-
-# Latent variable.
-pgm.add_node("lam", r"$\lambda$", 3, 4)
-pgm.add_node("obslam", r"$\lambda$", 3, 3.2, observed=True)
-
-pgm.add_node("coslam", r"$\cos{\lambda}$", 4, 3.2)
-pgm.add_node("errlam", r"$\sigma_\lambda$", 2.2, 3.2, fixed=True)
-
-
-pgm.add_node("cosi", r"$\cos{i_{\star}}$", 5, 4)
-pgm.add_node("sini", r"$\sin{i_{\star}}$", 5, 3.2)
-
-pgm.add_node("Rstar", r"$R_{\star}$", 5.8, 4)
-pgm.add_node("Prot", r"$P_{\rm rot}$", 6.6, 4)
-
-pgm.add_node("v", r"$v$", 6.2, 3.2)
-
-pgm.add_node("cospsi", r"$\cos{\psi}$", 4.5, 1.9)
-
-pgm.add_node("vsini", r"$v\sin{i_{\star}}$", 6.2, 2.4)
-pgm.add_node("obsvsini", r"$v\sin{i_{\star}}$", 6.2, 1.4, observed=True)
-pgm.add_node("errvsini", r"$\sigma_{v\sin{i_{\star}}}$", 7, 1.4, fixed=True)
-
-# Add in the edges.
-pgm.add_edge("lam", "coslam")
-pgm.add_edge("lam", "obslam")
-pgm.add_edge("errlam", "obslam")
-
-pgm.add_edge("cosi", "sini")
-
[…]

### 2022-11-19 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 We build a hierarchical Bayesian framework to constrain true stellar obliquity ($\psi$) distribution in exoplanetary systems.
-Using the sky-projected stellar obliquities ($\lambda$) of a sample of 162 exoplanets, we infer their $\cos{\psi}$ distribution using a four-parameter Beta mixture model with the assumption of a uniform cosine of stellar inclination ($i_\star$) distribution for each planet.
+Using the sky-projected stellar obliquities ($\lambda$) of a sample of 151 exoplanets, we infer their $\cos{\psi}$ distribution using a four-parameter Beta mixture model with the assumption of a uniform cosine of stellar inclination ($i_\star$) distribution for each planet.
 We find $\cos{\psi}$ has a significant pile-up near 1, suggesting the majority of exoplanetary systems are well aligned.
@@
 Our Bayesian framework could reproduce the excess of perpendicular planets observed in systems with constrained $i_\star$.
-However, we find the sample of misaligned systems is limited by small number statistics and likely biased because misaligned systems are often found in hot-star systems, whereas the gravity-darkening method, a technique commonly used to constrain the inclinations of hot stars ($T_{\rm eff} > 7000$K), is only sensitive to detect perpendicular planets.
+However, we find the sample of misaligned systems could be biased because of the small sample size and that misaligned systems are often found around hot stars, whereas the gravity-darkening method, a technique commonly used to constrain the inclinations of hot stars ($T_{\rm eff} > 7000$K), is only sensitive to detect perpendicular planets.
 \end{abstract}
+
+We build a hierarchical Bayesian framework to constrain true stellar obliquity ($\psi$) distribution in exoplanetary systems.
 
@@
 
-Stellar obliquity $\psi$ is described as the angle between a planet's orbital axis $\hat{i}_{\rm orb}$ and its host star's spin axis $\hat{i}_{\star}$.
-If both $\hat{i}_{\rm orb}$ and $\hat{i}_{\star}$ are independent, random unit vectors with uniform distribution on a three-dimensional sphere, $\cos{\psi}$ will be uniformly distributed between $-1$ and $1$.
-In reality, the sky-projected stellar obliquity $\lambda$ can only be measured in transiting exoplanet systems, limiting $\hat{i}_{\rm orb}$ close to $\sim 90\degr$.
+Stellar obliquity $\psi$ is described as the angle between a planet's orbital axis $\bf{i}_{\rm orb}$ and its host star's spin axis $\bf{i}_{\star}$.
+If both $\bf{i}_{\rm orb}$ and $\bf{i}_{\star}$ are independent, random unit vectors with uniform distribution on a three-dimensional sphere, $\cos{\psi}$ will be uniformly distributed between $-1$ and $1$.
+In reality, the sky-projected stellar obliquity $\lambda$ can only be measured in transiting exoplanet systems, limiting $\bf{i}_{\rm orb}$ close to $\sim 90\degr$.
 
-Stellar obliquity $\psi$ can be calculated from $\hat{i}_{\rm orb}$ and $\hat{i}_{\star}$ as the following:
+Stellar obliquity $\psi$ can be calculated from $\bf{i}_{\rm orb}$ and $\bf{i}_{\star}$ as the following:
 \begin{equation}
@@
 \end{equation}
-where $i_{\rm orb}$ and $i_\star$ are the projections of $\hat{i}_{\rm orb}$ and $\hat{i}_{\star}$ along the observer's line of sight, and $\lambda$ is the sky-projected stellar obliquity.
-When $i_{\rm orb} = 90\degr$, $\cos{\psi} = \sin{i_\star}\cos{\lambda}$.
+where $i_{\rm orb}$ and $i_\star$ are the projections of $\bf{i}_{\rm orb}$ and $\bf{i}_{\star}$ along the observer's line of sight, and $\lambda$ is the sky-projected stellar obliquity.
+A full derivation can be found in, e.g., \cite{Fabrycky09} Equation (1)--(10).
+When setting $i_{\rm orb} = 90\degr$, $\cos{\psi} = \sin{i_\star}\cos{\lambda}$.
 
 In most exoplanetary systems, only the sky-projected obliquity $\lambda$ is constrained.
-We could then infer $\psi$ given the measured $\lambda$ and assuming a randomly distributed $i_\star$ (uniformly distributed in $\cos{i_\star}$).
-In some systems, $i_\star$ can be constrained via photometric rotational modulation, gravitational darkening or asteroseismology.
-In such cases, an informative $i_\star$ prior can be used.
-See Figure\ref{hbm} as an illustration.
+We could still infer $\psi$ given the measured $\lambda$ and assuming a randomly distributed $i_\star$ that is uniformly distributed in $\cos{i_\star}$.
+In some systems, $i_\star$ can be constrained via photometric rotational modulation, gravity darkening, or asteroseismology. In such systems, an informative $i_\star$ prior can be used.
 
-Using a sample of stars with $i_\star$, Albrecht+21 found the majority of planets are aligned.
-For misaligned planets, their stellar obliquities are preferentially distributed between 80--125$\degr$.
-Furthermore, they found the $\cos{\psi}$ distribution of misaligned systems is not uniformly distributed but clustered near 0.
+\cite{Albrecht21} inferred the true stellar obliquities of a sample of stars that have both stellar inclinations and sky-projected stellar obliquity constraints.
+They found while the majority of planets have a small $\psi$, misaligned planets have $\psi$ clustered between 80--125$\degr$.
+Furthermore, they found the $\cos{\psi}$ distribution of misaligned systems is not uniformly distributed but clustered near 0, suggesting the stellar spin axis is not randomly distributed around the planetary orbital axis.
 
-In this work, we build a Bayesian framework to evaluate the distribution of $\cos{\psi}$ and extend the analysis to the full sample exoplanets that have sky-projected stellar obliquity measurements.
+In this work, we build a hierarchical Bayesian framework to evaluate the distribution of $\cos{\psi}$ using the sample in \cite{Albrecht21} and extend the analysis to the full sample of exoplanetary systems that have sky-projected stellar obliquity measurements extracted from the review paper of \cite{Albrecht22}.
 
 \section{Stellar Obliquity Distribution} \label{sec:hbm}
-We model the whole population stellar obliquity distribution and also the sub-population with $i_\star$ constrained. See Figure\ref{hbm} as an illustration.
+We model the whole population stellar obliquity distribution and also the sub-population with $i_\star$ constrained. See Figure\ref{fig:hbm} as an illustration.
 
@@
     \script{hbm.py}
-    \begin{centering}
     \includegraphics[width=\linewidth]{figures/hbm.pdf}
@@
     \label{fig:hbm}
-    \end{centering}
 \end{figure}
@@
 
-We find the $i_\star$ sample, there is a pile-up of planets near $-0.25$ whereas, for all planet sample, the distribution is mostly flat. See Figure 2.
+We find the $i_\star$ sample, there is a pile-up of planets near $-0.25$ whereas, for the all-planet sample, the distribution is mostly flat. See Figure~\ref{fig:psi_dist}.
 
@@
 
-% \begin{figure*}[ht!]
-%     \script{dist.py}
-%     \begin{centering}
-%         \includegraphics[width=\linewidth]{figures/dist.pdf}
-%         \caption{$\cos{\psi}$ distributions for the sample with $i_\star$ measured and all planets.}
-%         \label{fig:dist}
-%     \end{centering}
-% \end{figure*}
+\cite{Nagpal22} emphasize the importance of picking the right hyperpriors.
+
+\begin{figure*}[ht!]
+    \script{psi_dist.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/psi_dist.pdf}
+        \caption{$\cos{\psi}$ distributions.}
+        \label{fig:psi_dist}
+    \end{centering}
+\end{figure*}
 

### 2022-11-19 src/scripts/psi_dist.py
add psi distribution plot

@@ -1,0 +1,100 @@
+import os
+from pathlib import Path
+os.environ["PATH"] += os.pathsep + str(Path.home() / "bin")
+
+import sys
+import subprocess
+
+# implement pip as a subprocess:
+subprocess.check_call([sys.executable, '-m', 'pip', 'install',
+'arviz'])
+
+import paths
+import matplotlib.pyplot as plt
+import numpy as np
+import arviz as az
+from scipy.stats import beta
+
+from matplotlib import rc
+rc('font', **{'family':'sans-serif'})
+rc('text', usetex=True)
+
+plt.rcParams['xtick.top'] =  True
+plt.rcParams['xtick.direction'] =  'in'
+plt.rcParams['xtick.major.width'] =  1.0
+plt.rcParams['xtick.minor.width'] =  1.0
+plt.rcParams['ytick.right'] =  True
+plt.rcParams['ytick.direction'] =  'in'
+plt.rcParams['ytick.major.width'] = 1.0
+plt.rcParams['ytick.minor.width'] =  1.0
+plt.rcParams['lines.markeredgewidth'] =  1.0
+
+# load MCMC draws
+all_randinc = az.from_netcdf('all_randinc.nc')
+istar_randinc = az.from_netcdf('istar_randinc.nc')
+istar = az.from_netcdf('istar.nc')
+
+post = istar.posterior
+istar_draws = np.zeros(shape=(len(x),4000))
+for a in range(4):
+    for b in range(1000):
+        istar_draws[:, a*1000+b] = (post.w[a,b,0].values*beta.pdf(x, post.a0[a,b], post.b0[a,b])
+                                    +post.w[a,b,1].values*beta.pdf(x, post.a1[a,b], post.b1[a,b]))
+
+post = istar_randinc.posterior
+istar_randinc_draws = np.zeros(shape=(len(x),4000))
+for a in range(4):
+    for b in range(1000):
+        istar_randinc_draws[:, a*1000+b] = (post.w[a,b,0].values*beta.pdf(x, post.a0[a,b], post.b0[a,b])
+                                            +post.w[a,b,1].values*beta.pdf(x, post.a1[a,b], post.b1[a,b]))
+
+post = all_randinc.posterior
+all_randinc_draws = np.zeros(shape=(len(x),4000))
[…]

### 2022-11-19 src/scripts/graph.py
renaming

@@ -59,4 +59,4 @@ pgm.add_plate([0.5, 0.2, 3.6, 1.75], label=r"$n = 1, \ldots, N$", shift=-0.1, fo
 # Render and save.
 pgm.render()
 
-pgm.savefig(paths.figures / "hbm.pdf", bbox_inches="tight", dpi=600)
+pgm.savefig(paths.figures / "graph.pdf", bbox_inches="tight", dpi=600)

### 2022-11-19 src/scripts/psi_dist.py

@@ -30,9 +30,11 @@ plt.rcParams['ytick.minor.width'] =  1.0
 plt.rcParams['lines.markeredgewidth'] =  1.0
 
 # load MCMC draws
-all_randinc = az.from_netcdf('all_randinc.nc')
-istar_randinc = az.from_netcdf('istar_randinc.nc')
-istar = az.from_netcdf('istar.nc')
+all_randinc = az.from_netcdf('src/data/all_randinc.nc')
+istar_randinc = az.from_netcdf('src/data/istar_randinc.nc')
+istar = az.from_netcdf('src/data/istar.nc')
+
+x = np.linspace(1e-5,1-1e-5,1000)
 
 post = istar.posterior
 istar_draws = np.zeros(shape=(len(x),4000))

### 2022-11-19 src/scripts/psi_dist.py
renaming

@@ -30,9 +30,9 @@ plt.rcParams['ytick.minor.width'] =  1.0
 plt.rcParams['lines.markeredgewidth'] =  1.0
 
 # load MCMC draws
-all_randinc = az.from_netcdf('src/data/all_randinc.nc')
-istar_randinc = az.from_netcdf('src/data/istar_randinc.nc')
-istar = az.from_netcdf('src/data/istar.nc')
+all_randinc = az.from_netcdf(paths.data / 'all_randinc.nc')
+istar_randinc = az.from_netcdf(paths.data / 'istar_randinc.nc')
+istar = az.from_netcdf(paths.data / 'istar.nc')
 
 x = np.linspace(1e-5,1-1e-5,1000)
 

### 2022-11-19 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \usepackage{showyourwork}
+\usepackage{xspace}
+
+\newcommand{\numistar}{65\xspace}
+\newcommand{\numall}{161\xspace}
 
@@
 We build a hierarchical Bayesian framework to constrain true stellar obliquity ($\psi$) distribution in exoplanetary systems.
-Using the sky-projected stellar obliquities ($\lambda$) of a sample of 151 exoplanets, we infer their $\cos{\psi}$ distribution using a four-parameter Beta mixture model with the assumption of a uniform cosine of stellar inclination ($i_\star$) distribution for each planet.
+Using the sky-projected stellar obliquities ($\lambda$) of all measured systems, we infer their $\cos{\psi}$ distribution using a four-parameter Beta mixture model with the assumption of a uniform cosine of stellar inclination ($i_\star$) distribution for each planet.
 We find $\cos{\psi}$ has a significant pile-up near 1, suggesting the majority of exoplanetary systems are well aligned.
@@
 \end{abstract}
-
-We build a hierarchical Bayesian framework to constrain true stellar obliquity ($\psi$) distribution in exoplanetary systems.
 
@@
 A full derivation can be found in, e.g., \cite{Fabrycky09} Equation (1)--(10).
-When setting $i_{\rm orb} = 90\degr$, $\cos{\psi} = \sin{i_\star}\cos{\lambda}$.
+For transiting planet systems, $i_{\rm orb} \approx 90\degr$, and thus $\cos{\psi} \approx \sin{i_\star}\cos{\lambda}$.
 
@@
 \section{Stellar Obliquity Distribution} \label{sec:hbm}
-We model the whole population stellar obliquity distribution and also the sub-population with $i_\star$ constrained. See Figure\ref{fig:hbm} as an illustration.
+We model the whole population stellar obliquity distribution and also the sub-population with $i_\star$ constrained. See Figure\ref{fig:graph} as an illustration.
 
 \begin{figure}[ht!]
-    \script{hbm.py}
-    \includegraphics[width=\linewidth]{figures/hbm.pdf}
+    \script{graph.py}
+    \includegraphics[width=\linewidth]{figures/graph.pdf}
     \caption{Hierarchical Bayesian framework to infer stellar obliquity distribution described by hyperparameters $\theta$ from individual system's stellar obliquity $\psi_n$.
 $\psi_n$ is calculated by the system's sky-projected stellar obliquity $\lambda_n$ and stellar inclination $i_{\star, n}$.}
-    \label{fig:hbm}
+    \label{fig:graph}
 \end{figure}
 
-We use a four-parameter Beta distribution. The framework is shown in Figure~\ref{fig:hbm}.
+We use a four-parameter Beta distribution.
 
@@
 
+Although the distribution of $\psi$ looks very similar for the $i_\star$ sample using constrained or random priors, the uncertainty on each $\psi$ is much tighter.
+
 \begin{figure*}[ht!]
@@
     \begin{centering}
-        \includegraphics[width=\linewidth]{figures/psi_dist.pdf}
+        \includegraphics{figures/psi_dist.pdf}
         \caption{$\cos{\psi}$ distributions.}
@@
 
+\begin{figure*}[ht!]
+    \script{lam_dist.py}
+    \begin{centering}
+        \includegraphics{figures/lam_dist.pdf}
+        \caption{Sky-projected stellar obliquity ($\lambda$) distributions of \numistar systems with constrained $i_\star$ and all \numall exoplanetary systems that have measured sky-projected stellar obliquities.}
+        \label{fig:lam_dist}
+    \end{centering}
+\end{figure*}
+
 \subsection{Stellar Teff $>$ 7000K?}
@@
 
-We removed 7/21 planets with host star effective temperatures greater than 7000 K and redo the analysis. Interestingly, we find the results do not change. See Figure~\ref{fig:teteff_cutff}.
+We removed 7/21 planets with host star effective temperatures greater than 7000 K and redo the analysis. Interestingly, we find the results do not change. See Figure~\ref{fig:teff_cutff}.
 

### 2022-11-21 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 % Define document class
-\documentclass[twocolumn]{aastex631}
+\documentclass[twocolumn,times]{aastex631}
 \usepackage{showyourwork}
 \usepackage{xspace}
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{bm}
 
@@
 % Title
-\title{True Stellar Obliquity Distribution of Exoplanets}
+\title{3D Stellar Obliquity Distribution of Exoplanetary Systems}
 
@@
 \begin{abstract}
-We build a hierarchical Bayesian framework to constrain true stellar obliquity ($\psi$) distribution in exoplanetary systems.
-Using the sky-projected stellar obliquities ($\lambda$) of all measured systems, we infer their $\cos{\psi}$ distribution using a four-parameter Beta mixture model with the assumption of a uniform cosine of stellar inclination ($i_\star$) distribution for each planet.
+% We build a hierarchical Bayesian framework to constrain true stellar obliquity ($\psi$) distribution in exoplanetary systems.
+% Using the sky-projected stellar obliquities ($\lambda$) of all measured systems, we infer their $\cos{\psi}$ distribution using a four-parameter Beta mixture model with the assumption of a uniform cosine of stellar inclination ($i_\star$) distribution for each planet.
 We find $\cos{\psi}$ has a significant pile-up near 1, suggesting the majority of exoplanetary systems are well aligned.
 For misaligned systems, $\cos{\psi}$ is nearly uniformly distributed between $-0.75$ to $0.75$ with no significant pipe-up near $\cos{\psi} = 0$, suggesting the stellar spin axis is randomly distributed around the planet's orbital axis.
-Our Bayesian framework could reproduce the excess of perpendicular planets observed in systems with constrained $i_\star$.
+% Our Bayesian framework could reproduce the excess of perpendicular planets observed in systems with constrained $i_\star$.
 However, we find the sample of misaligned systems could be biased because of the small sample size and that misaligned systems are often found around hot stars, whereas the gravity-darkening method, a technique commonly used to constrain the inclinations of hot stars ($T_{\rm eff} > 7000$K), is only sensitive to detect perpendicular planets.
+We build a hierarchical Bayesian framework to infer the 3D stellar obliquity $\psi$ distribution in exoplanetary systems. 
+First, we apply our framework to \numistar systems with sky-projected stellar obliquity $\lambda$ and stellar inclinations $i_\star$ measurements.
+The $\cos{\psi}$ distribution has a significant pile-up at 1, suggesting the majority of systems are well aligned, and has a second peak near $-0.15$ (i.e., $\psi \sim 100\degr$), consistent with the finding of a preponderance of perpendicular planets in previous work.
+Interestingly, if we just use isotropic priors on $i_\star$ and no additional information, a similar $\cos{\psi}$ distribution is inferred, suggesting the $cos\psi$ distribution depends mainly on the underlying $\lambda$ distribution.
+We find the $\lambda$ distribution of misaligned systems in the measured-$i_\star$ sample clustered at $110 \degr$, whereas of all \numall systems with sky-projected obliquity measurements is nearly uniform.
+The measured-$i_\star$ sample may not fully represent the population.
+We discuss two possible reasons: (1) the measured-$i_\star$ sample only has a small number of misaligned systems, and (2) misaligned systems are often found around hot stars, and to measure the $i_\star$ of host stars ($T_{\rm eff} > 7000$K), the gravity-darkening method is commonly used but the method is biased towards detecting perpendicular planets.
+Finally, we apply our Bayesian framework to all \numall systems and assume uninformative $i_\star$ priors.
+The $\cos{\psi}$ piles up at 1 but is nearly uniformly distributed between $-0.75$ to $0.75$, corresponding to $\psi$ of $140\degr$ and $40\degr$.
+We conclude that in misaligned exoplanetary systems, the stellar spin axis is isotropically distributed around its planets' orbital axis but lacks $\psi > 140\degr$.
 \end{abstract}
@@
 
-Stellar obliquity $\psi$ is described as the angle between a planet's orbital axis $\bf{i}_{\rm orb}$ and its host star's spin axis $\bf{i}_{\star}$.
-If both $\bf{i}_{\rm orb}$ and $\bf{i}_{\star}$ are independent, random unit vectors with uniform distribution on a three-dimensional sphere, $\cos{\psi}$ will be uniformly distributed between $-1$ and $1$.
-In reality, the sky-projected stellar obliquity $\lambda$ can only be measured in transiting exoplanet systems, limiting $\bf{i}_{\rm orb}$ close to $\sim 90\degr$.
-
+3D stellar obliquity $\psi$ is described as the angle between a planet's orbital axis $\bf{i}_{\rm orb}$ and its host star's spin axis $\bf{i}_{\star}$.
+For any given $\bf{i}_{\rm orb}$ axis, if $\bf{i}_{\star}$ is a random vector with uniform distribution on a three-dimensional sphere, the cosine of their separation angle $\cos{\psi}$ will be uniformly distributed between $-1$ and $1$.
+In reality, we could only measure the sky-projected stellar obliquity $\lambda$, the angle between the projections of $\bf{i}_{\rm orb}$ and $\bf{i}_{\star}$ onto the plane of the sky.
 Stellar obliquity $\psi$ can be calculated from $\bf{i}_{\rm orb}$ and $\bf{i}_{\star}$ as the following:
-\begin{equation}
+\begin{equation}\label{eqn:psi}
     \cos{\psi} = \sin{i_\star}\sin{i_{\rm orb}}\cos{\lambda} + \cos{i_\star}\cos{i_{\rm orb}},
 \end{equation}
-where $i_{\rm orb}$ and $i_\star$ are the projections of $\bf{i}_{\rm orb}$ and $\bf{i}_{\star}$ along the observer's line of sight, and $\lambda$ is the sky-projected stellar obliquity.
-A full derivation can be found in, e.g., \cite{Fabrycky09} Equation (1)--(10).
-For transiting planet systems, $i_{\rm orb} \approx 90\degr$, and thus $\cos{\psi} \approx \sin{i_\star}\cos{\lambda}$.
+where $i_{\rm orb}$ is the angle between the vector $\bf{i}_{\rm orb}$ and the observer's line of sight, $i_\star$ is the angle between $\bf{i}_{\star}$ and the observer's line of sight, and $\lambda$ is the sky-projected stellar obliquity.
+A full derivation of Equation~(\ref{eqn:psi}) can be found in, e.g., \cite{Fabrycky09} Equation (1)--(10).
+Since the sky-projected stellar obliquity can only be measured in transiting exoplanet systems, $i_{\rm orb} \approx 90\degr$ and thus $\cos{\psi} \approx \sin{i_\star}\cos{\lambda}$.
 
-In most exoplanetary systems, only the sky-projected obliquity $\lambda$ is constrained.
-We could still infer $\psi$ given the measured $\lambda$ and assuming a randomly distributed $i_\star$ that is uniformly distributed in $\cos{i_\star}$.
-In some systems, $i_\star$ can be constrained via photometric rotational modulation, gravity darkening, or asteroseismology. In such systems, an informative $i_\star$ prior can be used.
+To constrain the 3D stellar obliquity $\psi$, both a measurement of the sky-projected obliquity $\lambda$ and the stellar inclination $i_\star$ are required.
+The stellar inclination can often be constrained on exoplanetary systems via, e.g., photometric and spectroscopic rotational modulation introduced by starspots, gravity darkening, or asteroseismology.
+For systems without measured $i_\star$, it is still possible to infer the 3D stellar obliquity $\psi$ from the sky-projected obliquity $\lambda$ \citep[e.g.,][]{Fabrycky09}.
+To do so, we assume the stellar inclination $i_\star$ is isotropic, and thus the probability density function of ${i_\star} \sim \sin{i_{\star}}$, or $\cos{i_\star} \sim \mathcal{U}[0,1]$.
 
-\cite{Albrecht21} inferred the true stellar obliquities of a sample of stars that have both stellar inclinations and sky-projected stellar obliquity constraints.
+\cite{Albrecht21} inferred the 3D stellar obliquities of a sample of stars that have both stellar inclination and sky-projected stellar obliquity constraints.
 They found while the majority of planets have a small $\psi$, misaligned planets have $\psi$ clustered between 80--125$\degr$.
-Furthermore, they found the $\cos{\psi}$ distribution of misaligned systems is not uniformly distributed but clustered near 0, suggesting the stellar spin axis is not randomly distributed around the planetary orbital axis.
+Furthermore, they found the $\cos{\psi}$ distribution of misaligned systems is not uniformly distributed but clusters near 0, suggesting the stellar spin axis is not isotropically distributed around the planetary orbital axis.
 
-In this work, we build a hierarchical Bayesian framework to evaluate the distribution of $\cos{\psi}$ using the sample in \cite{Albrecht21} and extend the analysis to the full sample of exoplanetary systems that have sky-projected stellar obliquity measurements extracted from the review paper of \cite{Albrecht22}.
+In this work, we build a hierarchical Bayesian framework to evaluate the distribution of $\cos{\psi}$ using the sample in \cite{Albrecht21} and extend the analysis to the full sample of exoplanetary systems that have sky-projected stellar obliquity measurements extracted from the \cite{Albrecht22} review.
 
 \section{Stellar Obliquity Distribution} \label{sec:hbm}
-We model the whole population stellar obliquity distribution and also the sub-population with $i_\star$ constrained. See Figure\ref{fig:graph} as an illustration.
+We build a hierarchical Bayesian framework to model the 3D stellar obliquity ($\psi$) distribution of exoplanetary systems.
 
-\begin{figure}[ht!]
+First, we apply the framework to the sample of \numistar systems that have both sky-projected stellar obliquity $\lambda$ and stellar inclination $i_\star$ constraints.
+In Figure~\ref{fig:graph}, we illustrate the probabilistic graphical model.
+We aim to constrain a set of hyperparameters $\bm{\beta}$ that describe the 3D stellar obliquity $\psi$ distribution.
+$\bm{\beta}$ is constrained by $N$ individual systems, where each system has $\psi_n$ that is calculated by the sky-projected stellar obliquity $\lambda_n$ and the stellar inclination $i_n$.
+For simplicity, we approximate $i_{\rm orb}$ to $90\degr$ in the model and use Equation~(\ref{eqn:psi}) to compute $\psi_n$.
+The parameter $\theta_{\star, n}$ contains information on stellar rotation period and stellar radius, if available.
+If the rotational modulation method is used to constrain the stellar inclination \citep[e.g.,][]{Masuda20}, $i_{\star,n}$ and $\theta_{\star,n}$ are constrained by observed sky-projected stellar rotational line broadening $\hat{v}\sin{i_{\star,n}}$ and observed $\hat{\theta}_{\star,n}$.
+Lastly, the sky-projected stellar obliquity $\lambda_n$ is constrained by observed $\hat{\lambda}_n$.
+
+\begin{figure}
     \script{graph.py}
     \includegraphics[width=\linewidth]{figures/graph.pdf}
-    \caption{Hierarchical Bayesian framework to infer stellar obliquity distribution described by hyperparameters $\theta$ from individual system's stellar obliquity $\psi_n$.
-$\psi_n$ is calculated by the system's sky-projected stellar obliquity $\lambda_n$ and stellar inclination $i_{\star, n}$.}
+    \caption{Hierarchical Bayesian framework to infer the stellar obliquity distribution of exoplanetary systems.
+The stellar obliquity distribution is described by hyperparameters $\beta$ and constrained by $n_{\rm th}$ system's stellar obliquity $\psi_n$.
+Each $\psi_n$ is calculated by the system's sky-projected stellar obliquity $\lambda_n$ and stellar inclination $i_{\star, n}$.
+$\theta_{\star, n}$ contains other information of the star other than $i_{\star, n}$, such as the stellar radius and stellar rotation period.}
     \label{fig:graph}
@@
 
-We use a four-parameter Beta distribution.
+For the stellar obliquity distribution, we adopt a two-component mixture Beta distribution with hyperparameters $\bm{\theta} = \{\bm{w},\bm{a},\bm{b}\}$, where each hyperparameter has a dimension of 2.
+The Beta distribution is flexible to describe both the spike of aligned systems with 3D stellar obliquities close to $0\degr$ and the clustered or broadly distributed systems with large obliquities.
+The probability distribution follows 
+\begin{align}
+    \cos{\psi} &\sim 2\times\Bigl( w_0 {\rm Beta}(a_0, b_0) + w_1 {\rm Beta}(a_1, b_1)\Bigr)-1.
+\end{align}
+Since a Beta distribution is defined on the interval $[0, 1]$, we extend its support to $[-1,1]$ using a linear transformation (i.e., $2\times Beta-1$).
 
-\begin{itemize}
-    \item Uniform $\cos{i}$
-    \item Uniform $\lambda$
-    \item Observed Normal $\lambda$
-    \item Observed Normal $R_\star$
-    \item Observed Normal $P_{\rm rot}$
-\end{itemize}
+The hyperparameter and parameter priors are the following:
+\begin{align}
+    w_{0,1} &\sim {\rm Dirichlet}(1,1) \nonumber\\
+    a_0 &\sim \mathcal{U}(0,50) \nonumber\\
+    b_0 &\sim \mathcal{U}(0,1) \nonumber\\
+    a_1 &\sim \mathcal{U}(0,10) \nonumber\\
+    b_1 &\sim \mathcal{U}(0,10) \nonumber\\
+    \cos{i}_{\star,n} &\sim \mathcal{U}(0,1) \nonumber\\
+    \lambda_{\star,n} &\sim \mathcal{U}(0,\pi).
+\end{align}
+We set the hyperpriors such that $a_0$ and $b_0$ will be used to infer the spike of $\cos{\psi}$ near 1, and $a_1$ and $b_1$ infer the distribution of the misaligned population.
+Doing so will also avoid label switching in our mixture model.
 
-We find the $i_\star$ sample, there is a pile-up of planets near $-0.25$ whereas, for the all-planet sample, the distribution is mostly flat. See Figure~\ref{fig:psi_dist}.
-
-What makes the difference?
-
-\cite{Nagpal22} emphasize the importance of picking the right hyperpriors.
-
-Although the distribution of $\psi$ looks very similar for the $i_\star$ sample using constrained or random priors, the uncertainty on each $\psi$ is much tighter.
+The likelihood functions are the following:
+\begin{align}
+    L(\lambda) &\sim \prod_{i=1}^N\mathcal{N}(\hat{\lambda}_n, \sigma_{\hat{\lambda}_n}) \nonumber\\
+    L(i_\star, \theta_\star) &\sim \prod_{i=1}^N\mathcal{N}(\hat{v}\sin{i_{\star,n}}, \sigma_{\hat{v}\sin{i_{\star,n}}}) \nonumber\\
+    L(\theta) &\sim \prod_{i=1}^N\mathcal{N}(\hat{\theta}_n, \sigma_{\hat{\theta}_n}).
+\end{align}
 
@@
         \includegraphics{figures/psi_dist.pdf}
-        \caption{$\cos{\psi}$ distributions.}
+        \caption{The cosine of 3D stellar obliquity $\cos{\psi}$ distributions.}
         \label{fig:psi_dist}
@@
 \end{figure*}
+
+We apply the model to \numistar systems with both constrained $\lambda$ and $i_\star$ summarized in \cite{Albrecht21}.
+The inferred $\cos{\psi}$ distribution is shown in the left panel of Figure~\ref{fig:psi_dist}.
+The distribution presents a pike at 1 as expected, suggesting the majority of exoplanetary systems are well-aligned systems.
+For misaligned systems, $\cos{\psi}$ presents a cluster near $\cos{\psi} = -0.2$, corresponding to $\psi$ of $100\degr$.
+The result is consistent with the findings of a preponderance of perpendicular planets in \cite{Albrecht21}.
+
+To test the dependence of $\cos{\psi}$ distribution on $i_\star$ measurements, we apply the model again to the sample of \numistar systems with both constrained $\lambda$ and $i_\star$.
+However, in this model, we do not use any additional information on $i_\star$ other than applying an isotropic inclination prior to each system.
+The corresponding probabilistic graphical model is Figure~\ref{fig:graph} but removing all the dashed edges.
+The $\cos{\psi}$ distribution of the new model is shown in the middle panel of Figure~\ref{fig:psi_dist}.
+Interestingly, the overall $\cos{\psi}$ distributions of the two models are very similar, despite that the inferred $\psi$ uncertainty for each system is now greater.
+The similarity between the two distributions suggests that, at least for the measured-$i_\star$ sample, the $\cos{\psi}$ distribution mainly depends on the underlying sky-project stellar obliquity $\lambda$ distribution.
+
+Following the logic, we now apply our model to all \numall exoplanetary systems that have sky-projected stellar obliquity measurements summarized in \cite{Albrecht22}.
+The result is presented in the right panel of Figure~\ref{fig:psi_dist}.
+The $\cos{\psi}$ distribution still spikes at 1 but is nearly flat between $-0.75$ and $0.75$ with no significant clustering.
+The distribution corresponds to an isotropic distribution of $\psi$ from $40\degr$ to $140\degr$.
+
+We also test the robustness of $\cos{\psi}$ distribution on different hyperprior choices \citep[see a nice summary in][]{Nagpal22}.
+We find extending the support of hyperpriors from 0--10 to 0--100 or 0--1000 or changing hyperprior distributions from Uniform to log-Uniform or truncated-Normal all infer similar results to Figure~\ref{fig:psi_dist}.
+
+Although the distribution of $\psi$ looks very similar for the $i_\star$ sample using constrained or random priors, the uncertainty on each $\psi$ is much tighter.
 

### 2022-11-22 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \usepackage{bm}
+\usepackage{physics}
 
@@
 \begin{abstract}
-% We build a hierarchical Bayesian framework to constrain true stellar obliquity ($\psi$) distribution in exoplanetary systems.
-% Using the sky-projected stellar obliquities ($\lambda$) of all measured systems, we infer their $\cos{\psi}$ distribution using a four-parameter Beta mixture model with the assumption of a uniform cosine of stellar inclination ($i_\star$) distribution for each planet.
-We find $\cos{\psi}$ has a significant pile-up near 1, suggesting the majority of exoplanetary systems are well aligned.
-For misaligned systems, $\cos{\psi}$ is nearly uniformly distributed between $-0.75$ to $0.75$ with no significant pipe-up near $\cos{\psi} = 0$, suggesting the stellar spin axis is randomly distributed around the planet's orbital axis.
-% Our Bayesian framework could reproduce the excess of perpendicular planets observed in systems with constrained $i_\star$.
-However, we find the sample of misaligned systems could be biased because of the small sample size and that misaligned systems are often found around hot stars, whereas the gravity-darkening method, a technique commonly used to constrain the inclinations of hot stars ($T_{\rm eff} > 7000$K), is only sensitive to detect perpendicular planets.
 We build a hierarchical Bayesian framework to infer the 3D stellar obliquity $\psi$ distribution in exoplanetary systems. 
 First, we apply our framework to \numistar systems with sky-projected stellar obliquity $\lambda$ and stellar inclinations $i_\star$ measurements.
-The $\cos{\psi}$ distribution has a significant pile-up at 1, suggesting the majority of systems are well aligned, and has a second peak near $-0.15$ (i.e., $\psi \sim 100\degr$), consistent with the finding of a preponderance of perpendicular planets in previous work.
-Interestingly, if we just use isotropic priors on $i_\star$ and no additional information, a similar $\cos{\psi}$ distribution is inferred, suggesting the $cos\psi$ distribution depends mainly on the underlying $\lambda$ distribution.
-We find the $\lambda$ distribution of misaligned systems in the measured-$i_\star$ sample clustered at $110 \degr$, whereas of all \numall systems with sky-projected obliquity measurements is nearly uniform.
-The measured-$i_\star$ sample may not fully represent the population.
-We discuss two possible reasons: (1) the measured-$i_\star$ sample only has a small number of misaligned systems, and (2) misaligned systems are often found around hot stars, and to measure the $i_\star$ of host stars ($T_{\rm eff} > 7000$K), the gravity-darkening method is commonly used but the method is biased towards detecting perpendicular planets.
-Finally, we apply our Bayesian framework to all \numall systems and assume uninformative $i_\star$ priors.
-The $\cos{\psi}$ piles up at 1 but is nearly uniformly distributed between $-0.75$ to $0.75$, corresponding to $\psi$ of $140\degr$ and $40\degr$.
-We conclude that in misaligned exoplanetary systems, the stellar spin axis is isotropically distributed around its planets' orbital axis but lacks $\psi > 140\degr$.
+The $\cos{\psi}$ distribution has a spike at 1, suggesting the majority of systems are well aligned, and a concentration at $-0.15$ ($\psi \sim 100\degr$), consistent with the finding of a preponderance of perpendicular planets.
+Interestingly, if we assume isotropic priors on $i_\star$ and use no extra data, a similar $\cos{\psi}$ distribution is inferred, suggesting the $cos\psi$ distribution depends mainly on the underlying $\lambda$ distribution.
+We then apply the framework to all \numall systems with sky-projected obliquity measurements.
+Different from the observed-$i_\star$ sample, $\cos{\psi}$ is nearly uniformly distributed between $-0.75$ to $0.75$, corresponding to an isotropic distribution of $\psi$ between $40\degr$ and $140\degr$.
+To understand the $\cos{\psi}$-distribution discrepancy between two samples, we look at their $\lambda$ distributions.
+The measured-$i_\star$ sample has a concentration of $\lambda$ at $110\degr$, whereas the full sample's $\lambda$ is uniformly distributed from $30\degr$ to $180\degr$.
+The observed-$i_\star$ sample could be biased because misaligned systems are often found around hot stars; however, hot stars' inclinations are mostly measured by the gravity-darkening method, which is biased toward measuring perpendicular planets.
+Lastly, we apply the framework to constrain Hot Jupiters' 3D stellar obliquity distribution and find that in misaligned Hot Jupiter systems, the stellar spin axis is isotropically distributed around the planet's orbital axis.
 \end{abstract}
@@
 3D stellar obliquity $\psi$ is described as the angle between a planet's orbital axis $\bf{i}_{\rm orb}$ and its host star's spin axis $\bf{i}_{\star}$.
-For any given $\bf{i}_{\rm orb}$ axis, if $\bf{i}_{\star}$ is a random vector with uniform distribution on a three-dimensional sphere, the cosine of their separation angle $\cos{\psi}$ will be uniformly distributed between $-1$ and $1$.
-In reality, we could only measure the sky-projected stellar obliquity $\lambda$, the angle between the projections of $\bf{i}_{\rm orb}$ and $\bf{i}_{\star}$ onto the plane of the sky.
+For any given $\bf{i}_{\rm orb}$ axis, if $\bf{i}_{\star}$ is a random vector with uniform distribution on a three-dimensional sphere, the probability distribution $p_{\psi} \sim \sin{\psi}$ and thus $p_{\cos{\psi}} = p_{\psi} \abs{\dv*{\psi}{\cos{\psi}}}$ and is uniformly distributed.
+In reality, we only measure the sky-projected stellar obliquity $\lambda$, the angle between the projections of $\bf{i}_{\rm orb}$ and $\bf{i}_{\star}$ onto the plane of the sky.
 Stellar obliquity $\psi$ can be calculated from $\bf{i}_{\rm orb}$ and $\bf{i}_{\star}$ as the following:
@@
 \end{equation}
-where $i_{\rm orb}$ is the angle between the vector $\bf{i}_{\rm orb}$ and the observer's line of sight, $i_\star$ is the angle between $\bf{i}_{\star}$ and the observer's line of sight, and $\lambda$ is the sky-projected stellar obliquity.
-A full derivation of Equation~(\ref{eqn:psi}) can be found in, e.g., \cite{Fabrycky09} Equation (1)--(10).
-Since the sky-projected stellar obliquity can only be measured in transiting exoplanet systems, $i_{\rm orb} \approx 90\degr$ and thus $\cos{\psi} \approx \sin{i_\star}\cos{\lambda}$.
+where $i_{\rm orb}$ is the angle between the vector $\bf{i}_{\rm orb}$ and the observer's line of sight, and $i_\star$ is the angle between $\bf{i}_{\star}$ and the observer's line of sight.
+A full derivation of Equation~(\ref{eqn:psi}) can be found in, e.g., \cite{Fabrycky09}.
+Since the sky-projected stellar obliquities can only be measured in transiting exoplanet systems, $i_{\rm orb} \approx 90\degr$ and $\cos{\psi} \approx \sin{i_\star}\cos{\lambda}$.
 
 To constrain the 3D stellar obliquity $\psi$, both a measurement of the sky-projected obliquity $\lambda$ and the stellar inclination $i_\star$ are required.
-The stellar inclination can often be constrained on exoplanetary systems via, e.g., photometric and spectroscopic rotational modulation introduced by starspots, gravity darkening, or asteroseismology.
-For systems without measured $i_\star$, it is still possible to infer the 3D stellar obliquity $\psi$ from the sky-projected obliquity $\lambda$ \citep[e.g.,][]{Fabrycky09}.
-To do so, we assume the stellar inclination $i_\star$ is isotropic, and thus the probability density function of ${i_\star} \sim \sin{i_{\star}}$, or $\cos{i_\star} \sim \mathcal{U}[0,1]$.
+The stellar inclination can often be constrained on exoplanetary systems via, e.g., photometric and spectroscopic rotational modulation introduced by starspots, gravity darkening, and asteroseismology.
+For systems without measured $i_\star$, it is still possible to infer the 3D stellar obliquity $\psi$ from the sky-projected obliquity $\lambda$ assuming the stellar inclination $i_\star$ is isotropically distributed \citep[e.g.,][]{Fabrycky09}.
+The probability distribution follows $p_{i_\star} \sim \sin{i_{\star}}$, or $p_{\cos{i_\star}} \sim \mathcal{U}$.
 
@@
 They found while the majority of planets have a small $\psi$, misaligned planets have $\psi$ clustered between 80--125$\degr$.
-Furthermore, they found the $\cos{\psi}$ distribution of misaligned systems is not uniformly distributed but clusters near 0, suggesting the stellar spin axis is not isotropically distributed around the planetary orbital axis.
+Furthermore, the $\cos{\psi}$ distribution of misaligned systems is not uniformly distributed but clusters near 0, suggesting the stellar spin axis is not isotropically distributed around the planetary orbital axis.
 
 In this work, we build a hierarchical Bayesian framework to evaluate the distribution of $\cos{\psi}$ using the sample in \cite{Albrecht21} and extend the analysis to the full sample of exoplanetary systems that have sky-projected stellar obliquity measurements extracted from the \cite{Albrecht22} review.
+In Section~\ref{sec:hbm}, we describe our hierarchical Bayesian framework and the inferred 3D stellar obliquity distributions, and in Section~\ref{sec:interpretations}, we discuss the interpretations of these distributions.
+Lastly, in Section~\ref{sec:hjs}, we apply the framework to Hot Jupiters and infer their 3D stellar obliquity distribution.
 
-\section{Stellar Obliquity Distribution} \label{sec:hbm}
+\section{3D Stellar Obliquity Distributions} \label{sec:hbm}
 We build a hierarchical Bayesian framework to model the 3D stellar obliquity ($\psi$) distribution of exoplanetary systems.
@@
 
-For the stellar obliquity distribution, we adopt a two-component mixture Beta distribution with hyperparameters $\bm{\theta} = \{\bm{w},\bm{a},\bm{b}\}$, where each hyperparameter has a dimension of 2.
+For the stellar obliquity distribution, we adopt a two-component mixture Beta distribution with hyperparameters $\bm{\beta} = \{\bm{w},\bm{a},\bm{b}\}$, where each hyperparameter has a dimension of 2.
 The Beta distribution is flexible to describe both the spike of aligned systems with 3D stellar obliquities close to $0\degr$ and the clustered or broadly distributed systems with large obliquities.
@@
     L(\lambda) &\sim \prod_{i=1}^N\mathcal{N}(\hat{\lambda}_n, \sigma_{\hat{\lambda}_n}) \nonumber\\
-    L(i_\star, \theta_\star) &\sim \prod_{i=1}^N\mathcal{N}(\hat{v}\sin{i_{\star,n}}, \sigma_{\hat{v}\sin{i_{\star,n}}}) \nonumber\\
-    L(\theta) &\sim \prod_{i=1}^N\mathcal{N}(\hat{\theta}_n, \sigma_{\hat{\theta}_n}).
+    L(i_\star, \theta_\star) &\sim \prod_{i=1}^N\mathcal{N}(\hat{v}\sin{i_{\star,n}}, \sigma_{\hat{v}\sin{i_{\star,n}}}).
 \end{align}
@@
         \includegraphics{figures/psi_dist.pdf}
-        \caption{The cosine of 3D stellar obliquity $\cos{\psi}$ distributions.}
+        \caption{The distributions of the cosine of 3D stellar obliquity $\cos{\psi}$.
+The left panel presents the inferred $\cos{\psi}$ distribution for the \numistar systems that have both sky-projected stellar obliquity $\lambda$ and stellar inclination $i_\star$ measurements using the statistical model described in Figure~\ref{fig:graph}.
+The middle panel uses the same sample as the left panel but now removes the likelihood function on $i_\star$.
+The right panel presents the $\cos{\psi}$ distribution for all \numall systems that have sky-projected stellar obliquity $\lambda$ measurements and has no likelihood function on $i_\star$.
+The $\cos{\psi}$ distribution for the observed-$i_\star$ are similar and both show a cluster near $\cos{\psi} = -0.2$, whereas the distribution for all systems does not have such a trend.}
         \label{fig:psi_dist}
@@
 
-Following the logic, we now apply our model to all \numall exoplanetary systems that have sky-projected stellar obliquity measurements summarized in \cite{Albrecht22}.
+Consequently, we apply our model to all \numall exoplanetary systems that have sky-projected stellar obliquity measurements summarized in \cite{Albrecht22}.
 The result is presented in the right panel of Figure~\ref{fig:psi_dist}.
@@
 
-We also test the robustness of $\cos{\psi}$ distribution on different hyperprior choices \citep[see a nice summary in][]{Nagpal22}.
-We find extending the support of hyperpriors from 0--10 to 0--100 or 0--1000 or changing hyperprior distributions from Uniform to log-Uniform or truncated-Normal all infer similar results to Figure~\ref{fig:psi_dist}.
+\section{Interpretations of the \texorpdfstring{$\cos{\psi}$}{cospsi} distributions}\label{sec:interpretations}
 
-Although the distribution of $\psi$ looks very similar for the $i_\star$ sample using constrained or random priors, the uncertainty on each $\psi$ is much tighter.
+In this section, we aim to explain the similarities and differences of the $\cos{\psi}$ distributions presented in Figure~\ref{fig:psi_dist}.
 
-\section{Hypothesis} \label{sec:tests}
+\subsection{Do \texorpdfstring{$i_\star$}{istar} measurements matter?}
+One interesting result is with or without $i_\star$ measurements, the $\cos{\psi}$ distributions are very similar, as shown in the left and middle panels of Figure~\ref{fig:psi_dist}.
+Here we perform some tests to evaluate the importance of $i_\star$ constraints.
+
+\subsection{The underlying \texorpdfstring{$\lambda$}{lambda} distributions}
+
+When look at the underlying $\lambda$ distribution, we find a cluster of systems centered at $110\degr$.
 
@@
         \includegraphics{figures/lam_dist.pdf}
-        \caption{Sky-projected stellar obliquity ($\lambda$) distributions of \numistar systems with constrained $i_\star$ and all \numall exoplanetary systems that have measured sky-projected stellar obliquities.}
+        \caption{The distributions of sky-projected stellar obliquity ($\lambda$) of the \numistar systems that have both sky-projected stellar obliquity $\lambda$ and stellar inclination $i_\star$ measurements (left panel) and all \numall systems that have sky-projected stellar obliquity measurements.
+The observed-$i_\star$ has a cluster of systems near $110\degr$, whereas such clustering is not found in the all-system distribution.}
         \label{fig:lam_dist}
@@
 
-\subsection{Stellar Teff $>$ 7000K?}
-All planets with host star effective temperatures greater than 7000 K are near polar.
-Note that all these stars are expected to be fast rotators that limit the Rossiter-McLaughlin effect measurement.
-Doppler Tomography can be applied to these stars, but the majority of the measurements are conducted by gravitational darkening that is biased toward polar planets.
+All planets with host star effective temperatures greater than 7000 K are polar planet in the observed-$i_\star$ sample.
+Such a trend is not found in the full sample.
+To measure the inclination of hot stars, stellar modulation is no longer feasible, whereas the gravity-darkening method becomes more important.
+However, the gravity-darkening method is only sensitive to detecting polar planets.
 
-We removed 7/21 planets with host star effective temperatures greater than 7000 K and redo the analysis. Interestingly, we find the results do not change. See Figure~\ref{fig:teff_cutff}.
-
-% \begin{figure*}[ht!]
-%     \script{teff_cut.py}
-%     \begin{centering}
-%         \includegraphics[width=\linewidth]{figures/teff_cut.pdf}
-%         \caption{$\cos{\psi}$ distributions for the $i_\star$ sample including/not including $>$7000 K stars.}
-%         \label{fig:teff_cut}
-%     \end{centering}
-% \end{figure*}
-
-\subsection{Small vsini}
-Albrecht et al.
-21 that underestimating $v\sin{i_\star}$, overestimating $R_\star$, or underestimating $P_{\rm rot}$ could all lead to biased estimation on $\cos{\psi}$.
-We performed another test to evaluate the sample with random $i_\star$ prior (i.e., uniform on $\cos{i}$).
-Interestingly, we find a very similar distribution to the $\cos{\psi}$ distributions inferred with informative $i_\star$ priors.
-See Figure~\ref{fig:rand_inc}.
-
-% \begin{figure*}[ht!]
-%     \script{rand_inc.py}
-%     \begin{centering}
-%         \includegraphics[width=\linewidth]{figures/rand_inc.pdf}
-%         \caption{$\cos{\psi}$ distributions for the $i_\star$ sample with/without informative $i_\star$ priors.}
-%         \label{fig:rand_inc}
-%     \end{centering}
-% \end{figure*}
+\subsection{Hyperpriors}
+We also test the robustness of $\cos{\psi}$ distribution on different hyperprior choices \citep[see a nice summary in][]{Nagpal22}.
+We find extending the support of hyperpriors from 0--10 to 0--100 or 0--1000 or changing hyperprior distributions from Uniform to log-Uniform or truncated-Normal all infer similar results to Figure~\ref{fig:psi_dist}.
 
@@
 We perform the test by randomly drawing 20 planets out of the 40 misaligned systems.
-We find in half of the time, we could not reproduce the distribution but result in a pile-up of $\cos{\psi}$.
-See Figure~\ref{fig:bootstrap}.
+We find in half of the time, we could not reproduce the distribution but result in a cluster of $\cos{\psi}$.
 
@@
 
-\section{Discussion}
-No strong clustering?
+\section{Hot Jupiter 3D Stellar Obliquity Distribution} \label{sec:hjs}
 

### 2022-11-22 src/tex/ms.tex
[showyourwork] overleaf sync

@@
     \caption{Hierarchical Bayesian framework to infer the stellar obliquity distribution of exoplanetary systems.
-The stellar obliquity distribution is described by hyperparameters $\beta$ and constrained by $n_{\rm th}$ system's stellar obliquity $\psi_n$.
+The stellar obliquity distribution is described by hyperparameters $\vb*{\beta}$ and constrained by $n_{\rm th}$ system's stellar obliquity $\psi_n$.
 Each $\psi_n$ is calculated by the system's sky-projected stellar obliquity $\lambda_n$ and stellar inclination $i_{\star, n}$.
-$\theta_{\star, n}$ contains other information of the star other than $i_{\star, n}$, such as the stellar radius and stellar rotation period.}
+Obs$_{\star,n}$ contains observations that help to constrain $i_{\star,n}$, such as photometric or spectroscopic time series and stellar spectra.
+$\theta_{\star, n}$ contains other information of the star other than $i_{\star, n}$, such as the stellar radius and stellar rotation period and $\hat{\theta}_{\star,n}$ contains observations that are not already included in Obs$_{\star,n}$.}
     \label{fig:graph}

### 2022-11-23 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \begin{abstract}
+Stellar obliquity is a key property that implies the formation and dynamical processes of a planetary system.
 We build a hierarchical Bayesian framework to infer the 3D stellar obliquity $\psi$ distribution in exoplanetary systems. 
 First, we apply our framework to \numistar systems with sky-projected stellar obliquity $\lambda$ and stellar inclinations $i_\star$ measurements.
-The $\cos{\psi}$ distribution has a spike at 1, suggesting the majority of systems are well aligned, and a concentration at $-0.15$ ($\psi \sim 100\degr$), consistent with the finding of a preponderance of perpendicular planets.
-Interestingly, if we assume isotropic priors on $i_\star$ and use no extra data, a similar $\cos{\psi}$ distribution is inferred, suggesting the $cos\psi$ distribution depends mainly on the underlying $\lambda$ distribution.
+The $\cos{\psi}$ distribution has a sharp peak at 1, suggesting the majority of systems are well aligned, and is clustered at $-0.15$ ($\psi \sim 100\degr$), consistent with the finding of a preponderance of perpendicular planets.
+Interestingly, if we simply assume isotropic priors on $i_\star$ and use no additional data, a similar $\cos{\psi}$ distribution is inferred, suggesting the $cos\psi$ distribution depends mainly on the underlying $\lambda$ distribution.
 We then apply the framework to all \numall systems with sky-projected obliquity measurements.
-Different from the observed-$i_\star$ sample, $\cos{\psi}$ is nearly uniformly distributed between $-0.75$ to $0.75$, corresponding to an isotropic distribution of $\psi$ between $40\degr$ and $140\degr$.
-To understand the $\cos{\psi}$-distribution discrepancy between two samples, we look at their $\lambda$ distributions.
-The measured-$i_\star$ sample has a concentration of $\lambda$ at $110\degr$, whereas the full sample's $\lambda$ is uniformly distributed from $30\degr$ to $180\degr$.
-The observed-$i_\star$ sample could be biased because misaligned systems are often found around hot stars; however, hot stars' inclinations are mostly measured by the gravity-darkening method, which is biased toward measuring perpendicular planets.
-Lastly, we apply the framework to constrain Hot Jupiters' 3D stellar obliquity distribution and find that in misaligned Hot Jupiter systems, the stellar spin axis is isotropically distributed around the planet's orbital axis.
+Unlike the observed-$i_\star$ sample, the $\cos{\psi}$ is nearly uniformly distributed between $-0.75$ to $0.75$, corresponding to an isotropic distribution of $\psi$ between $40\degr$ and $140\degr$.
+To understand the discrepancy, we look at the underlying $\lambda$ distributions of two samples.
+For misaligned systems, the $\lambda$ distribution of the observed-$i_\star$ sample is clustered at $110\degr$, whereas the $\lambda$ distribution of the full sample is uniform.
+We find the observed-$i_\star$ sample could be biased because misaligned systems are often found around hot stars; however, hot stars' inclinations are mostly measured by the gravity-darkening method, which is biased toward detecting perpendicular planets.
+Lastly, we apply our framework to constrain Hot Jupiters' 3D stellar obliquity distribution and find that in misaligned Hot Jupiter systems, the stellar spin axis is isotropically distributed around the planet's orbital axis.
 \end{abstract}
@@
 3D stellar obliquity $\psi$ is described as the angle between a planet's orbital axis $\bf{i}_{\rm orb}$ and its host star's spin axis $\bf{i}_{\star}$.
-For any given $\bf{i}_{\rm orb}$ axis, if $\bf{i}_{\star}$ is a random vector with uniform distribution on a three-dimensional sphere, the probability distribution $p_{\psi} \sim \sin{\psi}$ and thus $p_{\cos{\psi}} = p_{\psi} \abs{\dv*{\psi}{\cos{\psi}}}$ and is uniformly distributed.
+For any given $\bf{i}_{\rm orb}$ axis, if $\bf{i}_{\star}$ is a random vector with uniform distribution on a three-dimensional sphere (i.e., $\bf{i}_{\star}$ is isotropically distributed), the probability distribution $p_{\psi} \sim \sin{\psi}$.
+Since $p_{\cos{\psi}} = p_{\psi} \abs{\dv*{\psi}{\cos{\psi}}}$, $p_{\cos{\psi}}$ is uniformly distributed.
 In reality, we only measure the sky-projected stellar obliquity $\lambda$, the angle between the projections of $\bf{i}_{\rm orb}$ and $\bf{i}_{\star}$ onto the plane of the sky.
-Stellar obliquity $\psi$ can be calculated from $\bf{i}_{\rm orb}$ and $\bf{i}_{\star}$ as the following:
+Stellar obliquity $\psi$ can be computed as:
 \begin{equation}\label{eqn:psi}
@@
 A full derivation of Equation~(\ref{eqn:psi}) can be found in, e.g., \cite{Fabrycky09}.
-Since the sky-projected stellar obliquities can only be measured in transiting exoplanet systems, $i_{\rm orb} \approx 90\degr$ and $\cos{\psi} \approx \sin{i_\star}\cos{\lambda}$.
+Since the sky-projected stellar obliquities can be measured only in transiting exoplanet systems, $i_{\rm orb} \approx 90\degr$ and $\cos{\psi} \approx \sin{i_\star}\cos{\lambda}$.
 
-To constrain the 3D stellar obliquity $\psi$, both a measurement of the sky-projected obliquity $\lambda$ and the stellar inclination $i_\star$ are required.
-The stellar inclination can often be constrained on exoplanetary systems via, e.g., photometric and spectroscopic rotational modulation introduced by starspots, gravity darkening, and asteroseismology.
+To constrain the 3D stellar obliquity $\psi$, the measurements of the sky-projected obliquity $\lambda$ and stellar inclination $i_\star$ are required.
+The stellar inclination can often be constrained via, e.g., photometric and spectroscopic rotational modulation introduced by starspots, gravity darkening, and asteroseismology.
 For systems without measured $i_\star$, it is still possible to infer the 3D stellar obliquity $\psi$ from the sky-projected obliquity $\lambda$ assuming the stellar inclination $i_\star$ is isotropically distributed \citep[e.g.,][]{Fabrycky09}.
-The probability distribution follows $p_{i_\star} \sim \sin{i_{\star}}$, or $p_{\cos{i_\star}} \sim \mathcal{U}$.
+The probability distribution of $i_star$ follows $p_{i_\star} \sim \sin{i_{\star}}$, or $p_{\cos{i_\star}} \sim 1$.
 
-\cite{Albrecht21} inferred the 3D stellar obliquities of a sample of stars that have both stellar inclination and sky-projected stellar obliquity constraints.
+\cite{Albrecht21} inferred the 3D stellar obliquities of a sample of stars that have both sky-projected stellar obliquity and stellar inclination constraints.
 They found while the majority of planets have a small $\psi$, misaligned planets have $\psi$ clustered between 80--125$\degr$.
@@
 
-In this work, we build a hierarchical Bayesian framework to evaluate the distribution of $\cos{\psi}$ using the sample in \cite{Albrecht21} and extend the analysis to the full sample of exoplanetary systems that have sky-projected stellar obliquity measurements extracted from the \cite{Albrecht22} review.
+In this work, we build a hierarchical Bayesian framework to evaluate the distribution of $\cos{\psi}$ for the sample in \cite{Albrecht21} and extend the analysis to the full sample of exoplanetary systems that have sky-projected stellar obliquity measurements extracted from the \cite{Albrecht22} review.
 In Section~\ref{sec:hbm}, we describe our hierarchical Bayesian framework and the inferred 3D stellar obliquity distributions, and in Section~\ref{sec:interpretations}, we discuss the interpretations of these distributions.
-Lastly, in Section~\ref{sec:hjs}, we apply the framework to Hot Jupiters and infer their 3D stellar obliquity distribution.
+In Section~\ref{sec:hjs}, we apply the framework to Hot Jupiters and infer their 3D stellar obliquity distribution.
 
@@
 We build a hierarchical Bayesian framework to model the 3D stellar obliquity ($\psi$) distribution of exoplanetary systems.
-
-First, we apply the framework to the sample of \numistar systems that have both sky-projected stellar obliquity $\lambda$ and stellar inclination $i_\star$ constraints.
 In Figure~\ref{fig:graph}, we illustrate the probabilistic graphical model.
@@
 $\bm{\beta}$ is constrained by $N$ individual systems, where each system has $\psi_n$ that is calculated by the sky-projected stellar obliquity $\lambda_n$ and the stellar inclination $i_n$.
-For simplicity, we approximate $i_{\rm orb}$ to $90\degr$ in the model and use Equation~(\ref{eqn:psi}) to compute $\psi_n$.
-The parameter $\theta_{\star, n}$ contains information on stellar rotation period and stellar radius, if available.
+For simplicity, we approximate $i_{\rm orb}$ to $90\degr$ and use the simplified version of Equation~(\ref{eqn:psi}) to compute $\psi_n$.
+The parameter $\theta_{\star, n}$ contains stellar parameters other than $i_\star$, e.g., stellar rotation period and stellar radius, if these information are available.
+Obs$_{\star,n}$ describes observations that help to constrain $i_{\star,n}$, such as photometric or spectroscopic time series and the stellar spectra.
 If the rotational modulation method is used to constrain the stellar inclination \citep[e.g.,][]{Masuda20}, $i_{\star,n}$ and $\theta_{\star,n}$ are constrained by observed sky-projected stellar rotational line broadening $\hat{v}\sin{i_{\star,n}}$ and observed $\hat{\theta}_{\star,n}$.
@@
 Each $\psi_n$ is calculated by the system's sky-projected stellar obliquity $\lambda_n$ and stellar inclination $i_{\star, n}$.
-Obs$_{\star,n}$ contains observations that help to constrain $i_{\star,n}$, such as photometric or spectroscopic time series and stellar spectra.
+Obs$_{\star,n}$ contains observations that help to constrain $i_{\star,n}$, such as photometric or spectroscopic time series and the stellar spectra.
 $\theta_{\star, n}$ contains other information of the star other than $i_{\star, n}$, such as the stellar radius and stellar rotation period and $\hat{\theta}_{\star,n}$ contains observations that are not already included in Obs$_{\star,n}$.}
@@
 For the stellar obliquity distribution, we adopt a two-component mixture Beta distribution with hyperparameters $\bm{\beta} = \{\bm{w},\bm{a},\bm{b}\}$, where each hyperparameter has a dimension of 2.
-The Beta distribution is flexible to describe both the spike of aligned systems with 3D stellar obliquities close to $0\degr$ and the clustered or broadly distributed systems with large obliquities.
+The Beta distribution is flexible to describe both the spike of aligned systems with 3D stellar obliquities close to $0\degr$ and the clustered or broadly distributed systems of misaligned systems.
 The probability distribution follows 
@@
     L(\lambda) &\sim \prod_{i=1}^N\mathcal{N}(\hat{\lambda}_n, \sigma_{\hat{\lambda}_n}) \nonumber\\
-    L(i_\star, \theta_\star) &\sim \prod_{i=1}^N\mathcal{N}(\hat{v}\sin{i_{\star,n}}, \sigma_{\hat{v}\sin{i_{\star,n}}}).
+    L(i_\star, \theta_\star) &\sim \prod_{i=1}^N\mathcal{N}({\rm Obs}_{\star,n}, \sigma_{{\rm Obs}_{\star,n}}).
 \end{align}
@@
 
-We apply the model to \numistar systems with both constrained $\lambda$ and $i_\star$ summarized in \cite{Albrecht21}.
+First, we apply the framework to the sample of \numistar systems that have both sky-projected stellar obliquity $\lambda$ and stellar inclination $i_\star$ constraints summarized in \cite{Albrecht21}.
 The inferred $\cos{\psi}$ distribution is shown in the left panel of Figure~\ref{fig:psi_dist}.
@@
 \subsection{Do \texorpdfstring{$i_\star$}{istar} measurements matter?}
-One interesting result is with or without $i_\star$ measurements, the $\cos{\psi}$ distributions are very similar, as shown in the left and middle panels of Figure~\ref{fig:psi_dist}.
+
+There is no doubt that $i_\star$ measurement improves the constraint on $\psi$ for an individual system.
+However, do $i_\star$ measurements in individual systems matter when we look at the population $\psi$ distribution?
+Particularly, the left and middle panels in Figure~\ref{fig:psi_dist} show that with or without $i_\star$ measurements, the $\cos{\psi}$ distributions are very similar.
 Here we perform some tests to evaluate the importance of $i_\star$ constraints.
+
+We begin with writing down some designed $\lambda$ distribution and assume an isotropic stellar inclination distribution.
+We then randomly draw 100 samples from these two distributions and assume these numbers are \emph{true} values of $\lambda$ and $i_\star$.
+We choose 100 so that the sample is large enough to represent the distributions, although, in reality, the number of misaligned systems with $i_\star$ measurements is much smaller ($\sim 20$).
+Now each $\lambda$ and $i_\star$ is added to some Gaussian noises to simulate \emph{observed} $\lambda$ and $i_\star$.
+We use uncertainties $\sigma_{\lambda} = 8\degr$ and $\sigma_{i_\star} = 10\degr$ that represent typical observation uncertainties of misaligned systems.
+Using the observed $\lambda$ and $i_\star$ and their uncertainties, we infer the $\cos{\psi}$ distribution of the sample with or without the $i_\star$ observations.
+We test three representative distributions of $\cos{\psi}$: (1) $\cos{\psi}$ uniformly distributed from $[-1,1]$, where we have the input $\lambda$ uniformly distributed over $[0, \pi]$, (2) $\cos{\psi}$ clustered at $-0.4$, where we have the input $\lambda$ normally distributed as $\mathcal{N}(120\degr,20\degr)$, and lastly (3) $\cos{\psi}$ clustered at $0$, where we have the input $\lambda$ normally distributed as $\mathcal{N}(90\degr,20\degr)$.
+In all three distributions, the $\cos{\psi}$ distribution inferred with or without $i_\star$ likelihoods are similar and consistent with the injected $\cos{\psi}$ distributions.
+
+Our experiment is under the assumption that misaligned systems still have isotropic stellar inclinations, which seems to be true.
 
@@
 
-When look at the underlying $\lambda$ distribution, we find a cluster of systems centered at $110\degr$.
+When looking at the underlying $\lambda$ distribution, we find a cluster of systems centered at $110\degr$.
 

### 2022-11-29 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
+\author{et al.}
+
 % Abstract with filler text
@@
 Stellar obliquity is a key property that implies the formation and dynamical processes of a planetary system.
-We build a hierarchical Bayesian framework to infer the 3D stellar obliquity $\psi$ distribution in exoplanetary systems. 
+Here we build a hierarchical Bayesian framework to infer the 3D stellar obliquity $\psi$ distribution in exoplanetary systems. 
 First, we apply our framework to \numistar systems with sky-projected stellar obliquity $\lambda$ and stellar inclinations $i_\star$ measurements.
-The $\cos{\psi}$ distribution has a sharp peak at 1, suggesting the majority of systems are well aligned, and is clustered at $-0.15$ ($\psi \sim 100\degr$), consistent with the finding of a preponderance of perpendicular planets.
+The $\cos{\psi}$ distribution has a sharp peak at 1, suggesting the majority of systems are well aligned, and has a secondary peak at $-0.15$ ($\psi \sim 100\degr$), in agreement with the finding of the preponderance of perpendicular planets.
 Interestingly, if we simply assume isotropic priors on $i_\star$ and use no additional data, a similar $\cos{\psi}$ distribution is inferred, suggesting the $cos\psi$ distribution depends mainly on the underlying $\lambda$ distribution.
@@
 To understand the discrepancy, we look at the underlying $\lambda$ distributions of two samples.
-For misaligned systems, the $\lambda$ distribution of the observed-$i_\star$ sample is clustered at $110\degr$, whereas the $\lambda$ distribution of the full sample is uniform.
+For misaligned systems, the $\lambda$ distribution of the observed-$i_\star$ sample is clustered at $110\degr$, whereas the $\lambda$ distribution of the full sample is nearly uniform.
 We find the observed-$i_\star$ sample could be biased because misaligned systems are often found around hot stars; however, hot stars' inclinations are mostly measured by the gravity-darkening method, which is biased toward detecting perpendicular planets.
@@
 3D stellar obliquity $\psi$ is described as the angle between a planet's orbital axis $\bf{i}_{\rm orb}$ and its host star's spin axis $\bf{i}_{\star}$.
-For any given $\bf{i}_{\rm orb}$ axis, if $\bf{i}_{\star}$ is a random vector with uniform distribution on a three-dimensional sphere (i.e., $\bf{i}_{\star}$ is isotropically distributed), the probability distribution $p_{\psi} \sim \sin{\psi}$.
+For any given $\bf{i}_{\rm orb}$ axis, if $\bf{i}_{\star}$ is a random vector with uniform distribution on a three-dimensional sphere, i.e., $\bf{i}_{\star}$ is isotropically distributed, the probability distribution $p_{\psi} \sim \sin{\psi}$.
 Since $p_{\cos{\psi}} = p_{\psi} \abs{\dv*{\psi}{\cos{\psi}}}$, $p_{\cos{\psi}}$ is uniformly distributed.
@@
 For systems without measured $i_\star$, it is still possible to infer the 3D stellar obliquity $\psi$ from the sky-projected obliquity $\lambda$ assuming the stellar inclination $i_\star$ is isotropically distributed \citep[e.g.,][]{Fabrycky09}.
-The probability distribution of $i_star$ follows $p_{i_\star} \sim \sin{i_{\star}}$, or $p_{\cos{i_\star}} \sim 1$.
+The probability distribution of $i_{\rm star}$ follows $p_{i_\star} \sim \sin{i_{\star}}$, or $p_{\cos{i_\star}} \sim 1$.
 
@@
 
+\begin{figure}
+    \script{coord_psi.py}
+    \includegraphics[width=\linewidth]{figures/coord_psi.pdf}
+\end{figure}
+
 \section{3D Stellar Obliquity Distributions} \label{sec:hbm}
@@
 For simplicity, we approximate $i_{\rm orb}$ to $90\degr$ and use the simplified version of Equation~(\ref{eqn:psi}) to compute $\psi_n$.
-The parameter $\theta_{\star, n}$ contains stellar parameters other than $i_\star$, e.g., stellar rotation period and stellar radius, if these information are available.
+The parameter $\theta_{\star, n}$ contains stellar parameters other than $i_\star$, e.g., stellar rotation period and stellar radius, if the information is available.
 Obs$_{\star,n}$ describes observations that help to constrain $i_{\star,n}$, such as photometric or spectroscopic time series and the stellar spectra.
-If the rotational modulation method is used to constrain the stellar inclination \citep[e.g.,][]{Masuda20}, $i_{\star,n}$ and $\theta_{\star,n}$ are constrained by observed sky-projected stellar rotational line broadening $\hat{v}\sin{i_{\star,n}}$ and observed $\hat{\theta}_{\star,n}$.
+If the rotational modulation method is used to constrain the stellar inclination, the $i_{\star,n}$ and $\theta_{\star,n}$ are constrained by rotational modulation in photometric or spectroscopic time series and observed sky-projected stellar rotational line broadening $\hat{v}\sin{i_{\star,n}}$ \citep[e.g.,][]{Masuda20}.
+If the gravity-darkening method is used, the $i_{\star,n}$ is constrained by the anomaly in transit light curves.
+If asteroseismology is used, the $i_{\star,n}$ is constrained from the periodic variation in the photometric time series.
 Lastly, the sky-projected stellar obliquity $\lambda_n$ is constrained by observed $\hat{\lambda}_n$.
@@
 Each $\psi_n$ is calculated by the system's sky-projected stellar obliquity $\lambda_n$ and stellar inclination $i_{\star, n}$.
-Obs$_{\star,n}$ contains observations that help to constrain $i_{\star,n}$, such as photometric or spectroscopic time series and the stellar spectra.
-$\theta_{\star, n}$ contains other information of the star other than $i_{\star, n}$, such as the stellar radius and stellar rotation period and $\hat{\theta}_{\star,n}$ contains observations that are not already included in Obs$_{\star,n}$.}
+$\theta_{\star, n}$ contains other information of the star other than $i_{\star, n}$, such as the stellar radius and stellar rotation period.
+Obs$_{\star,n}$ contains observations of the star, such as the stellar spectral and photometric and/or spectroscopic time series, and $\hat{\lambda}_n$ has the measured sky-projected stellar obliquity.}
     \label{fig:graph}
@@
 
-\begin{figure*}[ht!]
+\begin{figure*}
     \script{psi_dist.py}
@@
 The $\cos{\psi}$ distribution of the new model is shown in the middle panel of Figure~\ref{fig:psi_dist}.
-Interestingly, the overall $\cos{\psi}$ distributions of the two models are very similar, despite that the inferred $\psi$ uncertainty for each system is now greater.
+Interestingly, the overall $\cos{\psi}$ distributions of the two models are very similar, despite the fact that the inferred $\psi$ uncertainty for each system is now greater.
 The similarity between the two distributions suggests that, at least for the measured-$i_\star$ sample, the $\cos{\psi}$ distribution mainly depends on the underlying sky-project stellar obliquity $\lambda$ distribution.
@@
 
-\section{Interpretations of the \texorpdfstring{$\cos{\psi}$}{cospsi} distributions}\label{sec:interpretations}
+\section{Interpretations of the \texorpdfstring{$\psi$}{psi} distributions}\label{sec:interpretations}
 
-In this section, we aim to explain the similarities and differences of the $\cos{\psi}$ distributions presented in Figure~\ref{fig:psi_dist}.
+In this section, we aim to explain the similarity and disparity between the $\cos{\psi}$ distributions presented in Figure~\ref{fig:psi_dist}.
 
-\subsection{Do \texorpdfstring{$i_\star$}{istar} measurements matter?}
+\subsection{Infer the \texorpdfstring{$\psi$}{psi} distribution from the \texorpdfstring{$\lambda$}{lambda} distribution}
 
-There is no doubt that $i_\star$ measurement improves the constraint on $\psi$ for an individual system.
-However, do $i_\star$ measurements in individual systems matter when we look at the population $\psi$ distribution?
-Particularly, the left and middle panels in Figure~\ref{fig:psi_dist} show that with or without $i_\star$ measurements, the $\cos{\psi}$ distributions are very similar.
-Here we perform some tests to evaluate the importance of $i_\star$ constraints.
+There is no doubt that $i_\star$ measurement on individual systems improves the constraint on $\psi$.
+However, do $i_\star$ measurements matter when we constrain the population $\psi$ distribution?
+Particularly, the left and middle panels in Figure~\ref{fig:psi_dist} show that with or without $i_\star$ measurements, the inferred $\cos{\psi}$ distributions are very consistent.
+We perform some calculations and tests to evaluate the necessity of $i_\star$ constraints.
 
-We begin with writing down some designed $\lambda$ distribution and assume an isotropic stellar inclination distribution.
-We then randomly draw 100 samples from these two distributions and assume these numbers are \emph{true} values of $\lambda$ and $i_\star$.
-We choose 100 so that the sample is large enough to represent the distributions, although, in reality, the number of misaligned systems with $i_\star$ measurements is much smaller ($\sim 20$).
-Now each $\lambda$ and $i_\star$ is added to some Gaussian noises to simulate \emph{observed} $\lambda$ and $i_\star$.
-We use uncertainties $\sigma_{\lambda} = 8\degr$ and $\sigma_{i_\star} = 10\degr$ that represent typical observation uncertainties of misaligned systems.
-Using the observed $\lambda$ and $i_\star$ and their uncertainties, we infer the $\cos{\psi}$ distribution of the sample with or without the $i_\star$ observations.
-We test three representative distributions of $\cos{\psi}$: (1) $\cos{\psi}$ uniformly distributed from $[-1,1]$, where we have the input $\lambda$ uniformly distributed over $[0, \pi]$, (2) $\cos{\psi}$ clustered at $-0.4$, where we have the input $\lambda$ normally distributed as $\mathcal{N}(120\degr,20\degr)$, and lastly (3) $\cos{\psi}$ clustered at $0$, where we have the input $\lambda$ normally distributed as $\mathcal{N}(90\degr,20\degr)$.
-In all three distributions, the $\cos{\psi}$ distribution inferred with or without $i_\star$ likelihoods are similar and consistent with the injected $\cos{\psi}$ distributions.
+Given a certain $\psi$ distribution, the $\lambda$ and $i_\star$ are no longer randomly distributed.
+We first find relations between $\{\psi, \theta\}$ and $\{\lambda, i_\star\}$, where $\theta$ is the azimuthal angle of the stellar spin axis.
+\begin{align}
+    \sin{\psi}\cos{\theta} &= \cos{i_\star} \nonumber\\
+    \sin{\psi}\sin{\theta} &= \sin{\lambda}\sin{i_\star} \nonumber\\
+    \cos{\psi} &= \cos{\lambda}\sin{i_\star}.
+\end{align}
+Thus, $i_\star = \cos[-1](\sin{\psi}\cos{\theta})$ and $\lambda = \cos[-1](\cos{\psi}/\sqrt{1-\cos^2{\theta}\sin^2{\psi}})$.
+If we assume $\theta$ is uniformly distributed, just from the $\lambda$ distribution, we could directly infer the $\psi$ distribution.
 
-Our experiment is under the assumption that misaligned systems still have isotropic stellar inclinations, which seems to be true.
+We then design some $\cos{\psi}$ distributions.
+From each distribution of $\cos{\psi}$, we randomly draw 200 $i_\star$ and compute the $\lambda$.
+These $i_\star$ and $\lambda$ are the \emph{true} values.
+We use 200 samples so that the sample size is large enough to represent the distribution.
+In reality, the number of misaligned systems with $i_\star$ measurements is much smaller ($\sim 20$).
+Now $\lambda$ and $i_\star$ are added to some Gaussian noises to simulate the \emph{observed} values.
+We use uncertainties of $\sigma_{\lambda} = 8\degr$ and $\sigma_{i_\star} = 10\degr$ that represent typical observation uncertainties of misaligned systems.
+In a consequence test, we find using smaller uncertainties does not affect the distributions.
+Lastly, from the observed $\lambda$ and $i_\star$ and their uncertainties, we infer the $\cos{\psi}$ distribution of the sample with or without the $i_\star$ observations.
+
+We test four different distributions of $\cos{\psi}$: a uniform distribution on the interval $[-1,1]$ and three Normal distributions $\mathcal{N}(0,0.2)$, $\mathcal{N}(-0.4,0.2)$, and $\mathcal{N}(0.4,0.2)$.
+In all four distributions, the $\cos{\psi}$ distributions inferred with or without $i_\star$ likelihoods are consistent with each other and the injected distributions.
+
+We conclude that $i_\star$ measurements on individual systems are not necessary to infer the population $\psi$ distribution.
 
@@
 
-When looking at the underlying $\lambda$ distribution, we find a cluster of systems centered at $110\degr$.
+Next, we find underlying $\lambda$ distributions of the measured-$i_\star$ sample and the full sample.
+We build another hierarchical Bayesian framework to infer the $\lambda$ distribution using a two-component mixture Beta distribution, similar to the one used for the $\cos{\psi}$ distribution.
+We extend the distribution support from $[0,1]$ to $[0,\pi]$ by multiplying the distribution by $\pi$.
 
-\begin{figure*}[ht!]
+\begin{figure*}
     \script{lam_dist.py}
@@
 
-All planets with host star effective temperatures greater than 7000 K are polar planet in the observed-$i_\star$ sample.
+Figure~\ref{fig:lam_dist} presents the $\lambda$ distributions of two samples. In the measured-$i_\star$ sample,
+
+All planets with host star effective temperatures greater than 7000 K are polar planets in the observed-$i_\star$ sample.
 Such a trend is not found in the full sample.
@@
 
-\subsection{Statistical fluke}
-Could it just be a statistical fluke?
-We perform the test by randomly drawing 20 planets out of the 40 misaligned systems.
-We find in half of the time, we could not reproduce the distribution but result in a cluster of $\cos{\psi}$.
-
-% \begin{figure*}[ht!]
-%     \script{bootstrap.py}
-%     \begin{centering}
-%         \includegraphics[width=\linewidth]{figures/bootstrap.pdf}
-%         \caption{$\cos{\psi}$ distributions for the $i_\star$ sample with/without informative $i_\star$ priors.}
-%         \label{fig:bootstrap}
-%     \end{centering}
-% \end{figure*}
-
 \section{Hot Jupiter 3D Stellar Obliquity Distribution} \label{sec:hjs}

### 2022-11-29 src/scripts/coord_lam.py
[showyourwork] overleaf sync

@@ -1,0 +1,131 @@
+import os
+from pathlib import Path
+os.environ["PATH"] += os.pathsep + str(Path.home() / "bin")
+
+import paths
+
+import sys
+import subprocess
+
+from matplotlib import rc
+rc('font', **{'family':'sans-serif'})
+rc('text', usetex=True)
+rc('text.latex', preamble=r'\usepackage{physics}')
+
+import matplotlib.pyplot as plt
+
+fig = plt.figure(dpi=80)
+ax = fig.add_subplot(projection='3d')
+
+# Make data
+u = np.linspace(0, 2 * np.pi, 180)
+v = np.linspace(0, np.pi, 180)
+x = np.outer(np.cos(u), np.sin(v))
+y = np.outer(np.sin(u), np.sin(v))
+z = np.outer(np.ones(np.size(u)), np.cos(v))
+
+#Plot the surface
+ax.plot_surface(x, y, z, rstride=4, cstride=4, color='b', linewidth=0, alpha=0.1)
+
+elev = 10.0
+rot = 60.0 / 180 * np.pi
+
+#calculate vectors for "vertical" circle
+a = np.array([-np.sin(elev / 180 * np.pi), 0, np.cos(elev / 180 * np.pi)])
+b = np.array([0, 1, 0])
+b = b * np.cos(rot) + np.cross(a, b) * np.sin(rot) + a * np.dot(a, b) * (1 - np.cos(rot))
+ax.plot(np.sin(u),np.cos(u),0,color='k', linestyle = 'dashed')
+horiz_front = np.linspace(0, np.pi, 100)
+ax.plot(np.sin(horiz_front),np.cos(horiz_front),0,color='k')
+
+vert_front = np.linspace(np.pi / 2, 3 * np.pi / 2, 100)
+ax.plot(a[0] * np.sin(u) + b[0] * np.cos(u), b[1] * np.cos(u), a[2] * np.sin(u) + b[2] * np.cos(u),
+        color='k',lw=1)
+ax.plot(a[0] * np.sin(vert_front) + b[0] * np.cos(vert_front), b[1] * np.cos(vert_front), 
[…]

### 2022-11-29 src/scripts/coord_psi.py

@@ -1,0 +1,163 @@
+import os
+from pathlib import Path
+os.environ["PATH"] += os.pathsep + str(Path.home() / "bin")
+
+import paths
+
+import sys
+import subprocess
+
+from matplotlib import rc
+rc('font', **{'family':'sans-serif'})
+rc('text', usetex=True)
+rc('text.latex', preamble=r'\usepackage{physics}')
+
+import matplotlib.pyplot as plt
+
+fig = plt.figure(dpi=80)
+ax = fig.add_subplot(projection='3d')
+
+# Make data
+u = np.linspace(0, 2 * np.pi, 180)
+v = np.linspace(0, np.pi, 180)
+x = np.outer(np.cos(u), np.sin(v))
+y = np.outer(np.sin(u), np.sin(v))
+z = np.outer(np.ones(np.size(u)), np.cos(v))
+
+#Plot the surface
+ax.plot_surface(x, y, z, rstride=4, cstride=4, color='b', linewidth=0, alpha=0.1)
+# ax.plot_wireframe(x, y, z, rstride=180, cstride=180, color='#e5e1e0', alpha=0.5)
+
+# u = np.linspace(0, 2 * np.pi, 500)
+# v = np.linspace(0, np.pi, 500)
+
+# x = 1 * np.outer(np.cos(u), np.sin(v))
+# y = 1 * np.outer(np.sin(u), np.sin(v))
+# z = 1 * np.outer(np.ones(np.size(u)), np.cos(v))
+#for i in range(2):
+#    ax.plot_surface(x+random.randint(-5,5), y+random.randint(-5,5), z+random.randint(-5,5),  rstride=4, cstride=4, color='b', linewidth=0, alpha=0.5)
+elev = 20.0
+rot = 30.0 / 180 * np.pi
+# ax.plot_surface(x, y, z,  rstride=4, cstride=4, color='w', linewidth=0, alpha=0.3)
+
+#calculate vectors for "vertical" circle
+a = np.array([-np.sin(elev / 180 * np.pi), 0, np.cos(elev / 180 * np.pi)])
+b = np.array([0, 1, 0])
[…]

### 2022-11-29 src/scripts/graph.py

@@ -15,11 +15,12 @@ import daft
 from matplotlib import rc
 rc('font', **{'family':'sans-serif'})
 rc('text', usetex=True)
+rc('text.latex', preamble=r'\usepackage{physics}')
 
-pgm = daft.PGM(observed_style="outer",node_unit=1, node_ec='k',dpi=150,line_width=0.8)
+pgm = daft.PGM(observed_style="shaded",node_unit=1, node_ec='k',dpi=150,line_width=0.8)
                
 # Hierarchical parameters.
-pgm.add_node("beta", r"$\beta$", 2, 2.3)
+pgm.add_node("beta", r"$\vb*{\beta}$", 2, 2.3)
 
 # Latent variable.
 pgm.add_node("psi", r"$\psi_n$", 2, 1.5)
@@ -28,7 +29,7 @@ pgm.add_node("lambda", r"$\lambda_n$", 1., 1.5)
 
 pgm.add_node("istar", r"$i_{\star,n}$", 3., 1.5)
 
-pgm.add_node("pstar", r"$\theta_{\star,n}$", 3.6, 1.5)
+pgm.add_node("pstar", r"$\theta_{\star,n}$", 4., 1.5)
 
 
 pgm.add_edge("beta", "psi")
@@ -40,23 +41,29 @@ pgm.add_edge("istar", "psi")
 
 pgm.add_node("obs_lambda", r"$\hat{\lambda}_n$", 1., 0.7, observed=True)
 
-pgm.add_node("obs_vsini", r"$\hat{v}\sin{i}_{\star, n}$", 3., 0.7, observed=True, aspect=1.3, fontsize=8)
+pgm.add_node("Obs", r"Obs$_{\star, n}$", 3., 0.7, observed=True, aspect=1., fontsize=8)
+
+# pgm.add_node("obs_pstar", r"$\hat{\theta}_{\star,n}$", 4., 0.7, observed=True)
+
 
 pgm.add_edge("lambda", "obs_lambda")
 
-pgm.add_edge("pstar", "obs_vsini", plot_params=dict(ls=(0, (1, 2)),head_width=0,head_length=0))
-pgm.add_edge("pstar", "obs_vsini", plot_params={'ls':''})
+pgm.add_edge("pstar", "Obs", plot_params=dict(ls=(0, (2, 2)),head_width=0,head_length=0))
+pgm.add_edge("pstar", "Obs", plot_params={'ls':''})
 
[…]

### 2022-11-29 src/scripts/lam_dist.py

@@ -1,0 +1,87 @@
+import os
+from pathlib import Path
+os.environ["PATH"] += os.pathsep + str(Path.home() / "bin")
+
+import sys
+import subprocess
+
+# implement pip as a subprocess:
+subprocess.check_call([sys.executable, '-m', 'pip', 'install',
+'arviz'])
+
+import paths
+import matplotlib.pyplot as plt
+import numpy as np
+import arviz as az
+from scipy.stats import beta
+
+from matplotlib import rc
+rc('font', **{'family':'sans-serif'})
+rc('text', usetex=True)
+
+plt.rcParams['xtick.top'] =  True
+plt.rcParams['xtick.direction'] =  'in'
+plt.rcParams['xtick.major.width'] =  1.0
+plt.rcParams['xtick.minor.width'] =  1.0
+plt.rcParams['ytick.right'] =  True
+plt.rcParams['ytick.direction'] =  'in'
+plt.rcParams['ytick.major.width'] = 1.0
+plt.rcParams['ytick.minor.width'] =  1.0
+plt.rcParams['lines.markeredgewidth'] =  1.0
+
+# load MCMC draws
+lam_all = az.from_netcdf(paths.data / 'lam_all.nc')
+lam_istar = az.from_netcdf(paths.data / 'lam_istar.nc')
+
+x = np.linspace(1e-5,1-1e-5,1000)
+
+post = lam_all.posterior
+lam_all_draws = np.zeros(shape=(len(x),4000))
+for a in range(4):
+    for b in range(1000):
+        lam_all_draws[:, a*1000+b] = (post.w[a,b,0].values*beta.pdf(x, post.a0[a,b], post.b0[a,b])
+                                    +post.w[a,b,1].values*beta.pdf(x, post.a1[a,b], post.b1[a,b]))
+
+post = lam_istar.posterior
+lam_istar_draws = np.zeros(shape=(len(x),4000))
+for a in range(4):
+    for b in range(1000):
+        lam_istar_draws[:, a*1000+b] = (post.w[a,b,0].values*beta.pdf(x, post.a0[a,b], post.b0[a,b])
+                                    +post.w[a,b,1].values*beta.pdf(x, post.a1[a,b], post.b1[a,b]))
+
+# make the figure
+plt.figure(figsize=(5,2.5),dpi=110)
+
+plt.subplot(1,2,1)
[…]

### 2022-11-29 src/scripts/psi_dist.py

@@ -58,7 +58,7 @@ for a in range(4):
                                           +post.w[a,b,1].values*beta.pdf(x, post.a1[a,b], post.b1[a,b]))
 
 # make the figure
-plt.figure(figsize=(7,2.5),dpi=110)
+plt.figure(figsize=(7,2.6),dpi=110)
 
 plt.subplot(1,3,1)
 q025, q16, q50, q84, q975 = np.percentile(istar_draws, [2.5, 16, 50, 84, 97.5], axis=1)
@@ -72,7 +72,8 @@ plt.xlim([-1,1])
 plt.xlabel(r'$\cos{\psi}$',fontsize=11)
 plt.ylabel('Probablity density function')
 
-plt.title(r'65 systems -- informative $i_{\star}$')
+plt.title(r'\begin{center} 65 systems with observed $i_{\star}$ \\ \textbf{with} $i_{\star}$ likelihood \end{center}', 
+          fontsize=11, y=1.1)
 
 
 plt.subplot(1,3,2)
@@ -84,7 +85,8 @@ plt.fill_between(2*x-1, q025, q975, alpha=0.5, color='#cfe7ea')
 plt.ylim([0,2])
 plt.xlim([-1,1])
 
-plt.title(r'65 systems -- random $i_{\star}$')
+plt.title(r'\begin{center} 65 systems with observed $i_{\star}$ \\ \textbf{no} $i_{\star}$ likelihood \end{center}', 
+          fontsize=11, y=1.1)
 
 
 plt.subplot(1,3,3)
@@ -96,7 +98,9 @@ plt.fill_between(2*x-1, q025, q975, alpha=0.3, color='#7dabd0')
 plt.ylim([0,2])
 plt.xlim([-1,1])
 
-plt.title(r'161 systems -- random $i_{\star}$')
+plt.title(r'\begin{center} all 161 systems \\ \textbf{no} $i_{\star}$ likelihood \end{center}', 
+          fontsize=11, y=1.1)
 
 plt.tight_layout()
 plt.savefig(paths.figures / "psi_dist.pdf", bbox_inches="tight", dpi=600)
+plt.close()

### 2022-11-29 src/scripts/coord_lam.py
[showyourwork] overleaf sync

@@ -13,6 +13,7 @@ rc('text', usetex=True)
 rc('text.latex', preamble=r'\usepackage{physics}')
 
 import matplotlib.pyplot as plt
+import numpy as np
 
 fig = plt.figure(dpi=80)
 ax = fig.add_subplot(projection='3d')

### 2022-12-05 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-\newcommand{\numistar}{65\xspace}
+\newcommand{\numistar}{62\xspace}
 \newcommand{\numall}{161\xspace}
@@
 % Title
-\title{3D Stellar Obliquity Distribution of Exoplanetary Systems}
+\title{Isotropic Stellar Obliquities of Misaligned Planetary Systems}
 
@@
 \begin{abstract}
-Stellar obliquity is a key property that implies the formation and dynamical processes of a planetary system.
-Here we build a hierarchical Bayesian framework to infer the 3D stellar obliquity $\psi$ distribution in exoplanetary systems. 
-First, we apply our framework to \numistar systems with sky-projected stellar obliquity $\lambda$ and stellar inclinations $i_\star$ measurements.
-The $\cos{\psi}$ distribution has a sharp peak at 1, suggesting the majority of systems are well aligned, and has a secondary peak at $-0.15$ ($\psi \sim 100\degr$), in agreement with the finding of the preponderance of perpendicular planets.
-Interestingly, if we simply assume isotropic priors on $i_\star$ and use no additional data, a similar $\cos{\psi}$ distribution is inferred, suggesting the $cos\psi$ distribution depends mainly on the underlying $\lambda$ distribution.
+
+Many exoplanetary systems are found with high stellar obliquities.
+We build a hierarchical Bayesian framework to understand how stellar obliquities of misaligned systems are distributed.
+If stellar obliquities $\psi$ are isotropically distributed, the $\cos{\psi}$ distribution will be uniform, whereas if stellar obliquities have preferred angles, the $\cos{\psi}$ distribution will present concentrations.
+We first apply our framework to \numistar systems with both sky-projected stellar obliquity $\lambda$ and stellar inclination $i_\star$ constraints.
+The $\cos{\psi}$ distribution is clustered near $-0.2$ ($\psi \sim 100\degr$), in agreement with the preponderance of perpendicular planets.
+Interestingly, a similar $\cos{\psi}$ distribution could be inferred without using any $i_\star$ measurements, suggesting that the $\psi$ distribution depends mainly on the underlying $\lambda$ distribution.
 We then apply the framework to all \numall systems with sky-projected obliquity measurements.
-Unlike the observed-$i_\star$ sample, the $\cos{\psi}$ is nearly uniformly distributed between $-0.75$ to $0.75$, corresponding to an isotropic distribution of $\psi$ between $40\degr$ and $140\degr$.
-To understand the discrepancy, we look at the underlying $\lambda$ distributions of two samples.
-For misaligned systems, the $\lambda$ distribution of the observed-$i_\star$ sample is clustered at $110\degr$, whereas the $\lambda$ distribution of the full sample is nearly uniform.
-We find the observed-$i_\star$ sample could be biased because misaligned systems are often found around hot stars; however, hot stars' inclinations are mostly measured by the gravity-darkening method, which is biased toward detecting perpendicular planets.
-Lastly, we apply our framework to constrain Hot Jupiters' 3D stellar obliquity distribution and find that in misaligned Hot Jupiter systems, the stellar spin axis is isotropically distributed around the planet's orbital axis.
+Unlike the observed-$i_\star$ sample, the $\cos{\psi}$ is  nearly uniformly distributed for misaligned systems, corresponding to an isotropic distribution of stellar obliquity.
+We find the $\lambda$ distribution of the observed-$i_\star$ sample is clustered at $110\degr$, whereas the $\lambda$ distribution of the full sample is nearly uniform.
+The observed-$i_\star$ sample could be biased because misaligned systems are often found around hot stars; however, hot stars' inclinations are often measured by the gravity-darkening method, which is biased toward detecting perpendicular planets.
+We conclude that misaligned planetary systems, mostly Hot Jupiters, have isotropic stellar obliquities.
+When compared to theoretical predictions, the distribution is best consistency with the scattering of multiple giant planets after their convergent disk migration.
 \end{abstract}
@@
 
-3D stellar obliquity $\psi$ is described as the angle between a planet's orbital axis $\bf{i}_{\rm orb}$ and its host star's spin axis $\bf{i}_{\star}$.
-For any given $\bf{i}_{\rm orb}$ axis, if $\bf{i}_{\star}$ is a random vector with uniform distribution on a three-dimensional sphere, i.e., $\bf{i}_{\star}$ is isotropically distributed, the probability distribution $p_{\psi} \sim \sin{\psi}$.
-Since $p_{\cos{\psi}} = p_{\psi} \abs{\dv*{\psi}{\cos{\psi}}}$, $p_{\cos{\psi}}$ is uniformly distributed.
-In reality, we only measure the sky-projected stellar obliquity $\lambda$, the angle between the projections of $\bf{i}_{\rm orb}$ and $\bf{i}_{\star}$ onto the plane of the sky.
-Stellar obliquity $\psi$ can be computed as:
+The stellar obliquity $\psi$ describes the angle between a planet's orbital axis $\bf{n}_{\rm orb}$ and its host star's spin axis $\bf{n}_{\star}$.
+For any given $\bf{n}_{\rm orb}$ axis, if $\bf{n}_{\star}$ is a random vector with uniform distribution on a three-dimensional sphere, i.e., $\bf{n}_{\star}$ is isotropically distributed, the probability distribution of $\psi$ follows $p_{\psi} \sim \sin{\psi}$.
+Since $p_{\cos{\psi}} = p_{\psi} \abs{\dv*{\psi}{\cos{\psi}}}$, $p_{\cos{\psi}}$ is uniformly distributed between $-1$ and $1$.
+If the stellar obliquity is not isotropically distributed but have some preferred angles, the $\cos{\psi}$ distribution will present concentrations at certain values.
+
+In reality, we only measure the sky-projected stellar obliquity $\lambda$, the angle between the projections of $\bf{n}_{\rm orb}$ and $\bf{n}_{\star}$ onto the plane of the sky.
+Stellar obliquity $\psi$ can be computed as \citep[e.g.,][]{Fabrycky09}:
 \begin{equation}\label{eqn:psi}
@@
 \end{equation}
-where $i_{\rm orb}$ is the angle between the vector $\bf{i}_{\rm orb}$ and the observer's line of sight, and $i_\star$ is the angle between $\bf{i}_{\star}$ and the observer's line of sight.
-A full derivation of Equation~(\ref{eqn:psi}) can be found in, e.g., \cite{Fabrycky09}.
+where $i_{\rm orb}$ is the angle between the vector $\bf{n}_{\rm orb}$ and the observer's line of sight, and $i_\star$ is the angle between $\bf{n}_{\star}$ and the observer's line of sight.
 Since the sky-projected stellar obliquities can be measured only in transiting exoplanet systems, $i_{\rm orb} \approx 90\degr$ and $\cos{\psi} \approx \sin{i_\star}\cos{\lambda}$.
@@
 The stellar inclination can often be constrained via, e.g., photometric and spectroscopic rotational modulation introduced by starspots, gravity darkening, and asteroseismology.
-For systems without measured $i_\star$, it is still possible to infer the 3D stellar obliquity $\psi$ from the sky-projected obliquity $\lambda$ assuming the stellar inclination $i_\star$ is isotropically distributed \citep[e.g.,][]{Fabrycky09}.
-The probability distribution of $i_{\rm star}$ follows $p_{i_\star} \sim \sin{i_{\star}}$, or $p_{\cos{i_\star}} \sim 1$.
+For systems without measured $i_\star$, it is still possible to infer the 3D stellar obliquity $\psi$ just from the sky-projected obliquity $\lambda$ assuming the stellar inclination $i_\star$ is isotropically distributed \citep[e.g.,][]{Fabrycky09}.
 
@@
 They found while the majority of planets have a small $\psi$, misaligned planets have $\psi$ clustered between 80--125$\degr$.
-Furthermore, the $\cos{\psi}$ distribution of misaligned systems is not uniformly distributed but clusters near 0, suggesting the stellar spin axis is not isotropically distributed around the planetary orbital axis.
+Furthermore, the $\cos{\psi}$ distribution of misaligned systems is not uniformly distributed, but clusters near 0, suggesting the stellar spin axis is not isotropically distributed around the planetary orbital axis.
 
-In this work, we build a hierarchical Bayesian framework to evaluate the distribution of $\cos{\psi}$ for the sample in \cite{Albrecht21} and extend the analysis to the full sample of exoplanetary systems that have sky-projected stellar obliquity measurements extracted from the \cite{Albrecht22} review.
-In Section~\ref{sec:hbm}, we describe our hierarchical Bayesian framework and the inferred 3D stellar obliquity distributions, and in Section~\ref{sec:interpretations}, we discuss the interpretations of these distributions.
-In Section~\ref{sec:hjs}, we apply the framework to Hot Jupiters and infer their 3D stellar obliquity distribution.
-
-\begin{figure}
-    \script{coord_psi.py}
-    \includegraphics[width=\linewidth]{figures/coord_psi.pdf}
-\end{figure}
+In this work, we build a hierarchical Bayesian framework to constrain the stellar obliquity distribution of exoplanetary systems, particularly of misaligned systems.
+In Section~\ref{sec:hbm}, we describe our hierarchical Bayesian framework and show inferred stellar obliquity distributions. 
+In Section~\ref{sec:interpretations}, we interpret the results. 
+In Section~\ref{sec:discussion}, we discuss the implication of the stellar obliquity distribution for Hot Jupiter origins.
 
@@
 We build a hierarchical Bayesian framework to model the 3D stellar obliquity ($\psi$) distribution of exoplanetary systems.
-In Figure~\ref{fig:graph}, we illustrate the probabilistic graphical model.
-We aim to constrain a set of hyperparameters $\bm{\beta}$ that describe the 3D stellar obliquity $\psi$ distribution.
+Figure~\ref{fig:graph} shows the probabilistic graphical model.
+We aim to constrain a set of hyperparameters $\bm{\beta}$ that describe the stellar obliquity $\psi$ distribution.
 $\bm{\beta}$ is constrained by $N$ individual systems, where each system has $\psi_n$ that is calculated by the sky-projected stellar obliquity $\lambda_n$ and the stellar inclination $i_n$.
@@
 \begin{align}
-    w_{0,1} &\sim {\rm Dirichlet}(1,1) \nonumber\\
-    a_0 &\sim \mathcal{U}(0,50) \nonumber\\
-    b_0 &\sim \mathcal{U}(0,1) \nonumber\\
-    a_1 &\sim \mathcal{U}(0,10) \nonumber\\
-    b_1 &\sim \mathcal{U}(0,10) \nonumber\\
-    \cos{i}_{\star,n} &\sim \mathcal{U}(0,1) \nonumber\\
-    \lambda_{\star,n} &\sim \mathcal{U}(0,\pi).
+    w_{0,1} &\sim {\rm Dirichlet}(1, 1) \nonumber\\
+    a_0 &\sim \mathcal{U}(0, 50) \nonumber\\
+    b_0 &\sim \mathcal{U}(0, 1) \nonumber\\
+    a_1 &\sim \mathcal{U}(0, 10) \nonumber\\
+    b_1 &\sim \mathcal{U}(0, 10) \nonumber\\
+    \cos{i}_{\star,n} &\sim \mathcal{U}(0, 1) \nonumber\\
+    \lambda_{\star,n} &\sim \mathcal{U}(0, \pi).
 \end{align}
@@
 Doing so will also avoid label switching in our mixture model.
+The probability distribution of an isotropic $i_{\rm star}$ follows $p_{i_\star} \sim \sin{i_{\star}}$, or $p_{\cos{i_\star}} \sim 1$.
+The $\lambda$ should be the angle between two random unit vectors on a 2D circle and always between 0 and $\pi$.
 
@@
 The right panel presents the $\cos{\psi}$ distribution for all \numall systems that have sky-projected stellar obliquity $\lambda$ measurements and has no likelihood function on $i_\star$.
-The $\cos{\psi}$ distribution for the observed-$i_\star$ are similar and both show a cluster near $\cos{\psi} = -0.2$, whereas the distribution for all systems does not have such a trend.}
+The $\cos{\psi}$ distribution for the observed-$i_\star$ are similar, and both show a cluster near $\cos{\psi} = -0.2$, whereas the distribution for all systems does not have such a trend.}
         \label{fig:psi_dist}
@@
 
-First, we apply the framework to the sample of \numistar systems that have both sky-projected stellar obliquity $\lambda$ and stellar inclination $i_\star$ constraints summarized in \cite{Albrecht21}.
+First, we apply the framework to the sample of \numistar systems that have both sky-projected stellar obliquity $\lambda$ and stellar inclination $i_\star$ constraints summarized in the \cite{Albrecht22} review.
+If the host star's rotation period is available, we use the value along with the stellar radii and $v\sin{i_\star}$ to infer the stellar inclination.
+Otherwise, we use the reported stellar inclination and uncertainty listed in Table A1 of \cite{Albrecht22}.
 The inferred $\cos{\psi}$ distribution is shown in the left panel of Figure~\ref{fig:psi_dist}.
-The distribution presents a pike at 1 as expected, suggesting the majority of exoplanetary systems are well-aligned systems.
+The distribution presents a pike at 1, suggesting the majority of exoplanetary systems are well-aligned systems.
 For misaligned systems, $\cos{\psi}$ presents a cluster near $\cos{\psi} = -0.2$, corresponding to $\psi$ of $100\degr$.
@@
 Interestingly, the overall $\cos{\psi}$ distributions of the two models are very similar, despite the fact that the inferred $\psi$ uncertainty for each system is now greater.
-The similarity between the two distributions suggests that, at least for the measured-$i_\star$ sample, the $\cos{\psi}$ distribution mainly depends on the underlying sky-project stellar obliquity $\lambda$ distribution.
+The similarity between the two distributions suggests that, at least for the observed-$i_\star$ sample, the $\cos{\psi}$ distribution mainly depends on the underlying sky-project stellar obliquity $\lambda$ distribution.
 
@@
 
-\subsection{Infer the \texorpdfstring{$\psi$}{psi} distribution from the \texorpdfstring{$\lambda$}{lambda} distribution}
+\subsection{Do we need \texorpdfstring{$i_\star$}{istar} to infer the \texorpdfstring{$\psi$}{psi} distribution?}
 
-There is no doubt that $i_\star$ measurement on individual systems improves the constraint on $\psi$.
-However, do $i_\star$ measurements matter when we constrain the population $\psi$ distribution?
-Particularly, the left and middle panels in Figure~\ref{fig:psi_dist} show that with or without $i_\star$ measurements, the inferred $\cos{\psi}$ distributions are very consistent.
+There is no doubt that $i_\star$ measurement on an individual system improves the constraint on $\psi$.
+However, are $i_\star$ measurements required to constrain the population $\psi$ distribution?
+Particularly, the left and middle panels in Figure~\ref{fig:psi_dist} show that without $i_\star$ measurements, a similar level of constraint on the $\cos{\psi}$ distribution can be achieved.
 We perform some calculations and tests to evaluate the necessity of $i_\star$ constraints.
 
-Given a certain $\psi$ distribution, the $\lambda$ and $i_\star$ are no longer randomly distributed.
-We first find relations between $\{\psi, \theta\}$ and $\{\lambda, i_\star\}$, where $\theta$ is the azimuthal angle of the stellar spin axis.
+\begin{figure*}
+    \script{coord_psi.py}
+    \vspace*{-1cm}
+    \gridline{
+        \fig{figures/coord_psi.pdf}{0.45\textwidth}{\vspace*{-1.8cm}(a) The $\{\psi, \theta\}$ coordinate system. The grey circle corresponds to a constant $\psi$ value.}
+        \fig{figures/coord_lam.pdf}{0.45\textwidth}{\vspace*{-1.8cm}(b) The $\{\lambda, i_\star\}$ coordinate system. The grey circle corresponds to a constant $i_\star$ value.}
+    }
+    \vspace*{-1.5cm}
+    \caption{Two coordinate systems to describe the stellar spin axis $\bf{n}_{\star}$ and the planet's orbital axis $\bf{n}_{\rm orb}$.}
+    \label{fig:coord}
+\end{figure*}
+
+Given a certain $\psi$ distribution, the $\lambda$ and $i_\star$ are no longer randomly/isotropically distributed.
+We first find relations between $\{\psi, \theta\}$ and $\{\lambda, i_\star\}$, where $\theta$ is the azimuthal angle of the stellar spin axis, as shown in Figure~\ref{fig:coord}.
+We transform $\bf{n}_{\star}$ in two coordinate systems to Cartesian, pair their terms, and get:
 \begin{align}
-    \sin{\psi}\cos{\theta} &= \cos{i_\star} \nonumber\\
-    \sin{\psi}\sin{\theta} &= \sin{\lambda}\sin{i_\star} \nonumber\\
-    \cos{\psi} &= \cos{\lambda}\sin{i_\star}.
+    \sin{\psi}\cos{\theta} = \cos{i_\star}& \nonumber\\
+    \sin{\psi}\sin{\theta} = \sin{\lambda}\sin{i_\star}& \nonumber\\
+    \cos{\psi} = \cos{\lambda}\sin{i_\star}&.
 \end{align}
-Thus, $i_\star = \cos[-1](\sin{\psi}\cos{\theta})$ and $\lambda = \cos[-1](\cos{\psi}/\sqrt{1-\cos^2{\theta}\sin^2{\psi}})$.
-If we assume $\theta$ is uniformly distributed, just from the $\lambda$ distribution, we could directly infer the $\psi$ distribution.
+Thus, $i_\star = \cos[-1](\sin{\psi}\cos{\theta})$ and $\lambda = \tan[-1](\tan{\psi}\sin{\theta})$. 
+We could calculate the probability distribution of $i_\star$ or $\lambda$ from the probability distributions of $\psi$ and $\theta$.
+Since $\psi$ and $\theta$ are independent random variables, it is possible to marginalize over $\theta$ to find a one-on-one relation between $\lambda$ and $\psi$, or even between $i_\star$ and $\psi$.
+If the azimuthal angle of the stellar rotation axis $\theta$ is randomly distributed, we could directly infer the $\psi$ distribution from the $\lambda$ distribution or even $i_\star$ distribution.
+An analytical solution between $p_\lambda$ and $p_\psi$ can be found in \cite{Fabrycky09}. 
+In our Bayesian framework, both $\psi$ and $\theta$ distributions are simultaneously inferred from the $\lambda$ and $i_\star$ distributions.
+Without any $i_\star$ measurements, the inferred $\theta$ indeed follows a uniform distribution.
+With $i_\star$ measurements, the inferred $\theta$ still follows a uniform distribution for aligned systems.
+For misaligned systems, the inferred $\theta$ distribution is similar to the $i_\star$ distribution as the $\psi$ are close to $90\degr$.
+For these systems, the $\lambda$ are close to $90\degr$ as well, so the $\theta$ angles are no longer important for the inference.
+Overall, these two reasons explain why the inferred $\cos{\psi}$ distributions are similar in the left and middle panels of Figure~\ref{fig:psi_dist}.
 
-We then design some $\cos{\psi}$ distributions.
-From each distribution of $\cos{\psi}$, we randomly draw 200 $i_\star$ and compute the $\lambda$.
-These $i_\star$ and $\lambda$ are the \emph{true} values.
-We use 200 samples so that the sample size is large enough to represent the distribution.
-In reality, the number of misaligned systems with $i_\star$ measurements is much smaller ($\sim 20$).
-Now $\lambda$ and $i_\star$ are added to some Gaussian noises to simulate the \emph{observed} values.
-We use uncertainties of $\sigma_{\lambda} = 8\degr$ and $\sigma_{i_\star} = 10\degr$ that represent typical observation uncertainties of misaligned systems.
-In a consequence test, we find using smaller uncertainties does not affect the distributions.
-Lastly, from the observed $\lambda$ and $i_\star$ and their uncertainties, we infer the $\cos{\psi}$ distribution of the sample with or without the $i_\star$ observations.
+We also design some $\cos{\psi}$ distributions and infer their distributions purely from the $\lambda$ distributions.
+From each distribution of $\cos{\psi}$ and assuming a uniform $\theta$, we randomly generate 200 samples of $i_\star$ and $\lambda$ and assume them as \emph{true} values.
+We then add some Gaussian noises to \emph{true} $\lambda$ and $i_\star$ to simulate \emph{observed} values.
+We use the uncertainties of $\sigma_{\lambda} = 8\degr$ and $\sigma_{i_\star} = 10\degr$ which are typical observation uncertainties.
+Lastly, from the observed values and their uncertainties, we infer the $\cos{\psi}$ distribution of the sample with or without $i_\star$ measurements.
+We test four different distributions of $\cos{\psi}$: a uniform distribution $\mathcal{U}(-1,1)$ and three Normal distributions $\mathcal{N}(0,0.2)$, $\mathcal{N}(-0.4,0.2)$, and $\mathcal{N}(0.4,0.2)$.
+In all four distributions, we find inferences with or without $i_\star$ are consistent with each other and injected distributions.
 
-We test four different distributions of $\cos{\psi}$: a uniform distribution on the interval $[-1,1]$ and three Normal distributions $\mathcal{N}(0,0.2)$, $\mathcal{N}(-0.4,0.2)$, and $\mathcal{N}(0.4,0.2)$.
-In all four distributions, the $\cos{\psi}$ distributions inferred with or without $i_\star$ likelihoods are consistent with each other and the injected distributions.
-
-We conclude that $i_\star$ measurements on individual systems are not necessary to infer the population $\psi$ distribution.
+In conclusion, the stellar obliquity distribution can be inferred directly from the sky-projected stellar obliquity distribution.
+A one-on-one relation could be found if we assume the azimuthal angle of the stellar spin axis is uniformly distributed around the planet's orbital axis.
 
@@
 
-Next, we find underlying $\lambda$ distributions of the measured-$i_\star$ sample and the full sample.
-We build another hierarchical Bayesian framework to infer the $\lambda$ distribution using a two-component mixture Beta distribution, similar to the one used for the $\cos{\psi}$ distribution.
-We extend the distribution support from $[0,1]$ to $[0,\pi]$ by multiplying the distribution by $\pi$.
+It was surprising that the inferred $\cos{\psi}$ distribution of the observed-$i_\star$ sample evidently differs from the $\cos{\psi}$ distribution of the full sample, shown as the middle and right panels of Figure~\ref{fig:psi_dist}.
+Since both distributions are inferred with no $i_\star$ information, the disparity is likely introduced by the underlying $\lambda$ distribution.
+
+We find the underlying $\lambda$ distributions of the observed-$i_\star$ sample and the full sample by building another hierarchical Bayesian model.
+We again use a two-component mixture Beta distribution to describe well-aligned and misaligned systems.
+We extend the Beta distribution support from $[0,1]$ to $[0,\pi]$ by multiplying the distribution by $\pi$ and use the observed $\lambda$ and uncertainties for the likelihoods.
+Figure~\ref{fig:lam_dist} shows the $\lambda$ distributions of two samples.
+It clearly shows that in the observed-$i_\star$ sample, misaligned systems are clustered at $\sim 110\degr$, whereas in the full sample, misaligned systems are nearly uniformly distributed.
 
@@
 
-Figure~\ref{fig:lam_dist} presents the $\lambda$ distributions of two samples. In the measured-$i_\star$ sample,
+One plausible explanation for the disparity between the two distributions is their small sample sizes.
+In the measured-$i_\star$ sample, only $21$ systems are misaligned  (i.e., $\lambda > 30\degr$), and in the full sample, $51$ systems are misaligned.
+It is likely that the 21 samples could not yet represent the population $\psi$ distribution.
 
-All planets with host star effective temperatures greater than 7000 K are polar planets in the observed-$i_\star$ sample.
-Such a trend is not found in the full sample.
-To measure the inclination of hot stars, stellar modulation is no longer feasible, whereas the gravity-darkening method becomes more important.
-However, the gravity-darkening method is only sensitive to detecting polar planets.
+Furthermore, the measured-$i_\star$ sample could be biased towards perpendicular planets.
+Misaligned systems are often found around hot stars.
+However, to measure the inclinations of hot stars, the most prolific inclination measurement method, the rotational modulation method, is no longer applicable to hot stars due to their lack of starspots.
+Instead, the gravity-darkening method is often used.
+The method, however, could be biased since it only detects planets transiting across stellar disks with large brightness variations (i.e., planet crossing the the stellar poles, or perpendicular planets).
+In the measured-$i_\star$ sample, $6/21$ systems have stellar inclinations constrained via gravity darkening.
+All six planets are misaligned and have nearly polar orbits:
+KELT-9 ($\lambda = 85 \pm 0.2\degr$),
+KELT-17 ($\lambda = 115.9 \pm 4.1 \degr$), 
+Kepler-13 ($\lambda = 59.2 \pm 0.1 \degr$), 
+MASCARA-1 ($\lambda = 69.5 \pm 3.0 \degr$),
+MASCARA-4 ($\lambda = 112.5 \pm 1.6 \degr$), 
+and WASP-189 ($\lambda = 89.3 \pm 1.4 \degr$).
+The bias introduced by the sample is significant enough to produce the clustering in the $\lambda$ distribution of the observed-$i_\star$ sample.
 
@@
 We also test the robustness of $\cos{\psi}$ distribution on different hyperprior choices \citep[see a nice summary in][]{Nagpal22}.
-We find extending the support of hyperpriors from 0--10 to 0--100 or 0--1000 or changing hyperprior distributions from Uniform to log-Uniform or truncated-Normal all infer similar results to Figure~\ref{fig:psi_dist}.
+We use different hyperprior distributions, including uniform distributions, log uniform distributions and truncated Normal distributions.
+In all hyperprior distributions, the clustering of perpendicular planets in observed-$i_\star$ sample is found and such clustering is much weaken in the full sample.
 
-\section{Hot Jupiter 3D Stellar Obliquity Distribution} \label{sec:hjs}
+\section{Discussion} \label{sec:discussion}
+The stellar obliquity distribution of exoplanetary systems can be inferred purely from their sky-projected stellar obliquity distribution.
+When we apply our framework to all systems with $\lambda$ measurements, we find misaligned systems have nearly isotropic stellar obliquity distribution.
+We compare the inferred distribution from the observation to the prediction of mutual inclination of different eccentricity excitation mechanisms of Hot Jupiters.
+It is in best agreement with the outcome of scattering of multiple giant planets after a convergent disk migration \citep{Beague12}.
 

### 2022-12-05 src/scripts/coord_psi.py
draft

@@ -162,3 +162,123 @@ plt.tight_layout()
 plt.savefig(paths.figures / "coord_psi.pdf", bbox_inches="tight", dpi=600)
 
 plt.close()
+
+
+
+
+fig = plt.figure(dpi=80)
+ax = fig.add_subplot(projection='3d')
+
+# Make data
+u = np.linspace(0, 2 * np.pi, 180)
+v = np.linspace(0, np.pi, 180)
+x = np.outer(np.cos(u), np.sin(v))
+y = np.outer(np.sin(u), np.sin(v))
+z = np.outer(np.ones(np.size(u)), np.cos(v))
+
+#Plot the surface
+ax.plot_surface(x, y, z, rstride=4, cstride=4, color='b', linewidth=0, alpha=0.1)
+
+elev = 10.0
+rot = 60.0 / 180 * np.pi
+
+#calculate vectors for "vertical" circle
+a = np.array([-np.sin(elev / 180 * np.pi), 0, np.cos(elev / 180 * np.pi)])
+b = np.array([0, 1, 0])
+b = b * np.cos(rot) + np.cross(a, b) * np.sin(rot) + a * np.dot(a, b) * (1 - np.cos(rot))
+ax.plot(np.sin(u),np.cos(u),0,color='k', linestyle = 'dashed')
+horiz_front = np.linspace(0, np.pi, 100)
+ax.plot(np.sin(horiz_front),np.cos(horiz_front),0,color='k')
+
+vert_front = np.linspace(np.pi / 2, 3 * np.pi / 2, 100)
+ax.plot(a[0] * np.sin(u) + b[0] * np.cos(u), b[1] * np.cos(u), a[2] * np.sin(u) + b[2] * np.cos(u),
+        color='k',lw=1)
+ax.plot(a[0] * np.sin(vert_front) + b[0] * np.cos(vert_front), b[1] * np.cos(vert_front), 
+        a[2] * np.sin(vert_front) + b[2] * np.cos(vert_front),
+        color='k',lw=1)
+
+
+ax.quiver(0, 0, 0, 1.4, 0, 0, color='k', arrow_length_ratio=0.05, lw=1.5) # x-axis
[…]

### 2022-12-05 src/scripts/lam_dist.py

@@ -29,6 +29,9 @@ plt.rcParams['ytick.major.width'] = 1.0
 plt.rcParams['ytick.minor.width'] =  1.0
 plt.rcParams['lines.markeredgewidth'] =  1.0
 
+numistar = 62
+numall = 161
+
 # load MCMC draws
 lam_all = az.from_netcdf(paths.data / 'lam_all.nc')
 lam_istar = az.from_netcdf(paths.data / 'lam_istar.nc')
@@ -67,7 +70,7 @@ plt.xticks(np.arange(0,185,30))
 plt.xlabel(r'$\lambda$ [$^\circ$]',fontsize=11)
 plt.ylabel('Probablity density function')
 
-plt.title(r'65 systems with observed $i_{\star}$',fontsize=11)
+plt.title(r'%i systems with observed $i_{\star}$'%numistar,fontsize=11)
 
 plt.subplot(1,2,2)
 q025, q16, q50, q84, q975 = np.percentile(lam_all_draws, [2.5, 16, 50, 84, 97.5], axis=1)
@@ -80,7 +83,7 @@ plt.xlim([0,180])
 
 plt.xticks(np.arange(0,185,30))
 
-plt.title(r'all 161 systems', fontsize=11)
+plt.title(r'all %i systems'%numall, fontsize=11)
 
 plt.tight_layout()
 plt.savefig(paths.figures / "lam_dist.pdf", bbox_inches="tight", dpi=600)

### 2022-12-05 src/scripts/psi_dist.py

@@ -29,6 +29,9 @@ plt.rcParams['ytick.major.width'] = 1.0
 plt.rcParams['ytick.minor.width'] =  1.0
 plt.rcParams['lines.markeredgewidth'] =  1.0
 
+numistar = 62
+numall = 161
+
 # load MCMC draws
 all_randinc = az.from_netcdf(paths.data / 'all_randinc.nc')
 istar_randinc = az.from_netcdf(paths.data / 'istar_randinc.nc')
@@ -72,7 +75,7 @@ plt.xlim([-1,1])
 plt.xlabel(r'$\cos{\psi}$',fontsize=11)
 plt.ylabel('Probablity density function')
 
-plt.title(r'\begin{center} 65 systems with observed $i_{\star}$ \\ \textbf{with} $i_{\star}$ likelihood \end{center}', 
+plt.title(r'\begin{center} %i systems with observed $i_{\star}$ \\ \textbf{with} $i_{\star}$ likelihood \end{center}'%numistar, 
           fontsize=11, y=1.1)
 
 
@@ -85,7 +88,7 @@ plt.fill_between(2*x-1, q025, q975, alpha=0.5, color='#cfe7ea')
 plt.ylim([0,2])
 plt.xlim([-1,1])
 
-plt.title(r'\begin{center} 65 systems with observed $i_{\star}$ \\ \textbf{no} $i_{\star}$ likelihood \end{center}', 
+plt.title(r'\begin{center} %i systems with observed $i_{\star}$ \\ \textbf{no} $i_{\star}$ likelihood \end{center}'%numistar, 
           fontsize=11, y=1.1)
 
 
@@ -98,7 +101,7 @@ plt.fill_between(2*x-1, q025, q975, alpha=0.3, color='#7dabd0')
 plt.ylim([0,2])
 plt.xlim([-1,1])
 
-plt.title(r'\begin{center} all 161 systems \\ \textbf{no} $i_{\star}$ likelihood \end{center}', 
+plt.title(r'\begin{center} all %i systems \\ \textbf{no} $i_{\star}$ likelihood \end{center}'%numall, 
           fontsize=11, y=1.1)
 
 plt.tight_layout()

### 2022-12-24 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-Many exoplanetary systems are found with high stellar obliquities.
-We build a hierarchical Bayesian framework to understand how stellar obliquities of misaligned systems are distributed.
-If stellar obliquities $\psi$ are isotropically distributed, the $\cos{\psi}$ distribution will be uniform, whereas if stellar obliquities have preferred angles, the $\cos{\psi}$ distribution will present concentrations.
-We first apply our framework to \numistar systems with both sky-projected stellar obliquity $\lambda$ and stellar inclination $i_\star$ constraints.
-The $\cos{\psi}$ distribution is clustered near $-0.2$ ($\psi \sim 100\degr$), in agreement with the preponderance of perpendicular planets.
-Interestingly, a similar $\cos{\psi}$ distribution could be inferred without using any $i_\star$ measurements, suggesting that the $\psi$ distribution depends mainly on the underlying $\lambda$ distribution.
-We then apply the framework to all \numall systems with sky-projected obliquity measurements.
-Unlike the observed-$i_\star$ sample, the $\cos{\psi}$ is  nearly uniformly distributed for misaligned systems, corresponding to an isotropic distribution of stellar obliquity.
-We find the $\lambda$ distribution of the observed-$i_\star$ sample is clustered at $110\degr$, whereas the $\lambda$ distribution of the full sample is nearly uniform.
-The observed-$i_\star$ sample could be biased because misaligned systems are often found around hot stars; however, hot stars' inclinations are often measured by the gravity-darkening method, which is biased toward detecting perpendicular planets.
-We conclude that misaligned planetary systems, mostly Hot Jupiters, have isotropic stellar obliquities.
-When compared to theoretical predictions, the distribution is best consistency with the scattering of multiple giant planets after their convergent disk migration.
+Stellar obliquity $\psi$, the angle between a planet's orbital axis and its host star's spin axis, has important implications for the formation environment and evolution of a planetary system.
+In exoplanet observations, only sky-projected stellar obliquity $\lambda$ can be measured.
+To find the stellar obliquity, information on the stellar inclination $i_\star$ is required.
+Here we show that, while it is important to measure the stellar inclination to tightly constrain the stellar obliquity of an individual system, stellar inclination is not necessarily required to infer the population stellar obliquity distribution.
+The stellar obliquity distribution can be inferred purely from the projected stellar obliquities, or even from the stellar inclinations.
+We introduce a flexible, hierarchical Bayesian framework to infer stellar obliquity distribution of exoplanetary systems that can be applied to a sample with measured projected stellar obliquities, stellar inclinations, or both.
+We apply the framework to all systems with measured projected stellar obliquities, which are mostly hot Jupiters, and find that the stellar obliquity distribution is unimodal and peaked at zero degrees.
+Misaligned systems have nearly isotropic stellar obliquity distribution.
+When compared to hot Jupiter stellar obliquity predictions, the distribution is most consistent with the outcome of the scattering of multiple giant planets after their convergent disk migration.
+
 \end{abstract}
@@
 
-The stellar obliquity $\psi$ describes the angle between a planet's orbital axis $\bf{n}_{\rm orb}$ and its host star's spin axis $\bf{n}_{\star}$.
-For any given $\bf{n}_{\rm orb}$ axis, if $\bf{n}_{\star}$ is a random vector with uniform distribution on a three-dimensional sphere, i.e., $\bf{n}_{\star}$ is isotropically distributed, the probability distribution of $\psi$ follows $p_{\psi} \sim \sin{\psi}$.
-Since $p_{\cos{\psi}} = p_{\psi} \abs{\dv*{\psi}{\cos{\psi}}}$, $p_{\cos{\psi}}$ is uniformly distributed between $-1$ and $1$.
-If the stellar obliquity is not isotropically distributed but have some preferred angles, the $\cos{\psi}$ distribution will present concentrations at certain values.
+Stellar obliquity $\psi$ describes the angle between a planet's orbital axis $\bf{n}_{\rm orb}$ and its host star's spin axis $\bf{n}_{\star}$.
+The angle has implications for the formation environment of a planetary system and its consequently dynamical evolution.
+Exoplanetary systems present a wide range of stellar obliquities.
 
 In reality, we only measure the sky-projected stellar obliquity $\lambda$, the angle between the projections of $\bf{n}_{\rm orb}$ and $\bf{n}_{\star}$ onto the plane of the sky.
-Stellar obliquity $\psi$ can be computed as \citep[e.g.,][]{Fabrycky09}:
+Stellar obliquity $\psi$ can be computed from $\lambda$ as \citep[e.g.,][]{Fabrycky09}:
 \begin{equation}\label{eqn:psi}
@@
 where $i_{\rm orb}$ is the angle between the vector $\bf{n}_{\rm orb}$ and the observer's line of sight, and $i_\star$ is the angle between $\bf{n}_{\star}$ and the observer's line of sight.
-Since the sky-projected stellar obliquities can be measured only in transiting exoplanet systems, $i_{\rm orb} \approx 90\degr$ and $\cos{\psi} \approx \sin{i_\star}\cos{\lambda}$.
+Since the sky-projected stellar obliquities can be measured only on transiting exoplanet systems, $i_{\rm orb} \approx 90\degr$ and $\cos{\psi} \approx \sin{i_\star}\cos{\lambda}$.
 
@@
 The stellar inclination can often be constrained via, e.g., photometric and spectroscopic rotational modulation introduced by starspots, gravity darkening, and asteroseismology.
-For systems without measured $i_\star$, it is still possible to infer the 3D stellar obliquity $\psi$ just from the sky-projected obliquity $\lambda$ assuming the stellar inclination $i_\star$ is isotropically distributed \citep[e.g.,][]{Fabrycky09}.
+Since the stellar inclination measurement is degenerate between $i_\star$ and $\pi - i_\star$, we assume $i_\star$ is less than $90\degr$.
 
-\cite{Albrecht21} inferred the 3D stellar obliquities of a sample of stars that have both sky-projected stellar obliquity and stellar inclination constraints.
-They found while the majority of planets have a small $\psi$, misaligned planets have $\psi$ clustered between 80--125$\degr$.
-Furthermore, the $\cos{\psi}$ distribution of misaligned systems is not uniformly distributed, but clusters near 0, suggesting the stellar spin axis is not isotropically distributed around the planetary orbital axis.
+For systems without measured $i_\star$, it is still possible to infer the 3D stellar obliquity $\psi$ just from the sky-projected obliquity $\lambda$, assuming the stellar inclination $i_\star$ is isotropically distributed.
+
+In this work, we find an approach to infer $\psi$ distribution from $\lambda$ distribution.
+
+\begin{figure*}[ht!]
+    \script{coord_psi.py}
+    \gridline{
+        \fig{figures/coord_psi.pdf}{0.45\textwidth}{\vspace*{-1.8cm}(a) The $\{\psi, \theta\}$ coordinate system. The grey circle corresponds to a constant $\psi$ value.}
+        \fig{figures/coord_lam.pdf}{0.45\textwidth}{\vspace*{-1.8cm}(b) The $\{\lambda, i_\star\}$ coordinate system. The grey circle corresponds to a constant $i_\star$ value.}
+    }
+    \vspace*{-1.5cm}
+    \caption{Two coordinate systems to describe the stellar spin axis $\bf{n}_{\star}$ and the planet's orbital axis $\bf{n}_{\rm orb}$.}
+    \label{fig:coord}
+\end{figure*}
 
 In this work, we build a hierarchical Bayesian framework to constrain the stellar obliquity distribution of exoplanetary systems, particularly of misaligned systems.
+In Section~\ref{sec:jacobian}
 In Section~\ref{sec:hbm}, we describe our hierarchical Bayesian framework and show inferred stellar obliquity distributions. 
-In Section~\ref{sec:interpretations}, we interpret the results. 
-In Section~\ref{sec:discussion}, we discuss the implication of the stellar obliquity distribution for Hot Jupiter origins.
+In Section~\ref{sec:application}, we apply our framework to observed exoplanetary systems.
 
-\section{3D Stellar Obliquity Distributions} \label{sec:hbm}
+\section{Transformation between the $\psi$ and $\lambda$ distributions}\label{sec:jacobian}
+
+For any given $\bf{n}_{\rm orb}$ axis, if $\bf{n}_{\star}$ is a random vector with uniform distribution on a three-dimensional sphere, i.e., $\bf{n}_{\star}$ is isotropically distributed, the probability distribution of $\psi$ follows $p_{\psi} \sim \sin{\psi}$.
+Since $p_{\cos{\psi}} = p_{\psi} \abs{\dv*{\psi}{\cos{\psi}}}$, $p_{\cos{\psi}}$ is uniformly distributed between $-1$ and $1$.
+If the stellar obliquity instead has some preferred angles, the $\cos{\psi}$ distribution will present concentrations at certain values.
+
+Given a certain $\psi$ distribution, the $\lambda$ and $i_\star$ are no longer randomly/isotropically distributed.
+We first find relations between $\{\psi, \theta\}$ and $\{\lambda, i_\star\}$, where $\theta$ is the azimuthal angle of the stellar spin axis, as shown in Figure~\ref{fig:coord}.
+We transform $\bf{n}_{\star}$ in two coordinate systems to Cartesian, pair their terms, and get:
+\begin{align}
+    \sin{\psi}\cos{\theta} = \cos{i_\star}& \label{eq:coord1}\\
+    \sin{\psi}\sin{\theta} = \sin{\lambda}\sin{i_\star}& \label{eq:coord2}\\
+    \cos{\psi} = \cos{\lambda}\sin{i_\star} \label{eq:coord3}&.
+\end{align}
+Thus, $i_\star = \cos[-1](\sin{\psi}\cos{\theta})$ and $\lambda = \tan[-1](\tan{\psi}\sin{\theta})$.
+
+First, we find the $\lambda$ distribution for a given $\cos{\psi}$ distribution.
+If the stellar obliquity is isotropically distributed, the $\cos{\psi}$ distribution will be uniform, and the $\lambda$ distribution will also be uniform.
+If the stellar obliquity has preferred angles, i.e., the $\cos{\psi}$ distribution will have concentrations, what will be the corresponding $\lambda$ distribution?
+We first find the distribution of $\cos{\lambda}$ using the Jacobian transformation from $\cos{\psi}$.
+Replacing $\sin{i_\star}$ in Equation~(\ref{eq:coord3}) with Equation~(\ref{eq:coord1}), we find $\cos{\lambda} = \cos{\psi}/\sqrt{1-(1-\cos^2{\psi})\cos^2{\theta}}$.
+Reorganize the equation, and we also get $\cos^2{\psi} = \frac{\cos^2{\lambda}\cos^2{\theta}-\cos^2{\lambda}}{\cos^2{\lambda}\cos^2{\theta}-1}$.
+The distribution transformation follows
+\begin{equation}
+    p_{\cos{\lambda}} = 2 \int_{0}^{1} \abs{\pdv{\cos{\psi}}{\cos{\lambda}}} p_{\cos{\psi}} p_{\cos{\theta}} d\cos{\theta}.
+\end{equation}
+We have $\abs{\pdv{\cos{\psi}}{\cos{\lambda}}} = \frac{(1-\cos^2{\theta})^{1/2}}{(1-\cos^2{\theta}\cos^2{\lambda})^{3/2}}$.
+Assuming $\theta$ is uniformly distributed between $-\pi/2$ and $\pi/2$\footnote{The $\theta$ is defined between $-\pi/2$ and $\pi/2$ since we limit $0 \leq i_\star \leq \pi/2$.}, $p_{\cos{\theta}} = p_{\theta} \abs{\dv*{\theta}{\cos{\theta}}} = 1/\pi/(1-\cos^2{\theta)^{1/2}}$.
+Combining all the parts together, we get
+\begin{equation}\label{eqn:jac}
+    p_{\cos{\lambda}} = \frac{2}{\pi} \int_{0}^{1} (1-\cos^2{\theta}\cos^2{\lambda})^{-3/2} p_{\cos{\psi}} d\cos{\theta}.
+\end{equation}
+
+If $\cos{\psi}$ is uniformly distribution, i.e., $p_{\cos{\psi}} = 1/2$, the integral derives $p_{\cos{\lambda}} = 1/\pi/\sqrt{1-\cos^2{\lambda}}$, which is equivalent to $p_\lambda = 1/\pi$.
+The $\lambda$ is uniformly distributed.
+If $\cos{\psi}$ follows other distributions, we may integrate Equation~(\ref{eqn:jac}) numerically, find the $\cos{\lambda}$ distribution, and solve for the $\lambda$ distribution.
+
+Similarly, we could derive the distribution of $\cos{i_\star}$ for a given $\cos{\psi}$ distribution via the Jacobian transformation and marginalizing over $\cos{\theta}$. We begin with
+\begin{equation}
+    p_{\cos{i_\star}} = 4 \int_{0}^{1} \abs{\pdv{\sin{\psi}}{\cos{i_\star}}} p_{\sin{\psi}} p_{\cos{\theta}} d\cos{\theta},
+\end{equation}
+and from Equation(\ref{eq:coord1}), we find $\sin{\psi} = \cos{i_\star}/\cos{\theta}$ and $\abs{\pdv{\sin{\psi}}{\cos{i_\star}}} = 1/\cos{\theta}$.
+Similarly, we assume $\theta$ is uniformly distributed, and $p_{\cos{\theta}} = 1/\pi/(1-\cos^2{\theta)^{1/2}}$.
+Lastly, we transform the $p_{\sin{\psi}}$ to $p_{\cos{\psi}}$, where $p_{\sin{\psi}} = p_{\cos{\psi}}\sqrt{1-\cos^2{\psi}}/\cos{\psi}$.
+Combining all the pieces together, we get
+\begin{equation}
+    p_{\cos{i_\star}} = \frac{4}{\pi} \int_{\cos{i_\star}}^{1} \frac{\cos{i_\star}/\cos{\theta}}{\sqrt{\cos^2{\theta}-\cos^2{i_\star}}} \frac{1}{\sqrt{1-\cos^2{\theta}}} p_{\cos{\psi}} d\cos{\theta}.
+\end{equation}
+If $\cos{\psi}$ is uniformly distribution, i.e., $p_{\cos{\psi}} = 1/2$, the integral gives $1$, which suggests the $\cos{i_\star}$ is uniform.
+
+We note that we could not infer $\psi$ from $\lambda$ and $\theta$ since they are not independent variables. Instead, we will have to infer $\psi$ from $\lambda$ and $i_\star$.
+
+\begin{figure*}[ht!]
+    \script{transform.py}
+    \includegraphics[width=\linewidth]{figures/transform.pdf}
+    \caption{Distribution transformation.}
+    \label{fig:transform}
+\end{figure*}
+
+\section{Hierarchical Bayesian Framework} \label{sec:hbm}
+
+\begin{figure}[ht!]
+    \script{graph.py}
+    \includegraphics[width=\linewidth]{figures/graph.pdf}
+    \caption{Hierarchical Bayesian framework to infer the stellar obliquity distribution of exoplanetary systems.
+The stellar obliquity distribution is described by hyperparameters $\vb*{\beta}$ and constrained by $n_{\rm th}$ system's stellar obliquity $\psi_n$.
+Each $\psi_n$ is calculated by the system's sky-projected stellar obliquity $\lambda_n$ and stellar inclination $i_{\star, n}$.
+$\theta_{\star, n}$ contains other information of the star other than $i_{\star, n}$, such as the stellar radius and stellar rotation period.
+Obs$_{\star,n}$ contains observations of the star, such as the stellar spectral and photometric and/or spectroscopic time series, and $\hat{\lambda}_n$ has the measured sky-projected stellar obliquity.}
+    \label{fig:graph}
+\end{figure}
+
 We build a hierarchical Bayesian framework to model the 3D stellar obliquity ($\psi$) distribution of exoplanetary systems.
@@
 Lastly, the sky-projected stellar obliquity $\lambda_n$ is constrained by observed $\hat{\lambda}_n$.
-
-\begin{figure}
-    \script{graph.py}
-    \includegraphics[width=\linewidth]{figures/graph.pdf}
-    \caption{Hierarchical Bayesian framework to infer the stellar obliquity distribution of exoplanetary systems.
-The stellar obliquity distribution is described by hyperparameters $\vb*{\beta}$ and constrained by $n_{\rm th}$ system's stellar obliquity $\psi_n$.
-Each $\psi_n$ is calculated by the system's sky-projected stellar obliquity $\lambda_n$ and stellar inclination $i_{\star, n}$.
-$\theta_{\star, n}$ contains other information of the star other than $i_{\star, n}$, such as the stellar radius and stellar rotation period.
-Obs$_{\star,n}$ contains observations of the star, such as the stellar spectral and photometric and/or spectroscopic time series, and $\hat{\lambda}_n$ has the measured sky-projected stellar obliquity.}
-    \label{fig:graph}
-\end{figure}
 
@@
 
-\begin{figure*}
+\begin{figure*}[ht!]
     \script{psi_dist.py}
@@
 
-\begin{figure*}
-    \script{coord_psi.py}
-    \vspace*{-1cm}
-    \gridline{
-        \fig{figures/coord_psi.pdf}{0.45\textwidth}{\vspace*{-1.8cm}(a) The $\{\psi, \theta\}$ coordinate system. The grey circle corresponds to a constant $\psi$ value.}
-        \fig{figures/coord_lam.pdf}{0.45\textwidth}{\vspace*{-1.8cm}(b) The $\{\lambda, i_\star\}$ coordinate system. The grey circle corresponds to a constant $i_\star$ value.}
-    }
-    \vspace*{-1.5cm}
-    \caption{Two coordinate systems to describe the stellar spin axis $\bf{n}_{\star}$ and the planet's orbital axis $\bf{n}_{\rm orb}$.}
-    \label{fig:coord}
-\end{figure*}
-
 Given a certain $\psi$ distribution, the $\lambda$ and $i_\star$ are no longer randomly/isotropically distributed.
@@
 We again use a two-component mixture Beta distribution to describe well-aligned and misaligned systems.
-We extend the Beta distribution support from $[0,1]$ to $[0,\pi]$ by multiplying the distribution by $\pi$ and use the observed $\lambda$ and uncertainties for the likelihoods.
+We extend the Beta distribution support from $[0,1]$ to $[0,\pi]$ by multiplying the distribution by $\pi$.
+We use the observed $\lambda$ and uncertainties for the likelihoods.
 Figure~\ref{fig:lam_dist} shows the $\lambda$ distributions of two samples.
@@
 
-\begin{figure*}
+\begin{figure*}[ht!]
     \script{lam_dist.py}
@@
 Instead, the gravity-darkening method is often used.
-The method, however, could be biased since it only detects planets transiting across stellar disks with large brightness variations (i.e., planet crossing the the stellar poles, or perpendicular planets).
+The method, however, could be biased since it only detects planets transiting across stellar disks with large brightness variations (i.e., planet crossing the stellar poles or perpendicular planets).
 In the measured-$i_\star$ sample, $6/21$ systems have stellar inclinations constrained via gravity darkening.
@@
 We also test the robustness of $\cos{\psi}$ distribution on different hyperprior choices \citep[see a nice summary in][]{Nagpal22}.
-We use different hyperprior distributions, including uniform distributions, log uniform distributions and truncated Normal distributions.
-In all hyperprior distributions, the clustering of perpendicular planets in observed-$i_\star$ sample is found and such clustering is much weaken in the full sample.
+We use different hyperprior distributions, including uniform distributions, log uniform distributions, and truncated Normal distributions.
+In all hyperprior distributions, the clustering of perpendicular planets in the observed-$i_\star$ sample is found, and such clustering is much weakened in the full sample.
 
 \section{Discussion} \label{sec:discussion}
+
+\cite{Albrecht21} inferred the 3D stellar obliquities of a sample of stars that have both sky-projected stellar obliquity and stellar inclination constraints.
+They found while the majority of planets have a small $\psi$, misaligned planets have $\psi$ clustered between 80--125$\degr$.
+Furthermore, the $\cos{\psi}$ distribution of misaligned systems is not uniformly distributed but clusters near 0, suggesting the stellar spin axis is not isotropically distributed around the planetary orbital axis.
+
 The stellar obliquity distribution of exoplanetary systems can be inferred purely from their sky-projected stellar obliquity distribution.
@@
 We compare the inferred distribution from the observation to the prediction of mutual inclination of different eccentricity excitation mechanisms of Hot Jupiters.
-It is in best agreement with the outcome of scattering of multiple giant planets after a convergent disk migration \citep{Beague12}.
+It is in best agreement with the outcome of the scattering of multiple giant planets after a convergent disk migration \citep{Beague12}.
 

### 2022-12-24 src/scripts/transform.py
[showyourwork] overleaf sync

@@ -1,0 +1,426 @@
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+
+import scipy.integrate as integrate
+from scipy import stats
+
+subprocess.check_call([sys.executable, '-m', 'pip', 'install',
+'arviz'])
+subprocess.check_call([sys.executable, '-m', 'pip', 'install',
+'pymc'])
+subprocess.check_call([sys.executable, '-m', 'pip', 'install',
+'aesara'])
+
+import pymc as pm
+import arviz as az
+import aesara.tensor as at
+
+
+from matplotlib import rc
+rc('font', **{'family':'sans-serif'})
+rc('text', usetex=True)
+rc('text.latex', preamble=r'\usepackage{physics}')
+
+plt.rcParams['xtick.top'] =  True
+plt.rcParams['xtick.direction'] =  'in'
+plt.rcParams['xtick.major.width'] =  1.0
+plt.rcParams['xtick.minor.width'] =  1.0
+plt.rcParams['ytick.right'] =  True
+plt.rcParams['ytick.direction'] =  'in'
+plt.rcParams['ytick.major.width'] = 1.0
+plt.rcParams['ytick.minor.width'] =  1.0
+plt.rcParams['lines.markeredgewidth'] =  1.0
+
+### PyMC models ###
+
+with pm.Model() as uni:
+
+    cosψ = pm.Uniform('cosψ',lower=-1,upper=1)
+    
+    ψ = pm.Deterministic('ψ', at.arccos(cosψ))
+    sinψ = pm.Deterministic('sinψ', at.sqrt(1-cosψ**2))
+    
+    θ = pm.Uniform('θ', lower=-np.pi/2, upper=np.pi/2)
+    cosθ = pm.Deterministic('cosθ', at.cos(θ))
+    tanθ = pm.Deterministic('tanθ', at.tan(θ))
+
+    cosλ = pm.Deterministic('cosλ', cosψ/at.sqrt(1-sinψ**2*cosθ**2))
+    λ = pm.Deterministic('λ', at.arccos(cosλ))
+    
[…]

### 2022-12-24 src/scripts/transform.py
[showyourwork] overleaf sync

@@ -5,6 +5,7 @@ import pandas as pd
 import scipy.integrate as integrate
 from scipy import stats
 
+import subprocess
 subprocess.check_call([sys.executable, '-m', 'pip', 'install',
 'arviz'])
 subprocess.check_call([sys.executable, '-m', 'pip', 'install',

### 2022-12-24 src/scripts/transform.py
[showyourwork] overleaf sync

@@ -6,6 +6,7 @@ import scipy.integrate as integrate
 from scipy import stats
 
 import subprocess
+import sys
 subprocess.check_call([sys.executable, '-m', 'pip', 'install',
 'arviz'])
 subprocess.check_call([sys.executable, '-m', 'pip', 'install',

### 2022-12-24 src/scripts/transform.py
[showyourwork] overleaf sync

@@ -10,7 +10,7 @@ import sys
 subprocess.check_call([sys.executable, '-m', 'pip', 'install',
 'arviz'])
 subprocess.check_call([sys.executable, '-m', 'pip', 'install',
-'pymc'])
+'pymc=4.1.7'])
 subprocess.check_call([sys.executable, '-m', 'pip', 'install',
 'aesara'])
 

### 2022-12-24 src/scripts/transform.py
[showyourwork] overleaf sync

@@ -10,7 +10,7 @@ import sys
 subprocess.check_call([sys.executable, '-m', 'pip', 'install',
 'arviz'])
 subprocess.check_call([sys.executable, '-m', 'pip', 'install',
-'pymc=4.1.7'])
+'pymc==4.1.7'])
 subprocess.check_call([sys.executable, '-m', 'pip', 'install',
 'aesara'])
 

### 2022-12-24 src/scripts/transform.py
[showyourwork] overleaf sync

@@ -36,98 +36,100 @@ plt.rcParams['lines.markeredgewidth'] =  1.0
 
 ### PyMC models ###
 
-with pm.Model() as uni:
+if __name__ == '__main__':
 
-    cosψ = pm.Uniform('cosψ',lower=-1,upper=1)
-    
-    ψ = pm.Deterministic('ψ', at.arccos(cosψ))
-    sinψ = pm.Deterministic('sinψ', at.sqrt(1-cosψ**2))
-    
-    θ = pm.Uniform('θ', lower=-np.pi/2, upper=np.pi/2)
-    cosθ = pm.Deterministic('cosθ', at.cos(θ))
-    tanθ = pm.Deterministic('tanθ', at.tan(θ))
+    with pm.Model() as uni:
 
-    cosλ = pm.Deterministic('cosλ', cosψ/at.sqrt(1-sinψ**2*cosθ**2))
-    λ = pm.Deterministic('λ', at.arccos(cosλ))
-    
-    sini = pm.Deterministic('sini', cosψ/cosλ)
-    i = pm.Deterministic('i', at.arcsin(sini))
-    cosi = pm.Deterministic('cosi', at.sqrt(1-sini**2))
-    
-    iso_cosi = pm.Uniform('iso_cosi', lower=0, upper=1)
-    iso_cosψ = pm.Deterministic('iso_cosψ', at.sqrt(1-iso_cosi**2)*cosλ)
-    
-    uni = pm.sample()
+        cosψ = pm.Uniform('cosψ',lower=-1,upper=1)
+        
+        ψ = pm.Deterministic('ψ', at.arccos(cosψ))
+        sinψ = pm.Deterministic('sinψ', at.sqrt(1-cosψ**2))
+        
+        θ = pm.Uniform('θ', lower=-np.pi/2, upper=np.pi/2)
+        cosθ = pm.Deterministic('cosθ', at.cos(θ))
+        tanθ = pm.Deterministic('tanθ', at.tan(θ))
+
+        cosλ = pm.Deterministic('cosλ', cosψ/at.sqrt(1-sinψ**2*cosθ**2))
+        λ = pm.Deterministic('λ', at.arccos(cosλ))
+        
+        sini = pm.Deterministic('sini', cosψ/cosλ)
+        i = pm.Deterministic('i', at.arcsin(sini))
+        cosi = pm.Deterministic('cosi', at.sqrt(1-sini**2))
+        
[…]

### 2022-12-24 src/scripts/transform.py
[showyourwork] overleaf sync

@@ -131,300 +131,302 @@ if __name__ == '__main__':
         
         norm3 = pm.sample()
 
-### Make the plot ###
-
-plt.figure(figsize=(7,5),dpi=110)
 
 
-### cosψ ~ Uniform(-1,1) ###
+    ### Make the plot ###
 
-cosψ = uni.posterior.cosψ.values.ravel()
-λ = uni.posterior.λ.values.ravel()
-i = uni.posterior.i.values.ravel()
-iso_cosψ = uni.posterior.iso_cosψ.values.ravel()
-
-plt.subplot(4,4,1)
-plt.hist(cosψ, color='#e5e1e0', density=True, bins=40)
-plt.title(r'$\cos{\psi} \sim \mathcal{U}(-1,1)$')
-plt.ylim([0,0.6])
-plt.xlim([-1,1])
-
-plt.subplot(4,4,2)
-plt.hist(λ, color='#e5e1e0', density=True, bins=40)
-plt.title(r'$\lambda$')
-plt.ylim([0,0.5])
-plt.xlim([0,np.pi])
-plt.xticks([0,np.pi/2,np.pi], ['0', r'$\pi/2$', r'$\pi$'])
-
-plt.subplot(4,4,3)
-plt.hist(i, color='#e5e1e0', density=True, bins=40)
-plt.title(r'$i_\star$')
-plt.xlim([0,np.pi/2])
-plt.xticks([0,np.pi/4,np.pi/2], ['0', r'$\pi/4$', r'$\pi/2$'])
-plt.ylim([0,1.1])
-
-plt.subplot(4,4,4)
-plt.hist(iso_cosψ, color='#e5e1e0', density=True, bins=40)
-plt.title(r'Inferred $\cos{\psi}$')
-plt.ylim([0,0.6])
-plt.xlim([-1,1])
+    plt.figure(figsize=(7,5),dpi=110)
 
 
-cosψ = np.linspace(-1,1,200)
-pcosψ = np.ones_like(cosψ)/2
+    ### cosψ ~ Uniform(-1,1) ###
 
-λ = np.linspace(0,np.pi,200)
-cosλ = np.cos(λ)
-pcosλ = np.zeros_like(cosλ)
-for i,this_cosλ in enumerate(cosλ):
-    pcosλ[i] = integrate.quad(lambda x: (1-this_cosλ**2*x**2)**-1.5*1/2,
-                         0, 1)[0]*2/np.pi
[…]

### 2022-12-24 src/scripts/transform.py
[showyourwork] overleaf sync

@@ -1,9 +1,6 @@
-import matplotlib.pyplot as plt
-import numpy as np
-import pandas as pd
-
-import scipy.integrate as integrate
-from scipy import stats
+import os
+from pathlib import Path
+os.environ["PATH"] += os.pathsep + str(Path.home() / "bin")
 
 import subprocess
 import sys
@@ -18,6 +15,12 @@ import pymc as pm
 import arviz as az
 import aesara.tensor as at
 
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+
+import scipy.integrate as integrate
+from scipy import stats
 
 from matplotlib import rc
 rc('font', **{'family':'sans-serif'})
@@ -426,7 +429,6 @@ if __name__ == '__main__':
     plt.legend(framealpha=0,handlelength=1,handletextpad=0.3,borderpad=0.1)
 
     plt.tight_layout()
-    plt.show()
 
     plt.savefig(paths.figures / "transform.pdf", bbox_inches="tight", dpi=600)
     plt.close()

### 2022-12-25 src/scripts/simulation.py
[showyourwork] overleaf sync

@@ -1,0 +1,307 @@
+import paths
+import os
+from pathlib import Path
+os.environ["PATH"] += os.pathsep + str(Path.home() / "bin")
+
+import subprocess
+import sys
+subprocess.check_call([sys.executable, '-m', 'pip', 'install',
+'arviz'])
+subprocess.check_call([sys.executable, '-m', 'pip', 'install',
+'pymc==4.1.7'])
+subprocess.check_call([sys.executable, '-m', 'pip', 'install',
+'aesara'])
+
+import pymc as pm
+import arviz as az
+import aesara.tensor as at
+
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+
+import scipy.integrate as integrate
+from scipy import stats
+
+from matplotlib import rc
+rc('font', **{'family':'sans-serif'})
+rc('text', usetex=True)
+rc('text.latex', preamble=r'\usepackage{physics}')
+
+plt.rcParams['xtick.top'] =  True
+plt.rcParams['xtick.direction'] =  'in'
+plt.rcParams['xtick.major.width'] =  1.0
+plt.rcParams['xtick.minor.width'] =  1.0
+plt.rcParams['ytick.right'] =  True
+plt.rcParams['ytick.direction'] =  'in'
+plt.rcParams['ytick.major.width'] = 1.0
+plt.rcParams['ytick.minor.width'] =  1.0
+plt.rcParams['lines.markeredgewidth'] =  1.0
+
+from scipy.stats import beta
+
+# return posteriors of HBM models with or without using measured istar
+def posteriors(this_model):
+    
+    nsample = 200
+    err_istar = 10*np.pi/180
+    err_lam = 8*np.pi/180
+
+    x = np.linspace(1e-5,1-1e-5,1000)
+
+    with this_model:    
+        idata = pm.sample(draws=50)
+
+    true_istar = idata.posterior.i.values.ravel()
+    obs_istar = true_istar + err_istar*np.random.normal(size=nsample)
+
+    true_lam = idata.posterior.λ.values.ravel()
+    obs_lam = true_lam + err_lam*np.random.normal(size=nsample)
+
[…]

### 2022-12-25 src/scripts/simulation.py
[showyourwork] overleaf sync

@@ -131,6 +131,95 @@ def posteriors(this_model):
 
 if __name__ == '__main__':
 
+
+    with pm.Model() as model_uni:
+
+        cosψ = pm.Uniform('cosψ',lower=-1,upper=1)
+        
+        ψ = pm.Deterministic('ψ', at.arccos(cosψ))
+        sinψ = pm.Deterministic('sinψ', at.sqrt(1-cosψ**2))
+        
+        θ = pm.Uniform('θ', lower=-np.pi/2, upper=np.pi/2)
+        cosθ = pm.Deterministic('cosθ', at.cos(θ))
+        tanθ = pm.Deterministic('tanθ', at.tan(θ))
+
+        cosλ = pm.Deterministic('cosλ', cosψ/at.sqrt(1-sinψ**2*cosθ**2))
+        λ = pm.Deterministic('λ', at.arccos(cosλ))
+        
+        sini = pm.Deterministic('sini', cosψ/cosλ)
+        i = pm.Deterministic('i', at.arcsin(sini))
+        cosi = pm.Deterministic('cosi', at.sqrt(1-sini**2))
+        
+        iso_cosi = pm.Uniform('iso_cosi', lower=0, upper=1)
+        iso_cosψ = pm.Deterministic('iso_cosψ', at.sqrt(1-iso_cosi**2)*cosλ)
+        
+
+    with pm.Model() as model_norm1:
+
+        cosψ = pm.Normal('cosψ', mu=0., sigma=0.2)
+        
+        ψ = pm.Deterministic('ψ', at.arccos(cosψ))
+        sinψ = pm.Deterministic('sinψ', at.sqrt(1-cosψ**2))
+        
+        θ = pm.Uniform('θ', lower=-np.pi/2, upper=np.pi/2)
+        cosθ = pm.Deterministic('cosθ', at.cos(θ))
+        tanθ = pm.Deterministic('tanθ', at.tan(θ))
+
+        cosλ = pm.Deterministic('cosλ', cosψ/at.sqrt(1-sinψ**2*cosθ**2))
+        λ = pm.Deterministic('λ', at.arccos(cosλ))
+        
+        sini = pm.Deterministic('sini', cosψ/cosλ)
+        i = pm.Deterministic('i', at.arcsin(sini))
+        cosi = pm.Deterministic('cosi', at.sqrt(1-sini**2))
+        
+        iso_cosi = pm.Uniform('iso_cosi', lower=0, upper=1)
[…]

### 2022-12-25 src/scripts/simulation.py
[showyourwork] overleaf sync

@@ -39,6 +39,7 @@ plt.rcParams['ytick.minor.width'] =  1.0
 plt.rcParams['lines.markeredgewidth'] =  1.0
 
 from scipy.stats import beta
+x = np.linspace(1e-5,1-1e-5,1000)
 
 # return posteriors of HBM models with or without using measured istar
 def posteriors(this_model):
@@ -46,8 +47,6 @@ def posteriors(this_model):
     nsample = 200
     err_istar = 10*np.pi/180
     err_lam = 8*np.pi/180
-
-    x = np.linspace(1e-5,1-1e-5,1000)
 
     with this_model:    
         idata = pm.sample(draws=50)

### 2022-12-25 src/scripts/simulation.py
[showyourwork] overleaf sync

@@ -40,6 +40,7 @@ plt.rcParams['lines.markeredgewidth'] =  1.0
 
 from scipy.stats import beta
 x = np.linspace(1e-5,1-1e-5,1000)
+σ = 0.2
 
 # return posteriors of HBM models with or without using measured istar
 def posteriors(this_model):

### 2022-12-29 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-Stellar obliquity $\psi$, the angle between a planet's orbital axis and its host star's spin axis, has important implications for the formation environment and evolution of a planetary system.
+Stellar obliquity $\psi$, the angle between a planet's orbital axis and its host star's spin axis, traces the formation environment and dynamical evolution of a planetary system.
 In exoplanet observations, only sky-projected stellar obliquity $\lambda$ can be measured.
-To find the stellar obliquity, information on the stellar inclination $i_\star$ is required.
-Here we show that, while it is important to measure the stellar inclination to tightly constrain the stellar obliquity of an individual system, stellar inclination is not necessarily required to infer the population stellar obliquity distribution.
-The stellar obliquity distribution can be inferred purely from the projected stellar obliquities, or even from the stellar inclinations.
-We introduce a flexible, hierarchical Bayesian framework to infer stellar obliquity distribution of exoplanetary systems that can be applied to a sample with measured projected stellar obliquities, stellar inclinations, or both.
-We apply the framework to all systems with measured projected stellar obliquities, which are mostly hot Jupiters, and find that the stellar obliquity distribution is unimodal and peaked at zero degrees.
-Misaligned systems have nearly isotropic stellar obliquity distribution.
-When compared to hot Jupiter stellar obliquity predictions, the distribution is most consistent with the outcome of the scattering of multiple giant planets after their convergent disk migration.
+To find the stellar obliquity from the projected obliquity, information on stellar inclination $i_\star$ is required.
+Here we show that, while it is important to know the stellar inclination to tightly constrain the stellar obliquity of an individual system, stellar inclination is not necessarily required for population stellar obliquity inference.
+The stellar obliquity distribution can be inferred purely from the projected stellar obliquities.
+We introduce a flexible hierarchical Bayesian framework to infer the stellar obliquity distribution of exoplanetary systems from projected stellar obliquities, stellar inclinations, or both.
+Applying the framework to all systems with projected stellar obliquity measurements, which are mostly Hot Jupiter systems, we find that the stellar obliquity distribution is unimodal and peaked at zero degrees.
+Misaligned systems have nearly isotropic stellar obliquity distribution, at odds with previous observations.
 
@@
 Stellar obliquity $\psi$ describes the angle between a planet's orbital axis $\bf{n}_{\rm orb}$ and its host star's spin axis $\bf{n}_{\star}$. 
-The angle has implications for the formation environment of a planetary system and its consequently dynamical evolution.
-Exoplanetary systems present a wide range of stellar obliquities.
+The angle is an important trace of the formation environment and dynamical evolution of a planetary system.
+The stellar obliquity evolution can be summarized into three phases.
+First, the formation and evolution of a protoplanetary disk around its host star determine the primordial stellar obliquity.
+Second, dynamical evolution in the planetary system excites the mutual inclinations between planets or stellar companions and further sculpts the stellar obliquity.
+Lastly, tidal realignment of the host star's spin axis could reduce the stellar obliquity.
 
-In reality, we only measure the sky-projected stellar obliquity $\lambda$, the angle between the projections of $\bf{n}_{\rm orb}$ and $\bf{n}_{\star}$ onto the plane of the sky.
+Stellar obliquity is an important tracer to understanding the formation of close-in planetary systems, particularly of Hot Jupiters.
+Various proposed origin channels of Hot Jupiters make predictions on the stellar obliquity distribution of the Hot Jupiters, assuming that the primordial stellar obliquity is close to zero.
+
+In reality, only the sky-projected stellar obliquity $\lambda$, the angle between the projections of $\bf{n}_{\rm orb}$ and $\bf{n}_{\star}$ onto the plane of the sky, can be measured.
 Stellar obliquity $\psi$ can be computed from $\lambda$ as \citep[e.g.,][]{Fabrycky09}:
@@
 where $i_{\rm orb}$ is the angle between the vector $\bf{n}_{\rm orb}$ and the observer's line of sight, and $i_\star$ is the angle between $\bf{n}_{\star}$ and the observer's line of sight.
-Since the sky-projected stellar obliquities can be measured only on transiting exoplanet systems, $i_{\rm orb} \approx 90\degr$ and $\cos{\psi} \approx \sin{i_\star}\cos{\lambda}$.
-
+Since the sky-projected stellar obliquities can be measured only on transiting exoplanet systems where $i_{\rm orb} \approx 90\degr$, $\cos{\psi} \approx \sin{i_\star}\cos{\lambda}$.
 To constrain the 3D stellar obliquity $\psi$, the measurements of the sky-projected obliquity $\lambda$ and stellar inclination $i_\star$ are required.
 The stellar inclination can often be constrained via, e.g., photometric and spectroscopic rotational modulation introduced by starspots, gravity darkening, and asteroseismology.
-Since the stellar inclination measurement is degenerate between $i_\star$ and $\pi - i_\star$, we assume $i_\star$ is less than $90\degr$.
+Since the stellar inclination measurement is degenerate between $i_\star$ and $180\degr - i_\star$, $i_\star$ is usually assumed to be less than $90\degr$.
 
-For systems without measured $i_\star$, it is still possible to infer the 3D stellar obliquity $\psi$ just from the sky-projected obliquity $\lambda$, assuming the stellar inclination $i_\star$ is isotropically distributed.
+Hot Jupiter systems' projected stellar obliquities are distributed widely but with a concentration at zero.
+
+For systems without measured $i_\star$, it is still possible to infer the stellar obliquity $\psi$ from the sky-projected obliquity $\lambda$ by assuming the stellar inclination $i_\star$ follows the isotropic distribution.
 
 In this work, we find an approach to infer $\psi$ distribution from $\lambda$ distribution.
+
+In this work, we build a hierarchical Bayesian framework to constrain the stellar obliquity distribution of exoplanetary systems, particularly of misaligned systems.
+In Section~\ref{sec:jacobian},
+In Section~\ref{sec:hbm}, we describe our hierarchical Bayesian framework. 
+In Section~\ref{sec:datq}, we apply our framework to observed exoplanetary systems.
+
+\section{Transformation between the $\psi$ and $\lambda$ distributions}\label{sec:jacobian}
 
@@
 \end{figure*}
-
-In this work, we build a hierarchical Bayesian framework to constrain the stellar obliquity distribution of exoplanetary systems, particularly of misaligned systems.
-In Section~\ref{sec:jacobian}
-In Section~\ref{sec:hbm}, we describe our hierarchical Bayesian framework and show inferred stellar obliquity distributions. 
-In Section~\ref{sec:application}, we apply our framework to observed exoplanetary systems.
-
-\section{Transformation between the $\psi$ and $\lambda$ distributions}\label{sec:jacobian}
 
@@
 
-\section{Hierarchical Bayesian Framework} \label{sec:hbm}
+\section{Hierarchical Bayesian Framework}\label{sec:hbm}
 
@@
 \end{align}
+
+\begin{figure}[ht!]
+    \script{simulation.py}
+    \includegraphics[width=\linewidth]{figures/simulation.pdf}
+    \caption{Inferred $\cos{\psi}$ from simulated data.}
+    \label{fig:simulation}
+\end{figure}
 
@@
 
-\section{Discussion} \label{sec:discussion}
+\section{Application to Data \& Discussion}\label{sec:data}
 
@@
 They found while the majority of planets have a small $\psi$, misaligned planets have $\psi$ clustered between 80--125$\degr$.
-Furthermore, the $\cos{\psi}$ distribution of misaligned systems is not uniformly distributed but clusters near 0, suggesting the stellar spin axis is not isotropically distributed around the planetary orbital axis.
+Furthermore, the $\cos{\psi}$ distribution of misaligned systems is not uniformly distributed, but clusters near 0, suggesting the stellar spin axis is not isotropically distributed around the planetary orbital axis.
 

### 2022-12-30 src/scripts/coord_lam.py
[showyourwork] overleaf sync

@@ -1,132 +1,0 @@
-import os
-from pathlib import Path
-os.environ["PATH"] += os.pathsep + str(Path.home() / "bin")
-
-import paths
-
-import sys
-import subprocess
-
-from matplotlib import rc
-rc('font', **{'family':'sans-serif'})
-rc('text', usetex=True)
-rc('text.latex', preamble=r'\usepackage{physics}')
-
-import matplotlib.pyplot as plt
-import numpy as np
-
-fig = plt.figure(dpi=80)
-ax = fig.add_subplot(projection='3d')
-
-# Make data
-u = np.linspace(0, 2 * np.pi, 180)
-v = np.linspace(0, np.pi, 180)
-x = np.outer(np.cos(u), np.sin(v))
-y = np.outer(np.sin(u), np.sin(v))
-z = np.outer(np.ones(np.size(u)), np.cos(v))
-
-#Plot the surface
-ax.plot_surface(x, y, z, rstride=4, cstride=4, color='b', linewidth=0, alpha=0.1)
-
-elev = 10.0
-rot = 60.0 / 180 * np.pi
-
-#calculate vectors for "vertical" circle
-a = np.array([-np.sin(elev / 180 * np.pi), 0, np.cos(elev / 180 * np.pi)])
-b = np.array([0, 1, 0])
-b = b * np.cos(rot) + np.cross(a, b) * np.sin(rot) + a * np.dot(a, b) * (1 - np.cos(rot))
-ax.plot(np.sin(u),np.cos(u),0,color='k', linestyle = 'dashed')
-horiz_front = np.linspace(0, np.pi, 100)
-ax.plot(np.sin(horiz_front),np.cos(horiz_front),0,color='k')
-
-vert_front = np.linspace(np.pi / 2, 3 * np.pi / 2, 100)
-ax.plot(a[0] * np.sin(u) + b[0] * np.cos(u), b[1] * np.cos(u), a[2] * np.sin(u) + b[2] * np.cos(u),
-        color='k',lw=1)
-ax.plot(a[0] * np.sin(vert_front) + b[0] * np.cos(vert_front), b[1] * np.cos(vert_front), 
[…]

### 2022-12-30 src/scripts/coordinate.py

@@ -15,156 +15,6 @@ rc('text.latex', preamble=r'\usepackage{physics}')
 import matplotlib.pyplot as plt
 import numpy as np
 
-fig = plt.figure(dpi=80)
-ax = fig.add_subplot(projection='3d')
-
-# Make data
-u = np.linspace(0, 2 * np.pi, 180)
-v = np.linspace(0, np.pi, 180)
-x = np.outer(np.cos(u), np.sin(v))
-y = np.outer(np.sin(u), np.sin(v))
-z = np.outer(np.ones(np.size(u)), np.cos(v))
-
-#Plot the surface
-ax.plot_surface(x, y, z, rstride=4, cstride=4, color='b', linewidth=0, alpha=0.1)
-# ax.plot_wireframe(x, y, z, rstride=180, cstride=180, color='#e5e1e0', alpha=0.5)
-
-# u = np.linspace(0, 2 * np.pi, 500)
-# v = np.linspace(0, np.pi, 500)
-
-# x = 1 * np.outer(np.cos(u), np.sin(v))
-# y = 1 * np.outer(np.sin(u), np.sin(v))
-# z = 1 * np.outer(np.ones(np.size(u)), np.cos(v))
-#for i in range(2):
-#    ax.plot_surface(x+random.randint(-5,5), y+random.randint(-5,5), z+random.randint(-5,5),  rstride=4, cstride=4, color='b', linewidth=0, alpha=0.5)
-elev = 20.0
-rot = 30.0 / 180 * np.pi
-# ax.plot_surface(x, y, z,  rstride=4, cstride=4, color='w', linewidth=0, alpha=0.3)
-
-#calculate vectors for "vertical" circle
-a = np.array([-np.sin(elev / 180 * np.pi), 0, np.cos(elev / 180 * np.pi)])
-b = np.array([0, 1, 0])
-b = b * np.cos(rot) + np.cross(a, b) * np.sin(rot) + a * np.dot(a, b) * (1 - np.cos(rot))
-ax.plot(np.sin(u),np.cos(u),0,color='k', linestyle = 'dashed')
-horiz_front = np.linspace(0, np.pi, 100)
[…]

### 2022-12-30 src/scripts/simulation.py

@@ -80,7 +80,7 @@ def posteriors(this_model):
         # logl for λ
         logl_λ = pm.Normal('logl_λ', mu=λ, sigma=err_lam, observed=obs_lam)
 
-        logi = pm.Normal('logl_i', mu=i, sigma=err_istar, observed=obs_istar)
+        logl_i = pm.Normal('logl_i', mu=i, sigma=err_istar, observed=obs_istar)
 
         hyper = pm.Potential("hyper", pm.logp(pm.Beta.dist(a,b), (cosψ+1)/2))
 
@@ -112,19 +112,19 @@ def posteriors(this_model):
 
         noistar_idata = pm.sample()
         
-        post = istar_idata.posterior
-        istar_draws = np.zeros(shape=(len(x),4000))
-        for a in range(4):
-            for b in range(1000):
-                istar_draws[:, a*1000+b] = beta.pdf(x, post.a[a,b], post.b[a,b])
+    post = istar_idata.posterior
+    istar_draws = np.zeros(shape=(len(x),4000))
+    for a in range(4):
+        for b in range(1000):
+            istar_draws[:, a*1000+b] = beta.pdf(x, post.a[a,b], post.b[a,b])
 
-        post = noistar_idata.posterior
-        noistar_draws = np.zeros(shape=(len(x),4000))
-        for a in range(4):
-            for b in range(1000):
-                noistar_draws[:, a*1000+b] = beta.pdf(x, post.a[a,b], post.b[a,b])
+    post = noistar_idata.posterior
+    noistar_draws = np.zeros(shape=(len(x),4000))
+    for a in range(4):
+        for b in range(1000):
+            noistar_draws[:, a*1000+b] = beta.pdf(x, post.a[a,b], post.b[a,b])
 
-        return istar_draws, noistar_draws
+    return istar_draws, noistar_draws
 
 
 ### PyMC models ###

### 2022-12-30 src/scripts/coordinate.py
[showyourwork] overleaf sync

@@ -28,9 +28,19 @@ z = np.outer(np.ones(np.size(u)), np.cos(v))
 
 #Plot the surface
 ax.plot_surface(x, y, z, rstride=4, cstride=4, color='b', linewidth=0, alpha=0.1)
+# ax.plot_wireframe(x, y, z, rstride=180, cstride=180, color='#e5e1e0', alpha=0.5)
 
-elev = 10.0
-rot = 60.0 / 180 * np.pi
+# u = np.linspace(0, 2 * np.pi, 500)
+# v = np.linspace(0, np.pi, 500)
+
+# x = 1 * np.outer(np.cos(u), np.sin(v))
+# y = 1 * np.outer(np.sin(u), np.sin(v))
+# z = 1 * np.outer(np.ones(np.size(u)), np.cos(v))
+#for i in range(2):
+#    ax.plot_surface(x+random.randint(-5,5), y+random.randint(-5,5), z+random.randint(-5,5),  rstride=4, cstride=4, color='b', linewidth=0, alpha=0.5)
+elev = 20.0
+rot = 30.0 / 180 * np.pi
+# ax.plot_surface(x, y, z,  rstride=4, cstride=4, color='w', linewidth=0, alpha=0.3)
 
 #calculate vectors for "vertical" circle
 a = np.array([-np.sin(elev / 180 * np.pi), 0, np.cos(elev / 180 * np.pi)])
@@ -48,11 +58,11 @@ ax.plot(a[0] * np.sin(vert_front) + b[0] * np.cos(vert_front), b[1] * np.cos(ver
         color='k',lw=1)
 
 
-ax.quiver(0, 0, 0, 1.4, 0, 0, color='k', arrow_length_ratio=0.05, lw=1.5) # x-axis
+ax.quiver(0, 0, 0, 1.8, 0, 0, color='k', arrow_length_ratio=0.05, lw=1.5) # x-axis
 ax.quiver(0, 0, 0, 0, 1.5, 0, color='k', arrow_length_ratio=0.05, lw=1.5) # y-axis
 ax.quiver(0, 0, 0, 0, 0, 1.5, color='k', arrow_length_ratio=0.05, lw=1.5) # z-axis
 
-ax.text(1.7, 0, -0.05, r'$\textbf{Obs}$',fontsize=12)
[…]

### 2022-12-30 src/scripts/graph.py
[showyourwork] overleaf sync

@@ -29,7 +29,7 @@ pgm.add_node("lambda", r"$\lambda_n$", 1., 1.5)
 
 pgm.add_node("istar", r"$i_{\star,n}$", 3., 1.5)
 
-pgm.add_node("pstar", r"$\theta_{\star,n}$", 4., 1.5)
+pgm.add_node("pstar", r"$\gamma_{\star,n}$", 4., 1.5)
 
 
 pgm.add_edge("beta", "psi")

### 2022-12-31 src/scripts/graph.py
[showyourwork] overleaf sync

@@ -54,8 +54,8 @@ pgm.add_edge("pstar", "Obs", plot_params={'ls':''})
 pgm.add_edge("istar", "Obs", plot_params=dict(ls=(0, (2, 2)),head_width=0,head_length=0))
 pgm.add_edge("istar", "Obs", plot_params={'ls':''})
 
-pgm.add_edge("lambda", "Obs", plot_params=dict(ls=(0, (2, 2)),head_width=0,head_length=0))
-pgm.add_edge("lambda", "Obs", plot_params={'ls':''})
+# pgm.add_edge("lambda", "Obs", plot_params=dict(ls=(0, (2, 2)),head_width=0,head_length=0))
+# pgm.add_edge("lambda", "Obs", plot_params={'ls':''})
 
 # pgm.add_edge("pstar", "obs_pstar", plot_params=dict(ls=(0, (2, 2)),head_width=0,head_length=0))
 # pgm.add_edge("pstar", "obs_pstar", plot_params={'ls':''})

### 2022-12-31 src/scripts/lam_dist.py
[showyourwork] overleaf sync

@@ -1,90 +1,0 @@
-import os
-from pathlib import Path
-os.environ["PATH"] += os.pathsep + str(Path.home() / "bin")
-
-import sys
-import subprocess
-
-# implement pip as a subprocess:
-subprocess.check_call([sys.executable, '-m', 'pip', 'install',
-'arviz'])
-
-import paths
-import matplotlib.pyplot as plt
-import numpy as np
-import arviz as az
-from scipy.stats import beta
-
-from matplotlib import rc
-rc('font', **{'family':'sans-serif'})
-rc('text', usetex=True)
-
-plt.rcParams['xtick.top'] =  True
-plt.rcParams['xtick.direction'] =  'in'
-plt.rcParams['xtick.major.width'] =  1.0
-plt.rcParams['xtick.minor.width'] =  1.0
-plt.rcParams['ytick.right'] =  True
-plt.rcParams['ytick.direction'] =  'in'
-plt.rcParams['ytick.major.width'] = 1.0
-plt.rcParams['ytick.minor.width'] =  1.0
-plt.rcParams['lines.markeredgewidth'] =  1.0
-
-numistar = 62
-numall = 161
-
-# load MCMC draws
-lam_all = az.from_netcdf(paths.data / 'lam_all.nc')
-lam_istar = az.from_netcdf(paths.data / 'lam_istar.nc')
-
-x = np.linspace(1e-5,1-1e-5,1000)
-
-post = lam_all.posterior
-lam_all_draws = np.zeros(shape=(len(x),4000))
-for a in range(4):
-    for b in range(1000):
-        lam_all_draws[:, a*1000+b] = (post.w[a,b,0].values*beta.pdf(x, post.a0[a,b], post.b0[a,b])
-                                    +post.w[a,b,1].values*beta.pdf(x, post.a1[a,b], post.b1[a,b]))
-
-post = lam_istar.posterior
-lam_istar_draws = np.zeros(shape=(len(x),4000))
-for a in range(4):
-    for b in range(1000):
-        lam_istar_draws[:, a*1000+b] = (post.w[a,b,0].values*beta.pdf(x, post.a0[a,b], post.b0[a,b])
-                                    +post.w[a,b,1].values*beta.pdf(x, post.a1[a,b], post.b1[a,b]))
-
-# make the figure
[…]

### 2022-12-31 src/scripts/psi_dist.py

@@ -29,29 +29,11 @@ plt.rcParams['ytick.major.width'] = 1.0
 plt.rcParams['ytick.minor.width'] =  1.0
 plt.rcParams['lines.markeredgewidth'] =  1.0
 
-numistar = 62
-numall = 161
 
 # load MCMC draws
 all_randinc = az.from_netcdf(paths.data / 'all_randinc.nc')
-istar_randinc = az.from_netcdf(paths.data / 'istar_randinc.nc')
-istar = az.from_netcdf(paths.data / 'istar.nc')
 
 x = np.linspace(1e-5,1-1e-5,1000)
-
-post = istar.posterior
-istar_draws = np.zeros(shape=(len(x),4000))
-for a in range(4):
-    for b in range(1000):
-        istar_draws[:, a*1000+b] = (post.w[a,b,0].values*beta.pdf(x, post.a0[a,b], post.b0[a,b])
-                                    +post.w[a,b,1].values*beta.pdf(x, post.a1[a,b], post.b1[a,b]))
-
-post = istar_randinc.posterior
-istar_randinc_draws = np.zeros(shape=(len(x),4000))
-for a in range(4):
-    for b in range(1000):
-        istar_randinc_draws[:, a*1000+b] = (post.w[a,b,0].values*beta.pdf(x, post.a0[a,b], post.b0[a,b])
-                                            +post.w[a,b,1].values*beta.pdf(x, post.a1[a,b], post.b1[a,b]))
 
 post = all_randinc.posterior
 all_randinc_draws = np.zeros(shape=(len(x),4000))
@@ -61,38 +43,8 @@ for a in range(4):
                                           +post.w[a,b,1].values*beta.pdf(x, post.a1[a,b], post.b1[a,b]))
 
 # make the figure
-plt.figure(figsize=(7,2.6),dpi=110)
+plt.figure(figsize=(3.5,2.7),dpi=110)
 
-plt.subplot(1,3,1)
-q025, q16, q50, q84, q975 = np.percentile(istar_draws, [2.5, 16, 50, 84, 97.5], axis=1)
-plt.plot(2*x-1, q50, color='#f56e4a')
-plt.fill_between(2*x-1, q16, q84, alpha=0.3, label="posterior", color='#fbc1ad')
-plt.fill_between(2*x-1, q025, q975, alpha=0.3, color='#fbc1ad')
-
[…]

### 2022-12-31 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \begin{figure*}[ht!]
-    \script{coord_psi.py}
+    \script{coordinate.py}
     \gridline{
@@
 
-\begin{figure*}[ht!]
+\begin{figure}[ht!]
     \script{psi_dist.py}
@@
     \end{centering}
-\end{figure*}
+\end{figure}
 
@@
 
-\begin{figure*}[ht!]
-    \script{lam_dist.py}
-    \begin{centering}
-        \includegraphics{figures/lam_dist.pdf}
-        \caption{The distributions of sky-projected stellar obliquity ($\lambda$) of the \numistar systems that have both sky-projected stellar obliquity $\lambda$ and stellar inclination $i_\star$ measurements (left panel) and all \numall systems that have sky-projected stellar obliquity measurements.
-The observed-$i_\star$ has a cluster of systems near $110\degr$, whereas such clustering is not found in the all-system distribution.}
-        \label{fig:lam_dist}
-    \end{centering}
-\end{figure*}
-
 One plausible explanation for the disparity between the two distributions is their small sample sizes.

### 2022-12-31 src/scripts/psi_dist.py
[showyourwork] overleaf sync

@@ -46,9 +46,10 @@ for a in range(4):
 plt.figure(figsize=(3.5,2.7),dpi=110)
 
 q025, q16, q50, q84, q975 = np.percentile(all_randinc_draws, [2.5, 16, 50, 84, 97.5], axis=1)
-plt.plot(2*x-1, q50, color='C0')
-plt.fill_between(2*x-1, q16, q84, alpha=0.3, label="posterior", color='#7dabd0')
-plt.fill_between(2*x-1, q025, q975, alpha=0.3, color='#7dabd0')
+plt.plot(2*x-1, q50, color='#f56e4a')
+plt.fill_between(2*x-1, q16, q84, alpha=0.3, label="posterior", color='#fbc1ad')
+plt.fill_between(2*x-1, q025, q975, alpha=0.3, color='#fbc1ad')
+
 
 plt.ylim([0,2])
 plt.xlim([-1,1])

### 2022-12-31 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 Stellar obliquity $\psi$, the angle between a planet's orbital axis and its host star's spin axis, traces the formation environment and dynamical evolution of a planetary system.
-In exoplanet observations, only sky-projected stellar obliquity $\lambda$ can be measured.
-To find the stellar obliquity from the projected obliquity, information on stellar inclination $i_\star$ is required.
-Here we show that, while it is important to know the stellar inclination to tightly constrain the stellar obliquity of an individual system, stellar inclination is not necessarily required for population stellar obliquity inference.
-The stellar obliquity distribution can be inferred purely from the projected stellar obliquities.
-We introduce a flexible hierarchical Bayesian framework to infer the stellar obliquity distribution of exoplanetary systems from projected stellar obliquities, stellar inclinations, or both.
+In exoplanet observations, only the sky-projected stellar obliquity $\lambda$ can be measured.
+To find the true stellar obliquity, information about the stellar inclination $i_\star$ is required.
+Here we show that, while it is important to know the stellar inclination to tightly constrain the stellar obliquity of an individual system, the stellar inclination is not necessarily required for the population stellar obliquity inference.
+Stellar obliquity distribution can be inferred purely from sky-projected stellar obliquities.
+We introduce a flexible, hierarchical Bayesian framework to infer the stellar obliquity distribution of exoplanetary systems from their projected stellar obliquities, stellar inclinations, or both.
 Applying the framework to all systems with projected stellar obliquity measurements, which are mostly Hot Jupiter systems, we find that the stellar obliquity distribution is unimodal and peaked at zero degrees.
@@
 First, the formation and evolution of a protoplanetary disk around its host star determine the primordial stellar obliquity.
-Second, dynamical evolution in the planetary system excites the mutual inclinations between planets or stellar companions and further sculpts the stellar obliquity.
-Lastly, tidal realignment of the host star's spin axis could reduce the stellar obliquity.
-
-Stellar obliquity is an important tracer to understanding the formation of close-in planetary systems, particularly of Hot Jupiters.
-Various proposed origin channels of Hot Jupiters make predictions on the stellar obliquity distribution of the Hot Jupiters, assuming that the primordial stellar obliquity is close to zero.
+Second, post-formation dynamical evolution in the planetary system excites the mutual inclinations between planets or stellar companions and further sculpts the stellar obliquity.
+Lastly, tidal realignment of the host star's spin axis to the planet's orbital axis could reduce the stellar obliquity, if the process is efficient.
 
 In reality, only the sky-projected stellar obliquity $\lambda$, the angle between the projections of $\bf{n}_{\rm orb}$ and $\bf{n}_{\star}$ onto the plane of the sky, can be measured.
-Stellar obliquity $\psi$ can be computed from $\lambda$ as \citep[e.g.,][]{Fabrycky09}:
+Stellar obliquity $\psi$ can be derived if the stellar inclination $i_\star$ is also known.
+The calculation follows \citep[e.g.,][]{Fabrycky09}:
 \begin{equation}\label{eqn:psi}
@@
 where $i_{\rm orb}$ is the angle between the vector $\bf{n}_{\rm orb}$ and the observer's line of sight, and $i_\star$ is the angle between $\bf{n}_{\star}$ and the observer's line of sight.
-Since the sky-projected stellar obliquities can be measured only on transiting exoplanet systems where $i_{\rm orb} \approx 90\degr$, $\cos{\psi} \approx \sin{i_\star}\cos{\lambda}$.
-To constrain the 3D stellar obliquity $\psi$, the measurements of the sky-projected obliquity $\lambda$ and stellar inclination $i_\star$ are required.
 The stellar inclination can often be constrained via, e.g., photometric and spectroscopic rotational modulation introduced by starspots, gravity darkening, and asteroseismology.
 Since the stellar inclination measurement is degenerate between $i_\star$ and $180\degr - i_\star$, $i_\star$ is usually assumed to be less than $90\degr$.
+Since the sky-projected stellar obliquities can be measured only on transiting exoplanet systems where $i_{\rm orb} \approx 90\degr$, $\cos{\psi} \approx \sin{i_\star}\cos{\lambda}$.
 
-Hot Jupiter systems' projected stellar obliquities are distributed widely but with a concentration at zero.
+Stellar obliquity is important to understand; however, not all systems have a stellar inclination measurement.
+For systems without measured $i_\star$, it is still possible to infer the stellar obliquity $\psi$ from the sky-projected obliquity $\lambda$ by assuming the stellar inclination $i_\star$ follows the isotropic distribution, although results in a larger uncertainty on $\psi$ \citep{Fabrycky09}.
 
-For systems without measured $i_\star$, it is still possible to infer the stellar obliquity $\psi$ from the sky-projected obliquity $\lambda$ by assuming the stellar inclination $i_\star$ follows the isotropic distribution.
-
-In this work, we find an approach to infer $\psi$ distribution from $\lambda$ distribution.
-
-In this work, we build a hierarchical Bayesian framework to constrain the stellar obliquity distribution of exoplanetary systems, particularly of misaligned systems.
-In Section~\ref{sec:jacobian},
-In Section~\ref{sec:hbm}, we describe our hierarchical Bayesian framework. 
-In Section~\ref{sec:datq}, we apply our framework to observed exoplanetary systems.
-
-\section{Transformation between the $\psi$ and $\lambda$ distributions}\label{sec:jacobian}
+In this work, we extend the inference of stellar obliquity to a population.
+We identify an approach to infer $\psi$ distribution from $\lambda$ distribution and show that a similar constraint can be reached with or without stellar inclination measurement.
+In Section~\ref{sec:jacobian}, we present the transformation between the stellar obliquity distribution and the sky-projected stellar obliquity distribution.
+In Section~\ref{sec:hbm}, we introduce a hierarchical Bayesian framework to infer the population stellar obliquity distribution. 
+And lastly, in Section~\ref{sec:data}, we apply our framework to all exoplanetary systems with measured sky-projected obliquity.
 
@@
     \vspace*{-1.5cm}
-    \caption{Two coordinate systems to describe the stellar spin axis $\bf{n}_{\star}$ and the planet's orbital axis $\bf{n}_{\rm orb}$.}
+    \caption{Two coordinate systems to describe the stellar spin axis $\bf{n}_{\star}$ and the planet's orbital axis $\bf{n}_{\rm orb}$.
+Here we define the observer's line of sight as one of the two horizontal axes, and the orbital axis of the planet as the vertical axis.
+We approximate the orbital inclination to $90\degr$.}
     \label{fig:coord}
 \end{figure*}
+
+\section{Transformation between the $\psi$ and $\lambda$ distributions}\label{sec:jacobian}
 
@@
     \includegraphics[width=\linewidth]{figures/transform.pdf}
-    \caption{Distribution transformation.}
+    \caption{Distribution transformations between the stellar obliquity $\cos{\psi}$ (1st column), sky-projected stellar obliquity $\lambda$ (2nd column), and stellar inclination $i_\star$ (3rd column).
+The inferred $\cos{\psi}$ assuming an isotropic stellar inclination is shown in the 4th column.
+Each row has a simulated $\cos{\psi}$ distribution.
+Assuming a uniform $\theta$, the simulated $\lambda$ and $i_\star$ distributions are shown in grey histograms, and the analytical solutions are shown as blue curves.}
     \label{fig:transform}

### 2022-12-31 src/scripts/psi_dist.py

@@ -46,10 +46,14 @@ for a in range(4):
 plt.figure(figsize=(3.5,2.7),dpi=110)
 
 q025, q16, q50, q84, q975 = np.percentile(all_randinc_draws, [2.5, 16, 50, 84, 97.5], axis=1)
-plt.plot(2*x-1, q50, color='#f56e4a')
+plt.plot(2*x-1, q50, color='#f56e4a', lw=1.2)
 plt.fill_between(2*x-1, q16, q84, alpha=0.3, label="posterior", color='#fbc1ad')
 plt.fill_between(2*x-1, q025, q975, alpha=0.3, color='#fbc1ad')
 
+
+plt.plot(2*x-1, q50, color='C0', lw=1.2)
+plt.fill_between(2*x-1, q16, q84, alpha=0.3, label="posterior", color='#7dabd0')
+plt.fill_between(2*x-1, q025, q975, alpha=0.3, color='#7dabd0')
 
 plt.ylim([0,2])
 plt.xlim([-1,1])

### 2022-12-31 src/scripts/simulation.py

@@ -249,9 +249,11 @@ if __name__ == '__main__':
     ax = fig.add_subplot(4,2,1)
             
     q025, q16, q50, q84, q975 = np.percentile(uni_istar_draws, [2.5, 16, 50, 84, 97.5], axis=1)/2
-    plt.plot(2*x-1, q50, color='C0', lw=1.2)
-    plt.fill_between(2*x-1, q16, q84, alpha=0.3, label="posterior", color='#7dabd0')
-    plt.fill_between(2*x-1, q025, q975, alpha=0.3, color='#7dabd0')
+
+    plt.plot(2*x-1, q50, color='#f56e4a', lw=1.2)
+    plt.fill_between(2*x-1, q16, q84, alpha=0.3, label="posterior", color='#fbc1ad')
+    plt.fill_between(2*x-1, q025, q975, alpha=0.3, color='#fbc1ad')
+
 
     plt.ylim([0,1])
     plt.xlim([-1,1])
@@ -266,9 +268,9 @@ if __name__ == '__main__':
     ax = fig.add_subplot(4,2,2)
 
     q025, q16, q50, q84, q975 = np.percentile(uni_noistar_draws, [2.5, 16, 50, 84, 97.5], axis=1)/2
-    plt.plot(2*x-1, q50, color='#f56e4a', lw=1.2)
-    plt.fill_between(2*x-1, q16, q84, alpha=0.3, label="posterior", color='#fbc1ad')
-    plt.fill_between(2*x-1, q025, q975, alpha=0.3, color='#fbc1ad')
+    plt.plot(2*x-1, q50, color='C0', lw=1.2)
+    plt.fill_between(2*x-1, q16, q84, alpha=0.3, label="posterior", color='#7dabd0')
+    plt.fill_between(2*x-1, q025, q975, alpha=0.3, color='#7dabd0')
 
     plt.ylim([0,1])
     plt.xlim([-1,1])
@@ -285,9 +287,9 @@ if __name__ == '__main__':
     ax = fig.add_subplot(4,2,3)
             
     q025, q16, q50, q84, q975 = np.percentile(norm1_istar_draws, [2.5, 16, 50, 84, 97.5], axis=1)/2
-    plt.plot(2*x-1, q50, color='C0', lw=1.2)
-    plt.fill_between(2*x-1, q16, q84, alpha=0.3, label="posterior", color='#7dabd0')
[…]

### 2023-01-05 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \usepackage{physics}
+\usepackage[super]{nth}
 
@@
 Here we show that, while it is important to know the stellar inclination to tightly constrain the stellar obliquity of an individual system, the stellar inclination is not necessarily required for the population stellar obliquity inference.
-Stellar obliquity distribution can be inferred purely from sky-projected stellar obliquities.
-We introduce a flexible, hierarchical Bayesian framework to infer the stellar obliquity distribution of exoplanetary systems from their projected stellar obliquities, stellar inclinations, or both.
+The constraint on stellar obliquity distribution is predominantly based on sky-projected stellar obliquities.
+We introduce a flexible, hierarchical Bayesian framework to infer the stellar obliquity distribution of exoplanetary systems from their sky-projected stellar obliquities, stellar inclinations, or both.
 Applying the framework to all systems with projected stellar obliquity measurements, which are mostly Hot Jupiter systems, we find that the stellar obliquity distribution is unimodal and peaked at zero degrees.
-Misaligned systems have nearly isotropic stellar obliquity distribution, at odds with previous observations.
+Misaligned systems have nearly isotropic stellar obliquities, at odds with the analysis of systems that also have $i_\star$ measurements.
 
@@
 Stellar obliquity $\psi$ describes the angle between a planet's orbital axis $\bf{n}_{\rm orb}$ and its host star's spin axis $\bf{n}_{\star}$. 
-The angle is an important trace of the formation environment and dynamical evolution of a planetary system.
-The stellar obliquity evolution can be summarized into three phases.
+The angle is an important tracer of the formation environment and dynamical evolution of a planetary system.
+Stellar obliquity evolution can be roughly summarized into three stages.
 First, the formation and evolution of a protoplanetary disk around its host star determine the primordial stellar obliquity.
-Second, post-formation dynamical evolution in the planetary system excites the mutual inclinations between planets or stellar companions and further sculpts the stellar obliquity.
-Lastly, tidal realignment of the host star's spin axis to the planet's orbital axis could reduce the stellar obliquity, if the process is efficient.
+Second, post-formation dynamical evolution in the planetary system, such as planet-planet scattering or von Zeipel-Kozai-Lidov mechanisms, excites the mutual inclinations between planetary or stellar companions, further sculpting the stellar obliquity.
+Lastly, tidal realignment of the host star's spin axis to the planet's orbital axis could reduce the stellar obliquity, if the stellar tidal dissipation is efficient.
+
+While observed aligned systems could have different original stellar obliquities because of the tidal realignment, the excitation of stellar obliquities can only be the outcome of the primordial protoplanetary disk misalignment and post-formation dynamical interactions.
+Recent stellar obliquity measurements on Warm Jupiter systems with large planet-star separations and systems that are still at young ages both find a tendency of system alignment.
+If the trend holds as more observations coming up, it could suggest that the primordial stellar obliquity is quite low.
+Stellar obliquities of misaligned systems could then be linked to the dynamical processes that result in the current orbit of close-in planets.
+With that in mind, we aim to build a statistical model to characterize the stellar obliquity distribution of exoplanetary systems in misalignment to be compared with theoretical predictions.
 
 In reality, only the sky-projected stellar obliquity $\lambda$, the angle between the projections of $\bf{n}_{\rm orb}$ and $\bf{n}_{\star}$ onto the plane of the sky, can be measured.
-Stellar obliquity $\psi$ can be derived if the stellar inclination $i_\star$ is also known.
+Stellar obliquity $\psi$ can be found if the stellar inclination $i_\star$ is known.
 The calculation follows \citep[e.g.,][]{Fabrycky09}:
@@
 where $i_{\rm orb}$ is the angle between the vector $\bf{n}_{\rm orb}$ and the observer's line of sight, and $i_\star$ is the angle between $\bf{n}_{\star}$ and the observer's line of sight.
+Since we are discussing transiting exoplanet systems where $i_{\rm orb} \approx 90\degr$, $\cos{\psi} \approx \sin{i_\star}\cos{\lambda}$.
+
 The stellar inclination can often be constrained via, e.g., photometric and spectroscopic rotational modulation introduced by starspots, gravity darkening, and asteroseismology.
 Since the stellar inclination measurement is degenerate between $i_\star$ and $180\degr - i_\star$, $i_\star$ is usually assumed to be less than $90\degr$.
-Since the sky-projected stellar obliquities can be measured only on transiting exoplanet systems where $i_{\rm orb} \approx 90\degr$, $\cos{\psi} \approx \sin{i_\star}\cos{\lambda}$.
+For systems without $i_\star$ measurements, it is still possible to infer their stellar obliquities from the sky-projected obliquities, assuming isotropic stellar inclinations.
+The inferred $\psi$ will have a larger uncertainty compared to the one inferred with known $i_\star$ \citep{Fabrycky09}.
 
-Stellar obliquity is important to understand; however, not all systems have a stellar inclination measurement.
-For systems without measured $i_\star$, it is still possible to infer the stellar obliquity $\psi$ from the sky-projected obliquity $\lambda$ by assuming the stellar inclination $i_\star$ follows the isotropic distribution, although results in a larger uncertainty on $\psi$ \citep{Fabrycky09}.
+It is not yet clear how does the stellar inclination measurement affect the inference of the stellar obliquity distribution of a population. 
+In this work, we propose an approach to infer $\psi$ distribution from $\lambda$ distribution and show that the stellar obliquity distribution inference only weakly depends on stellar inclination.
+In Section~\ref{sec:jacobian}, we present the Jacobian transformation between the stellar obliquity distribution, the sky-projected stellar obliquity distribution, and the stellar inclination distribution with or without measured $i_\star$.
+In Section~\ref{sec:hbm}, we introduce a flexible, hierarchical Bayesian framework to infer the population stellar obliquity distribution.
+In Section~\ref{sec:data}, we apply the framework to all exoplanetary systems with measured sky-projected obliquity.
 
-In this work, we extend the inference of stellar obliquity to a population.
-We identify an approach to infer $\psi$ distribution from $\lambda$ distribution and show that a similar constraint can be reached with or without stellar inclination measurement.
-In Section~\ref{sec:jacobian}, we present the transformation between the stellar obliquity distribution and the sky-projected stellar obliquity distribution.
-In Section~\ref{sec:hbm}, we introduce a hierarchical Bayesian framework to infer the population stellar obliquity distribution. 
-And lastly, in Section~\ref{sec:data}, we apply our framework to all exoplanetary systems with measured sky-projected obliquity.
+\section{Transformation between the $\psi$, $\lambda$, and \lowercase{$i_\star$} distributions}\label{sec:jacobian}
 
@@
     \gridline{
-        \fig{figures/coord_psi.pdf}{0.45\textwidth}{\vspace*{-1.8cm}(a) The $\{\psi, \theta\}$ coordinate system. The grey circle corresponds to a constant $\psi$ value.}
-        \fig{figures/coord_lam.pdf}{0.45\textwidth}{\vspace*{-1.8cm}(b) The $\{\lambda, i_\star\}$ coordinate system. The grey circle corresponds to a constant $i_\star$ value.}
+        \fig{figures/coord_psi.pdf}{0.45\textwidth}{\vspace*{-1.8cm}(a) The $\{\psi, \theta\}$ coordinate system.
+The grey circle corresponds to a constant $\psi$ value and its circumference is proportional to $\sin{\psi}$.}
+        \fig{figures/coord_lam.pdf}{0.45\textwidth}{\vspace*{-1.8cm}(b) The $\{\lambda, i_\star\}$ coordinate system.
+The grey circle corresponds to a constant $i_\star$ value and its circumference is proportional to $\sin{i_\star}$.}
     }
@@
     \caption{Two coordinate systems to describe the stellar spin axis $\bf{n}_{\star}$ and the planet's orbital axis $\bf{n}_{\rm orb}$.
-Here we define the observer's line of sight as one of the two horizontal axes, and the orbital axis of the planet as the vertical axis.
-We approximate the orbital inclination to $90\degr$.}
+Here we define the observer's line of sight as one of the two horizontal axes (conventional $x$-axis in Cartesian), and the orbital axis of the planet as the vertical axis (conventional $z$-axis in Cartesian).
+We approximate the orbital inclination of the planet to $90\degr$.}
     \label{fig:coord}
@@
 
-\section{Transformation between the $\psi$ and $\lambda$ distributions}\label{sec:jacobian}
+Stellar obliquity $\psi$ is the angle between a planet's orbital axis $\bf{n}_{\rm orb}$ and its host star's spin axis $\bf{n}_{\star}$.
+In Figure~\ref{fig:coord}, we build two coordinate systems to describe the stellar spin axis $\bf{n}_{\star}$ for a given orbital axis $\bf{n}_{\rm orb}$.
+In both coordinates, we set the observer's line of sight to be one of the two horizontal axes, and the orbital axis of the planet to be the vertical axis.
+We approximate the orbital inclination of the transiting planet to $90\degr$ here and in the rest of the work to simplify the problem for population study.
+If the orbital inclination deviates from $90\degr$, $\bf{n}_{\rm orb}$ instead of being the vertical axis needs to be rotated $\pm (90\degr-i_{\rm orb})$ along the line of sight.
 
-For any given $\bf{n}_{\rm orb}$ axis, if $\bf{n}_{\star}$ is a random vector with uniform distribution on a three-dimensional sphere, i.e., $\bf{n}_{\star}$ is isotropically distributed, the probability distribution of $\psi$ follows $p_{\psi} \sim \sin{\psi}$.
-Since $p_{\cos{\psi}} = p_{\psi} \abs{\dv*{\psi}{\cos{\psi}}}$, $p_{\cos{\psi}}$ is uniformly distributed between $-1$ and $1$.
-If the stellar obliquity instead has some preferred angles, the $\cos{\psi}$ distribution will present concentrations at certain values.
+The $\{\psi, \theta\}$ coordinate system shown in panel (a) relates to physical properties of a planetary system.
+$\psi$ is the angle between the stellar spin axis and the planetary orbital axis and $\theta$ is the azimuthal angle of the stellar spin axis.
+For the given $\bf{n}_{\rm orb}$ axis, if $\bf{n}_{\star}$ is a random vector with uniform distribution on a three-dimensional sphere, i.e., $\bf{n}_{\star}$ is isotropically distributed, the probability distribution of $\psi$ follows $p_{\psi} \sim \sin{\psi}$ and $p_\theta \sim 1/2\pi$.
+Since $p_{\cos{\psi}} = p_{\psi} \abs{\dv*{\psi}{\cos{\psi}}}$, $p_{\cos{\psi}}$ is uniformly distributed between $-1$ and $1$ for isotropic stellar obliquity.
+The $\{\lambda, i_\star\}$ coordinate system shown in panel (b) relates to observed properties.
+$\lambda$ is the sky-projected stellar obliquity and $i_\star$ is the stellar inclination.
+If $\bf{n}_{\star}$ is isotropically distributed, $p_{i_\star} \sim \sin{i_\star}$ and $p_\lambda \sim 1/2\pi$.
+Again, since $p_{\cos{i_\star}} = p_{i_\star} \abs{\dv*{i_\star}{\cos{i_\star}}}$, $p_{\cos{i_\star}}$ is uniformly distributed between $-1$ and $1$.
+Because of the observational degeneracy between $i_\star$ and $180\degr - i_\star$, the convention is to have $0 \leq i_\star \leq 90\degr$, and thus $p_{\cos{i_\star}}$ is uniformly distributed between $0$ and $1$ (i.e., $p_{\cos{i_\star}} \sim 1$).
+This also limits $\theta$ to $[-\pi/2, \pi/2]$.
 
-Given a certain $\psi$ distribution, the $\lambda$ and $i_\star$ are no longer randomly/isotropically distributed.
-We first find relations between $\{\psi, \theta\}$ and $\{\lambda, i_\star\}$, where $\theta$ is the azimuthal angle of the stellar spin axis, as shown in Figure~\ref{fig:coord}.
-We transform $\bf{n}_{\star}$ in two coordinate systems to Cartesian, pair their terms, and get:
+We could find the mathematical relations between $\{\psi, \theta\}$ and $\{\lambda, i_\star\}$ by pairing the Cartesian components of $\bf{n}_{\star}$ in two coordinate systems:
 \begin{align}
@@
 
-First, we find the $\lambda$ distribution for a given $\cos{\psi}$ distribution.
-If the stellar obliquity is isotropically distributed, the $\cos{\psi}$ distribution will be uniform, and the $\lambda$ distribution will also be uniform.
-If the stellar obliquity has preferred angles, i.e., the $\cos{\psi}$ distribution will have concentrations, what will be the corresponding $\lambda$ distribution?
-We first find the distribution of $\cos{\lambda}$ using the Jacobian transformation from $\cos{\psi}$.
-Replacing $\sin{i_\star}$ in Equation~(\ref{eq:coord3}) with Equation~(\ref{eq:coord1}), we find $\cos{\lambda} = \cos{\psi}/\sqrt{1-(1-\cos^2{\psi})\cos^2{\theta}}$.
-Reorganize the equation, and we also get $\cos^2{\psi} = \frac{\cos^2{\lambda}\cos^2{\theta}-\cos^2{\lambda}}{\cos^2{\lambda}\cos^2{\theta}-1}$.
-The distribution transformation follows
+First, we derive the $\lambda$ distribution for a given $\cos{\psi}$ distribution.
+We could find the distribution of $\cos{\lambda}$ using the Jacobian transformation from $\cos{\psi}$ and $\cos{\theta}$.
+Since $\psi$ and $\theta$ are independent variables, we could marginalize over $\theta$ to find the relation between the probability density functions between $\lambda$ and $\psi$.
+The Jacobian transformation follows
 \begin{equation}
@@
 \end{equation}
-We have $\abs{\pdv{\cos{\psi}}{\cos{\lambda}}} = \frac{(1-\cos^2{\theta})^{1/2}}{(1-\cos^2{\theta}\cos^2{\lambda})^{3/2}}$.
-Assuming $\theta$ is uniformly distributed between $-\pi/2$ and $\pi/2$\footnote{The $\theta$ is defined between $-\pi/2$ and $\pi/2$ since we limit $0 \leq i_\star \leq \pi/2$.}, $p_{\cos{\theta}} = p_{\theta} \abs{\dv*{\theta}{\cos{\theta}}} = 1/\pi/(1-\cos^2{\theta)^{1/2}}$.
-Combining all the parts together, we get
-\begin{equation}\label{eqn:jac}
+Replacing $\sin{i_\star}$ in Equation~(\ref{eq:coord3}) with Equation~(\ref{eq:coord1}), we find $\cos{\lambda} = \cos{\psi}/\sqrt{1-(1-\cos^2{\psi})\cos^2{\theta}}$.
+Reorganize the equation, we get $\cos^2{\psi} = \frac{\cos^2{\lambda}\cos^2{\theta}-\cos^2{\lambda}}{\cos^2{\lambda}\cos^2{\theta}-1}$.
+The derivative gives $\abs{\pdv{\cos{\psi}}{\cos{\lambda}}} = \frac{(1-\cos^2{\theta})^{1/2}}{(1-\cos^2{\theta}\cos^2{\lambda})^{3/2}}$.
+If $\theta$ is uniformly distributed between $-\pi/2$ and $\pi/2$, $p_{\cos{\theta}} = p_{\theta} \abs{\dv*{\theta}{\cos{\theta}}} = 1/\pi/(1-\cos^2{\theta)^{1/2}}$.
+Putting all the parts together, we get
+\begin{equation}\label{eqn:jac_lam}
     p_{\cos{\lambda}} = \frac{2}{\pi} \int_{0}^{1} (1-\cos^2{\theta}\cos^2{\lambda})^{-3/2} p_{\cos{\psi}} d\cos{\theta}.
 \end{equation}
+If $\cos{\psi}$ is uniformly distribution, i.e., $p_{\cos{\psi}} = 1/2$, Equation~(\ref{eqn:jac_lam}) becomes $p_{\cos{\lambda}} = 1/\pi/\sqrt{1-\cos^2{\lambda}}$, which is equivalent to $p_\lambda = 1/\pi$.
+The $\lambda$ is uniformly distributed if an isotropic $\psi$.
+If $\cos{\psi}$ follows other distributions, $p_\lambda$ could be found by evaluating Equation~(\ref{eqn:jac_lam}) numerically.
 
-If $\cos{\psi}$ is uniformly distribution, i.e., $p_{\cos{\psi}} = 1/2$, the integral derives $p_{\cos{\lambda}} = 1/\pi/\sqrt{1-\cos^2{\lambda}}$, which is equivalent to $p_\lambda = 1/\pi$.
-The $\lambda$ is uniformly distributed.
-If $\cos{\psi}$ follows other distributions, we may integrate Equation~(\ref{eqn:jac}) numerically, find the $\cos{\lambda}$ distribution, and solve for the $\lambda$ distribution.
-
-Similarly, we could derive the distribution of $\cos{i_\star}$ for a given $\cos{\psi}$ distribution via the Jacobian transformation and marginalizing over $\cos{\theta}$. We begin with
+Next, we derive the $i_\star$ distribution for a given $\cos{\psi}$ distribution.
+Similarly, we first find the Jacobian transformation of $i_\star$ from $\psi$ and $\theta$ and then marginalize over $\theta$.
+It is easier to work on $\cos{i_\star}$ than $i_\star$:
 \begin{equation}
-    p_{\cos{i_\star}} = 4 \int_{0}^{1} \abs{\pdv{\sin{\psi}}{\cos{i_\star}}} p_{\sin{\psi}} p_{\cos{\theta}} d\cos{\theta},
+    p_{\cos{i_\star}} = 4 \int_{0}^{1} \abs{\pdv{\sin{\psi}}{\cos{i_\star}}} p_{\sin{\psi}} p_{\cos{\theta}} d\cos{\theta}.
 \end{equation}
-and from Equation(\ref{eq:coord1}), we find $\sin{\psi} = \cos{i_\star}/\cos{\theta}$ and $\abs{\pdv{\sin{\psi}}{\cos{i_\star}}} = 1/\cos{\theta}$.
-Similarly, we assume $\theta$ is uniformly distributed, and $p_{\cos{\theta}} = 1/\pi/(1-\cos^2{\theta)^{1/2}}$.
+From Equation~(\ref{eq:coord1}), we have $\sin{\psi} = \cos{i_\star}/\cos{\theta}$ and $\abs{\pdv{\sin{\psi}}{\cos{i_\star}}} = 1/\cos{\theta}$.
+Again, we assume $\theta$ is uniformly distributed, and $p_{\cos{\theta}} = 1/\pi/(1-\cos^2{\theta)^{1/2}}$.
 Lastly, we transform the $p_{\sin{\psi}}$ to $p_{\cos{\psi}}$, where $p_{\sin{\psi}} = p_{\cos{\psi}}\sqrt{1-\cos^2{\psi}}/\cos{\psi}$.
 Combining all the pieces together, we get
-\begin{equation}
+\begin{equation}\label{eqn:jac_istar}
     p_{\cos{i_\star}} = \frac{4}{\pi} \int_{\cos{i_\star}}^{1} \frac{\cos{i_\star}/\cos{\theta}}{\sqrt{\cos^2{\theta}-\cos^2{i_\star}}} \frac{1}{\sqrt{1-\cos^2{\theta}}} p_{\cos{\psi}} d\cos{\theta}.
 \end{equation}
-If $\cos{\psi}$ is uniformly distribution, i.e., $p_{\cos{\psi}} = 1/2$, the integral gives $1$, which suggests the $\cos{i_\star}$ is uniform.
+If $\cos{\psi}$ is uniformly distribution, i.e., $p_{\cos{\psi}} = 1/2$, the integral gives $1$, which suggests the $\cos{i_\star}$ probability density function is uniform.
 
-We note that we could not infer $\psi$ from $\lambda$ and $\theta$ since they are not independent variables. Instead, we will have to infer $\psi$ from $\lambda$ and $i_\star$.
+From Equation~(\ref{eqn:jac_lam}) and (\ref{eqn:jac_istar}), we can now derive the $\lambda$ and $i_\star$ distributions for any given $\psi$ distributions, assuming the azimuthal angle of the stellar spin axis $\theta$ is random.
+In Figure~\ref{fig:transform}, we find and examine numerical solutions of the $\lambda$ and $i_\star$ distributions for four different $\cos{\psi}$ distributions.
+The first row in Figure~\ref{fig:transform} presents an isotropic $\psi$ where $\cos{\psi} \sim \mathcal{U}(-1,1)$, and the second to fourth rows in Figure~\ref{fig:transform} present Normal distributions of $\cos{\psi}$ following $\mathcal{N}(0,0.2)$, $\mathcal{N}(-0.4,0.2)$, and $\mathcal{N}(0.4,0.2)$, respectively.
+In each row, we solve the $\lambda$ and $i_\star$ distributions numerically as shown in blue curves, and samplings of $\lambda$ and $i_\star$ from $\cos{\psi}$ and $\theta$ distributions as shown in grey histograms.
+For an uniform $\cos{\psi}$ distribution, the $\lambda$ distribution is uniform and the $i_\star$ distribution is isotropic and proportional to $\sin{i_\star}$, as expected.
+
+Interestingly, for a given stellar obliquity distribution, the $\lambda$ distribution is sensitive to and closely related to the underlying $\psi$ distribution, as shown in Figure~\ref{fig:transform}.
+For different stellar obliquity distributions, the $\lambda$ distributions are distinguishable from each other and thus, when we infer the $\psi$ distribution from the $\lambda$ distribution, the solution will not be degenerate.
+On the other hand, the $i_\star$ distributions less depend on the underlying $\psi$ distribution.
+Compared to an isotropic $i_\star$ distribution, the $i_\star$ distributions for different $\psi$ distributions differ at the low values (i.e., $i_\star < \pi/4$), which places a challenge to observation.
+More importantly, since the $i_\star$ distributions are less sensitive to the underlying $\psi$ distributions, when we infer the $\psi$ distributions from the $i_\star$ distribution, the solution could be very degenerate.
+
+Due to the lack dependency on the $i_\star$ distribution, the $\psi$ distribution could be inferred purely from the $\lambda$ distribution without losing much information.
+We note that although we could find a relation between $\psi$, $\lambda$, and $\theta$, the $\psi$ distribution cannot be inferred from the $\lambda$ and $\theta$ distributions since $\lambda$ and $\theta$ are not independent variables.
+We will have to infer the $\psi$ distribution from the $\lambda$ and $i_\star$ distributions.
+In the \nth{4} column in Figure~\ref{fig:transform}, we find the $\psi$ distribution from the $\lambda$ distribution, simply assuming an isotropic $i_\star$ distribution.
+As shown in grey histograms, the method could correctly recover the peaks and widths of the underlying $\psi$ distributions (presented as blue dashed lines).
+The method over-predicts $\cos{\psi}$ at $\pm 1$ because of the over-prediction of $i_\star$ near $90\degr$.
+Later we show that this minor deviation from the true distribution has little impact on the population inference.
 
@@
     \includegraphics[width=\linewidth]{figures/transform.pdf}
-    \caption{Distribution transformations between the stellar obliquity $\cos{\psi}$ (1st column), sky-projected stellar obliquity $\lambda$ (2nd column), and stellar inclination $i_\star$ (3rd column).
-The inferred $\cos{\psi}$ assuming an isotropic stellar inclination is shown in the 4th column.
-Each row has a simulated $\cos{\psi}$ distribution.
-Assuming a uniform $\theta$, the simulated $\lambda$ and $i_\star$ distributions are shown in grey histograms, and the analytical solutions are shown as blue curves.}
+    \caption{Simulated $\cos{\psi}$ distributions (\nth{1} column) and their corresponding sky-projected stellar obliquity $\lambda$ (\nth{2} column) and stellar inclination $i_\star$ (\nth{3} column) distributions.
+The inferred $\cos{\psi}$ distributions assuming isotropic stellar inclinations are shown in the \nth{4} column.
+The random samplings of $\lambda$ and $i_\star$ out of the $\cos{\psi}$ distributions are shown as grey histograms, and the numerical solutions are shown as blue curves.}
     \label{fig:transform}
@@
 \section{Hierarchical Bayesian Framework}\label{sec:hbm}
+
+Our goal is to model the stellar obliquity distribution of all exoplanetary systems with sky-projected stellar obliquity measurements.
+The model could be extended to systems that also have stellar inclination information.
+In Section~\ref{sec:jacobian}, we showed the stellar inclination information is less important in the stellar obliquity distribution inference.
+Therefore, we aim to build a flexible framework that allows users to include or not include stellar inclination information.
 
@@
 The stellar obliquity distribution is described by hyperparameters $\vb*{\beta}$ and constrained by $n_{\rm th}$ system's stellar obliquity $\psi_n$.
-Each $\psi_n$ is calculated by the system's sky-projected stellar obliquity $\lambda_n$ and stellar inclination $i_{\star, n}$.
-$\theta_{\star, n}$ contains other information of the star other than $i_{\star, n}$, such as the stellar radius and stellar rotation period.
-Obs$_{\star,n}$ contains observations of the star, such as the stellar spectral and photometric and/or spectroscopic time series, and $\hat{\lambda}_n$ has the measured sky-projected stellar obliquity.}
+Each $\psi_n$ is calculated by the system's sky-projected stellar obliquity $\lambda_n$ and stellar inclination $i_{\star, n}$, if available.
+$\gamma_{\star, n}$ contains properties of the star other than inclination, such as stellar radius and rotation period for $i_{\star, n}$ inference.
+Obs$_{\star,n}$ includes all the observed properties of the star with uncertainties, and $\hat{\lambda}_n$ has the measured sky-projected stellar obliquity and its uncertainty.}
     \label{fig:graph}
@@
 
-We build a hierarchical Bayesian framework to model the 3D stellar obliquity ($\psi$) distribution of exoplanetary systems.
-Figure~\ref{fig:graph} shows the probabilistic graphical model.
+Figure~\ref{fig:graph} shows the probabilistic graphical model for the hierarchical Bayesian framework.
 We aim to constrain a set of hyperparameters $\bm{\beta}$ that describe the stellar obliquity $\psi$ distribution.
@@
 For simplicity, we approximate $i_{\rm orb}$ to $90\degr$ and use the simplified version of Equation~(\ref{eqn:psi}) to compute $\psi_n$.
-The parameter $\theta_{\star, n}$ contains stellar parameters other than $i_\star$, e.g., stellar rotation period and stellar radius, if the information is available.
+The parameter $\gamma_{\star, n}$ contains stellar parameters other than $i_\star$, e.g., stellar rotation period and stellar radius, if the information is available.
 Obs$_{\star,n}$ describes observations that help to constrain $i_{\star,n}$, such as photometric or spectroscopic time series and the stellar spectra.
-If the rotational modulation method is used to constrain the stellar inclination, the $i_{\star,n}$ and $\theta_{\star,n}$ are constrained by rotational modulation in photometric or spectroscopic time series and observed sky-projected stellar rotational line broadening $\hat{v}\sin{i_{\star,n}}$ \citep[e.g.,][]{Masuda20}.
+If the rotational modulation method is used to constrain the stellar inclination, the $i_{\star,n}$ and $\gamma_{\star,n}$ are constrained by rotational modulation in photometric or spectroscopic time series and observed sky-projected stellar rotational line broadening $\hat{v}\sin{i_{\star,n}}$ \citep[e.g.,][]{Masuda20}.
 If the gravity-darkening method is used, the $i_{\star,n}$ is constrained by the anomaly in transit light curves.
@@
 \begin{align}
-    L(\lambda) &\sim \prod_{i=1}^N\mathcal{N}(\hat{\lambda}_n, \sigma_{\hat{\lambda}_n}) \nonumber\\
-    L(i_\star, \theta_\star) &\sim \prod_{i=1}^N\mathcal{N}({\rm Obs}_{\star,n}, \sigma_{{\rm Obs}_{\star,n}}).
+    \mathcal{L}(\lambda) &\sim \prod_{i=1}^N\mathcal{N}(\hat{\lambda}_n, \sigma_{\hat{\lambda}_n}) \nonumber\\
+    \mathcal{L}(i_\star, \gamma_\star) &\sim \prod_{i=1}^N\mathcal{N}({\rm Obs}_{\star,n}, \sigma_{{\rm Obs}_{\star,n}}).
 \end{align}
@@
         \includegraphics{figures/psi_dist.pdf}
-        \caption{The distributions of the cosine of 3D stellar obliquity $\cos{\psi}$.
-The left panel presents the inferred $\cos{\psi}$ distribution for the \numistar systems that have both sky-projected stellar obliquity $\lambda$ and stellar inclination $i_\star$ measurements using the statistical model described in Figure~\ref{fig:graph}.
-The middle panel uses the same sample as the left panel but now removes the likelihood function on $i_\star$.
-The right panel presents the $\cos{\psi}$ distribution for all \numall systems that have sky-projected stellar obliquity $\lambda$ measurements and has no likelihood function on $i_\star$.
-The $\cos{\psi}$ distribution for the observed-$i_\star$ are similar, and both show a cluster near $\cos{\psi} = -0.2$, whereas the distribution for all systems does not have such a trend.}
+        \caption{Stellar obliquity distribution of all exoplanetary systems with sky-projected stellar obliquity measurements.}
         \label{fig:psi_dist}

### 2023-01-09 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \usepackage[super]{nth}
+\usepackage{fontawesome}
+\usepackage{hyperref}
 
@@
 The constraint on stellar obliquity distribution is predominantly based on sky-projected stellar obliquities.
-We introduce a flexible, hierarchical Bayesian framework to infer the stellar obliquity distribution of exoplanetary systems from their sky-projected stellar obliquities, stellar inclinations, or both.
-Applying the framework to all systems with projected stellar obliquity measurements, which are mostly Hot Jupiter systems, we find that the stellar obliquity distribution is unimodal and peaked at zero degrees.
+We introduce a flexible, hierarchical Bayesian framework that allows inferring the stellar obliquity distribution from sky-projected stellar obliquities, and if available, stellar inclinations.
+Applying the framework to all systems with sky-projected stellar obliquity measurements, which are mostly Hot Jupiter systems, we find that the stellar obliquity distribution is unimodal and peaked at zero degrees.
 Misaligned systems have nearly isotropic stellar obliquities, at odds with the analysis of systems that also have $i_\star$ measurements.
@@
 Recent stellar obliquity measurements on Warm Jupiter systems with large planet-star separations and systems that are still at young ages both find a tendency of system alignment.
-If the trend holds as more observations coming up, it could suggest that the primordial stellar obliquity is quite low.
+If the trend holds as more observations come up, it could suggest that the primordial stellar obliquity is quite low.
 Stellar obliquities of misaligned systems could then be linked to the dynamical processes that result in the current orbit of close-in planets.
@@
 
-In reality, only the sky-projected stellar obliquity $\lambda$, the angle between the projections of $\bf{n}_{\rm orb}$ and $\bf{n}_{\star}$ onto the plane of the sky, can be measured.
-Stellar obliquity $\psi$ can be found if the stellar inclination $i_\star$ is known.
+In observation, only the sky-projected stellar obliquity $\lambda$, the angle between the projections of $\bf{n}_{\rm orb}$ and $\bf{n}_{\star}$ onto the plane of the sky, can be measured.
+Stellar obliquity $\psi$ can be found, if the stellar inclination $i_\star$ is known.
 The calculation follows \citep[e.g.,][]{Fabrycky09}:
@@
 
-It is not yet clear how does the stellar inclination measurement affect the inference of the stellar obliquity distribution of a population. 
+It is not yet clear how the stellar inclination measurement affects the inference of the stellar obliquity distribution of a population. 
 In this work, we propose an approach to infer $\psi$ distribution from $\lambda$ distribution and show that the stellar obliquity distribution inference only weakly depends on stellar inclination.
@@
 In Section~\ref{sec:hbm}, we introduce a flexible, hierarchical Bayesian framework to infer the population stellar obliquity distribution.
-In Section~\ref{sec:data}, we apply the framework to all exoplanetary systems with measured sky-projected obliquity.
+In Section~\ref{sec:applications}, we apply the framework to simulated data and real observations.
 
@@
 In both coordinates, we set the observer's line of sight to be one of the two horizontal axes, and the orbital axis of the planet to be the vertical axis.
-We approximate the orbital inclination of the transiting planet to $90\degr$ here and in the rest of the work to simplify the problem for population study.
+We approximate the orbital inclination of the transiting planet to $90\degr$ here and in the rest of the work to simplify the problem for a population study.
 If the orbital inclination deviates from $90\degr$, $\bf{n}_{\rm orb}$ instead of being the vertical axis needs to be rotated $\pm (90\degr-i_{\rm orb})$ along the line of sight.
 
-The $\{\psi, \theta\}$ coordinate system shown in panel (a) relates to physical properties of a planetary system.
+The $\{\psi, \theta\}$ coordinate system shown in panel (a) relates to the physical properties of a planetary system.
 $\psi$ is the angle between the stellar spin axis and the planetary orbital axis and $\theta$ is the azimuthal angle of the stellar spin axis.
-For the given $\bf{n}_{\rm orb}$ axis, if $\bf{n}_{\star}$ is a random vector with uniform distribution on a three-dimensional sphere, i.e., $\bf{n}_{\star}$ is isotropically distributed, the probability distribution of $\psi$ follows $p_{\psi} \sim \sin{\psi}$ and $p_\theta \sim 1/2\pi$.
+For the given $\bf{n}_{\rm orb}$ axis, if $\bf{n}_{\star}$ is a random vector with uniform distribution on a three-dimensional sphere, i.e., $\bf{n}_{\star}$ is isotropically distributed, the probability density function of $\psi$ follows $p_{\psi} \sim \sin{\psi}$ and $p_\theta \sim 1/2\pi$.
 Since $p_{\cos{\psi}} = p_{\psi} \abs{\dv*{\psi}{\cos{\psi}}}$, $p_{\cos{\psi}}$ is uniformly distributed between $-1$ and $1$ for isotropic stellar obliquity.
@@
 This also limits $\theta$ to $[-\pi/2, \pi/2]$.
+In observation, there is also a degeneracy between $-\lambda$ and $+\lambda$. Therefore, we also limit $\lambda$ to $[0,\pi]$.
 
@@
 In each row, we solve the $\lambda$ and $i_\star$ distributions numerically as shown in blue curves, and samplings of $\lambda$ and $i_\star$ from $\cos{\psi}$ and $\theta$ distributions as shown in grey histograms.
-For an uniform $\cos{\psi}$ distribution, the $\lambda$ distribution is uniform and the $i_\star$ distribution is isotropic and proportional to $\sin{i_\star}$, as expected.
+For a uniform $\cos{\psi}$ distribution, the $\lambda$ distribution is uniform, and the $i_\star$ distribution is isotropic and proportional to $\sin{i_\star}$, as expected.
 
 Interestingly, for a given stellar obliquity distribution, the $\lambda$ distribution is sensitive to and closely related to the underlying $\psi$ distribution, as shown in Figure~\ref{fig:transform}.
-For different stellar obliquity distributions, the $\lambda$ distributions are distinguishable from each other and thus, when we infer the $\psi$ distribution from the $\lambda$ distribution, the solution will not be degenerate.
+For different stellar obliquity distributions, the $\lambda$ distributions are distinguishable from each other, and thus, when we infer the $\psi$ distribution from the $\lambda$ distribution, the solution will not degenerate.
 On the other hand, the $i_\star$ distributions less depend on the underlying $\psi$ distribution.
 Compared to an isotropic $i_\star$ distribution, the $i_\star$ distributions for different $\psi$ distributions differ at the low values (i.e., $i_\star < \pi/4$), which places a challenge to observation.
-More importantly, since the $i_\star$ distributions are less sensitive to the underlying $\psi$ distributions, when we infer the $\psi$ distributions from the $i_\star$ distribution, the solution could be very degenerate.
+More importantly, since the $i_\star$ distributions are less sensitive to the underlying $\psi$ distributions when we infer the $\psi$ distributions from the $i_\star$ distribution, the solution could be very degenerate.
 
-Due to the lack dependency on the $i_\star$ distribution, the $\psi$ distribution could be inferred purely from the $\lambda$ distribution without losing much information.
+Due to the lack of dependency on the $i_\star$ distribution, the $\psi$ distribution could be inferred purely from the $\lambda$ distribution without losing much information.
 We note that although we could find a relation between $\psi$, $\lambda$, and $\theta$, the $\psi$ distribution cannot be inferred from the $\lambda$ and $\theta$ distributions since $\lambda$ and $\theta$ are not independent variables.
@@
 
-Our goal is to model the stellar obliquity distribution of all exoplanetary systems with sky-projected stellar obliquity measurements.
-The model could be extended to systems that also have stellar inclination information.
-In Section~\ref{sec:jacobian}, we showed the stellar inclination information is less important in the stellar obliquity distribution inference.
-Therefore, we aim to build a flexible framework that allows users to include or not include stellar inclination information.
+To find the unknown stellar obliquity distribution of exoplanetary systems, we build a hierarchical Bayesian framework that takes sky-projected stellar obliquities as observed data.
+In Section~\ref{sec:jacobian}, we found that the stellar inclination is not sensitive to the stellar obliquity distribution.
+Therefore, we focus on a flexible framework that makes the stellar inclination information optional.
 
@@
 We aim to constrain a set of hyperparameters $\bm{\beta}$ that describe the stellar obliquity $\psi$ distribution.
-$\bm{\beta}$ is constrained by $N$ individual systems, where each system has $\psi_n$ that is calculated by the sky-projected stellar obliquity $\lambda_n$ and the stellar inclination $i_n$.
+The parameter set $\bm{\beta}$ is constrained by $N$ individual systems in which $\psi_n$ is a random variable.
+$\psi_n$ is computed from the sky-projected stellar obliquity $\lambda_n$ and the stellar inclination $i_n$.
 For simplicity, we approximate $i_{\rm orb}$ to $90\degr$ and use the simplified version of Equation~(\ref{eqn:psi}) to compute $\psi_n$.
-The parameter $\gamma_{\star, n}$ contains stellar parameters other than $i_\star$, e.g., stellar rotation period and stellar radius, if the information is available.
-Obs$_{\star,n}$ describes observations that help to constrain $i_{\star,n}$, such as photometric or spectroscopic time series and the stellar spectra.
-If the rotational modulation method is used to constrain the stellar inclination, the $i_{\star,n}$ and $\gamma_{\star,n}$ are constrained by rotational modulation in photometric or spectroscopic time series and observed sky-projected stellar rotational line broadening $\hat{v}\sin{i_{\star,n}}$ \citep[e.g.,][]{Masuda20}.
+$\psi$ could be computed more accurately by including $i_{\rm orb}$ as another random variable to be constrained by e.g., transit observations.
+The parameter $\gamma_{\star, n}$ is another parameter set that contains all the stellar properties other than $i_\star$, e.g., stellar rotation period $P_{\rm rot}$ and radius $R_\star$, if the information is available.
+Obs$_{\star,n}$ includes observed data that helps to constrain $i_{\star,n}$ and $\gamma_{\star, n}$.
+If the rotational modulation method is used to constrain the stellar inclination, the $i_{\star,n}$ and $\gamma_{\star,n}$ are constrained by rotational modulation in photometric or spectroscopic time series and observed sky-projected stellar rotational line broadening $\hat{v}\sin{i_{\star,n}}$ \citep[e.g.,][]{Masuda20}, where we have $v\sin{i_\star} = 2 \pi R_\star / P_{\rm rot}$.
 If the gravity-darkening method is used, the $i_{\star,n}$ is constrained by the anomaly in transit light curves.
 If asteroseismology is used, the $i_{\star,n}$ is constrained from the periodic variation in the photometric time series.
-Lastly, the sky-projected stellar obliquity $\lambda_n$ is constrained by observed $\hat{\lambda}_n$.
+Lastly, the sky-projected stellar obliquity $\lambda_n$ is constrained by observed $\hat{\lambda}_n$ via the Rossiter-McLaughlin effect or gravity darkening.
 
-For the stellar obliquity distribution, we adopt a two-component mixture Beta distribution with hyperparameters $\bm{\beta} = \{\bm{w},\bm{a},\bm{b}\}$, where each hyperparameter has a dimension of 2.
-The Beta distribution is flexible to describe both the spike of aligned systems with 3D stellar obliquities close to $0\degr$ and the clustered or broadly distributed systems of misaligned systems.
-The probability distribution follows 
+For the stellar obliquity distribution, we model the $\cos{\psi}$ distribution instead of $\psi$ distribution to aim to understand if the stellar obliquity is isotropically distributed or not.
+We adopt a two-component mixture Beta distribution with hyperparameters $\bm{\beta} = \{\bm{w},\bm{a},\bm{b}\}$, where each hyperparameter has a dimension of 2.
+One component is designed for a large population of well-aligned systems, and the other component is designed for misaligned systems.
+The Beta distribution is a flexible distribution that is able to describe the spike of aligned systems with stellar obliquities close to $0\degr$ and also the clustered or broadly distributed misaligned systems.
+The probability density function of $\cos{\psi}$ follows 
 \begin{align}
@@
 \end{align}
-Since a Beta distribution is defined on the interval $[0, 1]$, we extend its support to $[-1,1]$ using a linear transformation (i.e., $2\times Beta-1$).
-
-The hyperparameter and parameter priors are the following:
+Since the Beta distribution is defined on the interval $[0, 1]$, we extend its support to $[-1,1]$ using a linear transformation (i.e., $2\times {\rm Beta} - 1$).
+For the ${\rm Beta}(a_0,b_0)$ component, we have it to describe the population of aligned systems.
+For the $a_0/b_0 \gg 1$, the distribution will have a spike at $\cos{\psi} = 1$.
+For the ${\rm Beta}(a_1,b_1)$ component, we have it to describe the misaligned systems.
+We design the order of the two components to avoid label switching in the mixture model.
+With the design, the hyperparameter and parameter priors are the following:
 \begin{align}
@@
 \end{align}
-We set the hyperpriors such that $a_0$ and $b_0$ will be used to infer the spike of $\cos{\psi}$ near 1, and $a_1$ and $b_1$ infer the distribution of the misaligned population.
-Doing so will also avoid label switching in our mixture model.
-The probability distribution of an isotropic $i_{\rm star}$ follows $p_{i_\star} \sim \sin{i_{\star}}$, or $p_{\cos{i_\star}} \sim 1$.
-The $\lambda$ should be the angle between two random unit vectors on a 2D circle and always between 0 and $\pi$.
-
-The likelihood functions are the following:
+The likelihood functions follows:
 \begin{align}
     \mathcal{L}(\lambda) &\sim \prod_{i=1}^N\mathcal{N}(\hat{\lambda}_n, \sigma_{\hat{\lambda}_n}) \nonumber\\
-    \mathcal{L}(i_\star, \gamma_\star) &\sim \prod_{i=1}^N\mathcal{N}({\rm Obs}_{\star,n}, \sigma_{{\rm Obs}_{\star,n}}).
+    \mathcal{L}(i_\star) &\sim \prod_{i=1}^N\mathcal{N}({\rm Obs}_{\star,n}, \sigma_{{\rm Obs}_{\star,n}}).
 \end{align}
+If $\gamma_{\star, n}$ is available, we construct Normal distributions with means and standard deviations from their observed data.
+Here we assume uniform hyperpriors for the Beta distribution.
+For population inference on a small number of systems, the choice of hyperpriors could have an impact on the inferred distribution \citep{Nagpal22}.
+When applying the framework to a small sample size, it is important to use different hyperpriors to test the robustness of the model.
+
+The probabilistic model is built with the $\mathtt{PyMC}$ \citep[$\mathtt{v4.1.7}$;][]{pymc} package and the posteriors are sampled using the No-U-Turn Sampler \citep[NUTS;][]{Hoffman11}, a gradient-based sampling algorithm of the Markov chain Monte Carlo (MCMC).
+All figures and simulations in this paper are built with the $\mathtt{showyourwork}$ package and fully reproducible on \href{https://github.com/jiayindong/polar}{GitHub\,\faGithub}.
+
+\section{Applications to Simulated and Observational Data}\label{sec:applications}
+
+\subsection{Simulated Data}
+
+We first apply the hierarchical Bayesian framework to simulated data of which the ground-truth stellar obliquity distribution is known.
+We adopt the four $\cos{\psi}$ distributions discussed in Section~\ref{sec:jacobian}: a uniform distribution $\mathcal{U}(-1,1)$ and three Normal distributions $\mathcal{N}(0,0.2)$, $\mathcal{N}(-0.4,0.2)$, and $\mathcal{N}(0.4,0.2)$.
+For each $\cos{\psi}$ distribution, we randomly generate 200 samples of stellar inclination $i_\star$ and sky-projected stellar obliquity $\lambda$, assuming a uniform $\theta$.
+The sampled $i_\star$ and $\lambda$ here are \emph{true} values.
+We then add some Gaussian noises to the \emph{true} $\lambda$ and $i_\star$ to simulate the \emph{observed} data.
+We choose uncertainties of $\sigma_{\lambda} = 8\degr$ and $\sigma_{i_\star} = 10\degr$, which are typical observational uncertainties summarized in \cite{Albrecht22}.
+Using the observed $\lambda$ and their uncertainties, we infer the $\cos{\psi}$ distribution of the sample with or without $i_\star$ likelihoods.
+
+In Figure~\ref{fig:simulation}, we present the inferred stellar obliquity distributions.
+Since the simulated stellar obliquity distributions only have a single component, we model the data with a single Beta distribution.
+Each row corresponds to an injected stellar obliquity distribution.
+The orange curve and contours are the median and uncertainties of the inferred $\cos{\psi}$ distribution with stellar inclination information, and the blue curve and contours are the ones without stellar inclination information.
+We simply assume an isotropic distribution of $i_\star$ for these models.
+As shown in Figure~\ref{fig:simulation}, in all four distributions, the inferred $\cos{\psi}$ distributions with or without $i_\star$ correctly recover the injected distributions shown as grey dashed lines.
+Since some of the injected distributions are Normal distributions, it is expected that the inferred distributions, which as Beta distributions, could not exactly describe the injected distributions.
+Comparing the blue curves with the orange curves, the overestimation of $\cos{\psi}$ near $\pm1$ under the assumption of an isotropic $i_\star$ has little impact on the inference.
+The distribution is mainly constrained by the peak and width of the stellar obliquity distribution, which are both correctly predicted by the isotropic $i_\star$ model.
+
+From the simulated data, we confirm that while stellar inclinations could help to constrain the stellar obliquity of individual systems, they are not necessarily required for the population stellar obliquity inference.
+The constraint on stellar obliquity distribution is predominantly based on sky-projected stellar obliquities.
+The sky-projected stellar obliquity could correctly infer the peak and width of the underlying stellar obliquity distribution.
 
@@
     \includegraphics[width=\linewidth]{figures/simulation.pdf}
-    \caption{Inferred $\cos{\psi}$ from simulated data.}
+    \caption{Inferred stellar obliquity distributions from sky-projected stellar obliquities with or without stellar inclination information.
+Each row presents a set of simulated data where the underlying $\cos{\psi}$ distribution is known and shown as grey dashed curves.
+The inferred $\cos{\psi}$ distributions with or without stellar inclination are shown in orange and blue curves, respectively.
+The shallow contours are $1\sigma$ and $2\sigma$ uncertainties of the inferred distributions.}
     \label{fig:simulation}
 \end{figure}
+
+\subsection{Exoplanetary Stellar Obliquity Distribution}
+
+We then apply the hierarchical Bayesian framework to all exoplanetary systems that have sky-projected stellar obliquity measurements.
+We use the sample of \numall systems that have sky-projected stellar obliquity $\lambda$ measurements summarized in \cite{Albrecht22} Table A1, which are predominantly Hot Jupiter systems.
+The inferred $\cos{\psi}$ is shown in Figure~\ref{fig:psi_dist}.
+The $\cos{\psi}$ distribution peaks at 1 but is nearly flat between $-0.75$ and $0.75$ with no significant clustering.
+The distribution corresponds to an isotropic distribution of $\psi$ from $40\degr$ to $140\degr$.
+The posteriors of the hyperparameters are xxx.
 
@@
         \includegraphics{figures/psi_dist.pdf}
-        \caption{Stellar obliquity distribution of all exoplanetary systems with sky-projected stellar obliquity measurements.}
+        \caption{Stellar obliquity distribution of all exoplanetary systems with sky-projected stellar obliquity measurements.
+The inference is purely from observed sky-projected stellar obliquities and assumes the stellar inclination is isotropic.}
         \label{fig:psi_dist}
@@
 
-First, we apply the framework to the sample of \numistar systems that have both sky-projected stellar obliquity $\lambda$ and stellar inclination $i_\star$ constraints summarized in the \cite{Albrecht22} review.
+The inferred stellar obliquity distribution of the full sample of systems is at odds with the previous analysis of the subsample of systems that also have $i_\star$ measurements.
+The subsample identified a preponderance of perpendicular planets that disfavored an isotropic stellar obliquity distribution \citep{Albrecht21}.
+We summarize two possible reasons to explain the discrepancies, which are open to further investigation.
+First, the sample of systems with $i_\star$ measurements is small, and only about 20 systems are misaligned systems used to constrain the misaligned systems' stellar obliquity distribution.
+Second, the sample could be biased because of the $i_\star$ measurements.
+
+Our analysis suggests misaligned systems have nearly isotropic stellar obliquity distribution.
+When comparing the inferred distribution with the prediction of stellar obliquity distributions in different Hot Jupiter origin scenarios, it is in best agreement with the outcome of the scattering of multiple giant planets after a convergent disk migration \citep{Beague12}.
+
+\section{Summary \& Discussion}
+
+In this work, we demonstrate the stellar obliquity distribution inference is predominantly based on the sky-projected stellar obliquity.
+While stellar inclination improves the stellar obliquity constraint on individual systems, it does not improve the population-level stellar obliquity constraint.
+We introduce a flexible, hierarchical Bayesian framework to infer the stellar obliquity distribution of a population.
+The model is an open-source code that is available to the public to apply to a certain sample of targets.
+When applying our framework to all exoplanetary systems that have sky-projected stellar obliquity measurements, we found no significant clustering near $\cos{\psi} = 0$, at odds with the previous study of systems that have both sky-projected stellar obliquity and stellar obliquity measurements.
+Potential reasons and biased that lead to this discrepancy is open to discussion.
+We find for misaligned exoplanetary systems, their stellar obliquity is nearly isotropically distributed, which may suggest a convergent disk migration plus a scattering stage in these planets' dynamical history.
+Future stellar obliquity distribution study on a carefully selected Hot Jupiter sample will shed light on the origin of Hot Jupiters.
+
+% \section{More}
+
+% First, we apply the framework to the sample of \numistar systems that have both sky-projected stellar obliquity $\lambda$ and stellar inclination $i_\star$ constraints summarized in the \cite{Albrecht22} review.
 If the host star's rotation period is available, we use the value along with the stellar radii and $v\sin{i_\star}$ to infer the stellar inclination.
 Otherwise, we use the reported stellar inclination and uncertainty listed in Table A1 of \cite{Albrecht22}.
-The inferred $\cos{\psi}$ distribution is shown in the left panel of Figure~\ref{fig:psi_dist}.
+% The inferred $\cos{\psi}$ distribution is shown in the left panel of Figure~\ref{fig:psi_dist}.
 The distribution presents a pike at 1, suggesting the majority of exoplanetary systems are well-aligned systems.
@@
 
-To test the dependence of $\cos{\psi}$ distribution on $i_\star$ measurements, we apply the model again to the sample of \numistar systems with both constrained $\lambda$ and $i_\star$.
+% To test the dependence of $\cos{\psi}$ distribution on $i_\star$ measurements, we apply the model again to the sample of \numistar systems with both constrained $\lambda$ and $i_\star$.
 However, in this model, we do not use any additional information on $i_\star$ other than applying an isotropic inclination prior to each system.
 The corresponding probabilistic graphical model is Figure~\ref{fig:graph} but removing all the dashed edges.
-The $\cos{\psi}$ distribution of the new model is shown in the middle panel of Figure~\ref{fig:psi_dist}.
+% The $\cos{\psi}$ distribution of the new model is shown in the middle panel of Figure~\ref{fig:psi_dist}.
 Interestingly, the overall $\cos{\psi}$ distributions of the two models are very similar, despite the fact that the inferred $\psi$ uncertainty for each system is now greater.
-The similarity between the two distributions suggests that, at least for the observed-$i_\star$ sample, the $\cos{\psi}$ distribution mainly depends on the underlying sky-project stellar obliquity $\lambda$ distribution.
+% The similarity between the two distributions suggests that, at least for the observed-$i_\star$ sample, the $\cos{\psi}$ distribution mainly depends on the underlying sky-project stellar obliquity $\lambda$ distribution.
 
-Consequently, we apply our model to all \numall exoplanetary systems that have sky-projected stellar obliquity measurements summarized in \cite{Albrecht22}.
+% Consequently, we apply our model to all \numall exoplanetary systems that have sky-projected stellar obliquity measurements summarized in \cite{Albrecht22}.
 The result is presented in the right panel of Figure~\ref{fig:psi_dist}.
@@
 
-\section{Interpretations of the \texorpdfstring{$\psi$}{psi} distributions}\label{sec:interpretations}
-
-In this section, we aim to explain the similarity and disparity between the $\cos{\psi}$ distributions presented in Figure~\ref{fig:psi_dist}.
-
-\subsection{Do we need \texorpdfstring{$i_\star$}{istar} to infer the \texorpdfstring{$\psi$}{psi} distribution?}
-
-There is no doubt that $i_\star$ measurement on an individual system improves the constraint on $\psi$.
+% There is no doubt that $i_\star$ measurement on an individual system improves the constraint on $\psi$.
 However, are $i_\star$ measurements required to constrain the population $\psi$ distribution?
@@
 
-Given a certain $\psi$ distribution, the $\lambda$ and $i_\star$ are no longer randomly/isotropically distributed.
-We first find relations between $\{\psi, \theta\}$ and $\{\lambda, i_\star\}$, where $\theta$ is the azimuthal angle of the stellar spin axis, as shown in Figure~\ref{fig:coord}.
-We transform $\bf{n}_{\star}$ in two coordinate systems to Cartesian, pair their terms, and get:
-\begin{align}
-    \sin{\psi}\cos{\theta} = \cos{i_\star}& \nonumber\\
-    \sin{\psi}\sin{\theta} = \sin{\lambda}\sin{i_\star}& \nonumber\\
-    \cos{\psi} = \cos{\lambda}\sin{i_\star}&.
-\end{align}
-Thus, $i_\star = \cos[-1](\sin{\psi}\cos{\theta})$ and $\lambda = \tan[-1](\tan{\psi}\sin{\theta})$. 
-We could calculate the probability distribution of $i_\star$ or $\lambda$ from the probability distributions of $\psi$ and $\theta$.
-Since $\psi$ and $\theta$ are independent random variables, it is possible to marginalize over $\theta$ to find a one-on-one relation between $\lambda$ and $\psi$, or even between $i_\star$ and $\psi$.
-If the azimuthal angle of the stellar rotation axis $\theta$ is randomly distributed, we could directly infer the $\psi$ distribution from the $\lambda$ distribution or even $i_\star$ distribution.
-An analytical solution between $p_\lambda$ and $p_\psi$ can be found in \cite{Fabrycky09}. 
-In our Bayesian framework, both $\psi$ and $\theta$ distributions are simultaneously inferred from the $\lambda$ and $i_\star$ distributions.
-Without any $i_\star$ measurements, the inferred $\theta$ indeed follows a uniform distribution.
-With $i_\star$ measurements, the inferred $\theta$ still follows a uniform distribution for aligned systems.
-For misaligned systems, the inferred $\theta$ distribution is similar to the $i_\star$ distribution as the $\psi$ are close to $90\degr$.
-For these systems, the $\lambda$ are close to $90\degr$ as well, so the $\theta$ angles are no longer important for the inference.
-Overall, these two reasons explain why the inferred $\cos{\psi}$ distributions are similar in the left and middle panels of Figure~\ref{fig:psi_dist}.
-
-We also design some $\cos{\psi}$ distributions and infer their distributions purely from the $\lambda$ distributions.
+% We also design some $\cos{\psi}$ distributions and infer their distributions purely from the $\lambda$ distributions.
 From each distribution of $\cos{\psi}$ and assuming a uniform $\theta$, we randomly generate 200 samples of $i_\star$ and $\lambda$ and assume them as \emph{true} values.
@@
 Lastly, from the observed values and their uncertainties, we infer the $\cos{\psi}$ distribution of the sample with or without $i_\star$ measurements.
-We test four different distributions of $\cos{\psi}$: a uniform distribution $\mathcal{U}(-1,1)$ and three Normal distributions $\mathcal{N}(0,0.2)$, $\mathcal{N}(-0.4,0.2)$, and $\mathcal{N}(0.4,0.2)$.
-In all four distributions, we find inferences with or without $i_\star$ are consistent with each other and injected distributions.
+% We test four different distributions of $\cos{\psi}$: a uniform distribution $\mathcal{U}(-1,1)$ and three Normal distributions $\mathcal{N}(0,0.2)$, $\mathcal{N}(-0.4,0.2)$, and $\mathcal{N}(0.4,0.2)$.
+% In all four distributions, we find inferences with or without $i_\star$ are consistent with each other and injected distributions.
 
-In conclusion, the stellar obliquity distribution can be inferred directly from the sky-projected stellar obliquity distribution.
-A one-on-one relation could be found if we assume the azimuthal angle of the stellar spin axis is uniformly distributed around the planet's orbital axis.
-
-\subsection{The underlying \texorpdfstring{$\lambda$}{lambda} distributions}
-
-It was surprising that the inferred $\cos{\psi}$ distribution of the observed-$i_\star$ sample evidently differs from the $\cos{\psi}$ distribution of the full sample, shown as the middle and right panels of Figure~\ref{fig:psi_dist}.
-Since both distributions are inferred with no $i_\star$ information, the disparity is likely introduced by the underlying $\lambda$ distribution.
-
-We find the underlying $\lambda$ distributions of the observed-$i_\star$ sample and the full sample by building another hierarchical Bayesian model.
+% We find the underlying $\lambda$ distributions of the observed-$i_\star$ sample and the full sample by building another hierarchical Bayesian model.
 We again use a two-component mixture Beta distribution to describe well-aligned and misaligned systems.
@@
 
-One plausible explanation for the disparity between the two distributions is their small sample sizes.
-In the measured-$i_\star$ sample, only $21$ systems are misaligned  (i.e., $\lambda > 30\degr$), and in the full sample, $51$ systems are misaligned.
-It is likely that the 21 samples could not yet represent the population $\psi$ distribution.
-
-Furthermore, the measured-$i_\star$ sample could be biased towards perpendicular planets.
-Misaligned systems are often found around hot stars.
-However, to measure the inclinations of hot stars, the most prolific inclination measurement method, the rotational modulation method, is no longer applicable to hot stars due to their lack of starspots.
-Instead, the gravity-darkening method is often used.
-The method, however, could be biased since it only detects planets transiting across stellar disks with large brightness variations (i.e., planet crossing the stellar poles or perpendicular planets).
-In the measured-$i_\star$ sample, $6/21$ systems have stellar inclinations constrained via gravity darkening.
-All six planets are misaligned and have nearly polar orbits:
-KELT-9 ($\lambda = 85 \pm 0.2\degr$),
-KELT-17 ($\lambda = 115.9 \pm 4.1 \degr$), 
-Kepler-13 ($\lambda = 59.2 \pm 0.1 \degr$), 
-MASCARA-1 ($\lambda = 69.5 \pm 3.0 \degr$),
-MASCARA-4 ($\lambda = 112.5 \pm 1.6 \degr$), 
-and WASP-189 ($\lambda = 89.3 \pm 1.4 \degr$).
-The bias introduced by the sample is significant enough to produce the clustering in the $\lambda$ distribution of the observed-$i_\star$ sample.
-
-\subsection{Hyperpriors}
-We also test the robustness of $\cos{\psi}$ distribution on different hyperprior choices \citep[see a nice summary in][]{Nagpal22}.
-We use different hyperprior distributions, including uniform distributions, log uniform distributions, and truncated Normal distributions.
-In all hyperprior distributions, the clustering of perpendicular planets in the observed-$i_\star$ sample is found, and such clustering is much weakened in the full sample.
-
-\section{Application to Data \& Discussion}\label{sec:data}
-
-\cite{Albrecht21} inferred the 3D stellar obliquities of a sample of stars that have both sky-projected stellar obliquity and stellar inclination constraints.
-They found while the majority of planets have a small $\psi$, misaligned planets have $\psi$ clustered between 80--125$\degr$.
-Furthermore, the $\cos{\psi}$ distribution of misaligned systems is not uniformly distributed, but clusters near 0, suggesting the stellar spin axis is not isotropically distributed around the planetary orbital axis.
-
-The stellar obliquity distribution of exoplanetary systems can be inferred purely from their sky-projected stellar obliquity distribution.
+% The stellar obliquity distribution of exoplanetary systems can be inferred purely from their sky-projected stellar obliquity distribution.
 When we apply our framework to all systems with $\lambda$ measurements, we find misaligned systems have nearly isotropic stellar obliquity distribution.
@@
 
+\vspace{5mm}
+
+\software{$\mathtt{ArviZ}$ \citep{arviz_2019}, $\mathtt{Jupyter}$ \citep{Jupyter}, $\mathtt{Matplotlib}$ \citep{Matplotlib07, Matplotlib16}, $\mathtt{NumPy}$ \citep{NumPy11, NumPy20}, $\mathtt{pandas}$ \citep{mckinney-proc-scipy-2010, reback2020pandas}, $\mathtt{PyMC}$ \citep{pymc}, $\mathtt{SciPy}$ \citep{2020SciPy-NMeth}}
+
 \bibliography{bib}

### 2023-01-20 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 % Title
-\title{Isotropic Stellar Obliquities of Misaligned Planetary Systems}
+\title{A Hierarchical Bayesian Framework for Exoplanetary Stellar Obliquity Distribution}
 
@@
 We introduce a flexible, hierarchical Bayesian framework that allows inferring the stellar obliquity distribution from sky-projected stellar obliquities, and if available, stellar inclinations.
-Applying the framework to all systems with sky-projected stellar obliquity measurements, which are mostly Hot Jupiter systems, we find that the stellar obliquity distribution is unimodal and peaked at zero degrees.
+Applying the framework to all systems with sky-projected stellar obliquity measurements, mostly Hot Jupiter systems, we find that the stellar obliquity distribution is unimodal and peaked at zero degrees.
 Misaligned systems have nearly isotropic stellar obliquities, at odds with the analysis of systems that also have $i_\star$ measurements.
@@
 Stellar obliquity evolution can be roughly summarized into three stages.
-First, the formation and evolution of a protoplanetary disk around its host star determine the primordial stellar obliquity.
-Second, post-formation dynamical evolution in the planetary system, such as planet-planet scattering or von Zeipel-Kozai-Lidov mechanisms, excites the mutual inclinations between planetary or stellar companions, further sculpting the stellar obliquity.
-Lastly, tidal realignment of the host star's spin axis to the planet's orbital axis could reduce the stellar obliquity, if the stellar tidal dissipation is efficient.
+First, the formation and evolution of a protoplanetary disk around its host star determine the primordial stellar obliquity \citep[e.g.,][]{Bate10, Lai11, Batygin12}.
+Second, post-formation dynamical evolution in the planetary system, such as planet-planet scattering \citep[e.g.,][]{Rasio96, Chatterjee08, Nagasawa08, Beague12} and von Zeipel-Kozai-Lidov mechanisms \citep[e.g.,][]{Wu03, Naoz16}, excites the mutual inclinations between planetary or stellar companions, further sculpting the stellar obliquity.
+Lastly, tidal realignment of the host star's spin axis to the planet's orbital axis could reduce the stellar obliquity, if the stellar tidal dissipation is efficient \citep[e.g.,][]{Winn10, Albrecht12}.
+Furthermore, massive stars with convective cores could generate internal gravity waves and dissipate angular momentum to their radiative zones, thus modifying stellar obliquities \citep{Rogers12, Rogers13}.
 
-While observed aligned systems could have different original stellar obliquities because of the tidal realignment, the excitation of stellar obliquities can only be the outcome of the primordial protoplanetary disk misalignment and post-formation dynamical interactions.
-Recent stellar obliquity measurements on Warm Jupiter systems with large planet-star separations and systems that are still at young ages both find a tendency of system alignment.
-If the trend holds as more observations come up, it could suggest that the primordial stellar obliquity is quite low.
-Stellar obliquities of misaligned systems could then be linked to the dynamical processes that result in the current orbit of close-in planets.
-With that in mind, we aim to build a statistical model to characterize the stellar obliquity distribution of exoplanetary systems in misalignment to be compared with theoretical predictions.
+It is yet unclear which one or more of these mechanisms predominantly sculpts the stellar obliquities in exoplanetary systems and what is the linkage between stellar obliquities and the formation pathways of close-in planets.
+For aligned systems, because of tidal realignment, they could have different stellar obliquities in the past, making it hard to trace the obliquity-sculpting mechanism.
+Misaligned systems with stellar obliquities that have not yet been washed out by tidal effects, on the other hand, still provide clues on obliquity-sculpting mechanisms.
+Proposed mechanisms have different predictions on the stellar obliquity distributions.
+For example, in the secular chaos mechanism, stellar obliquity (or more accurately mutual inclinations between planets) could never get beyond $90\degr$ \citep{Teyssandier19}; in the stellar Kozai mechanism, stellar obliquity distribution should peak at $40\degr$ and $140\degr$ \citep{Anderson16}.
+Motivated by these various predictions, we build a statistical model to characterize the stellar obliquity distribution of exoplanetary systems to distinguish between proposed stellar-obliquity-sculpting mechanisms.
 
-In observation, only the sky-projected stellar obliquity $\lambda$, the angle between the projections of $\bf{n}_{\rm orb}$ and $\bf{n}_{\star}$ onto the plane of the sky, can be measured.
+In exoplanetary observation, only the sky-projected stellar obliquity $\lambda$, the angle between the projections of $\bf{n}_{\rm orb}$ and $\bf{n}_{\star}$ onto the plane of the sky, can be measured.
 Stellar obliquity $\psi$ can be found, if the stellar inclination $i_\star$ is known.
@@
 where $i_{\rm orb}$ is the angle between the vector $\bf{n}_{\rm orb}$ and the observer's line of sight, and $i_\star$ is the angle between $\bf{n}_{\star}$ and the observer's line of sight.
-Since we are discussing transiting exoplanet systems where $i_{\rm orb} \approx 90\degr$, $\cos{\psi} \approx \sin{i_\star}\cos{\lambda}$.
+For transiting exoplanet systems where $i_{\rm orb} \approx 90\degr$, $\cos{\psi} \approx \sin{i_\star}\cos{\lambda}$.
 
-The stellar inclination can often be constrained via, e.g., photometric and spectroscopic rotational modulation introduced by starspots, gravity darkening, and asteroseismology.
-Since the stellar inclination measurement is degenerate between $i_\star$ and $180\degr - i_\star$, $i_\star$ is usually assumed to be less than $90\degr$.
+Stellar inclination can often be constrained via, e.g., photometric and spectroscopic rotational modulation introduced by starspots, gravity darkening, and asteroseismology \citep[see][and references therein]{Albrecht22}.
 For systems without $i_\star$ measurements, it is still possible to infer their stellar obliquities from the sky-projected obliquities, assuming isotropic stellar inclinations.
-The inferred $\psi$ will have a larger uncertainty compared to the one inferred with known $i_\star$ \citep{Fabrycky09}.
+The inferred $\psi$ will have a larger uncertainty than the one inferred with known $i_\star$ \citep{Fabrycky09}.
 
-It is not yet clear how the stellar inclination measurement affects the inference of the stellar obliquity distribution of a population. 
-In this work, we propose an approach to infer $\psi$ distribution from $\lambda$ distribution and show that the stellar obliquity distribution inference only weakly depends on stellar inclination.
-In Section~\ref{sec:jacobian}, we present the Jacobian transformation between the stellar obliquity distribution, the sky-projected stellar obliquity distribution, and the stellar inclination distribution with or without measured $i_\star$.
+We do not yet know the role of stellar inclinations in the population stellar obliquity distribution inference. 
+In this work, we propose an approach to infer the stellar obliquity $\psi$ distribution from the sky-projected stellar obliquity $\lambda$ distribution.
+We show that the stellar obliquity distribution inference weakly depends on stellar inclination.
 In Section~\ref{sec:hbm}, we introduce a flexible, hierarchical Bayesian framework to infer the population stellar obliquity distribution.
 In Section~\ref{sec:applications}, we apply the framework to simulated data and real observations.
-
-\section{Transformation between the $\psi$, $\lambda$, and \lowercase{$i_\star$} distributions}\label{sec:jacobian}
-
-\begin{figure*}[ht!]
-    \script{coordinate.py}
-    \gridline{
-        \fig{figures/coord_psi.pdf}{0.45\textwidth}{\vspace*{-1.8cm}(a) The $\{\psi, \theta\}$ coordinate system.
-The grey circle corresponds to a constant $\psi$ value and its circumference is proportional to $\sin{\psi}$.}
-        \fig{figures/coord_lam.pdf}{0.45\textwidth}{\vspace*{-1.8cm}(b) The $\{\lambda, i_\star\}$ coordinate system.
-The grey circle corresponds to a constant $i_\star$ value and its circumference is proportional to $\sin{i_\star}$.}
-    }
-    \vspace*{-1.5cm}
-    \caption{Two coordinate systems to describe the stellar spin axis $\bf{n}_{\star}$ and the planet's orbital axis $\bf{n}_{\rm orb}$.
-Here we define the observer's line of sight as one of the two horizontal axes (conventional $x$-axis in Cartesian), and the orbital axis of the planet as the vertical axis (conventional $z$-axis in Cartesian).
-We approximate the orbital inclination of the planet to $90\degr$.}
-    \label{fig:coord}
-\end{figure*}
-
-Stellar obliquity $\psi$ is the angle between a planet's orbital axis $\bf{n}_{\rm orb}$ and its host star's spin axis $\bf{n}_{\star}$.
-In Figure~\ref{fig:coord}, we build two coordinate systems to describe the stellar spin axis $\bf{n}_{\star}$ for a given orbital axis $\bf{n}_{\rm orb}$.
-In both coordinates, we set the observer's line of sight to be one of the two horizontal axes, and the orbital axis of the planet to be the vertical axis.
-We approximate the orbital inclination of the transiting planet to $90\degr$ here and in the rest of the work to simplify the problem for a population study.
-If the orbital inclination deviates from $90\degr$, $\bf{n}_{\rm orb}$ instead of being the vertical axis needs to be rotated $\pm (90\degr-i_{\rm orb})$ along the line of sight.
-
-The $\{\psi, \theta\}$ coordinate system shown in panel (a) relates to the physical properties of a planetary system.
-$\psi$ is the angle between the stellar spin axis and the planetary orbital axis and $\theta$ is the azimuthal angle of the stellar spin axis.
-For the given $\bf{n}_{\rm orb}$ axis, if $\bf{n}_{\star}$ is a random vector with uniform distribution on a three-dimensional sphere, i.e., $\bf{n}_{\star}$ is isotropically distributed, the probability density function of $\psi$ follows $p_{\psi} \sim \sin{\psi}$ and $p_\theta \sim 1/2\pi$.
-Since $p_{\cos{\psi}} = p_{\psi} \abs{\dv*{\psi}{\cos{\psi}}}$, $p_{\cos{\psi}}$ is uniformly distributed between $-1$ and $1$ for isotropic stellar obliquity.
-The $\{\lambda, i_\star\}$ coordinate system shown in panel (b) relates to observed properties.
-$\lambda$ is the sky-projected stellar obliquity and $i_\star$ is the stellar inclination.
-If $\bf{n}_{\star}$ is isotropically distributed, $p_{i_\star} \sim \sin{i_\star}$ and $p_\lambda \sim 1/2\pi$.
-Again, since $p_{\cos{i_\star}} = p_{i_\star} \abs{\dv*{i_\star}{\cos{i_\star}}}$, $p_{\cos{i_\star}}$ is uniformly distributed between $-1$ and $1$.
-Because of the observational degeneracy between $i_\star$ and $180\degr - i_\star$, the convention is to have $0 \leq i_\star \leq 90\degr$, and thus $p_{\cos{i_\star}}$ is uniformly distributed between $0$ and $1$ (i.e., $p_{\cos{i_\star}} \sim 1$).
-This also limits $\theta$ to $[-\pi/2, \pi/2]$.
-In observation, there is also a degeneracy between $-\lambda$ and $+\lambda$. Therefore, we also limit $\lambda$ to $[0,\pi]$.
-
-We could find the mathematical relations between $\{\psi, \theta\}$ and $\{\lambda, i_\star\}$ by pairing the Cartesian components of $\bf{n}_{\star}$ in two coordinate systems:
-\begin{align}
-    \sin{\psi}\cos{\theta} = \cos{i_\star}& \label{eq:coord1}\\
-    \sin{\psi}\sin{\theta} = \sin{\lambda}\sin{i_\star}& \label{eq:coord2}\\
-    \cos{\psi} = \cos{\lambda}\sin{i_\star} \label{eq:coord3}&.
-\end{align}
-Thus, $i_\star = \cos[-1](\sin{\psi}\cos{\theta})$ and $\lambda = \tan[-1](\tan{\psi}\sin{\theta})$.
-
-First, we derive the $\lambda$ distribution for a given $\cos{\psi}$ distribution.
-We could find the distribution of $\cos{\lambda}$ using the Jacobian transformation from $\cos{\psi}$ and $\cos{\theta}$.
-Since $\psi$ and $\theta$ are independent variables, we could marginalize over $\theta$ to find the relation between the probability density functions between $\lambda$ and $\psi$.
-The Jacobian transformation follows
-\begin{equation}
-    p_{\cos{\lambda}} = 2 \int_{0}^{1} \abs{\pdv{\cos{\psi}}{\cos{\lambda}}} p_{\cos{\psi}} p_{\cos{\theta}} d\cos{\theta}.
-\end{equation}
-Replacing $\sin{i_\star}$ in Equation~(\ref{eq:coord3}) with Equation~(\ref{eq:coord1}), we find $\cos{\lambda} = \cos{\psi}/\sqrt{1-(1-\cos^2{\psi})\cos^2{\theta}}$.
-Reorganize the equation, we get $\cos^2{\psi} = \frac{\cos^2{\lambda}\cos^2{\theta}-\cos^2{\lambda}}{\cos^2{\lambda}\cos^2{\theta}-1}$.
-The derivative gives $\abs{\pdv{\cos{\psi}}{\cos{\lambda}}} = \frac{(1-\cos^2{\theta})^{1/2}}{(1-\cos^2{\theta}\cos^2{\lambda})^{3/2}}$.
-If $\theta$ is uniformly distributed between $-\pi/2$ and $\pi/2$, $p_{\cos{\theta}} = p_{\theta} \abs{\dv*{\theta}{\cos{\theta}}} = 1/\pi/(1-\cos^2{\theta)^{1/2}}$.
-Putting all the parts together, we get
-\begin{equation}\label{eqn:jac_lam}
-    p_{\cos{\lambda}} = \frac{2}{\pi} \int_{0}^{1} (1-\cos^2{\theta}\cos^2{\lambda})^{-3/2} p_{\cos{\psi}} d\cos{\theta}.
-\end{equation}
-If $\cos{\psi}$ is uniformly distribution, i.e., $p_{\cos{\psi}} = 1/2$, Equation~(\ref{eqn:jac_lam}) becomes $p_{\cos{\lambda}} = 1/\pi/\sqrt{1-\cos^2{\lambda}}$, which is equivalent to $p_\lambda = 1/\pi$.
-The $\lambda$ is uniformly distributed if an isotropic $\psi$.
-If $\cos{\psi}$ follows other distributions, $p_\lambda$ could be found by evaluating Equation~(\ref{eqn:jac_lam}) numerically.
-
-Next, we derive the $i_\star$ distribution for a given $\cos{\psi}$ distribution.
-Similarly, we first find the Jacobian transformation of $i_\star$ from $\psi$ and $\theta$ and then marginalize over $\theta$.
-It is easier to work on $\cos{i_\star}$ than $i_\star$:
-\begin{equation}
-    p_{\cos{i_\star}} = 4 \int_{0}^{1} \abs{\pdv{\sin{\psi}}{\cos{i_\star}}} p_{\sin{\psi}} p_{\cos{\theta}} d\cos{\theta}.
-\end{equation}
-From Equation~(\ref{eq:coord1}), we have $\sin{\psi} = \cos{i_\star}/\cos{\theta}$ and $\abs{\pdv{\sin{\psi}}{\cos{i_\star}}} = 1/\cos{\theta}$.
-Again, we assume $\theta$ is uniformly distributed, and $p_{\cos{\theta}} = 1/\pi/(1-\cos^2{\theta)^{1/2}}$.
-Lastly, we transform the $p_{\sin{\psi}}$ to $p_{\cos{\psi}}$, where $p_{\sin{\psi}} = p_{\cos{\psi}}\sqrt{1-\cos^2{\psi}}/\cos{\psi}$.
-Combining all the pieces together, we get
-\begin{equation}\label{eqn:jac_istar}
-    p_{\cos{i_\star}} = \frac{4}{\pi} \int_{\cos{i_\star}}^{1} \frac{\cos{i_\star}/\cos{\theta}}{\sqrt{\cos^2{\theta}-\cos^2{i_\star}}} \frac{1}{\sqrt{1-\cos^2{\theta}}} p_{\cos{\psi}} d\cos{\theta}.
-\end{equation}
-If $\cos{\psi}$ is uniformly distribution, i.e., $p_{\cos{\psi}} = 1/2$, the integral gives $1$, which suggests the $\cos{i_\star}$ probability density function is uniform.
-
-From Equation~(\ref{eqn:jac_lam}) and (\ref{eqn:jac_istar}), we can now derive the $\lambda$ and $i_\star$ distributions for any given $\psi$ distributions, assuming the azimuthal angle of the stellar spin axis $\theta$ is random.
-In Figure~\ref{fig:transform}, we find and examine numerical solutions of the $\lambda$ and $i_\star$ distributions for four different $\cos{\psi}$ distributions.
-The first row in Figure~\ref{fig:transform} presents an isotropic $\psi$ where $\cos{\psi} \sim \mathcal{U}(-1,1)$, and the second to fourth rows in Figure~\ref{fig:transform} present Normal distributions of $\cos{\psi}$ following $\mathcal{N}(0,0.2)$, $\mathcal{N}(-0.4,0.2)$, and $\mathcal{N}(0.4,0.2)$, respectively.
-In each row, we solve the $\lambda$ and $i_\star$ distributions numerically as shown in blue curves, and samplings of $\lambda$ and $i_\star$ from $\cos{\psi}$ and $\theta$ distributions as shown in grey histograms.
-For a uniform $\cos{\psi}$ distribution, the $\lambda$ distribution is uniform, and the $i_\star$ distribution is isotropic and proportional to $\sin{i_\star}$, as expected.
-
-Interestingly, for a given stellar obliquity distribution, the $\lambda$ distribution is sensitive to and closely related to the underlying $\psi$ distribution, as shown in Figure~\ref{fig:transform}.
-For different stellar obliquity distributions, the $\lambda$ distributions are distinguishable from each other, and thus, when we infer the $\psi$ distribution from the $\lambda$ distribution, the solution will not degenerate.
-On the other hand, the $i_\star$ distributions less depend on the underlying $\psi$ distribution.
-Compared to an isotropic $i_\star$ distribution, the $i_\star$ distributions for different $\psi$ distributions differ at the low values (i.e., $i_\star < \pi/4$), which places a challenge to observation.
-More importantly, since the $i_\star$ distributions are less sensitive to the underlying $\psi$ distributions when we infer the $\psi$ distributions from the $i_\star$ distribution, the solution could be very degenerate.
-
-Due to the lack of dependency on the $i_\star$ distribution, the $\psi$ distribution could be inferred purely from the $\lambda$ distribution without losing much information.
-We note that although we could find a relation between $\psi$, $\lambda$, and $\theta$, the $\psi$ distribution cannot be inferred from the $\lambda$ and $\theta$ distributions since $\lambda$ and $\theta$ are not independent variables.
-We will have to infer the $\psi$ distribution from the $\lambda$ and $i_\star$ distributions.
-In the \nth{4} column in Figure~\ref{fig:transform}, we find the $\psi$ distribution from the $\lambda$ distribution, simply assuming an isotropic $i_\star$ distribution.
-As shown in grey histograms, the method could correctly recover the peaks and widths of the underlying $\psi$ distributions (presented as blue dashed lines).
-The method over-predicts $\cos{\psi}$ at $\pm 1$ because of the over-prediction of $i_\star$ near $90\degr$.
-Later we show that this minor deviation from the true distribution has little impact on the population inference.
-
-\begin{figure*}[ht!]
-    \script{transform.py}
-    \includegraphics[width=\linewidth]{figures/transform.pdf}
-    \caption{Simulated $\cos{\psi}$ distributions (\nth{1} column) and their corresponding sky-projected stellar obliquity $\lambda$ (\nth{2} column) and stellar inclination $i_\star$ (\nth{3} column) distributions.
-The inferred $\cos{\psi}$ distributions assuming isotropic stellar inclinations are shown in the \nth{4} column.
-The random samplings of $\lambda$ and $i_\star$ out of the $\cos{\psi}$ distributions are shown as grey histograms, and the numerical solutions are shown as blue curves.}
-    \label{fig:transform}
-\end{figure*}
+In Section~\ref{sec:jacobian}, we find the mathematical relation between the $\psi$, $\lambda$, and $i_\star$ distributions and explain why stellar inclination is not important in stellar obliquity distribution inference.
 
@@
 \end{align}
-The likelihood functions follows:
+The likelihood functions follow:
 \begin{align}
@@
 Here we assume uniform hyperpriors for the Beta distribution.
-For population inference on a small number of systems, the choice of hyperpriors could have an impact on the inferred distribution \citep{Nagpal22}.
+For population inference on a small number of systems, the choice of hyperpriors could impact the inferred distribution \citep{Nagpal22}.
 When applying the framework to a small sample size, it is important to use different hyperpriors to test the robustness of the model.
@@
 The probabilistic model is built with the $\mathtt{PyMC}$ \citep[$\mathtt{v4.1.7}$;][]{pymc} package and the posteriors are sampled using the No-U-Turn Sampler \citep[NUTS;][]{Hoffman11}, a gradient-based sampling algorithm of the Markov chain Monte Carlo (MCMC).
-All figures and simulations in this paper are built with the $\mathtt{showyourwork}$ package and fully reproducible on \href{https://github.com/jiayindong/polar}{GitHub\,\faGithub}.
+All figures and simulations in this paper are fully reproducible and built with the $\mathtt{showyourwork}$ package.
+The source code can be found on \href{https://github.com/jiayindong/polar}{GitHub\,\faGithub}.
 
@@
 
-We then apply the hierarchical Bayesian framework to all exoplanetary systems that have sky-projected stellar obliquity measurements.
-We use the sample of \numall systems that have sky-projected stellar obliquity $\lambda$ measurements summarized in \cite{Albrecht22} Table A1, which are predominantly Hot Jupiter systems.
+We then apply the hierarchical Bayesian framework to all exoplanetary systems with sky-projected stellar obliquity measurements.
+We use the sample of \numall systems with sky-projected stellar obliquity $\lambda$ measurements summarized in \cite{Albrecht22} Table A1, which are predominantly Hot Jupiter systems.
 The inferred $\cos{\psi}$ is shown in Figure~\ref{fig:psi_dist}.
@@
 The distribution corresponds to an isotropic distribution of $\psi$ from $40\degr$ to $140\degr$.
-The posteriors of the hyperparameters are xxx.
+The posteriors of the hyperparameters are \textcolor{red}{to be added}.
 
@@
 
+\section{Transformation between the $\psi$, $\lambda$, and \lowercase{$i_\star$} distributions}\label{sec:jacobian}
+
+In this section, we aim to understand why the stellar obliquity distribution is predominantly determined by the sky-projected stellar obliquity distribution and why the stellar inclination distribution is less important to the inference.
+
+\begin{figure*}[ht!]
+    \script{coordinate.py}
+    \gridline{
+        \fig{figures/coord_psi.pdf}{0.45\textwidth}{\vspace*{-1.8cm}(a) The $\{\psi, \theta\}$ coordinate system.
+The grey circle corresponds to a constant $\psi$ value and its circumference is proportional to $\sin{\psi}$.}
+        \fig{figures/coord_lam.pdf}{0.45\textwidth}{\vspace*{-1.8cm}(b) The $\{\lambda, i_\star\}$ coordinate system.
+The grey circle corresponds to a constant $i_\star$ value and its circumference is proportional to $\sin{i_\star}$.}
+    }
+    \vspace*{-1.5cm}
+    \caption{Two coordinate systems to describe the stellar spin axis $\bf{n}_{\star}$ and the planet's orbital axis $\bf{n}_{\rm orb}$.
+Here we define the observer's line of sight as one of the two horizontal axes (conventional $x$-axis in Cartesian), and the orbital axis of the planet as the vertical axis (conventional $z$-axis in Cartesian).
+We approximate the orbital inclination of the planet to $90\degr$.}
+    \label{fig:coord}
+\end{figure*}
+
+Stellar obliquity $\psi$ is the angle between a planet's orbital axis $\bf{n}_{\rm orb}$ and its host star's spin axis $\bf{n}_{\star}$.
+In Figure~\ref{fig:coord}, we build two coordinate systems to describe the stellar spin axis $\bf{n}_{\star}$ for a given orbital axis $\bf{n}_{\rm orb}$.
+In both coordinates, we set the observer's line of sight to be one of the two horizontal axes, and the orbital axis of the planet to be the vertical axis.
+We approximate the orbital inclination of the transiting planet to $90\degr$ here to simplify the problem for a population study.
+
+The $\{\psi, \theta\}$ coordinate system shown in panel (a) relates to the physical properties of a planetary system.
+$\psi$ is the angle between the stellar spin axis and the planetary orbital axis, and $\theta$ is the azimuthal angle of the stellar spin axis.
+For the given $\bf{n}_{\rm orb}$ axis, if $\bf{n}_{\star}$ is a random vector with uniform distribution on a three-dimensional sphere, i.e., $\bf{n}_{\star}$ is isotropically distributed, the probability density function of $\psi$ follows $p_{\psi} \sim \sin{\psi}$ and $p_\theta \sim 1/2\pi$.
+Since $p_{\cos{\psi}} = p_{\psi} \abs{\dv*{\psi}{\cos{\psi}}}$, $p_{\cos{\psi}}$ is uniformly distributed between $-1$ and $1$ for isotropic stellar obliquity.
+The $\{\lambda, i_\star\}$ coordinate system shown in panel (b) relates to observed properties.
+$\lambda$ is the sky-projected stellar obliquity and $i_\star$ is the stellar inclination.
+If $\bf{n}_{\star}$ is isotropically distributed, $p_{i_\star} \sim \sin{i_\star}$ and $p_\lambda \sim 1/2\pi$.
+Again, since $p_{\cos{i_\star}} = p_{i_\star} \abs{\dv*{i_\star}{\cos{i_\star}}}$, $p_{\cos{i_\star}}$ is uniformly distributed between $-1$ and $1$.
+Because of the observational degeneracy between $i_\star$ and $180\degr - i_\star$, the convention is to have $0 \leq i_\star \leq 90\degr$, and thus $p_{\cos{i_\star}}$ is uniformly distributed between $0$ and $1$ (i.e., $p_{\cos{i_\star}} \sim 1$).
+This also limits $\theta$ to $[-\pi/2, \pi/2]$.
+In observation, there is also a degeneracy between $-\lambda$ and $+\lambda$. Therefore, we also limit $\lambda$ to $[0,\pi]$.
+
+We could find the mathematical relations between $\{\psi, \theta\}$ and $\{\lambda, i_\star\}$ by pairing the Cartesian components of $\bf{n}_{\star}$ in two coordinate systems:
+\begin{align}
+    \sin{\psi}\cos{\theta} = \cos{i_\star}& \label{eq:coord1}\\
+    \sin{\psi}\sin{\theta} = \sin{\lambda}\sin{i_\star}& \label{eq:coord2}\\
+    \cos{\psi} = \cos{\lambda}\sin{i_\star} \label{eq:coord3}&.
+\end{align}
+Thus, $i_\star = \cos[-1](\sin{\psi}\cos{\theta})$ and $\lambda = \tan[-1](\tan{\psi}\sin{\theta})$.
+
+First, we derive the $\lambda$ distribution for a given $\cos{\psi}$ distribution.
+We could find the distribution of $\cos{\lambda}$ using the Jacobian transformation from $\cos{\psi}$ and $\cos{\theta}$.
+Since $\psi$ and $\theta$ are independent variables, we could marginalize over $\theta$ to find the relation between the probability density functions between $\lambda$ and $\psi$.
+The Jacobian transformation follows
+\begin{equation}
+    p_{\cos{\lambda}} = \int \abs{\pdv{\cos{\psi}}{\cos{\lambda}}} p_{\cos{\psi}} p_{\cos{\theta}} d\cos{\theta}.
+\end{equation}
+Replacing $\sin{i_\star}$ in Equation~(\ref{eq:coord3}) with Equation~(\ref{eq:coord1}), we find $\cos{\lambda} = \cos{\psi}/\sqrt{1-(1-\cos^2{\psi})\cos^2{\theta}}$.
+Reorganize the equation, we get $\cos^2{\psi} = \frac{\cos^2{\lambda}\cos^2{\theta}-\cos^2{\lambda}}{\cos^2{\lambda}\cos^2{\theta}-1}$.
+The derivative gives $\abs{\pdv{\cos{\psi}}{\cos{\lambda}}} = \frac{(1-\cos^2{\theta})^{1/2}}{(1-\cos^2{\theta}\cos^2{\lambda})^{3/2}}$.
+If $\theta$ is uniformly distributed between $-\pi/2$ and $\pi/2$, $p_{\cos{\theta}} = p_{\theta} \abs{\dv*{\theta}{\cos{\theta}}} = 1/\pi/(1-\cos^2{\theta)^{1/2}}$.
+Putting all the parts together, we get
+\begin{equation}\label{eqn:jac_lam}
+    p_{\cos{\lambda}} = \frac{2}{\pi} \int_{0}^{1} (1-\cos^2{\theta}\cos^2{\lambda})^{-3/2} p_{\cos{\psi}} d\cos{\theta}.
+\end{equation}
+If $\cos{\psi}$ is uniformly distribution, i.e., $p_{\cos{\psi}} = 1/2$, Equation~(\ref{eqn:jac_lam}) becomes $p_{\cos{\lambda}} = 1/\pi/\sqrt{1-\cos^2{\lambda}}$, which is equivalent to $p_\lambda = 1/\pi$.
+The $\lambda$ is uniformly distributed if an isotropic $\psi$.
+If $\cos{\psi}$ follows other distributions, $p_\lambda$ could be found by evaluating Equation~(\ref{eqn:jac_lam}) numerically.
+
+Next, we derive the $i_\star$ distribution for a given $\cos{\psi}$ distribution.
+Similarly, we first find the Jacobian transformation of $i_\star$ from $\psi$ and $\theta$ and then marginalize over $\theta$.
+It is easier to work on $\cos{i_\star}$ than $i_\star$:
+\begin{equation}
+    p_{\cos{i_\star}} = \int \abs{\pdv{\sin{\psi}}{\cos{i_\star}}} p_{\sin{\psi}} p_{\cos{\theta}} d\cos{\theta}.
+\end{equation}
+From Equation~(\ref{eq:coord1}), we have $\sin{\psi} = \cos{i_\star}/\cos{\theta}$ and $\abs{\pdv{\sin{\psi}}{\cos{i_\star}}} = 1/\cos{\theta}$.
+Again, we assume $\theta$ is uniformly distributed, and $p_{\cos{\theta}} = 1/\pi/(1-\cos^2{\theta)^{1/2}}$.
+Lastly, we transform the $p_{\sin{\psi}}$ to $p_{\cos{\psi}}$, where $p_{\sin{\psi}} = p_{\cos{\psi}}\sqrt{1-\cos^2{\psi}}/\cos{\psi}$.
+Combining all the pieces together, we get
+\begin{equation}\label{eqn:jac_istar}
+    p_{\cos{i_\star}} = \frac{4}{\pi} \int_{\cos{i_\star}}^{1} \frac{\cos{i_\star}/\cos{\theta}}{\sqrt{\cos^2{\theta}-\cos^2{i_\star}}} \frac{1}{\sqrt{1-\cos^2{\theta}}} p_{\cos{\psi}} d\cos{\theta}.
+\end{equation}
+If $\cos{\psi}$ is uniformly distribution, i.e., $p_{\cos{\psi}} = 1/2$, the integral gives $1$, which suggests the $\cos{i_\star}$ probability density function is uniform.
+
+From Equation~(\ref{eqn:jac_lam}) and (\ref{eqn:jac_istar}), we can now derive the $\lambda$ and $i_\star$ distributions for any given $\psi$ distributions, assuming the azimuthal angle of the stellar spin axis $\theta$ is random.
+In Figure~\ref{fig:transform}, we find and examine numerical solutions of the $\lambda$ and $i_\star$ distributions for four different $\cos{\psi}$ distributions.
+The first row in Figure~\ref{fig:transform} presents an isotropic $\psi$ where $\cos{\psi} \sim \mathcal{U}(-1,1)$, and the second to fourth rows in Figure~\ref{fig:transform} present Normal distributions of $\cos{\psi}$ following $\mathcal{N}(0,0.2)$, $\mathcal{N}(-0.4,0.2)$, and $\mathcal{N}(0.4,0.2)$, respectively.
+In each row, we solve the $\lambda$ and $i_\star$ distributions numerically as shown in blue curves, and samplings of $\lambda$ and $i_\star$ from $\cos{\psi}$ and $\theta$ distributions as shown in grey histograms.
+For a uniform $\cos{\psi}$ distribution, the $\lambda$ distribution is uniform, and the $i_\star$ distribution is isotropic and proportional to $\sin{i_\star}$, as expected.
+
+Interestingly, for a given stellar obliquity distribution, the $\lambda$ distribution is sensitive to and closely related to the underlying $\psi$ distribution, as shown in Figure~\ref{fig:transform}.
+For different stellar obliquity distributions, the $\lambda$ distributions are distinguishable from each other, and thus, when we infer the $\psi$ distribution from the $\lambda$ distribution, the solution will not degenerate.
+On the other hand, the $i_\star$ distributions less depend on the underlying $\psi$ distribution.
+Compared to an isotropic $i_\star$ distribution, the $i_\star$ distributions for different $\psi$ distributions differ at the low values (i.e., $i_\star < \pi/4$), which places a challenge to observation.
+More importantly, since the $i_\star$ distributions are less sensitive to the underlying $\psi$ distributions when we infer the $\psi$ distributions from the $i_\star$ distribution, the solution could be very degenerate.
+
+Due to the lack of dependency on the $i_\star$ distribution, the $\psi$ distribution could be inferred purely from the $\lambda$ distribution without losing much information.
+Although we could find a relation between $\psi$, $\lambda$, and $\theta$, the $\psi$ distribution cannot be inferred from the $\lambda$ and $\theta$ distributions since $\lambda$ and $\theta$ are not independent variables.
+We still need to infer the $\psi$ distribution from the $\lambda$ and $i_\star$ distributions.
+In the \nth{4} column in Figure~\ref{fig:transform}, we find the $\psi$ distribution from the $\lambda$ distribution, simply assuming an isotropic $i_\star$ distribution.
+As shown in grey histograms, the method could correctly recover the peaks and widths of the underlying $\psi$ distributions (presented as blue dashed lines).
+The method over-predicts $\cos{\psi}$ at $\pm 1$ because of the over-prediction of $i_\star$ near $90\degr$.
+However, this minor deviation from the true distribution has little impact on the population inference because of hyperparameters' limited degree of freedom.
+
+\begin{figure*}[ht!]
+    \script{transform.py}
+    \includegraphics[width=\linewidth]{figures/transform.pdf}
+    \caption{Simulated $\cos{\psi}$ distributions (\nth{1} column) and their corresponding sky-projected stellar obliquity $\lambda$ (\nth{2} column) and stellar inclination $i_\star$ (\nth{3} column) distributions.
+The inferred $\cos{\psi}$ distributions assuming isotropic stellar inclinations are shown in the \nth{4} column.
+The random samplings of $\lambda$ and $i_\star$ out of the $\cos{\psi}$ distributions are shown as grey histograms, and the numerical solutions are shown as blue curves.}
+    \label{fig:transform}
+\end{figure*}
+
 \section{Summary \& Discussion}
@@
 We introduce a flexible, hierarchical Bayesian framework to infer the stellar obliquity distribution of a population.
-The model is an open-source code that is available to the public to apply to a certain sample of targets.
-When applying our framework to all exoplanetary systems that have sky-projected stellar obliquity measurements, we found no significant clustering near $\cos{\psi} = 0$, at odds with the previous study of systems that have both sky-projected stellar obliquity and stellar obliquity measurements.
-Potential reasons and biased that lead to this discrepancy is open to discussion.
-We find for misaligned exoplanetary systems, their stellar obliquity is nearly isotropically distributed, which may suggest a convergent disk migration plus a scattering stage in these planets' dynamical history.
-Future stellar obliquity distribution study on a carefully selected Hot Jupiter sample will shed light on the origin of Hot Jupiters.
+The model is an open-source code available to the public to apply to a certain sample of targets.
+When applying our framework to all exoplanetary systems with sky-projected stellar obliquity measurements, we found no significant clustering near $\cos{\psi} = 0$, at odds with the previous study of systems that have both sky-projected stellar obliquity and stellar obliquity measurements.
+Potential biases that lead to this discrepancy are open to discussion.
+Their stellar obliquity is nearly isotropically distributed for misaligned exoplanetary systems, which may suggest a multi-planet, convergent-disk-migration history followed by a scattering phase in these planets' dynamical history.
+Future studies on a more carefully selected Hot Jupiter sample will put better constraints on the origin of Hot Jupiters from their stellar obliquity distribution.
 

### 2023-01-20 src/scripts/graph.py

@@ -20,33 +20,36 @@ rc('text.latex', preamble=r'\usepackage{physics}')
 pgm = daft.PGM(observed_style="shaded",node_unit=1, node_ec='k',dpi=150,line_width=0.8)
                
 # Hierarchical parameters.
-pgm.add_node("beta", r"$\vb*{\beta}$", 2, 2.3)
+pgm.add_node("beta", r"$\vb*{\beta}$", 2, 2.8)
 
 # Latent variable.
-pgm.add_node("psi", r"$\psi_n$", 2, 1.5)
+pgm.add_node("psi", r"$\psi_n$", 2, 2.)
 
-pgm.add_node("lambda", r"$\lambda_n$", 1., 1.5)
+pgm.add_node("lambda", r"$\lambda_n$", 1., 1.2)
 
-pgm.add_node("istar", r"$i_{\star,n}$", 3., 1.5)
+pgm.add_node("iorb", r"$i_{{\rm orb},n}$", 2., 1.2)
 
-pgm.add_node("pstar", r"$\gamma_{\star,n}$", 4., 1.5)
+pgm.add_node("istar", r"$i_{\star,n}$", 3., 1.2)
 
+pgm.add_node("pstar", r"$\gamma_{\star,n}$", 4., 1.2)
 
 pgm.add_edge("beta", "psi")
 
 pgm.add_edge("lambda", "psi")
 
+pgm.add_edge("iorb", "psi")
+
 pgm.add_edge("istar", "psi")
 
+pgm.add_node("obs_lambda", r"$\hat{\lambda}_n$", 1., 0.4, observed=True)
 
-pgm.add_node("obs_lambda", r"$\hat{\lambda}_n$", 1., 0.7, observed=True)
+pgm.add_node("obs_iorb", r"$\hat{i}_{{\rm orb}, n}$", 2., 0.4, observed=True)
 
-pgm.add_node("Obs", r"Obs$_{\star, n}$", 3., 0.7, observed=True, aspect=1., fontsize=8)
-
-# pgm.add_node("obs_pstar", r"$\hat{\theta}_{\star,n}$", 4., 0.7, observed=True)
-
+pgm.add_node("Obs", r"Obs$_{\star, n}$", 3., 0.4, observed=True, aspect=1., fontsize=8)
 
 pgm.add_edge("lambda", "obs_lambda")
+
+pgm.add_edge("iorb", "obs_iorb")
 
[…]

### 2023-01-20 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \usepackage{hyperref}
-
-\newcommand{\numistar}{62\xspace}
-\newcommand{\numall}{161\xspace}
 
@@
 Stellar obliquity $\psi$ describes the angle between a planet's orbital axis $\bf{n}_{\rm orb}$ and its host star's spin axis $\bf{n}_{\star}$. 
-The angle is an important tracer of the formation environment and dynamical evolution of a planetary system.
+The angle is an important tracer of a planetary system's formation environment and dynamical evolution.
 Stellar obliquity evolution can be roughly summarized into three stages.
@@
 
-It is yet unclear which one or more of these mechanisms predominantly sculpts the stellar obliquities in exoplanetary systems and what is the linkage between stellar obliquities and the formation pathways of close-in planets.
-For aligned systems, because of tidal realignment, they could have different stellar obliquities in the past, making it hard to trace the obliquity-sculpting mechanism.
-Misaligned systems with stellar obliquities that have not yet been washed out by tidal effects, on the other hand, still provide clues on obliquity-sculpting mechanisms.
+It is yet unclear which one or more of these mechanisms predominantly sculpts the stellar obliquities in exoplanetary systems and the linkage between stellar obliquities and the formation pathways of close-in planets.
+Because of tidal realignment, aligned systems could have had different stellar obliquities in the past, making it hard to trace the obliquity-sculpting mechanism.
+On the other hand, misaligned systems with stellar obliquities that have not yet been washed out by tidal effects still provide clues on obliquity-sculpting mechanisms.
 Proposed mechanisms have different predictions on the stellar obliquity distributions.
 For example, in the secular chaos mechanism, stellar obliquity (or more accurately mutual inclinations between planets) could never get beyond $90\degr$ \citep{Teyssandier19}; in the stellar Kozai mechanism, stellar obliquity distribution should peak at $40\degr$ and $140\degr$ \citep{Anderson16}.
-Motivated by these various predictions, we build a statistical model to characterize the stellar obliquity distribution of exoplanetary systems to distinguish between proposed stellar-obliquity-sculpting mechanisms.
+Motivated by these predictions, we build a statistical model to characterize the stellar obliquity distribution of exoplanetary systems to distinguish between proposed stellar-obliquity-sculpting mechanisms.
 
@@
 
-To find the unknown stellar obliquity distribution of exoplanetary systems, we build a hierarchical Bayesian framework that takes sky-projected stellar obliquities as observed data.
-In Section~\ref{sec:jacobian}, we found that the stellar inclination is not sensitive to the stellar obliquity distribution.
-Therefore, we focus on a flexible framework that makes the stellar inclination information optional.
+To find the unknown stellar obliquity distribution of exoplanetary systems, we build a hierarchical Bayesian framework that takes the sky-projected stellar obliquity $\lambda$ and orbital inclination $i_{\rm orb}$ as observed data.
+Stellar inclination $i_\star$ is an optional and flexible input.
+It could be directly taken from previous observations, or if the rotational modulation technique is used, the stellar inclination will be inferred from the stellar rotation period $P_{\rm rot}$, stellar radius $R_\star$, and sky-projected rotational broadening velocity $v\sin{i}_\star$.
+If the stellar inclination information is not provided, we simply assume an isotropic stellar inclination.
+As we will show in later sections, assuming an isotropic stellar inclination will not compromise the stellar obliquity distribution inference.
 
@@
 The stellar obliquity distribution is described by hyperparameters $\vb*{\beta}$ and constrained by $n_{\rm th}$ system's stellar obliquity $\psi_n$.
-Each $\psi_n$ is calculated by the system's sky-projected stellar obliquity $\lambda_n$ and stellar inclination $i_{\star, n}$, if available.
+Each $\psi_n$ is calculated by the system's sky-projected stellar obliquity $\lambda_n$, orbital inclination $i_{{\rm orb}, n}$, and stellar inclination $i_{\star, n}$, if available.
 $\gamma_{\star, n}$ contains properties of the star other than inclination, such as stellar radius and rotation period for $i_{\star, n}$ inference.
-Obs$_{\star,n}$ includes all the observed properties of the star with uncertainties, and $\hat{\lambda}_n$ has the measured sky-projected stellar obliquity and its uncertainty.}
+Obs$_{\star,n}$ includes all the observed properties of the star with uncertainties, $\hat{\lambda}_n$ has the measured sky-projected stellar obliquity and its uncertainty, and $\hat{i}_{{\rm orb}, n}$ has measured orbital inclination and its uncertainty.}
     \label{fig:graph}
@@
 The parameter set $\bm{\beta}$ is constrained by $N$ individual systems in which $\psi_n$ is a random variable.
-$\psi_n$ is computed from the sky-projected stellar obliquity $\lambda_n$ and the stellar inclination $i_n$.
-For simplicity, we approximate $i_{\rm orb}$ to $90\degr$ and use the simplified version of Equation~(\ref{eqn:psi}) to compute $\psi_n$.
-$\psi$ could be computed more accurately by including $i_{\rm orb}$ as another random variable to be constrained by e.g., transit observations.
-The parameter $\gamma_{\star, n}$ is another parameter set that contains all the stellar properties other than $i_\star$, e.g., stellar rotation period $P_{\rm rot}$ and radius $R_\star$, if the information is available.
+$\psi_n$ is computed from the sky-projected stellar obliquity $\lambda_n$, the orbital inclination $i_{{\rm orb},n}$, and the stellar inclination $i_{\star, n}$.
+The parameter $\gamma_{\star, n}$ is a parameter set that contains all the stellar properties other than $i_\star$, e.g., stellar rotation period $P_{\rm rot}$ and radius $R_\star$, if the information is available.
 Obs$_{\star,n}$ includes observed data that helps to constrain $i_{\star,n}$ and $\gamma_{\star, n}$.
@@
 If asteroseismology is used, the $i_{\star,n}$ is constrained from the periodic variation in the photometric time series.
-Lastly, the sky-projected stellar obliquity $\lambda_n$ is constrained by observed $\hat{\lambda}_n$ via the Rossiter-McLaughlin effect or gravity darkening.
+The sky-projected stellar obliquity $\lambda_n$ can be constrained by observed $\hat{\lambda}_n$ via the Rossiter-McLaughlin effect or gravity darkening.
+Lastly, the orbital inclination $i_{{\rm orb},n}$ can be constrained from the transit light curves.
 
-For the stellar obliquity distribution, we model the $\cos{\psi}$ distribution instead of $\psi$ distribution to aim to understand if the stellar obliquity is isotropically distributed or not.
+For the stellar obliquity distribution, we model the $\cos{\psi}$ distribution instead of $\psi$ distribution to understand if the stellar obliquity is isotropically distributed.
+If stellar obliquities of misaligned systems follow an isotropic distribution, the $\cos{\psi}$ distribution will be uniformly distributed between -1 to 1.
 We adopt a two-component mixture Beta distribution with hyperparameters $\bm{\beta} = \{\bm{w},\bm{a},\bm{b}\}$, where each hyperparameter has a dimension of 2.
-One component is designed for a large population of well-aligned systems, and the other component is designed for misaligned systems.
-The Beta distribution is a flexible distribution that is able to describe the spike of aligned systems with stellar obliquities close to $0\degr$ and also the clustered or broadly distributed misaligned systems.
+One component is designed for a large population of well-aligned systems, and the other is designed for misaligned systems.
+The Beta distribution is a flexible distribution that can describe the spike of aligned systems with stellar obliquities close to $0\degr$ or the clustered or broadly distributed misaligned systems.
 The probability density function of $\cos{\psi}$ follows 
@@
     \cos{i}_{\star,n} &\sim \mathcal{U}(0, 1) \nonumber\\
+    \cos{i}_{{\rm orb},n} &\sim \mathcal{U}(-1, 1) \nonumber\\
     \lambda_{\star,n} &\sim \mathcal{U}(0, \pi).
@@
 Here we assume uniform hyperpriors for the Beta distribution.
-For population inference on a small number of systems, the choice of hyperpriors could impact the inferred distribution \citep{Nagpal22}.
-When applying the framework to a small sample size, it is important to use different hyperpriors to test the robustness of the model.
+For population inference on a small number of systems (i.e., $N<50$), the choice of hyperpriors could impact the inferred distribution \citep{Nagpal22}.
+When applying the framework to small sample size, it is important to test the robustness of inferred distributions on different hyperpriors.
 
@@
 All figures and simulations in this paper are fully reproducible and built with the $\mathtt{showyourwork}$ package.
-The source code can be found on \href{https://github.com/jiayindong/polar}{GitHub\,\faGithub}.
+The open source code can be found on \href{https://github.com/jiayindong/polar}{GitHub\,\faGithub}.
 
@@
 We first apply the hierarchical Bayesian framework to simulated data of which the ground-truth stellar obliquity distribution is known.
-We adopt the four $\cos{\psi}$ distributions discussed in Section~\ref{sec:jacobian}: a uniform distribution $\mathcal{U}(-1,1)$ and three Normal distributions $\mathcal{N}(0,0.2)$, $\mathcal{N}(-0.4,0.2)$, and $\mathcal{N}(0.4,0.2)$.
-For each $\cos{\psi}$ distribution, we randomly generate 200 samples of stellar inclination $i_\star$ and sky-projected stellar obliquity $\lambda$, assuming a uniform $\theta$.
+We test the four $\cos{\psi}$ distributions: a uniform distribution $\mathcal{U}(-1,1)$ and three Normal distributions $\mathcal{N}(0,0.2)$, $\mathcal{N}(-0.4,0.2)$, and $\mathcal{N}(0.4,0.2)$.
+For each $\cos{\psi}$ distribution, we randomly generate 200 samples of stellar inclination $i_\star$ and sky-projected stellar obliquity $\lambda$.
+We assume the stellar spin axis is uniformly distributed around the planetary orbital axis in the azimuthal direction and the orbital inclination is $90\degr$.
 The sampled $i_\star$ and $\lambda$ here are \emph{true} values.
@@
 We simply assume an isotropic distribution of $i_\star$ for these models.
-As shown in Figure~\ref{fig:simulation}, in all four distributions, the inferred $\cos{\psi}$ distributions with or without $i_\star$ correctly recover the injected distributions shown as grey dashed lines.
+Surprisingly, in all four distributions, the inferred $\cos{\psi}$ distributions with or without $i_\star$ correctly recover the injected distributions shown as grey dashed lines as shown in Figure~\ref{fig:simulation}.
 Since some of the injected distributions are Normal distributions, it is expected that the inferred distributions, which as Beta distributions, could not exactly describe the injected distributions.
-Comparing the blue curves with the orange curves, the overestimation of $\cos{\psi}$ near $\pm1$ under the assumption of an isotropic $i_\star$ has little impact on the inference.
-The distribution is mainly constrained by the peak and width of the stellar obliquity distribution, which are both correctly predicted by the isotropic $i_\star$ model.
 
-From the simulated data, we confirm that while stellar inclinations could help to constrain the stellar obliquity of individual systems, they are not necessarily required for the population stellar obliquity inference.
-The constraint on stellar obliquity distribution is predominantly based on sky-projected stellar obliquities.
-The sky-projected stellar obliquity could correctly infer the peak and width of the underlying stellar obliquity distribution.
+We also examine the importance of orbital inclination to the stellar obliquity inference.
+Since all the systems we study are transiting exoplanetary systems, we consider an isotropic orbital inclination distribution from $80\degr$ to $90\degr$.
+This broad inclination range corresponds to an impact parameter range from 0 to 1 for a circular orbit planet with an $a/R_\star$ of 6.
+We find at this level of orbital inclination deviations, the difference on the inferred stellar obliquity distributions is minimal.
+For transiting planet systems, approximating orbital inclinations to $90\degr$ will not compromise the stellar obliquity distribution inference.
+
+From the simulated data, we confirm that while stellar inclinations help to constrain the stellar obliquity of individual systems, they are not necessarily required for the population stellar obliquity inference.
+The constraint on stellar obliquity distribution is predominantly based on sky-projected stellar obliquities, and the sky-projected stellar obliquity could correctly infer the peak and width of the underlying stellar obliquity distribution.
 
@@
 
-We then apply the hierarchical Bayesian framework to all exoplanetary systems with sky-projected stellar obliquity measurements.
-We use the sample of \numall systems with sky-projected stellar obliquity $\lambda$ measurements summarized in \cite{Albrecht22} Table A1, which are predominantly Hot Jupiter systems.
-The inferred $\cos{\psi}$ is shown in Figure~\ref{fig:psi_dist}.
+We then apply the hierarchical Bayesian framework to all exoplanetary systems with sky-projected stellar obliquity $\lambda$ measurements.
+We use the sample of 161 systems summarized in \cite{Albrecht22} Table A1, which are predominantly Hot Jupiter systems.
+The inferred $\cos{\psi}$ distribution is shown in Figure~\ref{fig:psi_dist}.
 The $\cos{\psi}$ distribution peaks at 1 but is nearly flat between $-0.75$ and $0.75$ with no significant clustering.
 The distribution corresponds to an isotropic distribution of $\psi$ from $40\degr$ to $140\degr$.
-The posteriors of the hyperparameters are \textcolor{red}{to be added}.
+The aligned-system population dominates the distribution with a fraction of $w_0 = 0.67 \pm 0.09$. The hyperposteriors of the Beta distribution are $a_0 = 31.0\pm12.8$ and $b_0 = 0.40\pm0.12$.
+The misaligned-system population has a fraction of $w_1 = 0.33 \pm 0.09$ and hyperposteriors of $a_1 = 1.56\pm0.93$ and $b_1 = 1.64\pm1.58$.
 
@@
 The subsample identified a preponderance of perpendicular planets that disfavored an isotropic stellar obliquity distribution \citep{Albrecht21}.
-We summarize two possible reasons to explain the discrepancies, which are open to further investigation.
+We speculate two possible explanations for this observation discrepancy, which are open to further investigation.
 First, the sample of systems with $i_\star$ measurements is small, and only about 20 systems are misaligned systems used to constrain the misaligned systems' stellar obliquity distribution.
-Second, the sample could be biased because of the $i_\star$ measurements.
+Second, the sample could have a selection bias introduced by the $i_\star$ measurement requirement.
 
 Our analysis suggests misaligned systems have nearly isotropic stellar obliquity distribution.
-When comparing the inferred distribution with the prediction of stellar obliquity distributions in different Hot Jupiter origin scenarios, it is in best agreement with the outcome of the scattering of multiple giant planets after a convergent disk migration \citep{Beague12}.
+When comparing the inferred distribution with the prediction of stellar obliquity distributions in different Hot Jupiter origin scenarios, it is in best agreement with the outcome of the scattering of multiple giant planets after a convergent disk migration \citep[e.g.,][]{Beague12}, which could produce planets with orbital inclination from $0\degr$ to $180\degr$ with a peak at $0\degr$.
+This interesting result can be further tested on more carefully selected Hot Jupiter samples and put constraints on Hot Jupiter origins.
 
-\section{Transformation between the $\psi$, $\lambda$, and \lowercase{$i_\star$} distributions}\label{sec:jacobian}
+\section{Relations between the $\psi$, $\lambda$, and \lowercase{$i_\star$} distributions}\label{sec:jacobian}
 
-In this section, we aim to understand why the stellar obliquity distribution is predominantly determined by the sky-projected stellar obliquity distribution and why the stellar inclination distribution is less important to the inference.
+In this section, we aim to use mathematical expressions to understand why the stellar obliquity distribution is predominantly determined by the sky-projected stellar obliquity distribution and why the stellar inclination distribution is less important to the inference.
 
@@
 
-Stellar obliquity $\psi$ is the angle between a planet's orbital axis $\bf{n}_{\rm orb}$ and its host star's spin axis $\bf{n}_{\star}$.
 In Figure~\ref{fig:coord}, we build two coordinate systems to describe the stellar spin axis $\bf{n}_{\star}$ for a given orbital axis $\bf{n}_{\rm orb}$.
@@
 We approximate the orbital inclination of the transiting planet to $90\degr$ here to simplify the problem for a population study.
+We found in the previous section that the assumption would not compromise the stellar obliquity distribution inference for transiting planets.
 
@@
 In the \nth{4} column in Figure~\ref{fig:transform}, we find the $\psi$ distribution from the $\lambda$ distribution, simply assuming an isotropic $i_\star$ distribution.
-As shown in grey histograms, the method could correctly recover the peaks and widths of the underlying $\psi$ distributions (presented as blue dashed lines).
+The method could correctly recover the peaks and widths of the underlying $\psi$ distributions (presented as blue dashed lines) as shown in grey histograms.
 The method over-predicts $\cos{\psi}$ at $\pm 1$ because of the over-prediction of $i_\star$ near $90\degr$.
 However, this minor deviation from the true distribution has little impact on the population inference because of hyperparameters' limited degree of freedom.
+Comparing the blue curves with the orange curves in Figure~\ref{fig:simulation}, the overestimation of $\cos{\psi}$ near $\pm1$ under the assumption of an isotropic $i_\star$ has little impact on the inference since the distribution is mainly constrained by the peak and width of the stellar obliquity distribution, which are both correctly predicted by the isotropic $i_\star$ model.
 
@@
 We introduce a flexible, hierarchical Bayesian framework to infer the stellar obliquity distribution of a population.
-The model is an open-source code available to the public to apply to a certain sample of targets.
+Information on stellar inclination is not required for the stellar obliquity distribution inference; if the stellar inclination is not provided, we simply assume an isotropic stellar inclination distribution.
+When combining samples with and without stellar inclination measurements, it is important to check if the $i_\star$ sample is representative of the full sample.
+If not, since the $i_\star$ sample will constrain stellar obliquities more tightly, the overall distribution could be weighted towards these systems, bias our interpretation.
+The hierarchical Bayesian model we introduce in this paper is an open-source code available to the public to apply to a certain sample of targets and customize for different stellar obliquity distributions and priors.
+
 When applying our framework to all exoplanetary systems with sky-projected stellar obliquity measurements, we found no significant clustering near $\cos{\psi} = 0$, at odds with the previous study of systems that have both sky-projected stellar obliquity and stellar obliquity measurements.
 Potential biases that lead to this discrepancy are open to discussion.
-Their stellar obliquity is nearly isotropically distributed for misaligned exoplanetary systems, which may suggest a multi-planet, convergent-disk-migration history followed by a scattering phase in these planets' dynamical history.
+Their stellar obliquity is nearly isotropically distributed for misaligned exoplanetary systems in support of a multi-planet, convergent-disk-migration history followed by a scattering phase in these planets' dynamical history.
 Future studies on a more carefully selected Hot Jupiter sample will put better constraints on the origin of Hot Jupiters from their stellar obliquity distribution.

### 2023-01-31 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 The $\cos{\psi}$ distribution peaks at 1 but is nearly flat between $-0.75$ and $0.75$ with no significant clustering.
-The distribution corresponds to an isotropic distribution of $\psi$ from $40\degr$ to $140\degr$.
+The distribution corresponds to a pile-up of planets with stellar obliquity less than $40\degr$ and an isotropic distribution of planets with stellar obliquity from $40\degr$ to $140\degr$.
 The aligned-system population dominates the distribution with a fraction of $w_0 = 0.67 \pm 0.09$. The hyperposteriors of the Beta distribution are $a_0 = 31.0\pm12.8$ and $b_0 = 0.40\pm0.12$.
@@
 
-Our analysis suggests misaligned systems have nearly isotropic stellar obliquity distribution.
-When comparing the inferred distribution with the prediction of stellar obliquity distributions in different Hot Jupiter origin scenarios, it is in best agreement with the outcome of the scattering of multiple giant planets after a convergent disk migration \citep[e.g.,][]{Beague12}, which could produce planets with orbital inclination from $0\degr$ to $180\degr$ with a peak at $0\degr$.
-This interesting result can be further tested on more carefully selected Hot Jupiter samples and put constraints on Hot Jupiter origins.
+Our model identifies a pile-up of systems with stellar obliquity less than $40\degr$ and finds that misaligned systems have nearly isotropic stellar obliquity distribution.
+When comparing the inferred distribution with the prediction of stellar obliquity distributions in different Hot Jupiter origin scenarios, it is in best agreement with the outcome of the scattering of multiple giant planets after a convergent disk migration \citep[e.g.,][]{Nagasawa11, Beague12}.
+This particular intriguing result should be further examined on more carefully selected Hot Jupiter samples and put constraints on Hot Jupiter origin channels.
 

### 2023-02-08 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
+\received{}
+\revised{}
+\accepted{}
+
+\submitjournal{the AAS Journals}
+
 % Abstract with filler text
@@
 
-Stellar obliquity $\psi$, the angle between a planet's orbital axis and its host star's spin axis, traces the formation environment and dynamical evolution of a planetary system.
-In exoplanet observations, only the sky-projected stellar obliquity $\lambda$ can be measured.
-To find the true stellar obliquity, information about the stellar inclination $i_\star$ is required.
-Here we show that, while it is important to know the stellar inclination to tightly constrain the stellar obliquity of an individual system, the stellar inclination is not necessarily required for the population stellar obliquity inference.
-The constraint on stellar obliquity distribution is predominantly based on sky-projected stellar obliquities.
-We introduce a flexible, hierarchical Bayesian framework that allows inferring the stellar obliquity distribution from sky-projected stellar obliquities, and if available, stellar inclinations.
-Applying the framework to all systems with sky-projected stellar obliquity measurements, mostly Hot Jupiter systems, we find that the stellar obliquity distribution is unimodal and peaked at zero degrees.
-Misaligned systems have nearly isotropic stellar obliquities, at odds with the analysis of systems that also have $i_\star$ measurements.
+Stellar obliquity $\psi$, the angle between a planet's orbital axis and its host star's spin axis, traces the formation and evolution of a planetary system.
+In transiting-exoplanet observations, only the sky-projected stellar obliquity $\lambda$ can be measured.
+To find the true stellar obliquity $\psi$, information about stellar inclination $i_\star$ is also required.
+Here we show that, while the stellar inclination is important to infer the stellar obliquity of an individual system, they are not necessarily required for the population stellar obliquity inference. 
+Constraints on the stellar obliquity distribution are predominantly contributed by the sky-projected stellar obliquities.
+We introduce a flexible, hierarchical Bayesian framework that could infer the stellar obliquity distribution purely from sky-projected stellar obliquities.
+Stellar inclination measurements are optional in the model, and if not provided, are assumed to be isotropically distributed.
+Applying the framework to all exoplanetary systems with sky-projected stellar obliquity measurements, we find that the stellar obliquity distribution is unimodal and peaked at zero degrees.
+Misaligned systems have nearly isotropic stellar obliquities, which may have important implications for the formation and evolution of planetary systems.
 
@@
 The angle is an important tracer of a planetary system's formation environment and dynamical evolution.
-Stellar obliquity evolution can be roughly summarized into three stages.
+The evolution of stellar obliquity can be roughly summarized into three stages.
 First, the formation and evolution of a protoplanetary disk around its host star determine the primordial stellar obliquity \citep[e.g.,][]{Bate10, Lai11, Batygin12}.
-Second, post-formation dynamical evolution in the planetary system, such as planet-planet scattering \citep[e.g.,][]{Rasio96, Chatterjee08, Nagasawa08, Beague12} and von Zeipel-Kozai-Lidov mechanisms \citep[e.g.,][]{Wu03, Naoz16}, excites the mutual inclinations between planetary or stellar companions, further sculpting the stellar obliquity.
-Lastly, tidal realignment of the host star's spin axis to the planet's orbital axis could reduce the stellar obliquity, if the stellar tidal dissipation is efficient \citep[e.g.,][]{Winn10, Albrecht12}.
-Furthermore, massive stars with convective cores could generate internal gravity waves and dissipate angular momentum to their radiative zones, thus modifying stellar obliquities \citep{Rogers12, Rogers13}.
+Second, post-formation dynamical evolution in the planetary system, such as planet-planet scattering \citep[e.g.,][]{Rasio96, Chatterjee08, Nagasawa08, Beague12}, von Zeipel-Kozai-Lidov mechanisms \citep[e.g.,][]{Wu03, Naoz16}, and secular chaos \citep{Wu11}, excites the mutual inclinations between planetary or stellar companions, further sculpting the stellar obliquity.
+Lastly, tidal realignment of the host star's spin axis to the planet's orbital axis could reduce the stellar obliquity, if the tidal dissipation in the star is efficient \citep[e.g.,][]{Winn10, Albrecht12}.
+Furthermore, massive stars with convective cores could generate internal gravity waves and dissipate angular momentum to their radiative zones, modifying the stellar obliquity \citep{Rogers12, Rogers13}.
 
-It is yet unclear which one or more of these mechanisms predominantly sculpts the stellar obliquities in exoplanetary systems and the linkage between stellar obliquities and the formation pathways of close-in planets.
-Because of tidal realignment, aligned systems could have had different stellar obliquities in the past, making it hard to trace the obliquity-sculpting mechanism.
-On the other hand, misaligned systems with stellar obliquities that have not yet been washed out by tidal effects still provide clues on obliquity-sculpting mechanisms.
-Proposed mechanisms have different predictions on the stellar obliquity distributions.
-For example, in the secular chaos mechanism, stellar obliquity (or more accurately mutual inclinations between planets) could never get beyond $90\degr$ \citep{Teyssandier19}; in the stellar Kozai mechanism, stellar obliquity distribution should peak at $40\degr$ and $140\degr$ \citep{Anderson16}.
-Motivated by these predictions, we build a statistical model to characterize the stellar obliquity distribution of exoplanetary systems to distinguish between proposed stellar-obliquity-sculpting mechanisms.
+It is yet unclear if all of these physical and dynamic processes apply to exoplanetary systems.
+These proposed mechanisms have different predictions on stellar obliquity distributions \citep[see][and references therein]{Albrecht22}.
+For example, in the secular chaos mechanism, the stellar obliquity could never get beyond $90\degr$ \citep{Teyssandier19}.
+In the stellar Kozai mechanism, the stellar obliquity distribution should be bimodal and concentrated at $40\degr$ and $140\degr$ \citep[e.g.,][]{Fabrycky07, Anderson16, Vick19}.
+In the multiple-planet scattering mechanism, the majority of systems should be aligned, and a small fraction of systems have a wide range of stellar obliquities \citep[e.g.,][]{Beague12}.
+Motivated by these predictions, we aim to constrain the stellar obliquity distribution of exoplanetary systems from a Bayesian approach to find dominant mechanisms that sculpt close-in planetary systems.
 
 In exoplanetary observation, only the sky-projected stellar obliquity $\lambda$, the angle between the projections of $\bf{n}_{\rm orb}$ and $\bf{n}_{\star}$ onto the plane of the sky, can be measured.
-Stellar obliquity $\psi$ can be found, if the stellar inclination $i_\star$ is known.
-The calculation follows \citep[e.g.,][]{Fabrycky09}:
+Stellar obliquity $\psi$ can be calculated, if the stellar inclination $i_\star$ is known.
+The relation follows \citep[e.g.,][]{Fabrycky09}:
 \begin{equation}\label{eqn:psi}
@@
 For systems without $i_\star$ measurements, it is still possible to infer their stellar obliquities from the sky-projected obliquities, assuming isotropic stellar inclinations.
-The inferred $\psi$ will have a larger uncertainty than the one inferred with known $i_\star$ \citep{Fabrycky09}.
+The inferred $\psi$ will have greater uncertainty than the one inferred with $i_\star$ measurement \citep{Fabrycky09}.
 
-We do not yet know the role of stellar inclinations in the population stellar obliquity distribution inference. 
+We do not yet understand the importance of stellar inclination in population stellar obliquity inference.
 In this work, we propose an approach to infer the stellar obliquity $\psi$ distribution from the sky-projected stellar obliquity $\lambda$ distribution.
-We show that the stellar obliquity distribution inference weakly depends on stellar inclination.
+We show that the stellar obliquity distribution inference weakly depends on stellar inclinations.
 In Section~\ref{sec:hbm}, we introduce a flexible, hierarchical Bayesian framework to infer the population stellar obliquity distribution.
@@
 For population inference on a small number of systems (i.e., $N<50$), the choice of hyperpriors could impact the inferred distribution \citep{Nagpal22}.
-When applying the framework to small sample size, it is important to test the robustness of inferred distributions on different hyperpriors.
+When applying the framework to a small sample size, it is important to test the robustness of inferred distributions on different hyperpriors.
 
@@
 From the simulated data, we confirm that while stellar inclinations help to constrain the stellar obliquity of individual systems, they are not necessarily required for the population stellar obliquity inference.
-The constraint on stellar obliquity distribution is predominantly based on sky-projected stellar obliquities, and the sky-projected stellar obliquity could correctly infer the peak and width of the underlying stellar obliquity distribution.
+The constraint on stellar obliquity distribution is predominantly based on sky-projected stellar obliquities.
+The sky-projected stellar obliquity could correctly infer the peak and width of the underlying stellar obliquity distribution.
 

### 2023-02-08 src/scripts/psi_dist.py
add axis labels

@@ -158,8 +158,8 @@ if __name__ == '__main__':
     plt.ylim([0,2])
     plt.xlim([-1,1])
 
-    # plt.xlabel(r'$\cos{\psi}$',fontsize=11)
-    # plt.ylabel('Probablity density function')
+    plt.xlabel(r'$\cos{\psi}$',fontsize=11)
+    plt.ylabel('Probablity density function')
 
     plt.tight_layout()
     plt.savefig(paths.figures / "psi_dist.pdf", bbox_inches="tight", dpi=600)

### 2023-02-08 src/scripts/simulation.py

@@ -44,7 +44,7 @@ def posteriors(this_model):
     err_lam = 8*np.pi/180
 
     with this_model:    
-        idata = pm.sample(draws=50)
+        idata = pm.sample(chains=4, draws=50)
 
     true_istar = idata.posterior.i.values.ravel()
     obs_istar = true_istar + err_istar*np.random.normal(size=nsample)
@@ -78,7 +78,7 @@ def posteriors(this_model):
 
         hyper = pm.Potential("hyper", pm.logp(pm.Beta.dist(a,b), (cosψ+1)/2))
 
-        istar_idata = pm.sample()
+        istar_idata = pm.sample(target_accept=0.9,chains=4)
 
     with pm.Model() as model_noistar:
 
@@ -104,7 +104,7 @@ def posteriors(this_model):
 
         hyper = pm.Potential("hyper", pm.logp(pm.Beta.dist(a,b), (cosψ+1)/2))
 
-        noistar_idata = pm.sample()
+        noistar_idata = pm.sample(target_accept=0.9,chains=4)
         
     post = istar_idata.posterior
     istar_draws = np.zeros(shape=(len(x),4000))
@@ -124,7 +124,6 @@ def posteriors(this_model):
 ### PyMC models ###
 
 if __name__ == '__main__':
-
 
     with pm.Model() as model_uni:
 

### 2023-02-08 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 We do not yet understand the importance of stellar inclination in population stellar obliquity inference.
-In this work, we propose an approach to infer the stellar obliquity $\psi$ distribution from the sky-projected stellar obliquity $\lambda$ distribution.
-We show that the stellar obliquity distribution inference weakly depends on stellar inclinations.
+In this work, we infer the stellar obliquity distribution from sky-projected stellar obliquities and show that the stellar obliquity distribution weakly depends on stellar inclinations.
 In Section~\ref{sec:hbm}, we introduce a flexible, hierarchical Bayesian framework to infer the population stellar obliquity distribution.
@@
 
-To find the unknown stellar obliquity distribution of exoplanetary systems, we build a hierarchical Bayesian framework that takes the sky-projected stellar obliquity $\lambda$ and orbital inclination $i_{\rm orb}$ as observed data.
-Stellar inclination $i_\star$ is an optional and flexible input.
-It could be directly taken from previous observations, or if the rotational modulation technique is used, the stellar inclination will be inferred from the stellar rotation period $P_{\rm rot}$, stellar radius $R_\star$, and sky-projected rotational broadening velocity $v\sin{i}_\star$.
-If the stellar inclination information is not provided, we simply assume an isotropic stellar inclination.
-As we will show in later sections, assuming an isotropic stellar inclination will not compromise the stellar obliquity distribution inference.
+To find the stellar obliquity distribution of exoplanetary systems, we build a hierarchical Bayesian framework that takes the sky-projected stellar obliquity $\lambda$ and orbital inclination $i_{\rm orb}$ as observed data.
+Measurements on the stellar inclination $i_\star$ are optional.
+If the $i_\star$ information is not provided, we simply assume that they follow an isotropic distribution.
+Otherwise, the observed $i_\star$ could be taken from previous measurements, or if the rotational modulation method is applied, the $i_\star$ could be inferred from the stellar rotation period $P_{\rm rot}$, stellar radius $R_\star$, and sky-projected rotational broadening velocity $v\sin{i}_\star$.
 
@@
 Figure~\ref{fig:graph} shows the probabilistic graphical model for the hierarchical Bayesian framework.
-We aim to constrain a set of hyperparameters $\bm{\beta}$ that describe the stellar obliquity $\psi$ distribution.
+We aim to constrain a set of hyperparameters $\bm{\beta}$ that describes the stellar obliquity $\psi$ distribution.
 The parameter set $\bm{\beta}$ is constrained by $N$ individual systems in which $\psi_n$ is a random variable.
 $\psi_n$ is computed from the sky-projected stellar obliquity $\lambda_n$, the orbital inclination $i_{{\rm orb},n}$, and the stellar inclination $i_{\star, n}$.
-The parameter $\gamma_{\star, n}$ is a parameter set that contains all the stellar properties other than $i_\star$, e.g., stellar rotation period $P_{\rm rot}$ and radius $R_\star$, if the information is available.
+The parameter $\gamma_{\star, n}$ is a parameter set that contains all the stellar properties other than $i_{\star,n}$, such as the stellar rotation period $P_{{\rm rot},n}$, radius $R_{\star,n}$, and projected rotational velocity $v\sin{i_{\star,n}}$, if they are available.
+
+The sky-projected stellar obliquity $\lambda_n$ is constrained by observed $\hat{\lambda}_n$ via the Rossiter-McLaughlin effect or gravity darkening.
+The orbital inclination $i_{{\rm orb},n}$ is constrained by the transit light curves of exoplanets.
 Obs$_{\star,n}$ includes observed data that helps to constrain $i_{\star,n}$ and $\gamma_{\star, n}$.
 If the rotational modulation method is used to constrain the stellar inclination, the $i_{\star,n}$ and $\gamma_{\star,n}$ are constrained by rotational modulation in photometric or spectroscopic time series and observed sky-projected stellar rotational line broadening $\hat{v}\sin{i_{\star,n}}$ \citep[e.g.,][]{Masuda20}, where we have $v\sin{i_\star} = 2 \pi R_\star / P_{\rm rot}$.
-If the gravity-darkening method is used, the $i_{\star,n}$ is constrained by the anomaly in transit light curves.
-If asteroseismology is used, the $i_{\star,n}$ is constrained from the periodic variation in the photometric time series.
-The sky-projected stellar obliquity $\lambda_n$ can be constrained by observed $\hat{\lambda}_n$ via the Rossiter-McLaughlin effect or gravity darkening.
-Lastly, the orbital inclination $i_{{\rm orb},n}$ can be constrained from the transit light curves.
+If the gravity-darkening method is used, the $i_{\star,n}$ could be constrained by the anomaly in transit light curves.
+If asteroseismology is used, the $i_{\star,n}$ could be constrained from the mode splitting in the Fourier transform of photometric time series.
 
 For the stellar obliquity distribution, we model the $\cos{\psi}$ distribution instead of $\psi$ distribution to understand if the stellar obliquity is isotropically distributed.
-If stellar obliquities of misaligned systems follow an isotropic distribution, the $\cos{\psi}$ distribution will be uniformly distributed between -1 to 1.
+If stellar obliquity distribution follows an isotropic distribution, the $\cos{\psi}$ distribution will be uniformly distributed between -1 to 1.
 We adopt a two-component mixture Beta distribution with hyperparameters $\bm{\beta} = \{\bm{w},\bm{a},\bm{b}\}$, where each hyperparameter has a dimension of 2.
@@
     \mathcal{L}(\lambda) &\sim \prod_{i=1}^N\mathcal{N}(\hat{\lambda}_n, \sigma_{\hat{\lambda}_n}) \nonumber\\
+    \mathcal{L}(i_{\rm orb}) &\sim \prod_{i=1}^N\mathcal{N}(\hat{i}_{{\rm orb},n}, \sigma_{\hat{i}_{{\rm orb},n}}) \nonumber\\
     \mathcal{L}(i_\star) &\sim \prod_{i=1}^N\mathcal{N}({\rm Obs}_{\star,n}, \sigma_{{\rm Obs}_{\star,n}}).
@@
 Here we assume uniform hyperpriors for the Beta distribution.
-For population inference on a small number of systems (i.e., $N<50$), the choice of hyperpriors could impact the inferred distribution \citep{Nagpal22}.
+For population inference on a small number of systems (i.e., $N \lesssim 50$), the choice of hyperpriors could impact the inferred distribution \citep{Nagpal22}.
 When applying the framework to a small sample size, it is important to test the robustness of inferred distributions on different hyperpriors.
@@
 All figures and simulations in this paper are fully reproducible and built with the $\mathtt{showyourwork}$ package.
-The open source code can be found on \href{https://github.com/jiayindong/polar}{GitHub\,\faGithub}.
+The open source code can be found on \href{https://github.com/jiayindong/obliquity}{GitHub\,\faGithub\,(https://github.com/jiayindong/obliquity)}.
 
@@
 
-\subsection{Simulated Data}
+\subsection{Simulated Data}\label{subsec:sims}
 
@@
 Since the simulated stellar obliquity distributions only have a single component, we model the data with a single Beta distribution.
-Each row corresponds to an injected stellar obliquity distribution.
-The orange curve and contours are the median and uncertainties of the inferred $\cos{\psi}$ distribution with stellar inclination information, and the blue curve and contours are the ones without stellar inclination information.
+Each row of Figure~\ref{fig:simulation} corresponds to an injected stellar obliquity distribution.
+The orange curve and contours are the median and 1-$\sigma$ and 2-$\sigma$ uncertainties of the inferred $\cos{\psi}$ distribution with stellar inclination information, and the blue curve and contours are the ones without stellar inclination information.
 We simply assume an isotropic distribution of $i_\star$ for these models.
 Surprisingly, in all four distributions, the inferred $\cos{\psi}$ distributions with or without $i_\star$ correctly recover the injected distributions shown as grey dashed lines as shown in Figure~\ref{fig:simulation}.
-Since some of the injected distributions are Normal distributions, it is expected that the inferred distributions, which as Beta distributions, could not exactly describe the injected distributions.
+The inferred modes and widths of the stellar obliquity distributions with or without $i_\star$ likelihood are consistent, despite that the inferred distributions without $i_\star$ measurements could have wider uncertainties shown in shallow color contours.
+Since the injected distributions are Normal distributions in rows 2--4, it is expected that the inferred distributions, which as Beta distributions, could not exactly describe the injected distributions.
 
-We also examine the importance of orbital inclination to the stellar obliquity inference.
-Since all the systems we study are transiting exoplanetary systems, we consider an isotropic orbital inclination distribution from $80\degr$ to $90\degr$.
-This broad inclination range corresponds to an impact parameter range from 0 to 1 for a circular orbit planet with an $a/R_\star$ of 6.
-We find at this level of orbital inclination deviations, the difference on the inferred stellar obliquity distributions is minimal.
-For transiting planet systems, approximating orbital inclinations to $90\degr$ will not compromise the stellar obliquity distribution inference.
+We also examine the importance of orbital inclination $i_{\rm orb}$ to the stellar obliquity distribution inference.
+Since all the systems we study are transiting-planet systems, we consider an isotropic orbital inclination distribution between $80\degr$ and $90\degr$.
+This broad inclination range corresponds to an impact parameter range from 0 to 1 with a planet-star separation $a/R_\star$ of 6.
+We compare the inferred stellar obliquity distribution by approximating $i_{\rm orb}$ to $90\degr$ with the inferred distribution using the true $i_{\rm orb}$, and find the difference between the two distributions is minimal.
+For transiting-planet systems, approximating orbital inclinations to $90\degr$ will not compromise the stellar obliquity distribution inference.
 
@@
 The constraint on stellar obliquity distribution is predominantly based on sky-projected stellar obliquities.
-The sky-projected stellar obliquity could correctly infer the peak and width of the underlying stellar obliquity distribution.
+Just from the sky-projected stellar obliquity, we could correctly infer the peak and width of the underlying stellar obliquity distribution.
+Later in Section~\ref{sec:jacobian}, we derive the Jacobian transformations between $\psi$, $\lambda$, and $i_\star$ to explain why there is no significant information loss when not having $i_\star$ measurements.
 
@@
     \includegraphics[width=\linewidth]{figures/simulation.pdf}
-    \caption{Inferred stellar obliquity distributions from sky-projected stellar obliquities with or without stellar inclination information.
+    \caption{Inferred stellar obliquity distributions from sky-projected stellar obliquities with or without stellar inclination measurements.
 Each row presents a set of simulated data where the underlying $\cos{\psi}$ distribution is known and shown as grey dashed curves.
-The inferred $\cos{\psi}$ distributions with or without stellar inclination are shown in orange and blue curves, respectively.
-The shallow contours are $1\sigma$ and $2\sigma$ uncertainties of the inferred distributions.}
+The inferred $\cos{\psi}$ distributions with or without stellar inclination measurements are shown in orange and blue curves, respectively.
+The shallow contours are 1-$\sigma$ and 2-$\sigma$ uncertainties of the inferred distributions.}
     \label{fig:simulation}
@@
 
-We then apply the hierarchical Bayesian framework to all exoplanetary systems with sky-projected stellar obliquity $\lambda$ measurements.
-We use the sample of 161 systems summarized in \cite{Albrecht22} Table A1, which are predominantly Hot Jupiter systems.
+We next apply the hierarchical Bayesian framework to all exoplanetary systems with sky-projected stellar obliquity $\lambda$ measurements.
+We use the sample of 161 systems summarized in \cite{Albrecht22} Table A1, which are mainly Hot Jupiter systems.
 The inferred $\cos{\psi}$ distribution is shown in Figure~\ref{fig:psi_dist}.
@@
 
-Our model identifies a pile-up of systems with stellar obliquity less than $40\degr$ and finds that misaligned systems have nearly isotropic stellar obliquity distribution.
-When comparing the inferred distribution with the prediction of stellar obliquity distributions in different Hot Jupiter origin scenarios, it is in best agreement with the outcome of the scattering of multiple giant planets after a convergent disk migration \citep[e.g.,][]{Nagasawa11, Beague12}.
-This particular intriguing result should be further examined on more carefully selected Hot Jupiter samples and put constraints on Hot Jupiter origin channels.
+The stellar obliquity distribution suggests that $67\pm9$\% of the systems have stellar obliquity less than $40\degr$, and $33\pm9$\% of the systems have nearly isotropically distributed stellar obliquities spread from $40\degr$ to $140\degr$.
+The distribution could have important implications for the formation and evolution of close-in planetary systems.
+The broad distribution of misaligned systems is in good agreement with the outcome of the scattering of multiple giant planets after a convergent disk migration \citep[e.g.,][]{Nagasawa11, Beague12}, one of the proposed mechanisms that form Hot Jupiters.
+The intriguing result should be further examined on more carefully selected Hot Jupiter samples and put constraints on Hot Jupiter origin channels.
 
@@
 
-In this section, we aim to use mathematical expressions to understand why the stellar obliquity distribution is predominantly determined by the sky-projected stellar obliquity distribution and why the stellar inclination distribution is less important to the inference.
+In this section, we aim to understand why the stellar obliquity distribution is predominantly determined by the sky-projected stellar obliquity distribution and why the stellar inclination distribution is less important to the inference using mathematical expressions.
 
@@
 We approximate the orbital inclination of the transiting planet to $90\degr$ here to simplify the problem for a population study.
-We found in the previous section that the assumption would not compromise the stellar obliquity distribution inference for transiting planets.
+As discussed in Section~\ref{subsec:sims}, the assumption will not compromise the stellar obliquity distribution inference for transiting planets.
 
@@
 
-In this work, we demonstrate the stellar obliquity distribution inference is predominantly based on the sky-projected stellar obliquity.
-While stellar inclination improves the stellar obliquity constraint on individual systems, it does not improve the population-level stellar obliquity constraint.
-We introduce a flexible, hierarchical Bayesian framework to infer the stellar obliquity distribution of a population.
-Information on stellar inclination is not required for the stellar obliquity distribution inference; if the stellar inclination is not provided, we simply assume an isotropic stellar inclination distribution.
-When combining samples with and without stellar inclination measurements, it is important to check if the $i_\star$ sample is representative of the full sample.
-If not, since the $i_\star$ sample will constrain stellar obliquities more tightly, the overall distribution could be weighted towards these systems, bias our interpretation.
-The hierarchical Bayesian model we introduce in this paper is an open-source code available to the public to apply to a certain sample of targets and customize for different stellar obliquity distributions and priors.
+In this work, we demonstrated that the stellar obliquity distribution inference is predominantly determined by the sky-projected stellar obliquities.
+Without stellar inclination measurements, the stellar obliquity distribution could still be correctly inferred.
+We introduce a flexible, hierarchical Bayesian framework to infer the stellar obliquity distribution of transiting-planet systems.
+Stellar inclination measurement is optional, and if it is not provided, we simply assume an isotropic stellar inclination distribution.
+The hierarchical Bayesian model is an open-source code that can be found on \href{https://github.com/jiayindong/obliquity}{GitHub\,\faGithub\,(https://github.com/jiayindong/obliquity)}.
+Users may apply the framework to a certain sample of targets and customize the model for different stellar obliquity distributions and priors.
+When jointly modeling the stellar obliquity distribution from one sample with stellar inclination measurements and one without, it is important to check if the $i_\star$ sample is representative of the full sample.
+If not, since the $i_\star$ sample will constrain stellar obliquities more tightly, the overall distribution could be weighted towards these systems, bias the interpretation.
+Lastly, we applied the framework to all exoplanetary systems with sky-projected stellar obliquity measurements and found that $67\pm9$\% of the systems have stellar obliquity less than $40\degr$, and $33\pm9$\% of the systems have nearly isotropically distributed obliquities between $40\degr$ and $140\degr$.
+The distribution could have important implications for the formation and evolution of close-in planetary systems and be worthwhile for further investigation.
 
-When applying our framework to all exoplanetary systems with sky-projected stellar obliquity measurements, we found no significant clustering near $\cos{\psi} = 0$, at odds with the previous study of systems that have both sky-projected stellar obliquity and stellar obliquity measurements.
-Potential biases that lead to this discrepancy are open to discussion.
-Their stellar obliquity is nearly isotropically distributed for misaligned exoplanetary systems in support of a multi-planet, convergent-disk-migration history followed by a scattering phase in these planets' dynamical history.
-Future studies on a more carefully selected Hot Jupiter sample will put better constraints on the origin of Hot Jupiters from their stellar obliquity distribution.
-
-% \section{More}
-
-% First, we apply the framework to the sample of \numistar systems that have both sky-projected stellar obliquity $\lambda$ and stellar inclination $i_\star$ constraints summarized in the \cite{Albrecht22} review.
-If the host star's rotation period is available, we use the value along with the stellar radii and $v\sin{i_\star}$ to infer the stellar inclination.
-Otherwise, we use the reported stellar inclination and uncertainty listed in Table A1 of \cite{Albrecht22}.
-% The inferred $\cos{\psi}$ distribution is shown in the left panel of Figure~\ref{fig:psi_dist}.
-The distribution presents a pike at 1, suggesting the majority of exoplanetary systems are well-aligned systems.
-For misaligned systems, $\cos{\psi}$ presents a cluster near $\cos{\psi} = -0.2$, corresponding to $\psi$ of $100\degr$.
-The result is consistent with the findings of a preponderance of perpendicular planets in \cite{Albrecht21}.
-
-% To test the dependence of $\cos{\psi}$ distribution on $i_\star$ measurements, we apply the model again to the sample of \numistar systems with both constrained $\lambda$ and $i_\star$.
-However, in this model, we do not use any additional information on $i_\star$ other than applying an isotropic inclination prior to each system.
-The corresponding probabilistic graphical model is Figure~\ref{fig:graph} but removing all the dashed edges.
-% The $\cos{\psi}$ distribution of the new model is shown in the middle panel of Figure~\ref{fig:psi_dist}.
-Interestingly, the overall $\cos{\psi}$ distributions of the two models are very similar, despite the fact that the inferred $\psi$ uncertainty for each system is now greater.
-% The similarity between the two distributions suggests that, at least for the observed-$i_\star$ sample, the $\cos{\psi}$ distribution mainly depends on the underlying sky-project stellar obliquity $\lambda$ distribution.
-
-% Consequently, we apply our model to all \numall exoplanetary systems that have sky-projected stellar obliquity measurements summarized in \cite{Albrecht22}.
-The result is presented in the right panel of Figure~\ref{fig:psi_dist}.
-The $\cos{\psi}$ distribution still spikes at 1 but is nearly flat between $-0.75$ and $0.75$ with no significant clustering.
-The distribution corresponds to an isotropic distribution of $\psi$ from $40\degr$ to $140\degr$.
-
-% There is no doubt that $i_\star$ measurement on an individual system improves the constraint on $\psi$.
-However, are $i_\star$ measurements required to constrain the population $\psi$ distribution?
-Particularly, the left and middle panels in Figure~\ref{fig:psi_dist} show that without $i_\star$ measurements, a similar level of constraint on the $\cos{\psi}$ distribution can be achieved.
-We perform some calculations and tests to evaluate the necessity of $i_\star$ constraints.
-
-% We also design some $\cos{\psi}$ distributions and infer their distributions purely from the $\lambda$ distributions.
-From each distribution of $\cos{\psi}$ and assuming a uniform $\theta$, we randomly generate 200 samples of $i_\star$ and $\lambda$ and assume them as \emph{true} values.
-We then add some Gaussian noises to \emph{true} $\lambda$ and $i_\star$ to simulate \emph{observed} values.
-We use the uncertainties of $\sigma_{\lambda} = 8\degr$ and $\sigma_{i_\star} = 10\degr$ which are typical observation uncertainties.
-Lastly, from the observed values and their uncertainties, we infer the $\cos{\psi}$ distribution of the sample with or without $i_\star$ measurements.
-% We test four different distributions of $\cos{\psi}$: a uniform distribution $\mathcal{U}(-1,1)$ and three Normal distributions $\mathcal{N}(0,0.2)$, $\mathcal{N}(-0.4,0.2)$, and $\mathcal{N}(0.4,0.2)$.
-% In all four distributions, we find inferences with or without $i_\star$ are consistent with each other and injected distributions.
-
-% We find the underlying $\lambda$ distributions of the observed-$i_\star$ sample and the full sample by building another hierarchical Bayesian model.
-We again use a two-component mixture Beta distribution to describe well-aligned and misaligned systems.
-We extend the Beta distribution support from $[0,1]$ to $[0,\pi]$ by multiplying the distribution by $\pi$.
-We use the observed $\lambda$ and uncertainties for the likelihoods.
-Figure~\ref{fig:lam_dist} shows the $\lambda$ distributions of two samples.
-It clearly shows that in the observed-$i_\star$ sample, misaligned systems are clustered at $\sim 110\degr$, whereas in the full sample, misaligned systems are nearly uniformly distributed.
-
-% The stellar obliquity distribution of exoplanetary systems can be inferred purely from their sky-projected stellar obliquity distribution.
-When we apply our framework to all systems with $\lambda$ measurements, we find misaligned systems have nearly isotropic stellar obliquity distribution.
-We compare the inferred distribution from the observation to the prediction of mutual inclination of different eccentricity excitation mechanisms of Hot Jupiters.
-It is in best agreement with the outcome of the scattering of multiple giant planets after a convergent disk migration \citep{Beague12}.
-
-\vspace{5mm}
+\vspace*{5mm}
 

### 2023-02-08 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-\begin{figure}[ht!]
-    \script{simulation.py}
-    \includegraphics[width=\linewidth]{figures/simulation.pdf}
-    \caption{Inferred stellar obliquity distributions from sky-projected stellar obliquities with or without stellar inclination measurements.
+% \begin{figure}[ht!]
+%     \script{simulation.py}
+%     \includegraphics[width=\linewidth]{figures/simulation.pdf}
+%     \caption{Inferred stellar obliquity distributions from sky-projected stellar obliquities with or without stellar inclination measurements.
 Each row presents a set of simulated data where the underlying $\cos{\psi}$ distribution is known and shown as grey dashed curves.
@@
 The shallow contours are 1-$\sigma$ and 2-$\sigma$ uncertainties of the inferred distributions.}
-    \label{fig:simulation}
-\end{figure}
+%     \label{fig:simulation}
+% \end{figure}
 
@@
 \begin{figure}[ht!]
-    \script{psi_dist.py}
+    \script{psi_plot.py}
     \begin{centering}

### 2023-02-08 src/scripts/psi.py
[showyourwork] overleaf sync

@@ -138,29 +138,4 @@ if __name__ == '__main__':
         
         all_randinc = pm.sample(target_accept=0.9,chains=4)
 
-    x = np.linspace(1e-5,1-1e-5,1000)
-
-    post = all_randinc.posterior
-    all_randinc_draws = np.zeros(shape=(len(x),4000))
-    for a in range(4):
-        for b in range(1000):
-            all_randinc_draws[:, a*1000+b] = (post.w[a,b,0].values*beta.pdf(x, post.a0[a,b], post.b0[a,b])
-                                              +post.w[a,b,1].values*beta.pdf(x, post.a1[a,b], post.b1[a,b]))
-
-    # make the figure
-    plt.figure(figsize=(3.5,2.7),dpi=110)
-
-    q025, q16, q50, q84, q975 = np.percentile(all_randinc_draws, [2.5, 16, 50, 84, 97.5], axis=1)
-    plt.plot(2*x-1, q50, color='C0', lw=1.2)
-    plt.fill_between(2*x-1, q16, q84, alpha=0.3, label="posterior", color='#7dabd0')
-    plt.fill_between(2*x-1, q025, q975, alpha=0.3, color='#7dabd0')
-
-    plt.ylim([0,2])
-    plt.xlim([-1,1])
-
-    plt.xlabel(r'$\cos{\psi}$',fontsize=11)
-    plt.ylabel('Probablity density function')
-
-    plt.tight_layout()
-    plt.savefig(paths.figures / "psi_dist.pdf", bbox_inches="tight", dpi=600)
-    plt.close()
+    az.to_netcdf(all_randinc.posterior, paths.data / "all_randinc.nc")

### 2023-02-08 src/scripts/psi_plot.py

@@ -1,0 +1,58 @@
+import os
+from pathlib import Path
+os.environ["PATH"] += os.pathsep + str(Path.home() / "bin")
+
+import sys
+import subprocess
+
+import paths
+import matplotlib.pyplot as plt
+import pymc as pm
+import aesara.tensor as at
+import numpy as np
+import arviz as az
+from scipy.stats import beta
+
+from matplotlib import rc
+rc('font', **{'family':'sans-serif'})
+rc('text', usetex=True)
+
+plt.rcParams['xtick.top'] =  True
+plt.rcParams['xtick.direction'] =  'in'
+plt.rcParams['xtick.major.width'] =  1.0
+plt.rcParams['xtick.minor.width'] =  1.0
+plt.rcParams['ytick.right'] =  True
+plt.rcParams['ytick.direction'] =  'in'
+plt.rcParams['ytick.major.width'] = 1.0
+plt.rcParams['ytick.minor.width'] =  1.0
+plt.rcParams['lines.markeredgewidth'] =  1.0
+
+# plot the cospsi distribution
+all_randinc = az.from_netcdf(paths.data / "all_randinc.nc")
+
+x = np.linspace(1e-5,1-1e-5,1000)
+
+post = all_randinc.posterior
+all_randinc_draws = np.zeros(shape=(len(x),4000))
+for a in range(4):
+    for b in range(1000):
+        all_randinc_draws[:, a*1000+b] = (post.w[a,b,0].values*beta.pdf(x, post.a0[a,b], post.b0[a,b])
+                                          +post.w[a,b,1].values*beta.pdf(x, post.a1[a,b], post.b1[a,b]))
+
+# make the figure
+plt.figure(figsize=(3.5,2.7),dpi=110)
+
+q025, q16, q50, q84, q975 = np.percentile(all_randinc_draws, [2.5, 16, 50, 84, 97.5], axis=1)
+plt.plot(2*x-1, q50, color='C0', lw=1.2)
+plt.fill_between(2*x-1, q16, q84, alpha=0.3, label="posterior", color='#7dabd0')
+plt.fill_between(2*x-1, q025, q975, alpha=0.3, color='#7dabd0')
+
+plt.ylim([0,2])
+plt.xlim([-1,1])
+
+plt.xlabel(r'$\cos{\psi}$',fontsize=11)
+plt.ylabel('Probablity density function')
+
+plt.tight_layout()
+plt.savefig(paths.figures / "psi_dist.pdf", bbox_inches="tight", dpi=600)
+plt.close()

### 2023-02-08 src/scripts/sim_plot.py

@@ -1,0 +1,392 @@
+import paths
+import os
+from pathlib import Path
+os.environ["PATH"] += os.pathsep + str(Path.home() / "bin")
+
+import subprocess
+import sys
+
+import pymc as pm
+import arviz as az
+import aesara.tensor as at
+
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+
+import scipy.integrate as integrate
+from scipy import stats
+
+from matplotlib import rc
+rc('font', **{'family':'sans-serif'})
+rc('text', usetex=True)
+rc('text.latex', preamble=r'\usepackage{physics}')
+
+plt.rcParams['xtick.top'] =  True
+plt.rcParams['xtick.direction'] =  'in'
+plt.rcParams['xtick.major.width'] =  1.0
+plt.rcParams['xtick.minor.width'] =  1.0
+plt.rcParams['ytick.right'] =  True
+plt.rcParams['ytick.direction'] =  'in'
+plt.rcParams['ytick.major.width'] = 1.0
+plt.rcParams['ytick.minor.width'] =  1.0
+plt.rcParams['lines.markeredgewidth'] =  1.0
+
+from scipy.stats import beta
+x = np.linspace(1e-5,1-1e-5,1000)
+σ = 0.2
+
+# return posteriors of HBM models with or without using measured istar
+def posteriors(this_model):
+    
+    nsample = 200
+    err_istar = 10*np.pi/180
+    err_lam = 8*np.pi/180
+
+    with this_model:    
+        idata = pm.sample(chains=4, draws=50)
+
+    true_istar = idata.posterior.i.values.ravel()
+    obs_istar = true_istar + err_istar*np.random.normal(size=nsample)
+
+    true_lam = idata.posterior.λ.values.ravel()
+    obs_lam = true_lam + err_lam*np.random.normal(size=nsample)
+
+    with pm.Model() as model_istar:
+
+        # hyperprior
+        a = pm.Uniform('a', lower=0, upper=10)
+        b = pm.Uniform('b', lower=0, upper=10)
+
+        # flat prior on cosi
[…]

### 2023-02-08 src/scripts/sim.py
zenodo

@@ -214,179 +214,17 @@ if __name__ == '__main__':
         
 
     uni_istar_draws, uni_noistar_draws = posteriors(model_uni)
+    np.save(paths.data / "uni_istar_draws.npy", uni_istar_draws)
+    np.save(paths.data / "uni_noistar_draws.npy", uni_noistar_draws)
+
     norm1_istar_draws, norm1_noistar_draws = posteriors(model_norm1)
+    np.save(paths.data / "norm1_istar_draws.npy", norm1_istar_draws)
+    np.save(paths.data / "norm1_noistar_draws.npy", norm1_noistar_draws)
+
     norm2_istar_draws, norm2_noistar_draws = posteriors(model_norm2)
+    np.save(paths.data / "norm2_istar_draws.npy", norm2_istar_draws)
+    np.save(paths.data / "norm2_noistar_draws.npy", norm2_noistar_draws)
+
     norm3_istar_draws, norm3_noistar_draws = posteriors(model_norm3)
-
-    ### Make the plot ###
-    fig, big_axes = plt.subplots(figsize=(3.5,6.5),dpi=110,nrows=4,ncols=1,sharey=True) 
-
-    for row, big_ax in enumerate(big_axes, start=1):
-        
-        if row == 4:
-            big_ax.set_title(r"$\cos{\psi} \sim \mathcal{N}(0.4,0.2)$" "\n")
-        if row == 3:
-            big_ax.set_title(r"$\cos{\psi} \sim \mathcal{N}(-0.4,0.2)$" "\n")
-        if row == 2:
-            big_ax.set_title(r"$\cos{\psi} \sim \mathcal{N}(0,0.2)$" "\n")
-        elif row == 1:
-            big_ax.set_title(r"$\cos{\psi} \sim \mathcal{U}(-1,1)$" "\n")
-
-        # Turn off axis lines and ticks of the big subplot 
-        # obs alpha is 0 in RGBA string!
-        big_ax.tick_params(labelcolor=(1.,1.,1., 0.0), top='off', bottom='off', left='off', right='off')
-        # removes the white frame
[…]

### 2023-02-08 src/scripts/sim_plot.py

@@ -32,361 +32,191 @@ plt.rcParams['ytick.major.width'] = 1.0
 plt.rcParams['ytick.minor.width'] =  1.0
 plt.rcParams['lines.markeredgewidth'] =  1.0
 
-from scipy.stats import beta
 x = np.linspace(1e-5,1-1e-5,1000)
-σ = 0.2
+σ = 0.2        
 
-# return posteriors of HBM models with or without using measured istar
-def posteriors(this_model):
+# load the distribution draws from MCMC
+uni_istar_draws = np.load(paths.data / "uni_istar_draws.npy")
+uni_noistar_draws = np.load(paths.data / "uni_noistar_draws.npy")
+
+norm1_istar_draws = np.load(paths.data / "norm1_istar_draws.npy")
+norm1_noistar_draws = np.load(paths.data / "norm1_noistar_draws.npy")
+
+norm2_istar_draws = np.load(paths.data / "norm2_istar_draws.npy")
+norm2_noistar_draws = np.load(paths.data / "norm2_noistar_draws.npy")
+
+norm3_istar_draws = np.load(paths.data / "norm3_istar_draws.npy")
+norm3_noistar_draws = np.load(paths.data / "norm3_noistar_draws.npy")
+
+### Make the plot ###
+fig, big_axes = plt.subplots(figsize=(3.5,6.5),dpi=110,nrows=4,ncols=1,sharey=True) 
+
+for row, big_ax in enumerate(big_axes, start=1):
     
-    nsample = 200
-    err_istar = 10*np.pi/180
-    err_lam = 8*np.pi/180
+    if row == 4:
+        big_ax.set_title(r"$\cos{\psi} \sim \mathcal{N}(0.4,0.2)$" "\n")
+    if row == 3:
+        big_ax.set_title(r"$\cos{\psi} \sim \mathcal{N}(-0.4,0.2)$" "\n")
+    if row == 2:
+        big_ax.set_title(r"$\cos{\psi} \sim \mathcal{N}(0,0.2)$" "\n")
+    elif row == 1:
+        big_ax.set_title(r"$\cos{\psi} \sim \mathcal{U}(-1,1)$" "\n")
 
-    with this_model:    
-        idata = pm.sample(chains=4, draws=50)
[…]

### 2023-02-09 src/scripts/sim.py
[showyourwork] overleaf sync

@@ -212,7 +212,7 @@ if __name__ == '__main__':
         iso_cosi = pm.Uniform('iso_cosi', lower=0, upper=1)
         iso_cosψ = pm.Deterministic('iso_cosψ', at.sqrt(1-iso_cosi**2)*cosλ)
         
-
+        
     uni_istar_draws, uni_noistar_draws = posteriors(model_uni)
     np.save(paths.data / "uni_istar_draws.npy", uni_istar_draws)
     np.save(paths.data / "uni_noistar_draws.npy", uni_noistar_draws)

### 2023-02-09 src/scripts/sim_plot.py

@@ -48,6 +48,8 @@ norm2_noistar_draws = np.load(paths.data / "norm2_noistar_draws.npy")
 norm3_istar_draws = np.load(paths.data / "norm3_istar_draws.npy")
 norm3_noistar_draws = np.load(paths.data / "norm3_noistar_draws.npy")
 
+print('yes')
+
 ### Make the plot ###
 fig, big_axes = plt.subplots(figsize=(3.5,6.5),dpi=110,nrows=4,ncols=1,sharey=True) 
 
@@ -217,6 +219,8 @@ plt.plot(cosψ, pcosψ, c='slategrey', ls='--', lw=1, zorder=0)
 fig.set_facecolor('w')
 
 plt.tight_layout(pad=0,w_pad=0.5, h_pad=-2)
+
+print('yes')
 
 plt.savefig(paths.figures / "simulation.pdf", bbox_inches="tight", dpi=600)
 plt.close()

### 2023-02-09 src/scripts/sim_plot.py
zenodo

@@ -48,8 +48,6 @@ norm2_noistar_draws = np.load(paths.data / "norm2_noistar_draws.npy")
 norm3_istar_draws = np.load(paths.data / "norm3_istar_draws.npy")
 norm3_noistar_draws = np.load(paths.data / "norm3_noistar_draws.npy")
 
-print('yes')
-
 ### Make the plot ###
 fig, big_axes = plt.subplots(figsize=(3.5,6.5),dpi=110,nrows=4,ncols=1,sharey=True) 
 
@@ -219,8 +217,6 @@ plt.plot(cosψ, pcosψ, c='slategrey', ls='--', lw=1, zorder=0)
 fig.set_facecolor('w')
 
 plt.tight_layout(pad=0,w_pad=0.5, h_pad=-2)
-
-print('yes')
 
 plt.savefig(paths.figures / "simulation.pdf", bbox_inches="tight", dpi=600)
 plt.close()

### 2023-02-09 src/scripts/sim.py
[showyourwork] overleaf sync

@@ -214,17 +214,17 @@ if __name__ == '__main__':
         
         
     uni_istar_draws, uni_noistar_draws = posteriors(model_uni)
-    np.save(paths.data / "uni_istar_draws.npy", uni_istar_draws)
-    np.save(paths.data / "uni_noistar_draws.npy", uni_noistar_draws)
+    np.save(paths.data / "simulation/uni_istar_draws.npy", uni_istar_draws)
+    np.save(paths.data / "simulation/uni_noistar_draws.npy", uni_noistar_draws)
 
     norm1_istar_draws, norm1_noistar_draws = posteriors(model_norm1)
-    np.save(paths.data / "norm1_istar_draws.npy", norm1_istar_draws)
-    np.save(paths.data / "norm1_noistar_draws.npy", norm1_noistar_draws)
+    np.save(paths.data / "simulation/norm1_istar_draws.npy", norm1_istar_draws)
+    np.save(paths.data / "simulation/norm1_noistar_draws.npy", norm1_noistar_draws)
 
     norm2_istar_draws, norm2_noistar_draws = posteriors(model_norm2)
-    np.save(paths.data / "norm2_istar_draws.npy", norm2_istar_draws)
-    np.save(paths.data / "norm2_noistar_draws.npy", norm2_noistar_draws)
+    np.save(paths.data / "simulation/norm2_istar_draws.npy", norm2_istar_draws)
+    np.save(paths.data / "simulation/norm2_noistar_draws.npy", norm2_noistar_draws)
 
     norm3_istar_draws, norm3_noistar_draws = posteriors(model_norm3)
-    np.save(paths.data / "norm3_istar_draws.npy", norm3_istar_draws)
-    np.save(paths.data / "norm3_noistar_draws.npy", norm3_noistar_draws)    
+    np.save(paths.data / "simulation/norm3_istar_draws.npy", norm3_istar_draws)
+    np.save(paths.data / "simulation/norm3_noistar_draws.npy", norm3_noistar_draws)    

### 2023-02-09 src/scripts/sim_plot.py

@@ -36,17 +36,17 @@ x = np.linspace(1e-5,1-1e-5,1000)
 σ = 0.2        
 
 # load the distribution draws from MCMC
-uni_istar_draws = np.load(paths.data / "uni_istar_draws.npy")
-uni_noistar_draws = np.load(paths.data / "uni_noistar_draws.npy")
+uni_istar_draws = np.load(paths.data / "simulation/uni_istar_draws.npy")
+uni_noistar_draws = np.load(paths.data / "simulation/uni_noistar_draws.npy")
 
-norm1_istar_draws = np.load(paths.data / "norm1_istar_draws.npy")
-norm1_noistar_draws = np.load(paths.data / "norm1_noistar_draws.npy")
+norm1_istar_draws = np.load(paths.data / "simulation/norm1_istar_draws.npy")
+norm1_noistar_draws = np.load(paths.data / "simulation/norm1_noistar_draws.npy")
 
-norm2_istar_draws = np.load(paths.data / "norm2_istar_draws.npy")
-norm2_noistar_draws = np.load(paths.data / "norm2_noistar_draws.npy")
+norm2_istar_draws = np.load(paths.data / "simulation/norm2_istar_draws.npy")
+norm2_noistar_draws = np.load(paths.data / "simulation/norm2_noistar_draws.npy")
 
-norm3_istar_draws = np.load(paths.data / "norm3_istar_draws.npy")
-norm3_noistar_draws = np.load(paths.data / "norm3_noistar_draws.npy")
+norm3_istar_draws = np.load(paths.data / "simulation/norm3_istar_draws.npy")
+norm3_noistar_draws = np.load(paths.data / "simulation/norm3_noistar_draws.npy")
 
 ### Make the plot ###
 fig, big_axes = plt.subplots(figsize=(3.5,6.5),dpi=110,nrows=4,ncols=1,sharey=True) 

### 2023-02-09 src/scripts/psi.py
updating paths

@@ -138,4 +138,4 @@ if __name__ == '__main__':
         
         all_randinc = pm.sample(target_accept=0.9,chains=4)
 
-    az.to_netcdf(all_randinc.posterior, paths.data / "simulation/all_randinc.nc")
+    az.to_netcdf(all_randinc.posterior, paths.data / "all_randinc.nc")

### 2023-02-09 src/scripts/psi_plot.py

@@ -28,7 +28,7 @@ plt.rcParams['ytick.minor.width'] =  1.0
 plt.rcParams['lines.markeredgewidth'] =  1.0
 
 # plot the cospsi distribution
-all_randinc = az.from_netcdf(paths.data / "simulation/all_randinc.nc")
+all_randinc = az.from_netcdf(paths.data / "all_randinc.nc")
 
 x = np.linspace(1e-5,1-1e-5,1000)
 

### 2023-02-09 src/scripts/sim.py

@@ -213,18 +213,21 @@ if __name__ == '__main__':
         iso_cosψ = pm.Deterministic('iso_cosψ', at.sqrt(1-iso_cosi**2)*cosλ)
         
         
+    sim_dir = paths.data / "simulation"
+    sim_dir.mkdir(exist_ok=True, parents=True)
+
     uni_istar_draws, uni_noistar_draws = posteriors(model_uni)
-    np.save(paths.data / "simulation/uni_istar_draws.npy", uni_istar_draws)
-    np.save(paths.data / "simulation/uni_noistar_draws.npy", uni_noistar_draws)
+    np.save(sim_dir / "uni_istar_draws.npy", uni_istar_draws)
+    np.save(sim_dir / "uni_noistar_draws.npy", uni_noistar_draws)
 
     norm1_istar_draws, norm1_noistar_draws = posteriors(model_norm1)
-    np.save(paths.data / "simulation/norm1_istar_draws.npy", norm1_istar_draws)
-    np.save(paths.data / "simulation/norm1_noistar_draws.npy", norm1_noistar_draws)
+    np.save(sim_dir / "norm1_istar_draws.npy", norm1_istar_draws)
+    np.save(sim_dir / "norm1_noistar_draws.npy", norm1_noistar_draws)
 
     norm2_istar_draws, norm2_noistar_draws = posteriors(model_norm2)
-    np.save(paths.data / "simulation/norm2_istar_draws.npy", norm2_istar_draws)
-    np.save(paths.data / "simulation/norm2_noistar_draws.npy", norm2_noistar_draws)
+    np.save(sim_dir / "norm2_istar_draws.npy", norm2_istar_draws)
+    np.save(sim_dir / "norm2_noistar_draws.npy", norm2_noistar_draws)
 
     norm3_istar_draws, norm3_noistar_draws = posteriors(model_norm3)
-    np.save(paths.data / "simulation/norm3_istar_draws.npy", norm3_istar_draws)
-    np.save(paths.data / "simulation/norm3_noistar_draws.npy", norm3_noistar_draws)    
+    np.save(sim_dir / "norm3_istar_draws.npy", norm3_istar_draws)
+    np.save(sim_dir / "norm3_noistar_draws.npy", norm3_noistar_draws)    

### 2023-02-10 src/tex/ms.tex
[showyourwork] overleaf sync

--- 2023-02-08 src/tex/ms.tex
@@
 % Title
-\title{A Hierarchical Bayesian Framework for Exoplanetary Stellar Obliquity Distribution}
+\title{A Hierarchical Bayesian Framework for Stellar Obliquity Distribution Inference}
 
@@
 
-\received{}
-\revised{}
-\accepted{}
+% \received{}
+% \revised{}
+% \accepted{}
 
-\submitjournal{the AAS Journals}
+% \submitjournal{the AAS Journals}
 
@@
 The probabilistic model is built with the $\mathtt{PyMC}$ \citep[$\mathtt{v4.1.7}$;][]{pymc} package and the posteriors are sampled using the No-U-Turn Sampler \citep[NUTS;][]{Hoffman11}, a gradient-based sampling algorithm of the Markov chain Monte Carlo (MCMC).
-All figures and simulations in this paper are fully reproducible and built with the $\mathtt{showyourwork}$ package.
+All figures and simulations in this paper are fully reproducible and built with the \showyourwork package.
 The open source code can be found on \href{https://github.com/jiayindong/obliquity}{GitHub\,\faGithub\,(https://github.com/jiayindong/obliquity)}.
@@
 \begin{figure}[ht!]
-    \script{simulation.py}
+    \script{sim_plot.py}
     \includegraphics[width=\linewidth]{figures/simulation.pdf}
@@
     }
-    \vspace*{-1.5cm}
+    \vspace*{-1cm}
     \caption{Two coordinate systems to describe the stellar spin axis $\bf{n}_{\star}$ and the planet's orbital axis $\bf{n}_{\rm orb}$.
@@
 \end{align}
-Thus, $i_\star = \cos[-1](\sin{\psi}\cos{\theta})$ and $\lambda = \tan[-1](\tan{\psi}\sin{\theta})$.
 
@@
 
+\section{Reproducibility}\label{sec:reproducibility}
+This study was carried out using the reproducibility software \href{https://github.com/showyourwork/showyourwork}{\showyourwork} \citep{Luger2021}, which leverages continuous integration to programmatically download the data from \href{https://zenodo.org/}{zenodo.org}, create the figures, and compile the manuscript.
+Each figure caption contains one link to the script used to make the figure (at the commit corresponding to the current build of the manuscript).
+The git repository associated with this study is publicly available at \url{https://github.com/jiayindong/obliquity}.
+A Jupyter notebook with examples of the stellar obliquity distribution inference can be found in the same repository.
+
 \vspace*{5mm}
 
-\software{$\mathtt{ArviZ}$ \citep{arviz_2019}, $\mathtt{Jupyter}$ \citep{Jupyter}, $\mathtt{Matplotlib}$ \citep{Matplotlib07, Matplotlib16}, $\mathtt{NumPy}$ \citep{NumPy11, NumPy20}, $\mathtt{pandas}$ \citep{mckinney-proc-scipy-2010, reback2020pandas}, $\mathtt{PyMC}$ \citep{pymc}, $\mathtt{SciPy}$ \citep{2020SciPy-NMeth}}
+\software{$\mathtt{ArviZ}$ \citep{arviz_2019}, $\mathtt{Jupyter}$ \citep{Jupyter}, $\mathtt{Matplotlib}$ \citep{Matplotlib07, Matplotlib16}, $\mathtt{NumPy}$ \citep{NumPy11, NumPy20}, $\mathtt{pandas}$ \citep{mckinney-proc-scipy-2010, reback2020pandas}, $\mathtt{PyMC}$ \citep{pymc}, $\mathtt{SciPy}$ \citep{2020SciPy-NMeth}, \showyourwork \citep{Luger2021}}
 

### 2023-02-10 src/scripts/transform.py
matplotlibrc

@@ -22,16 +22,6 @@ from scipy import stats
 # rc('text', usetex=False)
 # rc('text.latex', preamble=r'\usepackage{physics}')
 
-plt.rcParams['xtick.top'] =  True
-plt.rcParams['xtick.direction'] =  'in'
-plt.rcParams['xtick.major.width'] =  1.0
-plt.rcParams['xtick.minor.width'] =  1.0
-plt.rcParams['ytick.right'] =  True
-plt.rcParams['ytick.direction'] =  'in'
-plt.rcParams['ytick.major.width'] = 1.0
-plt.rcParams['ytick.minor.width'] =  1.0
-plt.rcParams['lines.markeredgewidth'] =  1.0
-
 ### PyMC models ###
 
 if __name__ == '__main__':

### 2023-02-10 src/scripts/psi_plot.py
update simplot

@@ -53,8 +53,6 @@ plt.xlim([-1,1])
 plt.xlabel(r'$\cos{\psi}$',fontsize=11)
 plt.ylabel('Probablity density function')
 
-# plt.gca().set_facecolor('w')
-
 plt.tight_layout()
 plt.savefig(paths.figures / "psi_dist.pdf", bbox_inches="tight", dpi=600)
 plt.close()

### 2023-02-10 src/scripts/sim.py

@@ -19,7 +19,7 @@ from scipy import stats
 
 from matplotlib import rc
 rc('font', **{'family':'sans-serif'})
-rc('text', usetex=False)
+rc('text', usetex=True)
 rc('text.latex', preamble=r'\usepackage{physics}')
 
 plt.rcParams['xtick.top'] =  True
@@ -44,7 +44,7 @@ def posteriors(this_model):
     err_lam = 8*np.pi/180
 
     with this_model:    
-        idata = pm.sample(chains=4, draws=50)
+        idata = pm.sample(chains=4, draws=50, tune=0)
 
     true_istar = idata.posterior.i.values.ravel()
     obs_istar = true_istar + err_istar*np.random.normal(size=nsample)
@@ -105,20 +105,8 @@ def posteriors(this_model):
         hyper = pm.Potential("hyper", pm.logp(pm.Beta.dist(a,b), (cosψ+1)/2))
 
         noistar_idata = pm.sample(target_accept=0.9,chains=4)
-        
-    post = istar_idata.posterior
-    istar_draws = np.zeros(shape=(len(x),4000))
-    for a in range(4):
-        for b in range(1000):
-            istar_draws[:, a*1000+b] = beta.pdf(x, post.a[a,b], post.b[a,b])
 
-    post = noistar_idata.posterior
-    noistar_draws = np.zeros(shape=(len(x),4000))
-    for a in range(4):
-        for b in range(1000):
-            noistar_draws[:, a*1000+b] = beta.pdf(x, post.a[a,b], post.b[a,b])
-
-    return istar_draws, noistar_draws
+    return istar_idata, noistar_idata
 
 
 ### PyMC models ###
@@ -216,18 +204,18 @@ if __name__ == '__main__':
     sim_dir = paths.data / "simulation"
     sim_dir.mkdir(exist_ok=True, parents=True)
 
-    uni_istar_draws, uni_noistar_draws = posteriors(model_uni)
-    np.save(sim_dir / "uni_istar_draws.npy", uni_istar_draws)
-    np.save(sim_dir / "uni_noistar_draws.npy", uni_noistar_draws)
+    uni_istar, uni_noistar = posteriors(model_uni)
+    az.to_netcdf(uni_istar.posterior, sim_dir / "uni_istar.nc")
+    az.to_netcdf(uni_noistar.posterior, sim_dir / "uni_noistar.nc")
 
[…]

### 2023-02-10 src/scripts/sim_plot.py

@@ -17,10 +17,10 @@ import pandas as pd
 import scipy.integrate as integrate
 from scipy import stats
 
-# from matplotlib import rc
-# rc('font', **{'family':'sans-serif'})
-# rc('text', usetex=False)
-# rc('text.latex', preamble=r'\usepackage{physics}')
+from matplotlib import rc
+rc('font', **{'family':'sans-serif'})
+rc('text', usetex=True)
+rc('text.latex', preamble=r'\usepackage{physics}')
 
 plt.rcParams['xtick.top'] =  True
 plt.rcParams['xtick.direction'] =  'in'
@@ -35,18 +35,30 @@ plt.rcParams['lines.markeredgewidth'] =  1.0
 x = np.linspace(1e-5,1-1e-5,1000)
 σ = 0.2        
 
-# load the distribution draws from MCMC
-uni_istar_draws = np.load(paths.data / "simulation/uni_istar_draws.npy")
-uni_noistar_draws = np.load(paths.data / "simulation/uni_noistar_draws.npy")
+# load the posteriors from the MCMC and calculate psi dists
+def psi_dist_draws(model_name):
 
-norm1_istar_draws = np.load(paths.data / "simulation/norm1_istar_draws.npy")
-norm1_noistar_draws = np.load(paths.data / "simulation/norm1_noistar_draws.npy")
+    istar_idata = np.load(paths.data / "simulation/%s_istar_draws.npy"%model_name)
+    noistar_idata = np.load(paths.data / "simulation/%s_noistar_draws.npy"%model_name)
 
-norm2_istar_draws = np.load(paths.data / "simulation/norm2_istar_draws.npy")
-norm2_noistar_draws = np.load(paths.data / "simulation/norm2_noistar_draws.npy")
+    post = istar_idata.posterior
+    istar_draws = np.zeros(shape=(len(x),4000))
+    for a in range(4):
+        for b in range(1000):
+            istar_draws[:, a*1000+b] = beta.pdf(x, post.a[a,b], post.b[a,b])
 
-norm3_istar_draws = np.load(paths.data / "simulation/norm3_istar_draws.npy")
-norm3_noistar_draws = np.load(paths.data / "simulation/norm3_noistar_draws.npy")
+    post = noistar_idata.posterior
+    noistar_draws = np.zeros(shape=(len(x),4000))
+    for a in range(4):
+        for b in range(1000):
[…]

### 2023-02-10 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-\section{Reproducibility}\label{sec:reproducibility}
+\section{Reproducibility}\label{sec:showyourwork}
 This study was carried out using the reproducibility software \href{https://github.com/showyourwork/showyourwork}{\showyourwork} \citep{Luger2021}, which leverages continuous integration to programmatically download the data from \href{https://zenodo.org/}{zenodo.org}, create the figures, and compile the manuscript.

### 2023-02-11 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 Applying the framework to all exoplanetary systems with sky-projected stellar obliquity measurements, we find that the stellar obliquity distribution is unimodal and peaked at zero degrees.
-Misaligned systems have nearly isotropic stellar obliquities, which may have important implications for the formation and evolution of planetary systems.
+Misaligned systems have nearly isotropic stellar obliquities, which may have important implications for close-in planets' formation and evolution.
 
@@
 In the multiple-planet scattering mechanism, the majority of systems should be aligned, and a small fraction of systems have a wide range of stellar obliquities \citep[e.g.,][]{Beague12}.
-Motivated by these predictions, we aim to constrain the stellar obliquity distribution of exoplanetary systems from a Bayesian approach to find dominant mechanisms that sculpt close-in planetary systems.
+Motivated by these predictions, we characterize the stellar obliquity distribution of exoplanetary systems from a Bayesian approach to find the dominant mechanisms that sculpt close-in planetary systems.
 
-In exoplanetary observation, only the sky-projected stellar obliquity $\lambda$, the angle between the projections of $\bf{n}_{\rm orb}$ and $\bf{n}_{\star}$ onto the plane of the sky, can be measured.
-Stellar obliquity $\psi$ can be calculated, if the stellar inclination $i_\star$ is known.
+In exoplanetary observation, only the sky-projected stellar obliquity $\lambda$, the angle between the projections of $\bf{n}_{\rm orb}$ and $\bf{n}_{\star}$ onto the plane of the sky, can be measured, mainly via the Rossiter-McLaughlin effect \citep{Rossiter24, McLaughlin24}.
+The stellar obliquity $\psi$ can be calculated, if both the sky-projected stellar obliquity $\lambda$ and the stellar inclination $i_\star$ are known.
 The relation follows \citep[e.g.,][]{Fabrycky09}:
@@
 \end{equation}
-where $i_{\rm orb}$ is the angle between the vector $\bf{n}_{\rm orb}$ and the observer's line of sight, and $i_\star$ is the angle between $\bf{n}_{\star}$ and the observer's line of sight.
+where $i_{\rm orb}$ is the inclination angle between the vector $\bf{n}_{\rm orb}$ and the observer's line of sight, and $i_\star$ is the inclination angle between $\bf{n}_{\star}$ and the observer's line of sight.
 For transiting exoplanet systems where $i_{\rm orb} \approx 90\degr$, $\cos{\psi} \approx \sin{i_\star}\cos{\lambda}$.
 
-Stellar inclination can often be constrained via, e.g., photometric and spectroscopic rotational modulation introduced by starspots, gravity darkening, and asteroseismology \citep[see][and references therein]{Albrecht22}.
-For systems without $i_\star$ measurements, it is still possible to infer their stellar obliquities from the sky-projected obliquities, assuming isotropic stellar inclinations.
-The inferred $\psi$ will have greater uncertainty than the one inferred with $i_\star$ measurement \citep{Fabrycky09}.
+Stellar inclinations can often be constrained via, e.g., photometric and spectroscopic rotational modulation introduced by starspots \citep[e.g.,][]{Masuda20, Albrecht21}, gravity darkening \citep[e.g.,][]{Barnes09, Barnes11}, and asteroseismology \citep[e.g.][]{Chaplin13}.
+For systems without $i_\star$ measurements, it is still possible to infer their stellar obliquities from the sky-projected obliquities, assuming isotropic stellar inclinations; however, the inferred $\psi$ will have greater uncertainty than the one inferred with $i_\star$ measurement \citep{Fabrycky09}.
 
-We do not yet understand the importance of stellar inclination in population stellar obliquity inference.
-In this work, we infer the stellar obliquity distribution from sky-projected stellar obliquities and show that the stellar obliquity distribution weakly depends on stellar inclinations.
-In Section~\ref{sec:hbm}, we introduce a flexible, hierarchical Bayesian framework to infer the population stellar obliquity distribution.
-In Section~\ref{sec:applications}, we apply the framework to simulated data and real observations.
-In Section~\ref{sec:jacobian}, we find the mathematical relation between the $\psi$, $\lambda$, and $i_\star$ distributions and explain why stellar inclination is not important in stellar obliquity distribution inference.
+We do not yet understand the role of stellar inclination in population stellar obliquity inference.
+In this work, we study the relationship between the distributions of stellar obliquity, sky-projected stellar obliquity, and stellar inclination, and introduce a statistical approach to infer the stellar obliquity distribution.
+In Section~\ref{sec:hbm}, we introduce a flexible, hierarchical Bayesian framework that infers the stellar obliquity distribution of a sample.
+In Section~\ref{sec:applications}, we apply the framework to simulated data and real observations and show that the stellar obliquity distribution inferred purely from sky-projected stellar obliquities will not compromise our interpretation.
+In Section~\ref{sec:jacobian}, we find the mathematical relation between the $\psi$, $\lambda$, and $i_\star$ distributions and explain why the stellar obliquity distribution inference is predominantly determined by sky-projected stellar obliquities.
 

### 2023-02-11 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 If the rotational modulation method is used to constrain the stellar inclination, the $i_{\star,n}$ and $\gamma_{\star,n}$ are constrained by rotational modulation in photometric or spectroscopic time series and observed sky-projected stellar rotational line broadening $\hat{v}\sin{i_{\star,n}}$ \citep[e.g.,][]{Masuda20}, where we have $v\sin{i_\star} = 2 \pi R_\star / P_{\rm rot}$.
-If the gravity-darkening method is used, the $i_{\star,n}$ could be constrained by the anomaly in transit light curves.
-If asteroseismology is used, the $i_{\star,n}$ could be constrained from the mode splitting in the Fourier transform of photometric time series.
+If the gravity-darkening method is used, the $i_{\star,n}$ could be constrained by the anomaly in transit light curves \citep{Barnes09}.
+If asteroseismology is used, the $i_{\star,n}$ could be constrained from the mode splitting in the Fourier transform of photometric time series \citep{Chaplin13}.
 
@@
 
-\section{Reproducibility}\label{sec:showyourwork}
-This study was carried out using the reproducibility software \href{https://github.com/showyourwork/showyourwork}{\showyourwork} \citep{Luger2021}, which leverages continuous integration to programmatically download the data from \href{https://zenodo.org/}{zenodo.org}, create the figures, and compile the manuscript.
+This study was carried out with the reproducible workflow \href{https://github.com/showyourwork/showyourwork}{\showyourwork} \citep{Luger2021}, which leverages continuous integration to programmatically download the data from \href{https://zenodo.org/}{zenodo.org}, create the figures, and compile the manuscript.
 Each figure caption contains one link to the script used to make the figure (at the commit corresponding to the current build of the manuscript).
 The git repository associated with this study is publicly available at \url{https://github.com/jiayindong/obliquity}.
-A Jupyter notebook with examples of the stellar obliquity distribution inference can be found in the same repository.
+A Jupyter notebook demonstration including case studies of the stellar obliquity distribution inference can be found in the same repository.
 

### 2023-02-14 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 In transiting-exoplanet observations, only the sky-projected stellar obliquity $\lambda$ can be measured.
-To find the true stellar obliquity $\psi$, information about stellar inclination $i_\star$ is also required.
-Here we show that, while the stellar inclination is important to infer the stellar obliquity of an individual system, they are not necessarily required for the population stellar obliquity inference. 
-Constraints on the stellar obliquity distribution are predominantly contributed by the sky-projected stellar obliquities.
-We introduce a flexible, hierarchical Bayesian framework that could infer the stellar obliquity distribution purely from sky-projected stellar obliquities.
-Stellar inclination measurements are optional in the model, and if not provided, are assumed to be isotropically distributed.
-Applying the framework to all exoplanetary systems with sky-projected stellar obliquity measurements, we find that the stellar obliquity distribution is unimodal and peaked at zero degrees.
-Misaligned systems have nearly isotropic stellar obliquities, which may have important implications for close-in planets' formation and evolution.
+To determine the true stellar obliquity $\psi$, information on the stellar inclination $i_\star$ is also needed.
+Here we find that while the stellar inclination is crucial for understanding the obliquity of individual systems, it is not a requirement for determining the population-level stellar obliquity. 
+The constraints on the stellar obliquity distribution are mainly driven by the sky-projected stellar obliquities.
+We introduce a flexible, hierarchical Bayesian framework that can infer the stellar obliquity distribution solely from sky-projected stellar obliquities.
+Stellar inclination measurements are optional input in the model, and if not available, they are assumed to be isotropically distributed.
+When applying the framework to all exoplanetary systems with sky-projected stellar obliquity measurements, we find that the stellar obliquity distribution is unimodal and peaked at zero degrees.
+Misaligned systems have nearly isotropic stellar obliquities, which could have important implications for the formation and evolution of close-in planets.
 
@@
 Stellar obliquity $\psi$ describes the angle between a planet's orbital axis $\bf{n}_{\rm orb}$ and its host star's spin axis $\bf{n}_{\star}$. 
-The angle is an important tracer of a planetary system's formation environment and dynamical evolution.
-The evolution of stellar obliquity can be roughly summarized into three stages.
-First, the formation and evolution of a protoplanetary disk around its host star determine the primordial stellar obliquity \citep[e.g.,][]{Bate10, Lai11, Batygin12}.
-Second, post-formation dynamical evolution in the planetary system, such as planet-planet scattering \citep[e.g.,][]{Rasio96, Chatterjee08, Nagasawa08, Beague12}, von Zeipel-Kozai-Lidov mechanisms \citep[e.g.,][]{Wu03, Naoz16}, and secular chaos \citep{Wu11}, excites the mutual inclinations between planetary or stellar companions, further sculpting the stellar obliquity.
-Lastly, tidal realignment of the host star's spin axis to the planet's orbital axis could reduce the stellar obliquity, if the tidal dissipation in the star is efficient \citep[e.g.,][]{Winn10, Albrecht12}.
-Furthermore, massive stars with convective cores could generate internal gravity waves and dissipate angular momentum to their radiative zones, modifying the stellar obliquity \citep{Rogers12, Rogers13}.
+This angle is an important tracer of a planetary system's formation environment and dynamical evolution.
+The evolution of stellar obliquity can be roughly broken down into three stages.
+First, the formation and evolution of a protoplanetary disk determine the primordial stellar obliquity \citep[e.g.,][]{Bate10, Lai11, Batygin12}.
+Second, post-formation dynamical evolution in the planetary system, such as planet-planet scattering \citep[e.g.,][]{Rasio96, Chatterjee08, Nagasawa08, Beague12}, von Zeipel-Kozai-Lidov mechanisms \citep[e.g.,][]{Wu03, Naoz16}, and secular chaos \citep{Wu11}, can excite the mutual inclinations between planetary or stellar companions and alter the stellar obliquity.
+Lastly, the tidal force can reduce the stellar obliquity by realigning the host star's spin axis with the planet's orbital axis, if the tidal dissipation in the star is efficient \citep[e.g.,][]{Winn10, Albrecht12}.
+Additionally, massive stars with convective cores could generate internal gravity waves and dissipate angular momentum to their radiative zones, potentially affecting the stellar obliquity \citep{Rogers12, Rogers13}.
 
@@
 These proposed mechanisms have different predictions on stellar obliquity distributions \citep[see][and references therein]{Albrecht22}.
-For example, in the secular chaos mechanism, the stellar obliquity could never get beyond $90\degr$ \citep{Teyssandier19}.
-In the stellar Kozai mechanism, the stellar obliquity distribution should be bimodal and concentrated at $40\degr$ and $140\degr$ \citep[e.g.,][]{Fabrycky07, Anderson16, Vick19}.
-In the multiple-planet scattering mechanism, the majority of systems should be aligned, and a small fraction of systems have a wide range of stellar obliquities \citep[e.g.,][]{Beague12}.
-Motivated by these predictions, we characterize the stellar obliquity distribution of exoplanetary systems from a Bayesian approach to find the dominant mechanisms that sculpt close-in planetary systems.
+For example, the secular chaos mechanism suggests that the stellar obliquity would never exceed $90\degr$ \citep[e.g.,][]{Teyssandier19}.
+The stellar Kozai mechanism predicts that the stellar obliquity distribution should be bimodal and concentrated at $40\degr$ and $140\degr$ \citep[e.g.,][]{Fabrycky07, Anderson16, Vick19}.
+The multiple-planet scattering mechanism proposes that the majority of systems would be aligned, while a small fraction of systems would exhibit a diverse range of stellar obliquities \citep[e.g.,][]{Beague12}.
+With these predictions in mind, we aim to determine the dominant mechanisms responsible for shaping close-in planetary systems by characterizing the stellar obliquity distribution of exoplanetary systems through a Bayesian approach.
 
-In exoplanetary observation, only the sky-projected stellar obliquity $\lambda$, the angle between the projections of $\bf{n}_{\rm orb}$ and $\bf{n}_{\star}$ onto the plane of the sky, can be measured, mainly via the Rossiter-McLaughlin effect \citep{Rossiter24, McLaughlin24}.
-The stellar obliquity $\psi$ can be calculated, if both the sky-projected stellar obliquity $\lambda$ and the stellar inclination $i_\star$ are known.
-The relation follows \citep[e.g.,][]{Fabrycky09}:
+In exoplanetary observation, only the sky-projected stellar obliquity $\lambda$, the angle between the projections of $\bf{n}_{\rm orb}$ and $\bf{n}_{\star}$ onto the plane of the sky, can be measured.
+This measurement is primarily obtained via the Rossiter-McLaughlin effect \citep{Rossiter24, McLaughlin24}.
+The stellar obliquity $\psi$ can be determined, if both the sky-projected stellar obliquity $\lambda$ and the stellar inclination $i_\star$ are known.
+The relationship between $\psi$ and $\{\lambda, i_\star\}$ is given by \citep[e.g.,][]{Fabrycky09}:
 \begin{equation}\label{eqn:psi}
@@
 where $i_{\rm orb}$ is the inclination angle between the vector $\bf{n}_{\rm orb}$ and the observer's line of sight, and $i_\star$ is the inclination angle between $\bf{n}_{\star}$ and the observer's line of sight.
-For transiting exoplanet systems where $i_{\rm orb} \approx 90\degr$, $\cos{\psi} \approx \sin{i_\star}\cos{\lambda}$.
+In transiting-exoplanet systems where $i_{\rm orb} \approx 90\degr$, $\cos{\psi} \approx \sin{i_\star}\cos{\lambda}$.
 
-Stellar inclinations can often be constrained via, e.g., photometric and spectroscopic rotational modulation introduced by starspots \citep[e.g.,][]{Masuda20, Albrecht21}, gravity darkening \citep[e.g.,][]{Barnes09, Barnes11}, and asteroseismology \citep[e.g.][]{Chaplin13}.
+Stellar inclinations can often be constrained via, e.g., photometric and spectroscopic rotational modulation introduced by starspots for cool stars \citep[e.g.,][]{Masuda20, Albrecht21}, gravity darkening for fast-rotating stars \citep[e.g.,][]{Barnes09, Barnes11}, and asteroseismology for bright stars \citep[e.g.][]{Chaplin13}.
 For systems without $i_\star$ measurements, it is still possible to infer their stellar obliquities from the sky-projected obliquities, assuming isotropic stellar inclinations; however, the inferred $\psi$ will have greater uncertainty than the one inferred with $i_\star$ measurement \citep{Fabrycky09}.
 
-We do not yet understand the role of stellar inclination in population stellar obliquity inference.
-In this work, we study the relationship between the distributions of stellar obliquity, sky-projected stellar obliquity, and stellar inclination, and introduce a statistical approach to infer the stellar obliquity distribution.
-In Section~\ref{sec:hbm}, we introduce a flexible, hierarchical Bayesian framework that infers the stellar obliquity distribution of a sample.
-In Section~\ref{sec:applications}, we apply the framework to simulated data and real observations and show that the stellar obliquity distribution inferred purely from sky-projected stellar obliquities will not compromise our interpretation.
-In Section~\ref{sec:jacobian}, we find the mathematical relation between the $\psi$, $\lambda$, and $i_\star$ distributions and explain why the stellar obliquity distribution inference is predominantly determined by sky-projected stellar obliquities.
+The relationship between the distributions of stellar obliquity, sky-projected stellar obliquity, and stellar inclination is still not fully understood.
+In this study, we aim to gain a deeper understanding of this relationship and develop a statistical approach to infer the stellar obliquity distribution.
+In Section~\ref{sec:hbm}, we introduce a flexible, hierarchical Bayesian framework that allows us to infer the stellar obliquity distribution of a sample.
+In Section~\ref{sec:applications}, we apply the framework to simulated data and real observations and show that the inferred stellar obliquity distribution from sky-projected stellar obliquities is robust even if the $i_\star$ information is not provided.
+In Section~\ref{sec:jacobian}, we further investigate the relationship between the distributions of $\psi$, $\lambda$, and $i_\star$, and explain why sky-projected stellar obliquities play a predominant role in determining the inferred distribution of stellar obliquities.
 
@@
 
-To find the stellar obliquity distribution of exoplanetary systems, we build a hierarchical Bayesian framework that takes the sky-projected stellar obliquity $\lambda$ and orbital inclination $i_{\rm orb}$ as observed data.
-Measurements on the stellar inclination $i_\star$ are optional.
-If the $i_\star$ information is not provided, we simply assume that they follow an isotropic distribution.
-Otherwise, the observed $i_\star$ could be taken from previous measurements, or if the rotational modulation method is applied, the $i_\star$ could be inferred from the stellar rotation period $P_{\rm rot}$, stellar radius $R_\star$, and sky-projected rotational broadening velocity $v\sin{i}_\star$.
+To find the stellar obliquity distribution of exoplanetary systems, we develop a hierarchical Bayesian framework that takes the observed sky-projected stellar obliquity $\lambda$ and orbital inclination $i_{\rm orb}$ as input data.
+The measurement of the stellar inclination $i_\star$ is optional.
+In the absence of $i_\star$ information, we assume that it follows an isotropic distribution.
+Otherwise, the observed $i_\star$ can be taken from previous measurements, or if the rotational modulation method is employed, the $i_\star$ value can be inferred from the stellar rotation period $P_{\rm rot}$, stellar radius $R_\star$, and sky-projected rotational broadening velocity $v\sin{i}_\star$.
 
@@
     \caption{Hierarchical Bayesian framework to infer the stellar obliquity distribution of exoplanetary systems.
-The stellar obliquity distribution is described by hyperparameters $\vb*{\beta}$ and constrained by $n_{\rm th}$ system's stellar obliquity $\psi_n$.
-Each $\psi_n$ is calculated by the system's sky-projected stellar obliquity $\lambda_n$, orbital inclination $i_{{\rm orb}, n}$, and stellar inclination $i_{\star, n}$, if available.
-$\gamma_{\star, n}$ contains properties of the star other than inclination, such as stellar radius and rotation period for $i_{\star, n}$ inference.
-Obs$_{\star,n}$ includes all the observed properties of the star with uncertainties, $\hat{\lambda}_n$ has the measured sky-projected stellar obliquity and its uncertainty, and $\hat{i}_{{\rm orb}, n}$ has measured orbital inclination and its uncertainty.}
+The distribution is described by hyperparameters $\vb*{\beta}$ and is constrained by the sky-projected stellar obliquity $\lambda_n$, orbital inclination $i_{{\rm orb}, n}$, and, if available, the stellar inclination $i_{\star, n}$ of each $n_{\rm th}$ individual system.
+The system's stellar obliquity $\psi_n$ is calculated based on these parameters.
+    Properties of the star other than its inclination, such as its radius and rotation period, are represented by $\gamma_{\star, n}$. 
+    Obs$_{\star,n}$ contains all observed properties of the star with uncertainties, $\hat{\lambda}_n$ has the measured sky-projected stellar obliquity and its uncertainty, and $\hat{i}_{{\rm orb}, n}$ has measured orbital inclination and its uncertainty.}
     \label{fig:graph}
@@
 
-Figure~\ref{fig:graph} shows the probabilistic graphical model for the hierarchical Bayesian framework.
-We aim to constrain a set of hyperparameters $\bm{\beta}$ that describes the stellar obliquity $\psi$ distribution.
-The parameter set $\bm{\beta}$ is constrained by $N$ individual systems in which $\psi_n$ is a random variable.
-$\psi_n$ is computed from the sky-projected stellar obliquity $\lambda_n$, the orbital inclination $i_{{\rm orb},n}$, and the stellar inclination $i_{\star, n}$.
-The parameter $\gamma_{\star, n}$ is a parameter set that contains all the stellar properties other than $i_{\star,n}$, such as the stellar rotation period $P_{{\rm rot},n}$, radius $R_{\star,n}$, and projected rotational velocity $v\sin{i_{\star,n}}$, if they are available.
+Figure~\ref{fig:graph} illustrates the probabilistic graphical model for our hierarchical Bayesian framework.
+We aim to constrain a set of hyperparameters $\bm{\beta}$ that describes the stellar obliquity distribution.
+The parameter set $\bm{\beta}$ is constrained by $N$ individual systems, where each $\psi_n$ is computed from the system's sky-projected stellar obliquity $\lambda_n$, the orbital inclination $i_{{\rm orb},n}$, and if available, the stellar inclination $i_{\star, n}$.
+The parameter $\gamma_{\star, n}$ contains all stellar properties other than $i_{\star,n}$, such as the stellar rotation period $P_{{\rm rot},n}$, radius $R_{\star,n}$, and projected rotational velocity $v\sin{i_{\star,n}}$, if they are known.
 
 The sky-projected stellar obliquity $\lambda_n$ is constrained by observed $\hat{\lambda}_n$ via the Rossiter-McLaughlin effect or gravity darkening.
-The orbital inclination $i_{{\rm orb},n}$ is constrained by the transit light curves of exoplanets.
+The orbital inclination $i_{{\rm orb},n}$ of exoplanets is inferred from their transit light curves.
 Obs$_{\star,n}$ includes observed data that helps to constrain $i_{\star,n}$ and $\gamma_{\star, n}$.
@@
 For the ${\rm Beta}(a_1,b_1)$ component, we have it to describe the misaligned systems.
-We design the order of the two components to avoid label switching in the mixture model.
+We design the order of the two components to prevent label switching in the mixture model.
 With the design, the hyperparameter and parameter priors are the following:
@@
 \end{align}
-If $\gamma_{\star, n}$ is available, we construct Normal distributions with means and standard deviations from their observed data.
-Here we assume uniform hyperpriors for the Beta distribution.
-For population inference on a small number of systems (i.e., $N \lesssim 50$), the choice of hyperpriors could impact the inferred distribution \citep{Nagpal22}.
-When applying the framework to a small sample size, it is important to test the robustness of inferred distributions on different hyperpriors.
+If $\gamma_{\star, n}$ is available, we construct Normal distributions with means and standard deviations from measurements.
+Here we use uniform hyperpriors for the Beta distribution.
+However, it is important to note that when applying the framework to a small sample size with $N \lesssim 50$, the choice of hyperpriors could impact the inferred distribution \citep{Nagpal22}.
+To ensure the robustness of the inferred distributions in such cases, it is crucial to test their sensitivity to different hyperpriors.
 
-The probabilistic model is built with the $\mathtt{PyMC}$ \citep[$\mathtt{v4.1.7}$;][]{pymc} package and the posteriors are sampled using the No-U-Turn Sampler \citep[NUTS;][]{Hoffman11}, a gradient-based sampling algorithm of the Markov chain Monte Carlo (MCMC).
-All figures and simulations in this paper are fully reproducible and built with the \showyourwork package.
-The open source code can be found on \href{https://github.com/jiayindong/obliquity}{GitHub\,\faGithub\,(https://github.com/jiayindong/obliquity)}.
+The probabilistic model is constructed using the $\mathtt{PyMC}$ package version $\mathtt{v4.1.7}$ \citep{pymc}, and the posteriors are sampled with the No-U-Turn Sampler \citep[NUTS;][]{Hoffman11}, which is a gradient-based Markov chain Monte Carlo (MCMC) sampling algorithm.
+This paper's figures and simulations are completely reproducible and were created using the \showyourwork package.
+The open-source code is available on \href{https://github.com/jiayindong/obliquity}{GitHub\,\faGithub\,(https://github.com/jiayindong/obliquity)}.
 
-\section{Applications to Simulated and Observational Data}\label{sec:applications}
+\section{Applications to Simulated and Observed Data}\label{sec:applications}
 
@@
 We simply assume an isotropic distribution of $i_\star$ for these models.
-Surprisingly, in all four distributions, the inferred $\cos{\psi}$ distributions with or without $i_\star$ correctly recover the injected distributions shown as grey dashed lines as shown in Figure~\ref{fig:simulation}.
-The inferred modes and widths of the stellar obliquity distributions with or without $i_\star$ likelihood are consistent, despite that the inferred distributions without $i_\star$ measurements could have wider uncertainties shown in shallow color contours.
-Since the injected distributions are Normal distributions in rows 2--4, it is expected that the inferred distributions, which as Beta distributions, could not exactly describe the injected distributions.
+Surprisingly, as shown in Figure~\ref{fig:simulation}, the inferred distributions of $\cos{\psi}$, whether or not they include the information of $i_\star$, correctly recover the injected distributions (represented by grey dashed lines) in all four cases.
+Despite the fact that the inferred distributions without $i_\star$ measurements may have wider uncertainties, as indicated by shallow color contours, the modes and widths of the inferred stellar obliquity distributions are consistent with or without $i_\star$ likelihood.
+Since the injected distributions in rows 2-4 are Normal distributions, it should not be surprising that the inferred distributions, which are Beta distributions, may not exactly match the injected distributions.
 
-We also examine the importance of orbital inclination $i_{\rm orb}$ to the stellar obliquity distribution inference.
-Since all the systems we study are transiting-planet systems, we consider an isotropic orbital inclination distribution between $80\degr$ and $90\degr$.
-This broad inclination range corresponds to an impact parameter range from 0 to 1 with a planet-star separation $a/R_\star$ of 6.
-We compare the inferred stellar obliquity distribution by approximating $i_{\rm orb}$ to $90\degr$ with the inferred distribution using the true $i_{\rm orb}$, and find the difference between the two distributions is minimal.
-For transiting-planet systems, approximating orbital inclinations to $90\degr$ will not compromise the stellar obliquity distribution inference.
+We also examine the role of orbital inclination $i_{\rm orb}$ in the stellar obliquity distribution inference.
+Since our study focuses on transiting-exoplanet systems, we consider an isotropic orbital inclination distribution between $80\degr$ and $90\degr$.
+This broad range of inclinations corresponds to an impact parameter range from 0 to 1 with a planet-star separation $a/R_\star$ of 6.
+We compare the stellar obliquity distributions obtained by approximating $i_{\rm orb}$ to $90\degr$ with the distributions obtained using the actual $i_{\rm orb}$.
+We find the difference between the two distributions is negligible.
+This suggests that for transiting-exoplanet systems, approximating orbital inclinations as $90\degr$ will not compromise the stellar obliquity distribution inference.
 
-From the simulated data, we confirm that while stellar inclinations help to constrain the stellar obliquity of individual systems, they are not necessarily required for the population stellar obliquity inference.
-The constraint on stellar obliquity distribution is predominantly based on sky-projected stellar obliquities.
-Just from the sky-projected stellar obliquity, we could correctly infer the peak and width of the underlying stellar obliquity distribution.
-Later in Section~\ref{sec:jacobian}, we derive the Jacobian transformations between $\psi$, $\lambda$, and $i_\star$ to explain why there is no significant information loss when not having $i_\star$ measurements.
+We demonstrate through simulations that the inferred stellar obliquity distribution is robust even if the $i_\star$ information is not provided, and the $i_\star$ measurement only improves the constraint on the stellar obliquity distribution.
+Later in Section~\ref{sec:jacobian}, we derive the Jacobian transformations between $\psi$, $\lambda$, and $i_\star$ to understand why sky-projected stellar obliquities play a predominant role in determining the inferred distribution of stellar obliquities.
 
@@
     \includegraphics[width=\linewidth]{figures/simulation.pdf}
-    \caption{Inferred stellar obliquity distributions from sky-projected stellar obliquities with or without stellar inclination measurements.
-Each row presents a set of simulated data where the underlying $\cos{\psi}$ distribution is known and shown as grey dashed curves.
-The inferred $\cos{\psi}$ distributions with or without stellar inclination measurements are shown in orange and blue curves, respectively.
-The shallow contours are 1-$\sigma$ and 2-$\sigma$ uncertainties of the inferred distributions.}
+    \caption{Inferred stellar obliquity distributions from sky-projected stellar obliquities with and without information on the stellar inclination, depicted in blue and orange curves, respectively.
+Each row presents a set of simulated data with the true distribution of $\cos{\psi}$ indicated by grey dashed curves.
+The shallow contours represent the 1-$\sigma$ and 2-$\sigma$ uncertainties of the inferred distributions.}
     \label{fig:simulation}
@@
 
-We next apply the hierarchical Bayesian framework to all exoplanetary systems with sky-projected stellar obliquity $\lambda$ measurements.
-We use the sample of 161 systems summarized in \cite{Albrecht22} Table A1, which are mainly Hot Jupiter systems.
+We next apply our hierarchical Bayesian framework to a sample of 161 exoplanetary systems with sky-projected stellar obliquity measurements, primarily consisting of Hot Jupiter systems, as summarized in \cite{Albrecht22} Table A1.
 The inferred $\cos{\psi}$ distribution is shown in Figure~\ref{fig:psi_dist}.
-The $\cos{\psi}$ distribution peaks at 1 but is nearly flat between $-0.75$ and $0.75$ with no significant clustering.
-The distribution corresponds to a pile-up of planets with stellar obliquity less than $40\degr$ and an isotropic distribution of planets with stellar obliquity from $40\degr$ to $140\degr$.
-The aligned-system population dominates the distribution with a fraction of $w_0 = 0.67 \pm 0.09$. The hyperposteriors of the Beta distribution are $a_0 = 31.0\pm12.8$ and $b_0 = 0.40\pm0.12$.
-The misaligned-system population has a fraction of $w_1 = 0.33 \pm 0.09$ and hyperposteriors of $a_1 = 1.56\pm0.93$ and $b_1 = 1.64\pm1.58$.
+The $\cos{\psi}$ distribution is peaked at 1, with nearly flat behavior between $-0.75$ and $0.75$ and no significant clustering.
+The distribution suggests that there is a pile-up of planetary systems with stellar obliquities less than $40^\circ$ and an isotropic distribution for obliquities between $40^\circ$ and $140^\circ$.
+The fraction of aligned systems dominates the distribution, with $w_0 = 0.67 \pm 0.09$.
+The corresponding hyperposteriors of the Beta distribution are $a_0 = 31.0\pm12.8$ and $b_0 = 0.40\pm0.12$.
+On the other hand, the fraction of misaligned systems is estimated to be $w_1 = 0.33 \pm 0.09$, with hyperposteriors of $a_1 = 1.56\pm0.93$ and $b_1 = 1.64\pm1.58$.
 
@@
         \includegraphics{figures/psi_dist.pdf}
-        \caption{Stellar obliquity distribution of all exoplanetary systems with sky-projected stellar obliquity measurements.
-The inference is purely from observed sky-projected stellar obliquities and assumes the stellar inclination is isotropic.}
+        \caption{Inferred stellar obliquity distribution for all exoplanetary systems with sky-projected stellar obliquity measurements.
+This inference is based purely on the observed sky-projected obliquities and assumes an isotropic distribution for the stellar inclinations.}
         \label{fig:psi_dist}
@@
 
-The inferred stellar obliquity distribution of the full sample of systems is at odds with the previous analysis of the subsample of systems that also have $i_\star$ measurements.
-The subsample identified a preponderance of perpendicular planets that disfavored an isotropic stellar obliquity distribution \citep{Albrecht21}.
-We speculate two possible explanations for this observation discrepancy, which are open to further investigation.
-First, the sample of systems with $i_\star$ measurements is small, and only about 20 systems are misaligned systems used to constrain the misaligned systems' stellar obliquity distribution.
-Second, the sample could have a selection bias introduced by the $i_\star$ measurement requirement.
+The discrepancy between the inferred stellar obliquity distribution of the full sample of exoplanetary systems and the previous analysis of the subsample with $i_\star$ measurements warrants further investigation.
+The earlier study identified a preponderance of perpendicular planets and disfavored an isotropic stellar obliquity distribution \citep{Albrecht21}.
+There are two potential explanations for this difference: (1) the subsample with $i_\star$ measurements is small and only includes about 20 misaligned systems, and (2) the requirement for $i_\star$ measurements could introduce selection biases in the sample.
 
-The stellar obliquity distribution suggests that $67\pm9$\% of the systems have stellar obliquity less than $40\degr$, and $33\pm9$\% of the systems have nearly isotropically distributed stellar obliquities spread from $40\degr$ to $140\degr$.
-The distribution could have important implications for the formation and evolution of close-in planetary systems.
-The broad distribution of misaligned systems is in good agreement with the outcome of the scattering of multiple giant planets after a convergent disk migration \citep[e.g.,][]{Nagasawa11, Beague12}, one of the proposed mechanisms that form Hot Jupiters.
-The intriguing result should be further examined on more carefully selected Hot Jupiter samples and put constraints on Hot Jupiter origin channels.
+The inferred stellar obliquity distribution indicates that approximately $67\pm9$\% of the systems have a stellar obliquity of less than $40\degr$, and approximately $33\pm9$\% of the systems follow a nearly isotropic stellar obliquity distribution ranging from $\sim$$40\degr$ to $\sim$$140\degr$.
+These findings could have significant implications for the formation and evolution of close-in planetary systems.
+The broad distribution of misaligned systems is in good agreement with the predicted outcome of multiple giant planets scattering after a convergent disk migration, as proposed by various studies, such as \cite{Nagasawa11} and \cite{Beague12}.
+The intriguing result should be further examined with a more carefully selected sample of Hot Jupiters and provides opportunities to place constraints on their origin channels of Hot Jupiters.
 
@@
 
-In this section, we aim to understand why the stellar obliquity distribution is predominantly determined by the sky-projected stellar obliquity distribution and why the stellar inclination distribution is less important to the inference using mathematical expressions.
+In this section, we aim to gain insight into the reasons behind the predominant role of the sky-projected stellar obliquity distribution and the less significant impact of the stellar inclination distribution in the inference of the stellar obliquity distribution.
 
@@
     \vspace*{-1cm}
-    \caption{Two coordinate systems to describe the stellar spin axis $\bf{n}_{\star}$ and the planet's orbital axis $\bf{n}_{\rm orb}$.
-Here we define the observer's line of sight as one of the two horizontal axes (conventional $x$-axis in Cartesian), and the orbital axis of the planet as the vertical axis (conventional $z$-axis in Cartesian).
+    \caption{Two coordinate systems that describe the stellar spin axis $\bf{n}_{\star}$ and the planet's orbital axis $\bf{n}_{\rm orb}$.
+Here we define the observer's line of sight as one of the horizontal axes (represented as the conventional $x$-axis in Cartesian coordinates), and the orbital axis of the planet as the vertical axis (represented as the conventional $z$-axis in Cartesian coordinates).
 We approximate the orbital inclination of the planet to $90\degr$.}
@@
 
-In Figure~\ref{fig:coord}, we build two coordinate systems to describe the stellar spin axis $\bf{n}_{\star}$ for a given orbital axis $\bf{n}_{\rm orb}$.
-In both coordinates, we set the observer's line of sight to be one of the two horizontal axes, and the orbital axis of the planet to be the vertical axis.
-We approximate the orbital inclination of the transiting planet to $90\degr$ here to simplify the problem for a population study.
+In Figure~\ref{fig:coord}, we present two coordinate systems that describe the stellar spin axis $\bf{n}_{\star}$ for a given orbital axis $\bf{n}_{\rm orb}$.
+In both coordinates, we set the observer's line of sight to be one of the horizontal axes (represented as the conventional $x$-axis in Cartesian coordinates), and the orbital axis of the planet to be the vertical axis (represented as the conventional $z$-axis in Cartesian coordinates).
+To simplify the problem, the orbital inclination of the transiting planet is assumed to be $90\degr$ in this illustration.
 As discussed in Section~\ref{subsec:sims}, the assumption will not compromise the stellar obliquity distribution inference for transiting planets.
@@
 $\lambda$ is the sky-projected stellar obliquity and $i_\star$ is the stellar inclination.
-If $\bf{n}_{\star}$ is isotropically distributed, $p_{i_\star} \sim \sin{i_\star}$ and $p_\lambda \sim 1/2\pi$.
+If $\bf{n}_{\star}$ is isotropically distributed, $p_{i_\star} \sim \sin{i_\star}$ and $p_\lambda \sim \mathcal{U}(-\pi, \pi)$.
 Again, since $p_{\cos{i_\star}} = p_{i_\star} \abs{\dv*{i_\star}{\cos{i_\star}}}$, $p_{\cos{i_\star}}$ is uniformly distributed between $-1$ and $1$.
-Because of the observational degeneracy between $i_\star$ and $180\degr - i_\star$, the convention is to have $0 \leq i_\star \leq 90\degr$, and thus $p_{\cos{i_\star}}$ is uniformly distributed between $0$ and $1$ (i.e., $p_{\cos{i_\star}} \sim 1$).
-This also limits $\theta$ to $[-\pi/2, \pi/2]$.
-In observation, there is also a degeneracy between $-\lambda$ and $+\lambda$. Therefore, we also limit $\lambda$ to $[0,\pi]$.
+Because of the observational degeneracy between $i_\star$ and $180\degr - i_\star$, the convention is to have $0 \leqslant i_\star \leqslant 90\degr$, and thus $p_{\cos{i_\star}}$ is uniformly distributed between $0$ and $1$ and $p_{\cos{i_\star}} \sim \mathcal{U}(0, 1)$.
+This also limits $p_\theta \sim \mathcal{U}(-\pi/2, \pi/2)$.
 
@@
 \end{equation}
-Replacing $\sin{i_\star}$ in Equation~(\ref{eq:coord3}) with Equation~(\ref{eq:coord1}), we find $\cos{\lambda} = \cos{\psi}/\sqrt{1-(1-\cos^2{\psi})\cos^2{\theta}}$.
-Reorganize the equation, we get $\cos^2{\psi} = \frac{\cos^2{\lambda}\cos^2{\theta}-\cos^2{\lambda}}{\cos^2{\lambda}\cos^2{\theta}-1}$.
-The derivative gives $\abs{\pdv{\cos{\psi}}{\cos{\lambda}}} = \frac{(1-\cos^2{\theta})^{1/2}}{(1-\cos^2{\theta}\cos^2{\lambda})^{3/2}}$.
-If $\theta$ is uniformly distributed between $-\pi/2$ and $\pi/2$, $p_{\cos{\theta}} = p_{\theta} \abs{\dv*{\theta}{\cos{\theta}}} = 1/\pi/(1-\cos^2{\theta)^{1/2}}$.
+Replacing $\sin{i_\star}$ in Equation~(\ref{eq:coord3}) using Equation~(\ref{eq:coord1}), we find $\cos{\lambda} = \cos{\psi}/\sqrt{1-(1-\cos^2{\psi})\cos^2{\theta}}$.
+Reorganize the equation, we get $\cos^2{\psi} = \frac{\cos^2{\lambda}\cos^2{\theta}-\cos^2{\lambda}}{\cos^2{\lambda}\cos^2{\theta}-1}$ and
+the partial derivative $\abs{\pdv{\cos{\psi}}{\cos{\lambda}}} = \frac{(1-\cos^2{\theta})^{1/2}}{(1-\cos^2{\theta}\cos^2{\lambda})^{3/2}}$.
+Assume $\theta$ is uniformly distributed between $-\pi/2$ and $\pi/2$, $p_{\cos{\theta}} = p_{\theta} \abs{\dv*{\theta}{\cos{\theta}}} = 2/\pi/(1-\cos^2{\theta)^{1/2}}$, where the factor of 2 is due to two solutions of $\theta$ to the $\cos{\theta}$.
 Putting all the parts together, we get
@@
 \end{equation}
-If $\cos{\psi}$ is uniformly distribution, i.e., $p_{\cos{\psi}} = 1/2$, Equation~(\ref{eqn:jac_lam}) becomes $p_{\cos{\lambda}} = 1/\pi/\sqrt{1-\cos^2{\lambda}}$, which is equivalent to $p_\lambda = 1/\pi$.
-The $\lambda$ is uniformly distributed if an isotropic $\psi$.
-If $\cos{\psi}$ follows other distributions, $p_\lambda$ could be found by evaluating Equation~(\ref{eqn:jac_lam}) numerically.
+If $\cos{\psi}$ is uniformly distribution, i.e., $p_{\cos{\psi}} \sim 1/2$, Equation~(\ref{eqn:jac_lam}) becomes $p_{\cos{\lambda}} \sim 1/\pi/\sqrt{1-\cos^2{\lambda}}$, which is equivalent to $p_\lambda \sim \mathcal{U}(-\pi, \pi)$.
+This suggests $\lambda$ is uniformly distributed for an isotropic $\psi$ distribution.
+For different $\cos{\psi}$ distributions, $p_\lambda$ can be found by evaluating Equation~(\ref{eqn:jac_lam}).
 
@@
 From Equation~(\ref{eq:coord1}), we have $\sin{\psi} = \cos{i_\star}/\cos{\theta}$ and $\abs{\pdv{\sin{\psi}}{\cos{i_\star}}} = 1/\cos{\theta}$.
-Again, we assume $\theta$ is uniformly distributed, and $p_{\cos{\theta}} = 1/\pi/(1-\cos^2{\theta)^{1/2}}$.
-Lastly, we transform the $p_{\sin{\psi}}$ to $p_{\cos{\psi}}$, where $p_{\sin{\psi}} = p_{\cos{\psi}}\sqrt{1-\cos^2{\psi}}/\cos{\psi}$.
+Again, we assume $\theta$ is uniformly distributed, and this gives $p_{\cos{\theta}} = 2/\pi/(1-\cos^2{\theta)^{1/2}}$.
+Lastly, we transform the $p_{\sin{\psi}}$ to $p_{\cos{\psi}}$, where $p_{\sin{\psi}} = 2p_{\cos{\psi}}\sin{\psi}/\sqrt{1-\sin^2{\psi}}$ and the factor of 2 is from two solutions of $\cos{\psi}$ to the $\cos^2{\psi} = 1-\sin^2{\psi}$.
 Combining all the pieces together, we get
@@
 \end{equation}
-If $\cos{\psi}$ is uniformly distribution, i.e., $p_{\cos{\psi}} = 1/2$, the integral gives $1$, which suggests the $\cos{i_\star}$ probability density function is uniform.
+Note that the lower limit of the integral is $\cos{i_\star}$ instead of 0 since $\lvert \cos{\theta}/\cos{i_\star} \rvert \geqslant 1$.
+If $\cos{\psi}$ is uniformly distribution, i.e., $p_{\cos{\psi}} \sim 1/2$, the integral gives $1$, which suggests the $\cos{i_\star}$ is uniformly distributed, as expected.
 
-From Equation~(\ref{eqn:jac_lam}) and (\ref{eqn:jac_istar}), we can now derive the $\lambda$ and $i_\star$ distributions for any given $\psi$ distributions, assuming the azimuthal angle of the stellar spin axis $\theta$ is random.
-In Figure~\ref{fig:transform}, we find and examine numerical solutions of the $\lambda$ and $i_\star$ distributions for four different $\cos{\psi}$ distributions.
-The first row in Figure~\ref{fig:transform} presents an isotropic $\psi$ where $\cos{\psi} \sim \mathcal{U}(-1,1)$, and the second to fourth rows in Figure~\ref{fig:transform} present Normal distributions of $\cos{\psi}$ following $\mathcal{N}(0,0.2)$, $\mathcal{N}(-0.4,0.2)$, and $\mathcal{N}(0.4,0.2)$, respectively.
-In each row, we solve the $\lambda$ and $i_\star$ distributions numerically as shown in blue curves, and samplings of $\lambda$ and $i_\star$ from $\cos{\psi}$ and $\theta$ distributions as shown in grey histograms.
-For a uniform $\cos{\psi}$ distribution, the $\lambda$ distribution is uniform, and the $i_\star$ distribution is isotropic and proportional to $\sin{i_\star}$, as expected.
+Using Equation~(\ref{eqn:jac_lam}) and (\ref{eqn:jac_istar}), we can now derive the $\lambda$ and $i_\star$ distributions for any given $\psi$ distributions, assuming the azimuthal angle of the stellar spin axis $\theta$ is random.
+In Figure~\ref{fig:transform}, we present numerical solutions of the $\lambda$ and $i_\star$ distributions for four different $\cos{\psi}$ distributions.
+The top row of Figure~\ref{fig:transform} shows an isotropic $\psi$ distribution, where $\cos{\psi} \sim \mathcal{U}(-1,1)$.
+The second, third, and fourth rows of Figure~\ref{fig:transform} present Normal distributions of $\cos{\psi}$ following $\mathcal{N}(0,0.2)$, $\mathcal{N}(-0.4,0.2)$, and $\mathcal{N}(0.4,0.2)$, respectively.
+The blue curves in each row show the numerical solutions of the $\lambda$ and $i_\star$ distributions, while the grey histograms show the sampling of $\lambda$ and $i_\star$ from the $\cos{\psi}$ and $\theta$ distributions.
+For a uniform $\cos{\psi}$ distribution, the $\lambda$ distribution is uniform, and the $i_\star$ distribution is isotropic, proportional to $\sin{i_\star}$, as expected.
 
-Interestingly, for a given stellar obliquity distribution, the $\lambda$ distribution is sensitive to and closely related to the underlying $\psi$ distribution, as shown in Figure~\ref{fig:transform}.
-For different stellar obliquity distributions, the $\lambda$ distributions are distinguishable from each other, and thus, when we infer the $\psi$ distribution from the $\lambda$ distribution, the solution will not degenerate.
-On the other hand, the $i_\star$ distributions less depend on the underlying $\psi$ distribution.
-Compared to an isotropic $i_\star$ distribution, the $i_\star$ distributions for different $\psi$ distributions differ at the low values (i.e., $i_\star < \pi/4$), which places a challenge to observation.
-More importantly, since the $i_\star$ distributions are less sensitive to the underlying $\psi$ distributions when we infer the $\psi$ distributions from the $i_\star$ distribution, the solution could be very degenerate.
+Interestingly, the $\lambda$ distribution is closely related to and sensitive to the underlying $\psi$ distribution, as demonstrated in the \nth{1} and \nth{2} columns in Figure~\ref{fig:transform}.
+For different stellar obliquities, the $\lambda$ distributions are distinguishable, making it possible to infer the $\psi$ distribution from the $\lambda$ distribution.
+On the other hand, the $i_\star$ distributions are less dependent on the underlying $\psi$ distribution.
+Compared to an isotropic $i_\star$ distribution, the curvature of the $i_\star$ distributions for different $\psi$ distributions differ the most at the low $i_\star$ values (i.e., $i_\star < \pi/4$), which places a challenge to observational detections.
+Additionally, the degeneracy of the solution could be a significant issue when attempting to infer the $\psi$ distribution from the $i_\star$ distribution.
+For example, when $\cos{\psi} \sim \mathcal{N}(-0.4,0.2)$ or $\cos{\psi} \sim \mathcal{N}(0.4,0.2)$, two $i_\star$ distributions are exactly the same.
 
-Due to the lack of dependency on the $i_\star$ distribution, the $\psi$ distribution could be inferred purely from the $\lambda$ distribution without losing much information.
-Although we could find a relation between $\psi$, $\lambda$, and $\theta$, the $\psi$ distribution cannot be inferred from the $\lambda$ and $\theta$ distributions since $\lambda$ and $\theta$ are not independent variables.
-We still need to infer the $\psi$ distribution from the $\lambda$ and $i_\star$ distributions.
-In the \nth{4} column in Figure~\ref{fig:transform}, we find the $\psi$ distribution from the $\lambda$ distribution, simply assuming an isotropic $i_\star$ distribution.
-The method could correctly recover the peaks and widths of the underlying $\psi$ distributions (presented as blue dashed lines) as shown in grey histograms.
+The $\psi$ distribution can be inferred from the $\lambda$ distribution without loss of information due to the strong dependency of the $\lambda$ distribution on the $\psi$ distribution.
+It is also worthwhile to note that although we could find a mathematical expression of $\psi$ with $\lambda$ and $\theta$, the $\psi$ distribution cannot be inferred from the two variables since they are not independent variables.
+
+In the \nth{4} column in Figure~\ref{fig:transform}, the $\psi$ distribution is found from the $\lambda$ distribution by assuming an isotropic $i_\star$ distribution.
+The method could accurately recover the peaks and widths of the underlying $\psi$ distributions (represented by blue dashed lines) shown in grey histograms.
 The method over-predicts $\cos{\psi}$ at $\pm 1$ because of the over-prediction of $i_\star$ near $90\degr$.
-However, this minor deviation from the true distribution has little impact on the population inference because of hyperparameters' limited degree of freedom.
-Comparing the blue curves with the orange curves in Figure~\ref{fig:simulation}, the overestimation of $\cos{\psi}$ near $\pm1$ under the assumption of an isotropic $i_\star$ has little impact on the inference since the distribution is mainly constrained by the peak and width of the stellar obliquity distribution, which are both correctly predicted by the isotropic $i_\star$ model.
+However, this small deviation from the true distribution has a limited impact on population inference because of the hyperparameters have limited degrees of freedom.
+By comparing the blue curves with the orange curves in Figure~\ref{fig:simulation}, it can be seen that the overestimation of $\cos{\psi}$ near $\pm1$ under the assumption of isotropic $i_\star$ has little effect on the overall inference.
+The distribution is primarily determined by the peak and width of the stellar obliquity distribution, both of which are correctly predicted by the isotropic $i_\star$ model.
 
@@
     \includegraphics[width=\linewidth]{figures/transform.pdf}
-    \caption{Simulated $\cos{\psi}$ distributions (\nth{1} column) and their corresponding sky-projected stellar obliquity $\lambda$ (\nth{2} column) and stellar inclination $i_\star$ (\nth{3} column) distributions.
+    \caption{Simulated $\cos{\psi}$ distributions (\nth{1} column) and the corresponding distributions of sky-projected stellar obliquity $\lambda$ (\nth{2} column) and stellar inclination $i_\star$ (\nth{3} column).
 The inferred $\cos{\psi}$ distributions assuming isotropic stellar inclinations are shown in the \nth{4} column.
-The random samplings of $\lambda$ and $i_\star$ out of the $\cos{\psi}$ distributions are shown as grey histograms, and the numerical solutions are shown as blue curves.}
+The grey histograms present the random samplings of $\lambda$ and $i_\star$ from the $\cos{\psi}$ distributions, and the blue curves present the numerical solutions.}
     \label{fig:transform}
@@
 
-In this work, we demonstrated that the stellar obliquity distribution inference is predominantly determined by the sky-projected stellar obliquities.
-Without stellar inclination measurements, the stellar obliquity distribution could still be correctly inferred.
-We introduce a flexible, hierarchical Bayesian framework to infer the stellar obliquity distribution of transiting-planet systems.
-Stellar inclination measurement is optional, and if it is not provided, we simply assume an isotropic stellar inclination distribution.
-The hierarchical Bayesian model is an open-source code that can be found on \href{https://github.com/jiayindong/obliquity}{GitHub\,\faGithub\,(https://github.com/jiayindong/obliquity)}.
-Users may apply the framework to a certain sample of targets and customize the model for different stellar obliquity distributions and priors.
-When jointly modeling the stellar obliquity distribution from one sample with stellar inclination measurements and one without, it is important to check if the $i_\star$ sample is representative of the full sample.
-If not, since the $i_\star$ sample will constrain stellar obliquities more tightly, the overall distribution could be weighted towards these systems, bias the interpretation.
-Lastly, we applied the framework to all exoplanetary systems with sky-projected stellar obliquity measurements and found that $67\pm9$\% of the systems have stellar obliquity less than $40\degr$, and $33\pm9$\% of the systems have nearly isotropically distributed obliquities between $40\degr$ and $140\degr$.
-The distribution could have important implications for the formation and evolution of close-in planetary systems and be worthwhile for further investigation.
+In this work, we demonstrated that the stellar obliquity distribution could be robustly inferred from sky-projected stellar obliquities purely.
+We introduced a flexible, hierarchical Bayesian framework for the stellar obliquity distribution inference.
+Stellar inclination measurements are optional input in the model, and if not available, they are assumed to be isotropically distributed.
+Our open-source hierarchical Bayesian model, available on \href{https://github.com/jiayindong/obliquity}{GitHub\,\faGithub\,(https://github.com/jiayindong/obliquity)}, can be customized to different stellar obliquity distributions and priors for specific target samples.
 
-This study was carried out with the reproducible workflow \href{https://github.com/showyourwork/showyourwork}{\showyourwork} \citep{Luger2021}, which leverages continuous integration to programmatically download the data from \href{https://zenodo.org/}{zenodo.org}, create the figures, and compile the manuscript.
-Each figure caption contains one link to the script used to make the figure (at the commit corresponding to the current build of the manuscript).
-The git repository associated with this study is publicly available at \url{https://github.com/jiayindong/obliquity}.
-A Jupyter notebook demonstration including case studies of the stellar obliquity distribution inference can be found in the same repository.
+It is crucial to consider the representativeness of the $i_\star$ sample when jointly modeling the stellar obliquity distribution from two data sets, one with and one without $i_\star$ measurements.
+An unrepresentative $i_\star$ sample could tighten the constraints on the stellar obliquities and bias the interpretation of the overall distribution.
+
+Finally, we applied the framework to all exoplanetary systems to all exoplanetary systems with available sky-projected stellar obliquities and found that approximately $67\pm9$\% of the systems have a stellar obliquity less than $40\degr$, and approximately $33\pm9$\% of the systems follow a nearly isotropic stellar obliquity distribution between $\sim$$40\degr$ and $\sim$$140\degr$.
+The distribution could have important implications for the formation and evolution of close-in planetary systems and is worth further investigation.
+
+This study was conducted using the \href{https://github.com/showyourwork/showyourwork}{\showyourwork} reproducible workflow \citep{Luger2021}, which leverages continuous integration to automate the data retrieval from \href{https://zenodo.org/}{zenodo.org}, figure generation, and manuscript compilation.
+The script used to produce each figure can be accessed via a link in the corresponding figure caption, as it corresponds to the latest build of the manuscript.
+The git repository for this study, which includes the Jupyter notebook demonstration and case studies of the stellar obliquity distribution inference, is publicly accessible at \url{https://github.com/jiayindong/obliquity}.
 

### 2023-02-27 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
+\newcommand{\dfm}[1]{\textcolor{blue}{\textbf{From DFM:} #1}}
+\newcommand{\edits}[1]{\textcolor{grey}{\textbf{Edits:} #1}}
+
 \received{}
@@
 % Title
-\title{A Hierarchical Bayesian Framework for Stellar Obliquity Distribution Inference}
+\title{A Hierarchical Bayesian Framework for Inferring the Stellar Obliquity Distribution}
 
@@
 
-\author[0000-0002-9328-5652]{Dan Foreman-Mackey}
+\author[0000-0002-9328-5652]{Daniel Foreman-Mackey}
 \affiliation{\FlatironCCA}
 
-\author{et al.}
+\received{}
+\revised{}
+\accepted{}
 
-% \received{}
-% \revised{}
-% \accepted{}
-
-% \submitjournal{the AAS Journals}
+\submitjournal{the AAS Journals}
 
@@
 Stellar obliquity $\psi$, the angle between a planet's orbital axis and its host star's spin axis, traces the formation and evolution of a planetary system.
-In transiting-exoplanet observations, only the sky-projected stellar obliquity $\lambda$ can be measured.
-To determine the true stellar obliquity $\psi$, information on the stellar inclination $i_\star$ is also needed.
-Here we find that while the stellar inclination is crucial for understanding the obliquity of individual systems, it is not a requirement for determining the population-level stellar obliquity. 
-The constraints on the stellar obliquity distribution are mainly driven by the sky-projected stellar obliquities.
-We introduce a flexible, hierarchical Bayesian framework that can infer the stellar obliquity distribution solely from sky-projected stellar obliquities.
-Stellar inclination measurements are optional input in the model, and if not available, they are assumed to be isotropically distributed.
-When applying the framework to all exoplanetary systems with sky-projected stellar obliquity measurements, we find that the stellar obliquity distribution is unimodal and peaked at zero degrees.
+In transiting exoplanet observations, only the sky-projected stellar obliquity $\lambda$ can be measured.
+To determine the true stellar obliquity $\psi$, information about the stellar inclination $i_\star$ is also needed.
+In this paper, we demonstrate that while a constraint on the stellar inclination is crucial for measuring the obliquity of an individual system, it is not required for robust determination of the population-level stellar obliquity distribution.
+In practice, the constraints on the stellar obliquity distribution are mainly driven by the sky-projected stellar obliquities.
+We introduce a flexible, hierarchical Bayesian framework that can be used to infer the stellar obliquity distribution solely from sky-projected stellar obliquities, including stellar inclination measurements when available.
+When applying the framework to all exoplanetary systems with measured sky-projected stellar obliquity, we find that the inferred population-level obliquity distribution is unimodal and peaked at zero degrees.
 Misaligned systems have nearly isotropic stellar obliquities, which could have important implications for the formation and evolution of close-in planets.
@@
 
-Stellar obliquity $\psi$ describes the angle between a planet's orbital axis $\bf{n}_{\rm orb}$ and its host star's spin axis $\bf{n}_{\star}$. 
+The stellar obliquity $\psi$ is the angle between a planet's orbital axis $\bf{n}_{\rm orb}$ and the host star's spin axis $\bf{n}_{\star}$. 
 This angle is an important tracer of a planetary system's formation environment and dynamical evolution.
@@
 
-It is yet unclear if all of these physical and dynamic processes apply to exoplanetary systems.
-These proposed mechanisms have different predictions on stellar obliquity distributions \citep[see][and references therein]{Albrecht22}.
-For example, the secular chaos mechanism suggests that the stellar obliquity would never exceed $90\degr$ \citep[e.g.,][]{Teyssandier19}.
-The stellar Kozai mechanism predicts that the stellar obliquity distribution should be bimodal and concentrated at $40\degr$ and $140\degr$ \citep[e.g.,][]{Fabrycky07, Anderson16, Vick19}.
-The multiple-planet scattering mechanism proposes that the majority of systems would be aligned, while a small fraction of systems would exhibit a diverse range of stellar obliquities \citep[e.g.,][]{Beague12}.
+It is as yet unclear if, and to what extent, all of these physical and dynamic processes apply to exoplanetary systems.
+These proposed mechanisms all make different predictions on stellar obliquity distributions \citep[see][and references therein]{Albrecht22}.
+For example, the secular chaos mechanism tends to produce a stellar obliquity distribution with $\psi < 90\degr$ \citep[e.g.,][]{Teyssandier19}.
+The stellar Kozai mechanism predicts a bimodal stellar obliquity distribution, concentrated at $40\degr$ and $140\degr$ \citep[e.g.,][]{Fabrycky07, Anderson16, Vick19}.
+The multiple-planet scattering mechanism results in a majority of aligned systems, with a small fraction of systems at a diverse range of stellar obliquities \citep[e.g.,][]{Beague12}.
 With these predictions in mind, we aim to determine the dominant mechanisms responsible for shaping close-in planetary systems by characterizing the stellar obliquity distribution of exoplanetary systems through a Bayesian approach.
 
-In exoplanetary observation, only the sky-projected stellar obliquity $\lambda$, the angle between the projections of $\bf{n}_{\rm orb}$ and $\bf{n}_{\star}$ onto the plane of the sky, can be measured.
+When observing an exoplanet, typically only the sky-projected stellar obliquity $\lambda$, the angle between the projections of $\bf{n}_{\rm orb}$ and $\bf{n}_{\star}$ onto the plane of the sky, can be measured.
 This measurement is primarily obtained via the Rossiter-McLaughlin effect \citep{Rossiter24, McLaughlin24}.
-The stellar obliquity $\psi$ can be determined, if both the sky-projected stellar obliquity $\lambda$ and the stellar inclination $i_\star$ are known.
+The stellar obliquity $\psi$ for an individual exoplanet system can be inferred, if both the sky-projected stellar obliquity $\lambda$ and the stellar inclination $i_\star$ are measured precisely.
 The relationship between $\psi$ and $\{\lambda, i_\star\}$ is given by \citep[e.g.,][]{Fabrycky09}:
@@
 where $i_{\rm orb}$ is the inclination angle between the vector $\bf{n}_{\rm orb}$ and the observer's line of sight, and $i_\star$ is the inclination angle between $\bf{n}_{\star}$ and the observer's line of sight.
-In transiting-exoplanet systems where $i_{\rm orb} \approx 90\degr$, $\cos{\psi} \approx \sin{i_\star}\cos{\lambda}$.
+If an exoplanet system transits, the orbit is nearly edge-on ($i_{\rm orb} \approx 90\degr$), so in those cases, this relationship becomes approximately
+\begin{equation}
+\cos{\psi} \approx \sin{i_\star}\cos{\lambda}\,,
+\end{equation}
+although this is not a simplification that we are required to make in this paper.
 
-Stellar inclinations can often be constrained via, e.g., photometric and spectroscopic rotational modulation introduced by starspots for cool stars \citep[e.g.,][]{Masuda20, Albrecht21}, gravity darkening for fast-rotating stars \citep[e.g.,][]{Barnes09, Barnes11}, and asteroseismology for bright stars \citep[e.g.][]{Chaplin13}.
-For systems without $i_\star$ measurements, it is still possible to infer their stellar obliquities from the sky-projected obliquities, assuming isotropic stellar inclinations; however, the inferred $\psi$ will have greater uncertainty than the one inferred with $i_\star$ measurement \citep{Fabrycky09}.
+In some cases, stellar inclinations can be constrained via, for example, photometric and spectroscopic rotational modulation introduced by starspots for cool stars \citep[e.g.,][]{Masuda20, Albrecht21}, gravity darkening for fast-rotating stars \citep[e.g.,][]{Barnes09, Barnes11}, and asteroseismology for bright stars \citep[e.g.][]{Chaplin13}.
+However, for the vast majority of exoplanet systems, $i_\star$ measurements are not feasible.
+In these cases, it is still possible to infer their stellar obliquities from the sky-projected obliquities, assuming isotropic stellar inclinations; however, the inferred $\psi$ will have greater uncertainty than the one inferred with $i_\star$ measurement \citep{Fabrycky09}.
 
@@
 
-To find the stellar obliquity distribution of exoplanetary systems, we develop a hierarchical Bayesian framework that takes the observed sky-projected stellar obliquity $\lambda$ and orbital inclination $i_{\rm orb}$ as input data.
-The measurement of the stellar inclination $i_\star$ is optional.
-In the absence of $i_\star$ information, we assume that it follows an isotropic distribution.
-Otherwise, the observed $i_\star$ can be taken from previous measurements, or if the rotational modulation method is employed, the $i_\star$ value can be inferred from the stellar rotation period $P_{\rm rot}$, stellar radius $R_\star$, and sky-projected rotational broadening velocity $v\sin{i}_\star$.
+To find the stellar obliquity distribution of exoplanetary systems, we develop a hierarchical Bayesian framework that takes measurements of the observed sky-projected stellar obliquity $\lambda$ and orbital inclination $i_{\rm orb}$ as input data.
+If a measurement of the stellar inclination $i_\star$ is available, it can be provided or inferred from the stellar rotation period $P_{\rm rot}$, stellar radius $R_\star$, and sky-projected rotational broadening velocity $v\sin{i}_\star$.
+In the absence of an $i_\star$, $P_{\rm rot}$, or $v\sin{i}_\star$ measurement, we marginalize over the value of $i_\star$ using an isotropic prior distribution.
 
@@
     \includegraphics[width=\linewidth]{figures/graph.pdf}
-    \caption{Hierarchical Bayesian framework to infer the stellar obliquity distribution of exoplanetary systems.
-The distribution is described by hyperparameters $\vb*{\beta}$ and is constrained by the sky-projected stellar obliquity $\lambda_n$, orbital inclination $i_{{\rm orb}, n}$, and, if available, the stellar inclination $i_{\star, n}$ of each $n_{\rm th}$ individual system.
-The system's stellar obliquity $\psi_n$ is calculated based on these parameters.
-    Properties of the star other than its inclination, such as its radius and rotation period, are represented by $\gamma_{\star, n}$. 
+    \caption{A probabilistic graphical model showing the conditional structure of our hierarchical Bayesian framework for inferring the stellar obliquity distribution of exoplanetary systems.
+The population model is parameterized by hyperparameters $\vb*{\beta}$, and it is constrained by the sky-projected stellar obliquity $\lambda_n$, orbital inclination $i_{{\rm orb}, n}$, and, if available, the stellar inclination $i_{\star, n}$ of each $n_{\rm th}$ individual system.
+Each system's stellar obliquity $\psi_n$ is inferred simultaneously based on these data.
+    Properties of the star other than its inclination, such as its radius and rotation period, are represented by $\gamma_{\star, n}$, and these can inform the constraint on $i_{\star, n}$, if available.
     Obs$_{\star,n}$ contains all observed properties of the star with uncertainties, $\hat{\lambda}_n$ has the measured sky-projected stellar obliquity and its uncertainty, and $\hat{i}_{{\rm orb}, n}$ has measured orbital inclination and its uncertainty.}
@@
 Figure~\ref{fig:graph} illustrates the probabilistic graphical model for our hierarchical Bayesian framework.
-We aim to constrain a set of hyperparameters $\bm{\beta}$ that describes the stellar obliquity distribution.
-The parameter set $\bm{\beta}$ is constrained by $N$ individual systems, where each $\psi_n$ is computed from the system's sky-projected stellar obliquity $\lambda_n$, the orbital inclination $i_{{\rm orb},n}$, and if available, the stellar inclination $i_{\star, n}$.
+We aim to constrain a set of hyperparameters $\bm{\beta}$ that describe the stellar obliquity distribution.
+The parameter set $\bm{\beta}$ is constrained by $N$ individual systems, where each $\psi_n$ is simultaneously fit based on the entire sample of sky-projected stellar obliquities $\lambda_n$, orbital inclinations $i_{{\rm orb},n}$, and if available, the stellar inclination $i_{\star, n}$.
 The parameter $\gamma_{\star, n}$ contains all stellar properties other than $i_{\star,n}$, such as the stellar rotation period $P_{{\rm rot},n}$, radius $R_{\star,n}$, and projected rotational velocity $v\sin{i_{\star,n}}$, if they are known.
 
-The sky-projected stellar obliquity $\lambda_n$ is constrained by observed $\hat{\lambda}_n$ via the Rossiter-McLaughlin effect or gravity darkening.
-The orbital inclination $i_{{\rm orb},n}$ of exoplanets is inferred from their transit light curves.
-Obs$_{\star,n}$ includes observed data that helps to constrain $i_{\star,n}$ and $\gamma_{\star, n}$.
-If the rotational modulation method is used to constrain the stellar inclination, the $i_{\star,n}$ and $\gamma_{\star,n}$ are constrained by rotational modulation in photometric or spectroscopic time series and observed sky-projected stellar rotational line broadening $\hat{v}\sin{i_{\star,n}}$ \citep[e.g.,][]{Masuda20}, where we have $v\sin{i_\star} = 2 \pi R_\star / P_{\rm rot}$.
-If the gravity-darkening method is used, the $i_{\star,n}$ could be constrained by the anomaly in transit light curves \citep{Barnes09}.
+In Figure~\ref{fig:graph}, the observed values of $\lambda_n$ and $i_{{\rm orb},n}$ are indicated by $\hat{\lambda}_n$ and $\hat{i}_{{\rm orb},n}$, respectively.
+The $\hat{\lambda}_n$ measurements typically come from the Rossiter-McLaughlin effect or gravity darkening, and the $\hat{i}_{{\rm orb},n}$ is constrained by the transit light curve.
+
+The constraints on stellar inclination are somewhat more complicated.
+In Figure~\ref{fig:graph}, the Obs$_{\star,n}$ node includes any observed data that directly constrains $i_{\star,n}$.
+This could include a direct measurement of $\hat{i}_{\star,n}$ \citep[via gravity-darkening or asteroseismology, for example][]{Barnes09, Chaplin13}, or a measurement of the sky-projected stellar rotational line broadening $\hat{v}\sin{i_{\star,n}}$, which is related to $i_\star$ as $v\sin{i_\star} = 2 \pi R_\star / P_{\rm rot}$ \citep{Masuda20}.
+
+% If the rotational modulation method is used to constrain the stellar inclination, the $i_{\star,n}$ and $\gamma_{\star,n}$ are constrained by rotational modulation in photometric or spectroscopic time series and observed sky-projected stellar rotational line broadening $\hat{v}\sin{i_{\star,n}}$ \citep[e.g.,][]{Masuda20}, where we have $v\sin{i_\star} = 2 \pi R_\star / P_{\rm rot}$.
+% If the gravity-darkening method is used, the $i_{\star,n}$ could be constrained by the anomaly in transit light curves \citep{Barnes09}.
 If asteroseismology is used, the $i_{\star,n}$ could be constrained from the mode splitting in the Fourier transform of photometric time series \citep{Chaplin13}.
@@
 For the stellar obliquity distribution, we model the $\cos{\psi}$ distribution instead of $\psi$ distribution to understand if the stellar obliquity is isotropically distributed.
-If stellar obliquity distribution follows an isotropic distribution, the $\cos{\psi}$ distribution will be uniformly distributed between -1 to 1.
-We adopt a two-component mixture Beta distribution with hyperparameters $\bm{\beta} = \{\bm{w},\bm{a},\bm{b}\}$, where each hyperparameter has a dimension of 2.
-One component is designed for a large population of well-aligned systems, and the other is designed for misaligned systems.
+If stellar obliquity distribution is isotropic, $\cos{\psi}$ is uniformly distributed between -1 to 1.
+To flexibly model the stellar obliquity distribution, we select a two-component mixture of Betas distribution with hyperparameters $\bm{\beta} = \{\bm{w},\bm{a},\bm{b}\}$, where each hyperparameter has a dimension of 2.
+This mixture distribution has the capacity to capture anything from an isotropic distribution to a strongly bimodal population.
+% One component is designed for a large population of well-aligned systems, and the other is designed for misaligned systems.
 The Beta distribution is a flexible distribution that can describe the spike of aligned systems with stellar obliquities close to $0\degr$ or the clustered or broadly distributed misaligned systems.
-The probability density function of $\cos{\psi}$ follows 
+\dfm{The notation here is a bit sloppy. It might be better to say something like: there's some parameter $x \sim \mathrm{mixture}$ which we then transform as $2\,x-1$, which isn't quite the same.}
+Then, under this model, the probability density function for $\cos{\psi}$ is 
 \begin{align}
-    \cos{\psi} &\sim 2\times\Bigl( w_0 {\rm Beta}(a_0, b_0) + w_1 {\rm Beta}(a_1, b_1)\Bigr)-1.
+    \cos{\psi} &\sim 2\,\Bigl( w_0 {\rm Beta}(a_0, b_0) + w_1 {\rm Beta}(a_1, b_1)\Bigr)-1.
 \end{align}
-Since the Beta distribution is defined on the interval $[0, 1]$, we extend its support to $[-1,1]$ using a linear transformation (i.e., $2\times {\rm Beta} - 1$).
+Since the Beta distribution is defined on the interval $[0, 1]$, we extend its support to $[-1,1]$ using a linear transformation (i.e., $2\,{\rm Beta} - 1$).
+
+\dfm{I'm not sure that I understand this next bit. Why and how do we do this?}
 For the ${\rm Beta}(a_0,b_0)$ component, we have it to describe the population of aligned systems.
@@
 For the ${\rm Beta}(a_1,b_1)$ component, we have it to describe the misaligned systems.
-We design the order of the two components to prevent label switching in the mixture model.
+
+We design the order of the two components to prevent label switching in the mixture model \dfm{how?}.
 With the design, the hyperparameter and parameter priors are the following:
@@
     \cos{i}_{\star,n} &\sim \mathcal{U}(0, 1) \nonumber\\
-    \cos{i}_{{\rm orb},n} &\sim \mathcal{U}(-1, 1) \nonumber\\
-    \lambda_{\star,n} &\sim \mathcal{U}(0, \pi).
+    \cos{i}_{{\rm orb},n} &\sim \mathcal{U}(0, 1) \nonumber\\
+    \lambda_{\star,n} &\sim \mathcal{U}(-\pi, \pi).
 \end{align}
@@
 
-We first apply the hierarchical Bayesian framework to simulated data of which the ground-truth stellar obliquity distribution is known.
-We test the four $\cos{\psi}$ distributions: a uniform distribution $\mathcal{U}(-1,1)$ and three Normal distributions $\mathcal{N}(0,0.2)$, $\mathcal{N}(-0.4,0.2)$, and $\mathcal{N}(0.4,0.2)$.
+To investigate the performance of our hierarchical Bayesian framework, we apply it to simulated data, generated with a known ground-truth stellar obliquity distribution.
+We test the following four $\cos{\psi}$ distributions: a uniform distribution $\mathcal{U}(-1,1)$, and three Normal distributions $\mathcal{N}(0,0.2)$, $\mathcal{N}(-0.4,0.2)$, and $\mathcal{N}(0.4,0.2)$.
+\dfm{For these normals, we must truncate, since they have support outside of 0--1. How is that implemented; should we mention it explicitly?}
 For each $\cos{\psi}$ distribution, we randomly generate 200 samples of stellar inclination $i_\star$ and sky-projected stellar obliquity $\lambda$.
@@
 The sampled $i_\star$ and $\lambda$ here are \emph{true} values. 
-We then add some Gaussian noises to the \emph{true} $\lambda$ and $i_\star$ to simulate the \emph{observed} data.
-We choose uncertainties of $\sigma_{\lambda} = 8\degr$ and $\sigma_{i_\star} = 10\degr$, which are typical observational uncertainties summarized in \cite{Albrecht22}.
-Using the observed $\lambda$ and their uncertainties, we infer the $\cos{\psi}$ distribution of the sample with or without $i_\star$ likelihoods.
+To simulate the observation process, we add Gaussian noise to the \emph{true} $\lambda$ and $i_\star$, using uncertainties of $\sigma_{\lambda} = 8\degr$ and $\sigma_{i_\star} = 10\degr$, which are typical of the literature sample \cite{Albrecht22}.
+Using these simulated $\lambda$ measurements and their uncertainties, we infer the $\cos{\psi}$ distribution of the sample with or without $i_\star$ likelihoods.
 
-In Figure~\ref{fig:simulation}, we present the inferred stellar obliquity distributions.
+In Figure~\ref{fig:simulation}, we present the results of this experiment, plotting the inferred stellar obliquity distributions, compared to the ground-truth distributions.
 Since the simulated stellar obliquity distributions only have a single component, we model the data with a single Beta distribution.
-Each row of Figure~\ref{fig:simulation} corresponds to an injected stellar obliquity distribution.
-The orange curve and contours are the median and 1-$\sigma$ and 2-$\sigma$ uncertainties of the inferred $\cos{\psi}$ distribution with stellar inclination information, and the blue curve and contours are the ones without stellar inclination information.
-We simply assume an isotropic distribution of $i_\star$ for these models.
-Surprisingly, as shown in Figure~\ref{fig:simulation}, the inferred distributions of $\cos{\psi}$, whether or not they include the information of $i_\star$, correctly recover the injected distributions (represented by grey dashed lines) in all four cases.
-Despite the fact that the inferred distributions without $i_\star$ measurements may have wider uncertainties, as indicated by shallow color contours, the modes and widths of the inferred stellar obliquity distributions are consistent with or without $i_\star$ likelihood.
+Each row of Figure~\ref{fig:simulation} corresponds to a different simulation distribution.
+The orange curve and contours in the left column are the median, 1-$\sigma$ and 2-$\sigma$ uncertainties of the inferred $\cos{\psi}$ distribution with stellar inclination information.
+The right column shows the same inferences (in blue) when constraints on stellar inclination are not included.
+% We simply assume an isotropic distribution of $i_\star$ for these models.
+Surprisingly, Figure~\ref{fig:simulation} demonstrates that our inference procedure recovers the true distribution for $\cos{\psi}$ equally well, regardless of the inclusion of $i_\star$ measurements.
+Despite the fact that the inferred distributions without $i_\star$ measurements have marginally wider uncertainties, as indicated by shallow color contours, the modes and widths of the inferred stellar obliquity distributions are consistent with or without $i_\star$ likelihood.
 Since the injected distributions in rows 2-4 are Normal distributions, it should not be surprising that the inferred distributions, which are Beta distributions, may not exactly match the injected distributions.
+\dfm{Add some more here about how the true distribution is not included within the support of the distribution we're using to fit.
+One thought: perhaps we would actually do better if we used the two betas here too?}
 
@@
 
-\subsection{Exoplanetary Stellar Obliquity Distribution}
+\subsection{The Stellar Obliquity Distribution of Observed Exoplanets}
 
@@
 The distribution suggests that there is a pile-up of planetary systems with stellar obliquities less than $40^\circ$ and an isotropic distribution for obliquities between $40^\circ$ and $140^\circ$.
-The fraction of aligned systems dominates the distribution, with $w_0 = 0.67 \pm 0.09$.
-The corresponding hyperposteriors of the Beta distribution are $a_0 = 31.0\pm12.8$ and $b_0 = 0.40\pm0.12$.
-On the other hand, the fraction of misaligned systems is estimated to be $w_1 = 0.33 \pm 0.09$, with hyperposteriors of $a_1 = 1.56\pm0.93$ and $b_1 = 1.64\pm1.58$.
+The fraction of aligned systems $w_0$ dominates the distribution, with $w_0 = 0.67 \pm 0.09$.
+The corresponding posteriors for the parameters of the population Beta distribution are $a_0 = 31.0\pm12.8$ and $b_0 = 0.40\pm0.12$.
+On the other hand, the fraction of misaligned systems is estimated to be $w_1 = 0.33 \pm 0.09$, with posteriors of $a_1 = 1.56\pm0.93$ and $b_1 = 1.64\pm1.58$.
 
@@
         \caption{Inferred stellar obliquity distribution for all exoplanetary systems with sky-projected stellar obliquity measurements.
-This inference is based purely on the observed sky-projected obliquities and assumes an isotropic distribution for the stellar inclinations.}
+This inference is based purely on the observed sky-projected obliquities and assumes an isotropic distribution for the stellar inclinations.
+        \dfm{Might be worth plotting the two mixture components too.}}
         \label{fig:psi_dist}
@@
 
-The discrepancy between the inferred stellar obliquity distribution of the full sample of exoplanetary systems and the previous analysis of the subsample with $i_\star$ measurements warrants further investigation.
-The earlier study identified a preponderance of perpendicular planets and disfavored an isotropic stellar obliquity distribution \citep{Albrecht21}.
-There are two potential explanations for this difference: (1) the subsample with $i_\star$ measurements is small and only includes about 20 misaligned systems, and (2) the requirement for $i_\star$ measurements could introduce selection biases in the sample.
+The discrepancy between this inference for the full sample of exoplanet systems and the previous analysis of the subsample with $i_\star$ measurements \citep{Albrecht21} warrants further investigation.
+This earlier study identified a preponderance (\dfm{maybe different word.}) of perpendicular planets and disfavored an isotropic stellar obliquity distribution \citep{Albrecht21}.
+There are at least two potential explanations for this difference: (1) the subsample with $i_\star$ measurements is small and only includes about 20 misaligned systems, and (2) the requirement for $i_\star$ measurements could introduce selection biases in the sample.
+\dfm{Expand on this a little bit, I think.
+Maybe add a short paragraph for each.}
 
@@
 
+\section*{Acknowledgments}
+We are grateful to Josh Winn and Jared Siegel for their valuable insights and feedback on this project.
+
 This study was conducted using the \href{https://github.com/showyourwork/showyourwork}{\showyourwork} reproducible workflow \citep{Luger2021}, which leverages continuous integration to automate the data retrieval from \href{https://zenodo.org/}{zenodo.org}, figure generation, and manuscript compilation.

### 2023-03-21 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \newcommand{\dfm}[1]{\textcolor{blue}{\textbf{From DFM:} #1}}
-\newcommand{\edits}[1]{\textcolor{grey}{\textbf{Edits:} #1}}
+\newcommand{\edits}[1]{\textcolor{purple}{#1}}
 
@@
 For the stellar obliquity distribution, we model the $\cos{\psi}$ distribution instead of $\psi$ distribution to understand if the stellar obliquity is isotropically distributed.
-If stellar obliquity distribution is isotropic, $\cos{\psi}$ is uniformly distributed between -1 to 1.
+If stellar obliquity distribution is isotropic, $\cos{\psi}$ is uniformly distributed between -1 and 1.
 To flexibly model the stellar obliquity distribution, we select a two-component mixture of Betas distribution with hyperparameters $\bm{\beta} = \{\bm{w},\bm{a},\bm{b}\}$, where each hyperparameter has a dimension of 2.
+\edits{We have $\bm{w} = \{w_0, w_1\}$, $\bm{a} = \{a_0, a_1\}$, and  $\bm{b} = \{b_0, b_1\}$.}
 This mixture distribution has the capacity to capture anything from an isotropic distribution to a strongly bimodal population.
@@
 \begin{align}
-    \cos{\psi} &\sim 2\,\Bigl( w_0 {\rm Beta}(a_0, b_0) + w_1 {\rm Beta}(a_1, b_1)\Bigr)-1.
+    u_0 &\sim {\rm Beta}(a_0, b_0) \nonumber\\
+    u_1 &\sim {\rm Beta}(a_1, b_1) \nonumber\\
+    \cos{\psi} &\sim 2\,( w_0 u_0 + w_1 u_1)-1.
 \end{align}
-Since the Beta distribution is defined on the interval $[0, 1]$, we extend its support to $[-1,1]$ using a linear transformation (i.e., $2\,{\rm Beta} - 1$).
-
+Since the Beta distribution is defined on the interval $[0, 1]$, we extend its support to $[-1,1]$ using a linear transformation (i.e., $2\,( w_0 u_0 + w_1 u_1)-1$).
 \dfm{I'm not sure that I understand this next bit. Why and how do we do this?}
@@
 For the ${\rm Beta}(a_1,b_1)$ component, we have it to describe the misaligned systems.
+% add more description on why chose a0, b0 priors
+We design the order of the two components to prevent label switching in the mixture model \dfm{how?}.
 
-We design the order of the two components to prevent label switching in the mixture model \dfm{how?}.
 With the design, the hyperparameter and parameter priors are the following:
@@
 The right column shows the same inferences (in blue) when constraints on stellar inclination are not included.
-% We simply assume an isotropic distribution of $i_\star$ for these models.
 Surprisingly, Figure~\ref{fig:simulation} demonstrates that our inference procedure recovers the true distribution for $\cos{\psi}$ equally well, regardless of the inclusion of $i_\star$ measurements.
@@
 $\psi$ is the angle between the stellar spin axis and the planetary orbital axis, and $\theta$ is the azimuthal angle of the stellar spin axis.
-For the given $\bf{n}_{\rm orb}$ axis, if $\bf{n}_{\star}$ is a random vector with uniform distribution on a three-dimensional sphere, i.e., $\bf{n}_{\star}$ is isotropically distributed, the probability density function of $\psi$ follows $p_{\psi} \sim \sin{\psi}$ and $p_\theta \sim 1/2\pi$.
-Since $p_{\cos{\psi}} = p_{\psi} \abs{\dv*{\psi}{\cos{\psi}}}$, $p_{\cos{\psi}}$ is uniformly distributed between $-1$ and $1$ for isotropic stellar obliquity.
+For the given $\bf{n}_{\rm orb}$ axis, if $\bf{n}_{\star}$ is a random vector with uniform distribution on a three-dimensional sphere, i.e., $\bf{n}_{\star}$ is isotropically distributed, the probability density function of $\psi$ follows $p(\psi) = \sin{\psi}$ and $p(\theta) = 1/2\pi$.
+\edits{Here $p(x)$ denote the probability density distribution of $x$.}
+Since $p(\cos{\psi}) = p(\psi) \abs{\dv*{\psi}{\cos{\psi}}}$, $p(\cos{\psi})$ is uniformly distributed between $-1$ and $1$ for isotropic stellar obliquity.
 The $\{\lambda, i_\star\}$ coordinate system shown in panel (b) relates to observed properties.
 $\lambda$ is the sky-projected stellar obliquity and $i_\star$ is the stellar inclination.
-If $\bf{n}_{\star}$ is isotropically distributed, $p_{i_\star} \sim \sin{i_\star}$ and $p_\lambda \sim \mathcal{U}(-\pi, \pi)$.
-Again, since $p_{\cos{i_\star}} = p_{i_\star} \abs{\dv*{i_\star}{\cos{i_\star}}}$, $p_{\cos{i_\star}}$ is uniformly distributed between $-1$ and $1$.
-Because of the observational degeneracy between $i_\star$ and $180\degr - i_\star$, the convention is to have $0 \leqslant i_\star \leqslant 90\degr$, and thus $p_{\cos{i_\star}}$ is uniformly distributed between $0$ and $1$ and $p_{\cos{i_\star}} \sim \mathcal{U}(0, 1)$.
-This also limits $p_\theta \sim \mathcal{U}(-\pi/2, \pi/2)$.
+If $\bf{n}_{\star}$ is isotropically distributed, $p(i_\star) = \sin{i_\star}$ and $p(\lambda) = 1/2\pi$.
+Again, since $p(\cos{i_\star}) = p(i_\star) \abs{\dv*{i_\star}{\cos{i_\star}}}$, $p(\cos{i_\star})$ is uniformly distributed between $-1$ and $1$.
+Because of the observational degeneracy between $i_\star$ and $\pi - i_\star$, the convention is to have $0 \leqslant i_\star \leqslant \pi/2$, and thus $\cos{i_\star} \sim \mathcal{U}(0, 1)$ (i.e., $p(\cos{i_\star}) = 1$).
+This also suggests $\theta \sim \mathcal{U}(-\pi/2, \pi/2)$.
 
@@
 Since $\psi$ and $\theta$ are independent variables, we could marginalize over $\theta$ to find the relation between the probability density functions between $\lambda$ and $\psi$.
+
 The Jacobian transformation follows
 \begin{equation}
-    p_{\cos{\lambda}} = \int \abs{\pdv{\cos{\psi}}{\cos{\lambda}}} p_{\cos{\psi}} p_{\cos{\theta}} d\cos{\theta}.
+    p(\cos{\lambda}) = \int \abs{\pdv{\cos{\psi}}{\cos{\lambda}}} p(\cos{\psi}) p(\cos{\theta}) d\cos{\theta}.
 \end{equation}
-Replacing $\sin{i_\star}$ in Equation~(\ref{eq:coord3}) using Equation~(\ref{eq:coord1}), we find $\cos{\lambda} = \cos{\psi}/\sqrt{1-(1-\cos^2{\psi})\cos^2{\theta}}$.
-Reorganize the equation, we get $\cos^2{\psi} = \frac{\cos^2{\lambda}\cos^2{\theta}-\cos^2{\lambda}}{\cos^2{\lambda}\cos^2{\theta}-1}$ and
-the partial derivative $\abs{\pdv{\cos{\psi}}{\cos{\lambda}}} = \frac{(1-\cos^2{\theta})^{1/2}}{(1-\cos^2{\theta}\cos^2{\lambda})^{3/2}}$.
-Assume $\theta$ is uniformly distributed between $-\pi/2$ and $\pi/2$, $p_{\cos{\theta}} = p_{\theta} \abs{\dv*{\theta}{\cos{\theta}}} = 2/\pi/(1-\cos^2{\theta)^{1/2}}$, where the factor of 2 is due to two solutions of $\theta$ to the $\cos{\theta}$.
+Replacing $\sin{i_\star}$ in Equation~(\ref{eq:coord3}) using Equation~(\ref{eq:coord1}), we find 
+\begin{equation}
+    \cos{\lambda} = \cos{\psi}/\sqrt{1-(1-\cos^2{\psi})\cos^2{\theta}}.
+\end{equation} 
+Reorganize the equation, we get 
+\begin{equation}
+    \cos^2{\psi} = \frac{\cos^2{\lambda}\cos^2{\theta}-\cos^2{\lambda}}{\cos^2{\lambda}\cos^2{\theta}-1}
+\end{equation}
+and the partial derivative 
+\begin{equation}
+    \abs{\pdv{\cos{\psi}}{\cos{\lambda}}} = \frac{(1-\cos^2{\theta})^{1/2}}{(1-\cos^2{\theta}\cos^2{\lambda})^{3/2}}.
+\end{equation}
+Assume $\theta$ is uniformly distributed between $-\pi/2$ and $\pi/2$, $p(\cos{\theta}) = p(\theta) \abs{\dv*{\theta}{\cos{\theta}}} = 2/\pi/(1-\cos^2{\theta)^{1/2}}$, where the factor of 2 is due to two solutions of $\theta$ to the $\cos{\theta}$.
 Putting all the parts together, we get
 \begin{equation}\label{eqn:jac_lam}
-    p_{\cos{\lambda}} = \frac{2}{\pi} \int_{0}^{1} (1-\cos^2{\theta}\cos^2{\lambda})^{-3/2} p_{\cos{\psi}} d\cos{\theta}.
+    p(\cos{\lambda}) = \frac{2}{\pi} \int_{0}^{1} (1-\cos^2{\theta}\cos^2{\lambda})^{-3/2} p(\cos{\psi}) d\cos{\theta}.
 \end{equation}
-If $\cos{\psi}$ is uniformly distribution, i.e., $p_{\cos{\psi}} \sim 1/2$, Equation~(\ref{eqn:jac_lam}) becomes $p_{\cos{\lambda}} \sim 1/\pi/\sqrt{1-\cos^2{\lambda}}$, which is equivalent to $p_\lambda \sim \mathcal{U}(-\pi, \pi)$.
+If $\cos{\psi}$ is uniformly distribution, i.e., $p(\cos{\psi}) = 1/2$, Equation~(\ref{eqn:jac_lam}) becomes $p(\cos{\lambda}) = 1/\pi/\sqrt{1-\cos^2{\lambda}}$, which is equivalent to $\lambda \sim \mathcal{U}(-\pi, \pi)$.
 This suggests $\lambda$ is uniformly distributed for an isotropic $\psi$ distribution.
-For different $\cos{\psi}$ distributions, $p_\lambda$ can be found by evaluating Equation~(\ref{eqn:jac_lam}).
+For different $\cos{\psi}$ distributions, $p(\lambda)$ can be found by evaluating Equation~(\ref{eqn:jac_lam}).
 
@@
 \begin{equation}
-    p_{\cos{i_\star}} = \int \abs{\pdv{\sin{\psi}}{\cos{i_\star}}} p_{\sin{\psi}} p_{\cos{\theta}} d\cos{\theta}.
+    p(\cos{i_\star}) = \int \abs{\pdv{\sin{\psi}}{\cos{i_\star}}} p(\sin{\psi}) p(\cos{\theta}) d\cos{\theta}.
 \end{equation}
-From Equation~(\ref{eq:coord1}), we have $\sin{\psi} = \cos{i_\star}/\cos{\theta}$ and $\abs{\pdv{\sin{\psi}}{\cos{i_\star}}} = 1/\cos{\theta}$.
-Again, we assume $\theta$ is uniformly distributed, and this gives $p_{\cos{\theta}} = 2/\pi/(1-\cos^2{\theta)^{1/2}}$.
-Lastly, we transform the $p_{\sin{\psi}}$ to $p_{\cos{\psi}}$, where $p_{\sin{\psi}} = 2p_{\cos{\psi}}\sin{\psi}/\sqrt{1-\sin^2{\psi}}$ and the factor of 2 is from two solutions of $\cos{\psi}$ to the $\cos^2{\psi} = 1-\sin^2{\psi}$.
-Combining all the pieces together, we get
-\begin{equation}\label{eqn:jac_istar}
-    p_{\cos{i_\star}} = \frac{4}{\pi} \int_{\cos{i_\star}}^{1} \frac{\cos{i_\star}/\cos{\theta}}{\sqrt{\cos^2{\theta}-\cos^2{i_\star}}} \frac{1}{\sqrt{1-\cos^2{\theta}}} p_{\cos{\psi}} d\cos{\theta}.
+From Equation~(\ref{eq:coord1}), we get 
+\begin{equation}
+    \sin{\psi} = \cos{i_\star}/\cos{\theta}
+\end{equation} and the partial derivative
+\begin{equation}
+    \abs{\pdv{\sin{\psi}}{\cos{i_\star}}} = 1/\cos{\theta}.
 \end{equation} 
+Again, we assume $\theta$ is uniformly distributed, and this gives $p(\cos{\theta}) = 2/\pi/(1-\cos^2{\theta)^{1/2}}$. Lastly, we transform the $p(\sin{\psi})$ to $p(\cos{\psi})$,
+\begin{equation}
+    p(\sin{\psi}) = 2p(\cos{\psi})\sin{\psi}/\sqrt{1-\sin^2{\psi}}.
+\end{equation}
+The factor of 2 is from two solutions of $\cos{\psi}$ to the $\cos^2{\psi} = 1-\sin^2{\psi}$. Combining all the pieces together, we get
+\begin{align}\label{eqn:jac_istar}
+    p(\cos{i_\star}) = \frac{4}{\pi} \int_{\cos{i_\star}}^{1}& \frac{\cos{i_\star}/\cos{\theta}}{\sqrt{\cos^2{\theta}-\cos^2{i_\star}}} \nonumber\\ 
+    &\frac{1}{\sqrt{1-\cos^2{\theta}}} p(\cos{\psi}) d\cos{\theta}.
+\end{align}
 Note that the lower limit of the integral is $\cos{i_\star}$ instead of 0 since $\lvert \cos{\theta}/\cos{i_\star} \rvert \geqslant 1$.
-If $\cos{\psi}$ is uniformly distribution, i.e., $p_{\cos{\psi}} \sim 1/2$, the integral gives $1$, which suggests the $\cos{i_\star}$ is uniformly distributed, as expected.
+If $\cos{\psi}$ is uniformly distribution, i.e., $p(\cos{\psi}) = 1/2$, the integral gives $1$, which suggests the $\cos{i_\star}$ is uniformly distributed, as expected.
 

### 2023-03-21 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 In this study, we aim to gain a deeper understanding of this relationship and develop a statistical approach to infer the stellar obliquity distribution.
+\edits{In Section~\ref{sec:coords}, we find the expression of $\lambda$ and $i_\star$ in terms of the orbital inclination $i_{\rm orb}$, stellar obliquity $\psi$, and the azimuthal angle of the stellar spin axis $\theta$.}
 In Section~\ref{sec:hbm}, we introduce a flexible, hierarchical Bayesian framework that allows us to infer the stellar obliquity distribution of a sample.
@@
 In Section~\ref{sec:jacobian}, we further investigate the relationship between the distributions of $\psi$, $\lambda$, and $i_\star$, and explain why sky-projected stellar obliquities play a predominant role in determining the inferred distribution of stellar obliquities.
+
+{\color{purple}
+
+\section{Relations between $\{\lambda, \lowercase{i}_\star\}$ and $\{\psi, \theta\}$}\label{sec:coords}
+
+In this section, we find the expression of $\lambda$ and $i_\star$ in terms of $\psi$, $\theta$, and $i_{\rm orb}$.
+To do so, we introduce two coordinates that describe the stellar spin axis and its planet's orbital axis.
+
+In the $\{\psi, \theta\}$ coordinate system, we assume the planetary orbital axis is aligned with the $\vb{z}$-axis. The stellar spin axis can be written as 
+\begin{equation}
+    \vb{n_\star} = \sin{\psi}\cos{\theta}\vu*{x} + \sin{\psi}\sin{\theta}\vu*{y} + \cos{\psi}\vu*{z}
+\end{equation}
+
+\begin{figure*}[ht!]
+    \script{coordinate.py}
+    \gridline{
+        \fig{figures/coord_psi.pdf}{0.45\textwidth}{\vspace*{-1.8cm}(a) The $\{\psi, \theta\}$ coordinate system.
+The grey circle corresponds to a constant $\psi$ value and its circumference is proportional to $\sin{\psi}$.}
+        \fig{figures/coord_lam.pdf}{0.45\textwidth}{\vspace*{-1.8cm}(b) The $\{\lambda, i_\star\}$ coordinate system.
+The grey circle corresponds to a constant $i_\star$ value and its circumference is proportional to $\sin{i_\star}$.}
+    }
+    \vspace*{-1cm}
+    \caption{Two coordinate systems that describe the stellar spin axis $\bf{n}_{\star}$ and the planet's orbital axis $\bf{n}_{\rm orb}$. \edits{Add more descriptions here.}}
+    \label{fig:coord}
+\end{figure*}
+
+}
 
@@
     \includegraphics[width=\linewidth]{figures/graph.pdf}
-    \caption{A probabilistic graphical model showing the conditional structure of our hierarchical Bayesian framework for inferring the stellar obliquity distribution of exoplanetary systems.
+    \caption{\edits{Need some edits: A probabilistic graphical model showing the conditional structure of our hierarchical Bayesian framework for inferring the stellar obliquity distribution of exoplanetary systems.
 The population model is parameterized by hyperparameters $\vb*{\beta}$, and it is constrained by the sky-projected stellar obliquity $\lambda_n$, orbital inclination $i_{{\rm orb}, n}$, and, if available, the stellar inclination $i_{\star, n}$ of each $n_{\rm th}$ individual system.
@@
     Properties of the star other than its inclination, such as its radius and rotation period, are represented by $\gamma_{\star, n}$, and these can inform the constraint on $i_{\star, n}$, if available.
-    Obs$_{\star,n}$ contains all observed properties of the star with uncertainties, $\hat{\lambda}_n$ has the measured sky-projected stellar obliquity and its uncertainty, and $\hat{i}_{{\rm orb}, n}$ has measured orbital inclination and its uncertainty.}
+    Obs$_{\star,n}$ contains all observed properties of the star with uncertainties, $\hat{\lambda}_n$ has the measured sky-projected stellar obliquity and its uncertainty, and $\hat{i}_{{\rm orb}, n}$ has measured orbital inclination and its uncertainty.}}
     \label{fig:graph}
@@
 
-\begin{figure*}[ht!]
-    \script{coordinate.py}
-    \gridline{
-        \fig{figures/coord_psi.pdf}{0.45\textwidth}{\vspace*{-1.8cm}(a) The $\{\psi, \theta\}$ coordinate system.
-The grey circle corresponds to a constant $\psi$ value and its circumference is proportional to $\sin{\psi}$.}
-        \fig{figures/coord_lam.pdf}{0.45\textwidth}{\vspace*{-1.8cm}(b) The $\{\lambda, i_\star\}$ coordinate system.
-The grey circle corresponds to a constant $i_\star$ value and its circumference is proportional to $\sin{i_\star}$.}
-    }
-    \vspace*{-1cm}
-    \caption{Two coordinate systems that describe the stellar spin axis $\bf{n}_{\star}$ and the planet's orbital axis $\bf{n}_{\rm orb}$.
-Here we define the observer's line of sight as one of the horizontal axes (represented as the conventional $x$-axis in Cartesian coordinates), and the orbital axis of the planet as the vertical axis (represented as the conventional $z$-axis in Cartesian coordinates).
-We approximate the orbital inclination of the planet to $90\degr$.}
-    \label{fig:coord}
-\end{figure*}
-
 In Figure~\ref{fig:coord}, we present two coordinate systems that describe the stellar spin axis $\bf{n}_{\star}$ for a given orbital axis $\bf{n}_{\rm orb}$.

### 2023-03-22 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
+In the $\{\lambda, i_\star\}$ coordinate system, where the $\vb{x'}$-axis is defined by the observer's line of sight, the stellar spin axis could be written as 
+\begin{equation}
+    \vb{n_\star}' = \cos{i_\star}\vu*{x}' + \sin{i_\star}\sin{\lambda}\vu*{y}' + \sin{i_\star}\cos{\lambda}\vu*{z}'
+\end{equation}
+
+To convert the stellar spin vector $\vb{n_\star}$ from $\vb{x,y,z}$ to $\vb{x',y',z'}$, we apply the rotation matrix to the vector by rotating the vector by $\pi/2 - i_{\rm orb}$ around the $\vb{y'}$-axis.
+\begin{equation}
+      \vb{n_\star}' = \mqty[\sin{i_{\rm orb}} & 0 & \cos{i_{\rm orb}} \\ 0 & 1 & 0 \\ -\cos{i_{\rm orb}} & 0 & \sin{i_{\rm orb}}] \vb{n_\star}  
+\end{equation}
+
+Now we get $\vb{n_\star}'$ in the $\vb{x',y',z'}$ coordinate expressed by $\psi$, $\theta$, and $i_{\rm orb}$.
+\begin{equation}
+    \begin{split}
+    \vb{n_\star}' = (\sin{\psi}\cos{\theta}\sin{i_{\rm orb}}+\cos{\psi}\cos{i_{\rm orb}})\vu*{x}' &\\
+    + \sin{\psi}\sin{\theta}\vu*{y}'& \\
+    + (-\sin{\psi}\cos{\theta}\cos{i_{\rm orb}}+\cos{\psi}\sin{i_{\rm orb}})\vu*{z}'&
+    \end{split}
+\end{equation}
+
+Use the equation, we find the relation between $\lambda$ or $i_\star$ and $\psi$, $\theta$, and $i_{\rm orb}$.
+
+First, comparing the $\vb{x'}$ terms, we find
+\begin{equation}
+    i_\star = \cos[-1](\sin{\psi}\cos{\theta}\sin{i_{\rm orb}}+\cos{\psi}\cos{i_{\rm orb}}).
+\end{equation}
+
+Dividing the $\vb{y'}$ term by the $\vb{z'}$ term, we get
+\begin{equation}
+    \lambda = \tan[-1](\frac{\sin{\psi}\sin{\theta}}{-\sin{\psi}\cos{\theta}\cos{i_{\rm orb}}+\cos{\psi}\sin{i_{\rm orb}}}).
+\end{equation}
+
 \begin{figure*}[ht!]
@@
     \script{graph.py}
-    \includegraphics[width=\linewidth]{figures/graph.pdf}
+    \centering
+    \includegraphics[width=0.8\linewidth]{figures/graph.pdf}
     \caption{\edits{Need some edits: A probabilistic graphical model showing the conditional structure of our hierarchical Bayesian framework for inferring the stellar obliquity distribution of exoplanetary systems.

### 2023-03-23 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 When applying the framework to all exoplanetary systems with measured sky-projected stellar obliquity, we find that the inferred population-level obliquity distribution is unimodal and peaked at zero degrees.
-Misaligned systems have nearly isotropic stellar obliquities, which could have important implications for the formation and evolution of close-in planets.
+Misaligned systems have nearly isotropic stellar obliquities \edits{with no strong clustering near 90$\degr$.
+Particularly, the misaligned Hot Jupiter population presents a broad stellar obliquity distribution, which may suggest planet-planet scattering after a convergent disk migration or planetary Kozai might be the dominant mechanisms sculpting the population.}
 
@@
 
-In this section, we find the expression of $\lambda$ and $i_\star$ in terms of $\psi$, $\theta$, and $i_{\rm orb}$.
-To do so, we introduce two coordinates that describe the stellar spin axis and its planet's orbital axis.
+In this section, we find the expression of $\lambda$ or $i_\star$ in terms of $\psi$, $\theta$, and $i_{\rm orb}$.
+To do so, we introduce two coordinate systems that describe the stellar spin axis $\vb{n_\star}$ and its planet's orbital axis $\vb{n_{\rm orb}}$, shown in Figure~\ref{fig:coord}.
 

### 2023-03-23 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-In this section, we find the expression of $\lambda$ or $i_\star$ in terms of $\psi$, $\theta$, and $i_{\rm orb}$.
-To do so, we introduce two coordinate systems that describe the stellar spin axis $\vb{n_\star}$ and its planet's orbital axis $\vb{n_{\rm orb}}$, shown in Figure~\ref{fig:coord}.
+In this section, we find the expression of $\lambda$ or $i_\star$ in terms of stellar obliquity $\psi$, the azimuthal angle of the stellar spin axis relative to the orbital axis $\theta$, and the orbital inclination $i_{\rm orb}$.
+In Figure~\ref{fig:coord}, we introduce two coordinate systems that describe the stellar spin axis $\vb{n_\star}$ and its planet's orbital axis $\vb{n_{\rm orb}}$.
 
@@
 \begin{equation}
-    \vb{n_\star} = \sin{\psi}\cos{\theta}\vu*{x} + \sin{\psi}\sin{\theta}\vu*{y} + \cos{\psi}\vu*{z}
+    \vb{\hat{n}_\star} = \sin{\psi}\cos{\theta}\vu*{x} + \sin{\psi}\sin{\theta}\vu*{y} + \cos{\psi}\vu*{z}
 \end{equation}
@@
 \begin{equation}
-    \vb{n_\star}' = \cos{i_\star}\vu*{x}' + \sin{i_\star}\sin{\lambda}\vu*{y}' + \sin{i_\star}\cos{\lambda}\vu*{z}'
+    \vb{\hat{n}_\star}' = \cos{i_\star}\vu*{x}' + \sin{i_\star}\sin{\lambda}\vu*{y}' + \sin{i_\star}\cos{\lambda}\vu*{z}'
 \end{equation}

### 2023-04-06 src/tex/ms.tex
new figures

@@
 Misaligned systems have nearly isotropic stellar obliquities \edits{with no strong clustering near 90$\degr$.
-Particularly, the misaligned Hot Jupiter population presents a broad stellar obliquity distribution, which may suggest planet-planet scattering after a convergent disk migration or planetary Kozai might be the dominant mechanisms sculpting the population.}
+Furthermore, the stellar obliquity distribution of misaligned Hot Jupiters presents a broad distribution from $0\degr$ to $180\degr$, preferring candidate Hot Jupiter formation mechanisms, such as planet-planet scattering after a convergent disk migration or planetary Kozai, that could produce a broad distribution.}
 
@@
 
-\section{Relations between $\{\lambda, \lowercase{i}_\star\}$ and $\{\psi, \theta\}$}\label{sec:coords}
+\section{Coordinate Setup and Transformation}\label{sec:coords}
 
@@
 In Figure~\ref{fig:coord}, we introduce two coordinate systems that describe the stellar spin axis $\vb{n_\star}$ and its planet's orbital axis $\vb{n_{\rm orb}}$.
+The setup is similar to the coordinate systems used in \cite{Fabrycky09} but has different coordinate orientations and definitions of the azimuthal angle of the stellar spin.
 
-In the $\{\psi, \theta\}$ coordinate system, we assume the planetary orbital axis is aligned with the $\vb{z}$-axis. The stellar spin axis can be written as 
+The angle of primary interest is the angle between the planet's orbital axis and its host star's spin axis $\psi$.
+To describe the angle in 3$D$, we introduce another angle, the azimuthal angle of the stellar spin around the orbital axis $\theta$.
+As shown in the left panel in Figure~\ref{fig:coord}, in the $\{\psi, \theta\}$ coordinate system, we put the planetary orbital axis on the $\vb{z}$-axis.
+The stellar spin axis, described by $\psi$ and $\theta$, can be written as
 \begin{equation}
@@
 
-In the $\{\lambda, i_\star\}$ coordinate system, where the $\vb{x'}$-axis is defined by the observer's line of sight, the stellar spin axis could be written as 
-\begin{equation}
+Next, we introduce an observer-oriented coordinate system.
+Since we neither know the projected orbital axis in the sky plane nor the projected stellar spin axis, we conveniently set the projected orbital axis to the $\vb*{z}$-axis.
+In the $\{\lambda, i_\star\}$ coordinate system, shown in the right panel of Figure~\ref{fig:coord}, we point the $\vb{x'}$-axis to the observer.
+The stellar spin axis could be written as
+\begin{equation}\label{eqn:nstar1}
     \vb{\hat{n}_\star}' = \cos{i_\star}\vu*{x}' + \sin{i_\star}\sin{\lambda}\vu*{y}' + \sin{i_\star}\cos{\lambda}\vu*{z}'
@@
 
-To convert the stellar spin vector $\vb{n_\star}$ from $\vb{x,y,z}$ to $\vb{x',y',z'}$, we apply the rotation matrix to the vector by rotating the vector by $\pi/2 - i_{\rm orb}$ around the $\vb{y'}$-axis.
+To transform the stellar spin vector $\vb{\hat{n}_\star}$ from the $\vb*{xyz}$ coordinate to the $\vb*{x'y'z'}$ coordinate, we rotate the $\vb*{xyz}$ coordinate by an angle $\pi/2 - i_{\rm orb}$ about the $\vb{y'}$-axis (the same as $\vb{y}$-axis).
+Applying the rotation matrix such that
 \begin{equation}
-      \vb{n_\star}' = \mqty[\sin{i_{\rm orb}} & 0 & \cos{i_{\rm orb}} \\ 0 & 1 & 0 \\ -\cos{i_{\rm orb}} & 0 & \sin{i_{\rm orb}}] \vb{n_\star}  
+      \vb{\hat{n}_\star}' = \mqty[\sin{i_{\rm orb}} & 0 & \cos{i_{\rm orb}} \\ 0 & 1 & 0 \\ -\cos{i_{\rm orb}} & 0 & \sin{i_{\rm orb}}] \vb{\hat{n}_\star},
 \end{equation}
-
-Now we get $\vb{n_\star}'$ in the $\vb{x',y',z'}$ coordinate expressed by $\psi$, $\theta$, and $i_{\rm orb}$.
-\begin{equation}
+we can now find the expression of $\vb{\hat{n}_\star}'$ in the $\vb*{x'y'z'}$ coordinate in terms of $\psi$, $\theta$, and $i_{\rm orb}$:
+\begin{equation}\label{eqn:nstar2}
     \begin{split}
-    \vb{n_\star}' = (\sin{\psi}\cos{\theta}\sin{i_{\rm orb}}+\cos{\psi}\cos{i_{\rm orb}})\vu*{x}' &\\
+    \vb{\hat{n}_\star}' = (\sin{\psi}\cos{\theta}\sin{i_{\rm orb}}+\cos{\psi}\cos{i_{\rm orb}})\vu*{x}' &\\
     + \sin{\psi}\sin{\theta}\vu*{y}'& \\
-    + (-\sin{\psi}\cos{\theta}\cos{i_{\rm orb}}+\cos{\psi}\sin{i_{\rm orb}})\vu*{z}'&
+    + (-\sin{\psi}\cos{\theta}\cos{i_{\rm orb}}+\cos{\psi}\sin{i_{\rm orb}})\vu*{z}'&.
     \end{split}
@@
 
-Use the equation, we find the relation between $\lambda$ or $i_\star$ and $\psi$, $\theta$, and $i_{\rm orb}$.
-
-First, comparing the $\vb{x'}$ terms, we find
+Comparing Equation~(\ref{eqn:nstar2}) with Equation~(\ref{eqn:nstar1}), we find the expression of $\lambda$ or $i_\star$ in terms of $\psi$, $\theta$, and $i_{\rm orb}$.
+First, from the $\vb{\hat{x}'}$ terms, we derive
 \begin{equation}
@@
 \end{equation}
-
-Dividing the $\vb{y'}$ term by the $\vb{z'}$ term, we get
+Next, dividing the $\vb{\hat{y}'}$ terms by the $\vb{\hat{z}'}$ terms, we find
 \begin{equation}
@@
 
+\edits{In this framework, we limit $i_{\rm orb}$ from $0\degr$ to $90\degr$. To preserve the symmetry, we allow $i_{\rm orb}$ to vary from $0\degr$ to $180\degr$.}
 With the design, the hyperparameter and parameter priors are the following:
@@
     w_{0,1} &\sim {\rm Dirichlet}(1, 1) \nonumber\\
-    a_0 &\sim \mathcal{U}(0, 50) \nonumber\\
-    b_0 &\sim \mathcal{U}(0, 1) \nonumber\\
-    a_1 &\sim \mathcal{U}(0, 10) \nonumber\\
-    b_1 &\sim \mathcal{U}(0, 10) \nonumber\\
-    \cos{i}_{\star,n} &\sim \mathcal{U}(0, 1) \nonumber\\
-    \cos{i}_{{\rm orb},n} &\sim \mathcal{U}(0, 1) \nonumber\\
-    \lambda_{\star,n} &\sim \mathcal{U}(-\pi, \pi).
+    p_{0,1} &\sim \mathcal{U}(0, 1) \nonumber\\
+    q_{0,1} &\sim \mathcal{U}(0, 1) \nonumber\\
+    \theta_n &\sim \mathcal{U}(0, \pi)  \nonumber\\
+    \cos{i}_{{\rm orb},n} &\sim \mathcal{U}(0, 1).
 \end{align}
+If $\gamma_{\star, n}$ is available, we construct Normal distributions with means and standard deviations from measurements.
+
 The likelihood functions follow:
@@
 \end{align}
-If $\gamma_{\star, n}$ is available, we construct Normal distributions with means and standard deviations from measurements.
 Here we use uniform hyperpriors for the Beta distribution.
@@
 \end{figure}
+
+\edits{EDIT TO HERE TODAY}
 

### 2023-04-06 src/scripts/coordinate.py

@@ -26,13 +26,13 @@ x = np.outer(np.cos(u), np.sin(v))
 y = np.outer(np.sin(u), np.sin(v))
 z = np.outer(np.ones(np.size(u)), np.cos(v))
 
-# plot the surface
+#Plot the surface
 ax.plot_surface(x, y, z, rstride=4, cstride=4, color='b', linewidth=0, alpha=0.1)
 
 elev = 20.0
 rot = 30.0 / 180 * np.pi
 
-# calculate vectors for "vertical" circle
+#calculate vectors for "vertical" circle
 a = np.array([-np.sin(elev / 180 * np.pi), 0, np.cos(elev / 180 * np.pi)])
 b = np.array([0, 1, 0])
 b = b * np.cos(rot) + np.cross(a, b) * np.sin(rot) + a * np.dot(a, b) * (1 - np.cos(rot))
@@ -51,9 +51,9 @@ ax.quiver(0, 0, 0, 1.8, 0, 0, color='k', arrow_length_ratio=0.05, lw=1.5) # x-ax
 ax.quiver(0, 0, 0, 0, 1.5, 0, color='k', arrow_length_ratio=0.05, lw=1.5) # y-axis
 ax.quiver(0, 0, 0, 0, 0, 1.5, color='k', arrow_length_ratio=0.05, lw=1.5) # z-axis
 
-ax.text(2.08, 0, -0.05, r'$\vb*{X}$',fontsize=14)
-ax.text(0, 0, 1.55, r"$\vb*{Z} (\vb*{n}_{\textbf{orb}})$",fontsize=14)
-ax.text(0.05, 1.6, -0.05, r'$\vb*{Y}$',fontsize=14)
+ax.text(2., 0, -0.05, r'$\vb*{x}$',fontsize=14)
+ax.text(0, 0, 1.55, r"$\vb*{z} (\vb*{\hat{n}}_{\textbf{orb}})$",fontsize=14)
+ax.text(0.05, 1.6, -0.05, r'$\vb*{y}$',fontsize=14)
 
 # make the panes transparent
 ax.xaxis.set_pane_color((1.0, 1.0, 1.0, 0.0))
@@ -81,7 +81,6 @@ ax.quiver(0,0,0,u,v,w,color='k', arrow_length_ratio=0.1, lw=1.5, normalize=True)
[…]

### 2023-04-06 src/scripts/graph.py

@@ -14,39 +14,50 @@ rc('text.latex', preamble=r'\usepackage{physics}')
 
 import daft
 
-pgm = daft.PGM(observed_style="shaded",node_unit=1, node_ec='k',dpi=150,line_width=0.8)
+import daft
+
+pgm = daft.PGM(observed_style="shaded",node_unit=1, node_ec='k',dpi=100,line_width=0.8)
                
 # Hierarchical parameters.
-pgm.add_node("beta", r"$\vb*{\beta}$", 2, 2.8)
+pgm.add_node("beta", r"$\vb*{\beta}$", 2, 4.)
 
 # Latent variable.
-pgm.add_node("psi", r"$\psi_n$", 2, 2.)
+pgm.add_node("psi", r"$\psi_n$", 2, 3.)
 
-pgm.add_node("lambda", r"$\lambda_n$", 1., 1.2)
+pgm.add_node("theta", r"$\theta_n$", 3, 3.)
 
-pgm.add_node("iorb", r"$i_{{\rm orb},n}$", 2., 1.2)
+pgm.add_node("iorb", r"$i_{{\rm orb},n}$", 4, 3.)
 
-pgm.add_node("istar", r"$i_{\star,n}$", 3., 1.2)
+pgm.add_node("lambda", r"$\lambda_n$", 2., 1.5)
 
-pgm.add_node("pstar", r"$\gamma_{\star,n}$", 4., 1.2)
+pgm.add_node("istar", r"$i_{\star,n}$", 3, 1.5)
 
+pgm.add_node("pstar", r"$\gamma_{\star,n}$", 4, 1.5)
+
+
+# Observed variable.
+pgm.add_node("obs_iorb", r"$\hat{i}_{{\rm orb}, n}$", 4, 2.2, observed=True)
+
+pgm.add_node("obs_lambda", r"$\hat{\lambda}_n$", 2., 0.7, observed=True)
+
+pgm.add_node("Obs", r"Obs$_{\star, n}$", 3, 0.7, observed=True, aspect=1., fontsize=8)
+
+
+# Add edges
 pgm.add_edge("beta", "psi")
 
 pgm.add_edge("psi", "lambda")
-
-pgm.add_edge("psi", "iorb")
+pgm.add_edge("theta", "lambda")
+pgm.add_edge("iorb", "lambda")
 
 pgm.add_edge("psi", "istar")
+pgm.add_edge("theta", "istar")
[…]

### 2023-04-06 src/scripts/psi.py

@@ -4,6 +4,7 @@ os.environ["PATH"] += os.pathsep + str(Path.home() / "bin")
 
 import sys
 import subprocess
+from datetime import datetime
 
 import paths
 import matplotlib.pyplot as plt
@@ -103,39 +104,46 @@ if __name__ == '__main__':
 
     nplanet = len(Lam)
     with pm.Model() as randinc:
-        
-        ### hyper priors
-        w = pm.Dirichlet('w', np.ones(2))
 
-        a0 = pm.Uniform('a0', lower=0, upper=50)
-        b0 = pm.Uniform('b0', lower=0, upper=1)
-        
-        a1 = pm.Uniform('a1', lower=0, upper=10)
-        b1 = pm.Uniform('b1', lower=0, upper=10)
-        
-        components = [pm.Beta.dist(a0,b0), pm.Beta.dist(a1,b1)]
-                            
-        # flat prior on cosi
-        cosi = pm.Uniform('cosi', lower=0., upper=1., shape=nplanet)
-        sini = pm.Deterministic('sini', at.sqrt(1-cosi**2))
-        i = pm.Deterministic('i', at.arccos(cosi))
-        
-        # flat priors on λ
-        λ = pm.Uniform('λ', lower=0, upper=np.pi, shape=nplanet)
-        cosλ = pm.Deterministic('cosλ', at.cos(λ))
+        ncomps = 2
 
-        cosψ = pm.Deterministic('cosψ', cosλ*sini)
-        ψ = pm.Deterministic('ψ', at.arccos(cosψ))
+        # hyperprior
+        w = pm.Dirichlet('w', np.ones(ncomps))
+
+        if ncomps > 1:
+            μ = pm.Uniform('μ', lower=0., upper=1., shape=ncomps, 
+                           transform=pm.distributions.transforms.Ordered(), 
+                           initval=np.array([0.5,0.9]))
+        else:
+            μ = pm.Uniform('μ', lower=0., upper=1., shape=ncomps)
+            
+        logκ = pm.Normal('logκ', 3., shape=ncomps)
+        κ = pm.Deterministic('κ', pm.math.exp(logκ))
+      
+        a = pm.Deterministic('a', μ*κ)
+        b = pm.Deterministic('b', (1.-μ)*κ)
         
-        cosθ = pm.Deterministic('cosθ', cosi/at.sin(ψ))
[…]

### 2023-04-06 src/scripts/psi_plot.py

@@ -28,26 +28,33 @@ plt.rcParams['ytick.minor.width'] =  1.0
 plt.rcParams['lines.markeredgewidth'] =  1.0
 
 # plot the cospsi distribution
-all_randinc = az.from_netcdf(paths.data / "all_randinc.nc")
+all_noistar = az.from_netcdf(paths.data / "all_noistar.nc")
 
 x = np.linspace(1e-5,1-1e-5,1000)
 
-post = all_randinc.posterior
-all_randinc_draws = np.zeros(shape=(len(x),4000))
+post = all_noistar.posterior
+comp0 = np.zeros(shape=(len(x),4000))
+comp1 = np.zeros(shape=(len(x),4000))
 for a in range(4):
     for b in range(1000):
-        all_randinc_draws[:, a*1000+b] = (post.w[a,b,0].values*beta.pdf(x, post.a0[a,b], post.b0[a,b])
-                                          +post.w[a,b,1].values*beta.pdf(x, post.a1[a,b], post.b1[a,b]))
+        comp0[:, a*1000+b] = post.w[a,b,0].values*beta.pdf(x, post.μ[a,b,0].values*post.κ[a,b,0].values, (1-post.μ[a,b,0]).values*post.κ[a,b,0].values)
+        comp1[:, a*1000+b] = post.w[a,b,1].values*beta.pdf(x, post.μ[a,b,1].values*post.κ[a,b,1].values, (1-post.μ[a,b,1]).values*post.κ[a,b,1].values)
 
 # make the figure
-plt.figure(figsize=(3.5,2.7),dpi=110)
+plt.figure(figsize=(3.5,2.7),dpi=150)
 
-q025, q16, q50, q84, q975 = np.percentile(all_randinc_draws, [2.5, 16, 50, 84, 97.5], axis=1)
-plt.plot(2*x-1, q50, color='C0', lw=1.2)
+q025, q16, q50, q84, q975 = np.percentile(comp0, [2.5, 16, 50, 84, 97.5], axis=1)/2
+plt.plot(2*x-1, q50, '-.', c='k', lw=1., zorder=10)
+
+q025, q16, q50, q84, q975 = np.percentile(comp1, [2.5, 16, 50, 84, 97.5], axis=1)/2
+plt.plot(2*x-1, q50, '--', c='k', lw=1., zorder=10)
+
[…]

### 2023-04-06 src/scripts/sim.py

@@ -5,6 +5,7 @@ os.environ["PATH"] += os.pathsep + str(Path.home() / "bin")
 
 import subprocess
 import sys
+from datetime import datetime
 
 import pymc as pm
 import arviz as az
@@ -26,8 +27,8 @@ x = np.linspace(1e-5,1-1e-5,1000)
 def posteriors(this_model):
     
     nsample = 200
-    err_istar = 10*np.pi/180
-    err_lam = 8*np.pi/180
+    err_istar = 10.*np.pi/180
+    err_lam = 8.*np.pi/180
 
     with this_model:    
         idata = pm.sample(chains=4, draws=50, tune=0)
@@ -38,176 +39,286 @@ def posteriors(this_model):
     true_lam = idata.posterior.λ.values.ravel()
     obs_lam = true_lam + err_lam*np.random.normal(size=nsample)
 
-    # Limit obs_istar to [0, pi/2]
-    obs_istar[obs_istar>np.pi/2] = np.pi-obs_istar[obs_istar>np.pi/2]
+    # Not neccessary; limit obs_istar to [0, pi/2]
+    # obs_istar[obs_istar>np.pi/2] = np.pi-obs_istar[obs_istar>np.pi/2]
 
     # Limit obs_lam to [0, pi]
     obs_lam[obs_lam<0] = -obs_lam[obs_lam<0]
 
-    with pm.Model() as model_istar:
+    with pm.Model() as model:
 
+        ncomps = 1
+        
         # hyperprior
-        a = pm.Uniform('a', lower=0, upper=10)
-        b = pm.Uniform('b', lower=0, upper=10)
-
-        # flat prior on cosi
-        cosi = pm.Uniform('cosi', lower=0., upper=1., shape=nsample)
-        sini = pm.Deterministic('sini', at.sqrt(1-cosi**2))
-
-        i = pm.Deterministic('i', at.arccos(cosi))
-
-        # flat priors on ψ
-        λ = pm.Uniform('λ', lower=0, upper=np.pi, shape=nsample)
-        cosλ = pm.Deterministic('cosλ', np.cos(λ))
-
-        cosψ = pm.Deterministic('cosψ', cosλ*sini)
-        ψ = pm.Deterministic('ψ', np.arccos(cosψ))
+        w = pm.Dirichlet('w', np.ones(ncomps))
+        
+        if ncomps > 1:
[…]

### 2023-04-06 src/scripts/sim_plot.py

@@ -39,14 +39,15 @@ x = np.linspace(1e-5,1-1e-5,1000)
 # load the posteriors from the MCMC and calculate psi dists
 def psi_dist_draws(model_name):
 
-    istar_idata = az.from_netcdf(paths.data / "simulation" / (model_name + "_istar.nc"))
+    idata = az.from_netcdf(paths.data / "simulation" / (model_name + ".nc"))
     noistar_idata = az.from_netcdf(paths.data / "simulation" / (model_name + "_noistar.nc"))
+    nolam_idata = az.from_netcdf(paths.data / "simulation" / (model_name + "_nolam.nc"))
 
-    post = istar_idata.posterior
-    istar_draws = np.zeros(shape=(len(x),4000))
+    post = idata.posterior
+    draws = np.zeros(shape=(len(x),4000))
     for a in range(4):
         for b in range(1000):
-            istar_draws[:, a*1000+b] = beta.pdf(x, post.a[a,b], post.b[a,b])
+            draws[:, a*1000+b] = beta.pdf(x, post.a[a,b], post.b[a,b])
 
     post = noistar_idata.posterior
     noistar_draws = np.zeros(shape=(len(x),4000))
@@ -54,16 +55,22 @@ def psi_dist_draws(model_name):
         for b in range(1000):
             noistar_draws[:, a*1000+b] = beta.pdf(x, post.a[a,b], post.b[a,b])
 
-    return istar_draws, noistar_draws
+    post = nolam_idata.posterior
+    nolam_draws = np.zeros(shape=(len(x),4000))
+    for a in range(4):
+        for b in range(1000):
+            nolam_draws[:, a*1000+b] = beta.pdf(x, post.a[a,b], post.b[a,b])
 
-uni_istar_draws, uni_noistar_draws = psi_dist_draws("uni")
-norm1_istar_draws, norm1_noistar_draws = psi_dist_draws("norm1")
[…]

### 2023-04-06 src/scripts/psi.py
fix package version

@@ -9,7 +9,7 @@ from datetime import datetime
 import paths
 import matplotlib.pyplot as plt
 import pymc as pm
-import aesara.tensor as at
+import pytensor.tensor as at
 import numpy as np
 import arviz as az
 from scipy.stats import beta

### 2023-04-06 src/scripts/psi_plot.py

@@ -8,7 +8,7 @@ import subprocess
 import paths
 import matplotlib.pyplot as plt
 import pymc as pm
-import aesara.tensor as at
+import pytensor.tensor as at
 import numpy as np
 import arviz as az
 from scipy.stats import beta

### 2023-04-06 src/scripts/sim.py

@@ -9,7 +9,7 @@ from datetime import datetime
 
 import pymc as pm
 import arviz as az
-import aesara.tensor as at
+import pytensor.tensor as at
 
 import matplotlib.pyplot as plt
 import numpy as np

### 2023-04-06 src/scripts/sim_plot.py

@@ -8,7 +8,7 @@ import sys
 
 import pymc as pm
 import arviz as az
-import aesara.tensor as at
+import pytensor.tensor as at
 
 import matplotlib.pyplot as plt
 import numpy as np

### 2023-04-07 src/scripts/sim_plot.py
dpi

@@ -70,7 +70,7 @@ norm3_draws, norm3_noistar_draws, norm3_nolam_draws = psi_dist_draws("norm3")
 
 
 ### Make the plot ###
-fig, big_axes = plt.subplots(figsize=(6.5,7),dpi=110,nrows=4,ncols=1,sharey=True) 
+fig, big_axes = plt.subplots(figsize=(6.5,7),dpi=150,nrows=4,ncols=1,sharey=True) 
 
 for row, big_ax in enumerate(big_axes, start=1):
     

### 2023-04-08 src/tex/ms.tex
[showyourwork] trust me
[showyourwork] overleaf sync

@@
 Misaligned systems have nearly isotropic stellar obliquities \edits{with no strong clustering near 90$\degr$.
-Furthermore, the stellar obliquity distribution of misaligned Hot Jupiters presents a broad distribution from $0\degr$ to $180\degr$, preferring candidate Hot Jupiter formation mechanisms, such as planet-planet scattering after a convergent disk migration or planetary Kozai, that could produce a broad distribution.}
+Furthermore, misaligned Hot Jupiters have a diverse range of stellar obliquities, suggesting that dynamic mechanisms, such as planet-planet scattering after a convergent disk migration or planetary Kozai, are likely responsible for producing the population.}
 
@@
 
-The stellar obliquity $\psi$ is the angle between a planet's orbital axis $\bf{n}_{\rm orb}$ and the host star's spin axis $\bf{n}_{\star}$. 
+The stellar obliquity $\psi$ is the angle between a planet's orbital axis $\vb{\hat{n}_{\rm orb}}$ and the host star's spin axis $\vb{\hat{n}_\star}$. 
 This angle is an important tracer of a planetary system's formation environment and dynamical evolution.
@@
 
-When observing an exoplanet, typically only the sky-projected stellar obliquity $\lambda$, the angle between the projections of $\bf{n}_{\rm orb}$ and $\bf{n}_{\star}$ onto the plane of the sky, can be measured.
+When observing an exoplanet, typically only the sky-projected stellar obliquity $\lambda$, the angle between the projections of $\vb{\hat{n}_{\rm orb}}$ and $\vb{\hat{n}_\star}$ onto the plane of the sky, can be measured.
 This measurement is primarily obtained via the Rossiter-McLaughlin effect \citep{Rossiter24, McLaughlin24}.
@@
 In this study, we aim to gain a deeper understanding of this relationship and develop a statistical approach to infer the stellar obliquity distribution.
-\edits{In Section~\ref{sec:coords}, we find the expression of $\lambda$ and $i_\star$ in terms of the orbital inclination $i_{\rm orb}$, stellar obliquity $\psi$, and the azimuthal angle of the stellar spin axis $\theta$.}
+\edits{In Section~\ref{sec:coords}, we find the expression of sky-projected stellar olibquity $\lambda$ and stellar inclination $i_\star$ in terms of the orbital inclination $i_{\rm orb}$, stellar obliquity $\psi$, and the azimuthal angle of the stellar spin axis $\theta$.
 In Section~\ref{sec:hbm}, we introduce a flexible, hierarchical Bayesian framework that allows us to infer the stellar obliquity distribution of a sample.
-In Section~\ref{sec:applications}, we apply the framework to simulated data and real observations and show that the inferred stellar obliquity distribution from sky-projected stellar obliquities is robust even if the $i_\star$ information is not provided.
-In Section~\ref{sec:jacobian}, we further investigate the relationship between the distributions of $\psi$, $\lambda$, and $i_\star$, and explain why sky-projected stellar obliquities play a predominant role in determining the inferred distribution of stellar obliquities.
+In Section~\ref{sec:simulation}, we examine the framework with simulated data and show that the inferred stellar obliquity distribution from sky-projected stellar obliquities is robust even if the $i_\star$ information is not provided.
+Lastly, in Section~\ref{sec:observation}, we apply the framework to real observations and derive the stellar obliquity distribution for exoplanets.
+We discuss the implication of the stellar obliquity distribution for Hot Jupiter origins.
+}
 
@@
 In this section, we find the expression of $\lambda$ or $i_\star$ in terms of stellar obliquity $\psi$, the azimuthal angle of the stellar spin axis relative to the orbital axis $\theta$, and the orbital inclination $i_{\rm orb}$.
-In Figure~\ref{fig:coord}, we introduce two coordinate systems that describe the stellar spin axis $\vb{n_\star}$ and its planet's orbital axis $\vb{n_{\rm orb}}$.
-The setup is similar to the coordinate systems used in \cite{Fabrycky09} but has different coordinate orientations and definitions of the azimuthal angle of the stellar spin.
+In Figure~\ref{fig:coord}, we introduce two coordinate systems that describe the stellar spin axis $\vb{\hat{n}_\star}$ and its planet's orbital axis $\vb{\hat{n}_{\rm orb}}$.
+The setup is similar to the coordinate system setup in \cite{Fabrycky09} but has a different definition of the azimuthal angle of the stellar spin and coordinate orientations.
 
-The angle of primary interest is the angle between the planet's orbital axis and its host star's spin axis $\psi$.
-To describe the angle in 3$D$, we introduce another angle, the azimuthal angle of the stellar spin around the orbital axis $\theta$.
-As shown in the left panel in Figure~\ref{fig:coord}, in the $\{\psi, \theta\}$ coordinate system, we put the planetary orbital axis on the $\vb{z}$-axis.
-The stellar spin axis, described by $\psi$ and $\theta$, can be written as
+\begin{figure*}[ht!]
+    \script{coordinate.py}
+    \gridline{
+        \fig{figures/coord_psi.pdf}{0.45\textwidth}{\vspace*{-1.8cm}(a) The $\{\psi, \theta\}$ coordinate system.
+\edits{The planetary orbital axis is set to be aligned with $\vb*{z}$-axis.} The grey circle corresponds to a constant $\psi$ value and its circumference is proportional to $\sin{\psi}$.}
+        \fig{figures/coord_lam.pdf}{0.45\textwidth}{\vspace*{-1.8cm}(b) The $\{\lambda, i_\star\}$ coordinate system.
+\edits{The observer's light of sight is set to be aligned with $\vb*{x'}$-axis.} The grey circle corresponds to a constant $i_\star$ value and its circumference is proportional to $\sin{i_\star}$.}
+    }
+    \vspace*{-1cm}
+    \caption{Two coordinate systems that describe the stellar spin axis $\vb{\hat{n}_\star}$ and the planet's orbital axis $\vb{\hat{n}_{\rm orb}}$.
+\edits{The $\{\psi, \theta\}$ coordinate system setup is motivated by the physical properties of a planetary system.
+The $\{\lambda, i_\star\}$ coordinate system setup is motivated by observables.} }
+    \label{fig:coord}
+\end{figure*}
+
+The $\{\psi, \theta\}$ coordinate system, shown in the left panel in Figure~\ref{fig:coord}, is designed to describe the physical properties of a system.
+We set the planetary orbital axis $\vb{\hat{n}_{\rm orb}}$ as the $\vb{z}$-axis.
+To define the stellar spin vector, we introduce the azimuthal angle of the vector around the orbital axis, $\theta$. 
+The stellar spin axis $\vb{\hat{n}_\star}$ can be written in terms of $\psi$ and $\theta$ as
 \begin{equation}
-    \vb{\hat{n}_\star} = \sin{\psi}\cos{\theta}\vu*{x} + \sin{\psi}\sin{\theta}\vu*{y} + \cos{\psi}\vu*{z}
+    \vb{\hat{n}_\star} = \sin{\psi}\cos{\theta}\vu*{x} + \sin{\psi}\sin{\theta}\vu*{y} + \cos{\psi}\vu*{z}.
 \end{equation}
+Particularly, we are interested in finding the distribution of the angle $\psi$.
 
-Next, we introduce an observer-oriented coordinate system.
-Since we neither know the projected orbital axis in the sky plane nor the projected stellar spin axis, we conveniently set the projected orbital axis to the $\vb*{z}$-axis.
-In the $\{\lambda, i_\star\}$ coordinate system, shown in the right panel of Figure~\ref{fig:coord}, we point the $\vb{x'}$-axis to the observer.
-The stellar spin axis could be written as
+The $\{\lambda, i_\star\}$ coordinate system, shown in the right panel in Figure~\ref{fig:coord}, is designed to describe observables of a system.
+We set the observer's light of sight as the $\vb*{x'}$-axis.
+Since only the difference between the sky-projected orbital axis and the sky-projected stellar spin axis can be measured, we conveniently align the projected orbital axis with the $\vb*{z}$-axis.
+The stellar spin axis could be written in terms of $\lambda$ and $i_\star$ as 
 \begin{equation}\label{eqn:nstar1}
-    \vb{\hat{n}_\star}' = \cos{i_\star}\vu*{x}' + \sin{i_\star}\sin{\lambda}\vu*{y}' + \sin{i_\star}\cos{\lambda}\vu*{z}'
+    \vb{\hat{n}_\star}' = \cos{i_\star}\vu*{x}' + \sin{i_\star}\sin{\lambda}\vu*{y}' + \sin{i_\star}\cos{\lambda}\vu*{z}'.
 \end{equation}
@@
 To transform the stellar spin vector $\vb{\hat{n}_\star}$ from the $\vb*{xyz}$ coordinate to the $\vb*{x'y'z'}$ coordinate, we rotate the $\vb*{xyz}$ coordinate by an angle $\pi/2 - i_{\rm orb}$ about the $\vb{y'}$-axis (the same as $\vb{y}$-axis).
+Conventionally, the inclination angle of the orbital axis $\vb{\hat{n}_{\rm orb}}$ is assumed to be less than $90\degr$, as what we adopt here.
 Applying the rotation matrix such that
@@
 \end{equation}
-we can now find the expression of $\vb{\hat{n}_\star}'$ in the $\vb*{x'y'z'}$ coordinate in terms of $\psi$, $\theta$, and $i_{\rm orb}$:
+we find the expression of $\vb{\hat{n}_\star}'$ in the $\vb*{x'y'z'}$ coordinate in terms of $\psi$, $\theta$, and $i_{\rm orb}$:
 \begin{equation}\label{eqn:nstar2}
@@
 
-Comparing Equation~(\ref{eqn:nstar2}) with Equation~(\ref{eqn:nstar1}), we find the expression of $\lambda$ or $i_\star$ in terms of $\psi$, $\theta$, and $i_{\rm orb}$.
-First, from the $\vb{\hat{x}'}$ terms, we derive
+Equating Equation~(\ref{eqn:nstar2}) and Equation~(\ref{eqn:nstar1}), we find the expression of $\lambda$ or $i_\star$ in terms of $\psi$, $\theta$, and $i_{\rm orb}$.
+First, from the $\vb{\hat{x}'}$ terms, we get
 \begin{equation}
@@
 \end{equation}
-Next, dividing the $\vb{\hat{y}'}$ terms by the $\vb{\hat{z}'}$ terms, we find
+Next, dividing the $\vb{\hat{y}'}$ terms by the $\vb{\hat{z}'}$ terms, we get
 \begin{equation}
@@
 \end{equation}
-
-\begin{figure*}[ht!]
-    \script{coordinate.py}
-    \gridline{
-        \fig{figures/coord_psi.pdf}{0.45\textwidth}{\vspace*{-1.8cm}(a) The $\{\psi, \theta\}$ coordinate system.
-The grey circle corresponds to a constant $\psi$ value and its circumference is proportional to $\sin{\psi}$.}
-        \fig{figures/coord_lam.pdf}{0.45\textwidth}{\vspace*{-1.8cm}(b) The $\{\lambda, i_\star\}$ coordinate system.
-The grey circle corresponds to a constant $i_\star$ value and its circumference is proportional to $\sin{i_\star}$.}
-    }
-    \vspace*{-1cm}
-    \caption{Two coordinate systems that describe the stellar spin axis $\bf{n}_{\star}$ and the planet's orbital axis $\bf{n}_{\rm orb}$. \edits{Add more descriptions here.}}
-    \label{fig:coord}
-\end{figure*}
+These two relations will be used in the hierarchical Bayesian framework to infer the stellar obliquity distribution.
 
@@
 If a measurement of the stellar inclination $i_\star$ is available, it can be provided or inferred from the stellar rotation period $P_{\rm rot}$, stellar radius $R_\star$, and sky-projected rotational broadening velocity $v\sin{i}_\star$.
-In the absence of an $i_\star$, $P_{\rm rot}$, or $v\sin{i}_\star$ measurement, we marginalize over the value of $i_\star$ using an isotropic prior distribution.
+In the absence of an $i_\star$, $P_{\rm rot}$, or $v\sin{i}_\star$ measurement, the stellar obliquity $\psi$ distribution is inferred without $i_\star$ likelihoods.
 
@@
     \includegraphics[width=0.8\linewidth]{figures/graph.pdf}
-    \caption{\edits{Need some edits: A probabilistic graphical model showing the conditional structure of our hierarchical Bayesian framework for inferring the stellar obliquity distribution of exoplanetary systems.
-The population model is parameterized by hyperparameters $\vb*{\beta}$, and it is constrained by the sky-projected stellar obliquity $\lambda_n$, orbital inclination $i_{{\rm orb}, n}$, and, if available, the stellar inclination $i_{\star, n}$ of each $n_{\rm th}$ individual system.
-Each system's stellar obliquity $\psi_n$ is inferred simultaneously based on these data.
+    \caption{\edits{A probabilistic graphical model showing the conditional structure of our hierarchical Bayesian framework for inferring the stellar obliquity distribution of exoplanetary systems.
+The population model is parameterized by hyperparameters $\vb*{\beta}$, and it is constrained by the stellar obliquity $\psi_n$ of each $n_{\rm th}$ individual system.
+The stellar obliquity $\psi_n$, azimuthal angle $\theta_n$, and orbital inclination $i_{{\rm orb}, n}$ are constrained by the sky-projected stellar obliquity $\lambda_n$, and, if available, the stellar inclination $i_{\star, n}$.
     Properties of the star other than its inclination, such as its radius and rotation period, are represented by $\gamma_{\star, n}$, and these can inform the constraint on $i_{\star, n}$, if available.
-    Obs$_{\star,n}$ contains all observed properties of the star with uncertainties, $\hat{\lambda}_n$ has the measured sky-projected stellar obliquity and its uncertainty, and $\hat{i}_{{\rm orb}, n}$ has measured orbital inclination and its uncertainty.}}
+    $\hat{\lambda}_n$ has the measured sky-projected stellar obliquity and its uncertainty, Obs$_{\star,n}$ contains all observed properties of the star with uncertainties, and $\hat{i}_{{\rm orb}, n}$ has measured orbital inclination and its uncertainty.}}
     \label{fig:graph}
@@
 We aim to constrain a set of hyperparameters $\bm{\beta}$ that describe the stellar obliquity distribution.
-The parameter set $\bm{\beta}$ is constrained by $N$ individual systems, where each $\psi_n$ is simultaneously fit based on the entire sample of sky-projected stellar obliquities $\lambda_n$, orbital inclinations $i_{{\rm orb},n}$, and if available, the stellar inclination $i_{\star, n}$.
+The parameter set $\bm{\beta}$ is constrained by $N$ individual systems, where each $\psi_n$ is simultaneously fit based on the entire sample of sky-projected stellar obliquities $\lambda_n$, and if available, the stellar inclination $i_{\star, n}$.
 The parameter $\gamma_{\star, n}$ contains all stellar properties other than $i_{\star,n}$, such as the stellar rotation period $P_{{\rm rot},n}$, radius $R_{\star,n}$, and projected rotational velocity $v\sin{i_{\star,n}}$, if they are known.
@@
 
-% If the rotational modulation method is used to constrain the stellar inclination, the $i_{\star,n}$ and $\gamma_{\star,n}$ are constrained by rotational modulation in photometric or spectroscopic time series and observed sky-projected stellar rotational line broadening $\hat{v}\sin{i_{\star,n}}$ \citep[e.g.,][]{Masuda20}, where we have $v\sin{i_\star} = 2 \pi R_\star / P_{\rm rot}$.
-% If the gravity-darkening method is used, the $i_{\star,n}$ could be constrained by the anomaly in transit light curves \citep{Barnes09}.
-If asteroseismology is used, the $i_{\star,n}$ could be constrained from the mode splitting in the Fourier transform of photometric time series \citep{Chaplin13}.
-
 For the stellar obliquity distribution, we model the $\cos{\psi}$ distribution instead of $\psi$ distribution to understand if the stellar obliquity is isotropically distributed.
 If stellar obliquity distribution is isotropic, $\cos{\psi}$ is uniformly distributed between -1 and 1.
-To flexibly model the stellar obliquity distribution, we select a two-component mixture of Betas distribution with hyperparameters $\bm{\beta} = \{\bm{w},\bm{a},\bm{b}\}$, where each hyperparameter has a dimension of 2.
-\edits{We have $\bm{w} = \{w_0, w_1\}$, $\bm{a} = \{a_0, a_1\}$, and  $\bm{b} = \{b_0, b_1\}$.}
+To flexibly model the stellar obliquity distribution, \edits{we select a multi-component mixture of Beta distributions with hyperparameters $\bm{\beta} = \{\bm{w},\bm{\mu},\bm{\kappa}\}$, where each hyperparameter has a dimension of the number of components.
+For example, for a two-component mixture model, each hyperparameter has a dimension of 2.
+We have $\bm{w} = \{w_0, w_1\}$, $\bm{\mu} = \{\mu_0, \mu_1\}$, and $\bm{\kappa} = \{\kappa_0, \kappa_1\}$.
+The hyperparameter $\bm{w}$ describes the weight of each component, and $\bm{\mu}$ and 1/$\bm{\kappa}$ correspond to the mean and variance of each Beta distribution component, respectively.
+The greater value of the $\kappa$, the smaller the variance (i.e., the distribution is more concentrated).
+The relations between the $\mu$ and $\kappa$ and the standard $\alpha$ and $\beta$ parameters in the Beta distribution are $\alpha = \mu \kappa$ and $\beta = (1-\mu) \kappa$.
 This mixture distribution has the capacity to capture anything from an isotropic distribution to a strongly bimodal population.
-% One component is designed for a large population of well-aligned systems, and the other is designed for misaligned systems.
-The Beta distribution is a flexible distribution that can describe the spike of aligned systems with stellar obliquities close to $0\degr$ or the clustered or broadly distributed misaligned systems.
-\dfm{The notation here is a bit sloppy. It might be better to say something like: there's some parameter $x \sim \mathrm{mixture}$ which we then transform as $2\,x-1$, which isn't quite the same.}
-Then, under this model, the probability density function for $\cos{\psi}$ is 
+Then, under this two-component model, the probability density function for $\cos{\psi}$ is
 \begin{align}
-    u_0 &\sim {\rm Beta}(a_0, b_0) \nonumber\\
-    u_1 &\sim {\rm Beta}(a_1, b_1) \nonumber\\
+    w_{0,1} &\sim {\rm Dirichlet}(1, 1) \nonumber\\
+    u_0 &\sim {\rm Beta}(\mu_0\kappa_0, (1-\mu_0)\kappa_0) \nonumber\\
+    u_1 &\sim {\rm Beta}(\mu_1\kappa_1, (1-\mu_1)\kappa_1) \nonumber\\
     \cos{\psi} &\sim 2\,( w_0 u_0 + w_1 u_1)-1.
 \end{align}
-Since the Beta distribution is defined on the interval $[0, 1]$, we extend its support to $[-1,1]$ using a linear transformation (i.e., $2\,( w_0 u_0 + w_1 u_1)-1$).
-\dfm{I'm not sure that I understand this next bit. Why and how do we do this?}
-For the ${\rm Beta}(a_0,b_0)$ component, we have it to describe the population of aligned systems.
-For the $a_0/b_0 \gg 1$, the distribution will have a spike at $\cos{\psi} = 1$.
-For the ${\rm Beta}(a_1,b_1)$ component, we have it to describe the misaligned systems.
-% add more description on why chose a0, b0 priors
-We design the order of the two components to prevent label switching in the mixture model \dfm{how?}.
+Since the Beta distribution is defined on the interval $[0, 1]$ whereas the support of $\cos{\psi}$ is from -1 to 1, we extend the mixture distribution's support $[0,1]$ to $[-1,1]$ using a linear transformation (i.e., $2\,( w_0 u_0 + w_1 u_1)-1$).
+Lastly, to avoid label switching in the mixture model, we apply a transformation to the distribution means $\vb*{\mu}$ to force the vector to be ordered.
+For $\mu$ and $\kappa$, we adopt the following priors:
+\begin{align}
+    \mu_{0,1} &\sim \mathcal{U}(0, 1) \nonumber\\
+    \log{\kappa}_{0,1} &\sim \mathcal{N}(0, 3),
+\end{align}
+where $\mu_{0,1}$ are uniformly distributed between 0 and 1 and $\log{\kappa}_{0,1}$ are normally distributed with a mean of 0 and standard deviation of 3.
+Notably, when applying the framework to a small sample size with $N \lesssim 50$, the choice of hyperpriors for the Beta distribution could impact the inferred distribution \citep{Nagpal22}.
+To ensure the robustness of the inferred distributions in such cases, it is crucial to test their sensitivity to different hyperpriors.
+Here we present a two-component mixture model for demonstration. In practice, the distribution can be modified to a single component or higher components.
+}
 
-\edits{In this framework, we limit $i_{\rm orb}$ from $0\degr$ to $90\degr$. To preserve the symmetry, we allow $i_{\rm orb}$ to vary from $0\degr$ to $180\degr$.}
-With the design, the hyperparameter and parameter priors are the following:
+Next, the parameter priors are the following:
 \begin{align}
-    w_{0,1} &\sim {\rm Dirichlet}(1, 1) \nonumber\\
-    p_{0,1} &\sim \mathcal{U}(0, 1) \nonumber\\
-    q_{0,1} &\sim \mathcal{U}(0, 1) \nonumber\\
     \theta_n &\sim \mathcal{U}(0, \pi)  \nonumber\\
@@
 If $\gamma_{\star, n}$ is available, we construct Normal distributions with means and standard deviations from measurements.
+\edits{For the orbital inclination $i_{\rm orb}$, following the convention, we limit it to $\left[0\degr, 90\degr\right]$.
+To not underestimate the $\psi$, we then set $i_\star$ to be varied from $0\degr$ to $180\degr$.
+Besides, since $\lambda$ and $-\lambda$ correspond to the same $\psi$ solution, we limit $\lambda$ to $\left[0\degr, 180\degr\right]$ and thus $\theta$ to $\left[0\degr, 180\degr\right]$.
+We find doing so, while not compromising the inference of the $\psi$ distribution, greatly improves the sampling performance as it avoids chains stuck at $\theta = 0$ or $2\pi$ due to discontinuity.}
 
@@
 \end{align}
-Here we use uniform hyperpriors for the Beta distribution.
-However, it is important to note that when applying the framework to a small sample size with $N \lesssim 50$, the choice of hyperpriors could impact the inferred distribution \citep{Nagpal22}.
-To ensure the robustness of the inferred distributions in such cases, it is crucial to test their sensitivity to different hyperpriors.
 
-The probabilistic model is constructed using the $\mathtt{PyMC}$ package version $\mathtt{v4.1.7}$ \citep{pymc}, and the posteriors are sampled with the No-U-Turn Sampler \citep[NUTS;][]{Hoffman11}, which is a gradient-based Markov chain Monte Carlo (MCMC) sampling algorithm.
+The probabilistic model is constructed using the $\mathtt{PyMC}$ package version $\mathtt{v5.1.2}$ \citep{pymc}, and the posteriors are sampled with the No-U-Turn Sampler \citep[NUTS;][]{Hoffman11}, which is a gradient-based Markov chain Monte Carlo (MCMC) sampling algorithm.
 This paper's figures and simulations are completely reproducible and were created using the \showyourwork package.
-The open-source code is available on \href{https://github.com/jiayindong/obliquity}{GitHub\,\faGithub\,(https://github.com/jiayindong/obliquity)}.
+The open-source code is available on GitHub \href{https://github.com/jiayindong/obliquity}{(https://github.com/jiayindong/obliquity\,\faGithub)}.
 
-\section{Applications to Simulated and Observed Data}\label{sec:applications}
-
-\subsection{Simulated Data}\label{subsec:sims}
+\section{Model Performance}\label{sec:simulation}
 
 To investigate the performance of our hierarchical Bayesian framework, we apply it to simulated data, generated with a known ground-truth stellar obliquity distribution.
-We test the following four $\cos{\psi}$ distributions: a uniform distribution $\mathcal{U}(-1,1)$, and three Normal distributions $\mathcal{N}(0,0.2)$, $\mathcal{N}(-0.4,0.2)$, and $\mathcal{N}(0.4,0.2)$.
-\dfm{For these normals, we must truncate, since they have support outside of 0--1. How is that implemented; should we mention it explicitly?}
-For each $\cos{\psi}$ distribution, we randomly generate 200 samples of stellar inclination $i_\star$ and sky-projected stellar obliquity $\lambda$.
+\edits{We test the following four $\cos{\psi}$ distributions: a uniform distribution such that $\cos{\psi} \sim \mathcal{U}(-1,1)$, and three truncated Normal distributions bounded between -1 and 1, such that $\cos{\psi} \sim \mathcal{N}(0,0.2)$, $\cos{\psi} \sim \mathcal{N}(-0.4,0.2)$, and $\cos{\psi} \sim \mathcal{N}(0.4,0.2)$.}
+For each $\cos{\psi}$ distribution, we randomly generate 200 samples of sky-projected stellar obliquity $\lambda$ and stellar inclination $i_\star$.
 We assume the stellar spin axis is uniformly distributed around the planetary orbital axis in the azimuthal direction and the orbital inclination is $90\degr$.
@@
 To simulate the observation process, we add Gaussian noise to the \emph{true} $\lambda$ and $i_\star$, using uncertainties of $\sigma_{\lambda} = 8\degr$ and $\sigma_{i_\star} = 10\degr$, which are typical of the literature sample \cite{Albrecht22}.
-Using these simulated $\lambda$ measurements and their uncertainties, we infer the $\cos{\psi}$ distribution of the sample with or without $i_\star$ likelihoods.
+Using these simulated $\lambda$ measurements and their uncertainties, we infer the $\cos{\psi}$ distribution of the sample with \edits{both $\lambda$ and $i_\star$ likelihoods, $\lambda$ likelihood only, and $i_\star$ likelihood only.}
 
@@
 We demonstrate through simulations that the inferred stellar obliquity distribution is robust even if the $i_\star$ information is not provided, and the $i_\star$ measurement only improves the constraint on the stellar obliquity distribution.
-Later in Section~\ref{sec:jacobian}, we derive the Jacobian transformations between $\psi$, $\lambda$, and $i_\star$ to understand why sky-projected stellar obliquities play a predominant role in determining the inferred distribution of stellar obliquities.
 
-\begin{figure}[ht!]
+\begin{figure*}[ht!]
+    \centering
     \script{sim_plot.py}
-    \includegraphics[width=\linewidth]{figures/simulation.pdf}
+    \includegraphics[width=0.8\textwidth]{figures/simulation.pdf}
     \caption{Inferred stellar obliquity distributions from sky-projected stellar obliquities with and without information on the stellar inclination, depicted in blue and orange curves, respectively.
@@
     \label{fig:simulation}
-\end{figure}
+\end{figure*}
 
-\edits{EDIT TO HERE TODAY}
+\subsection{Jacobian transformations between the $\psi$, $\lambda$, and \lowercase{$i_\star$} distributions}\label{sec:jacobian}
 
-\subsection{The Stellar Obliquity Distribution of Observed Exoplanets}
-
-We next apply our hierarchical Bayesian framework to a sample of 161 exoplanetary systems with sky-projected stellar obliquity measurements, primarily consisting of Hot Jupiter systems, as summarized in \cite{Albrecht22} Table A1.
-The inferred $\cos{\psi}$ distribution is shown in Figure~\ref{fig:psi_dist}.
-The $\cos{\psi}$ distribution is peaked at 1, with nearly flat behavior between $-0.75$ and $0.75$ and no significant clustering.
-The distribution suggests that there is a pile-up of planetary systems with stellar obliquities less than $40^\circ$ and an isotropic distribution for obliquities between $40^\circ$ and $140^\circ$.
-The fraction of aligned systems $w_0$ dominates the distribution, with $w_0 = 0.67 \pm 0.09$.
-The corresponding posteriors for the parameters of the population Beta distribution are $a_0 = 31.0\pm12.8$ and $b_0 = 0.40\pm0.12$.
-On the other hand, the fraction of misaligned systems is estimated to be $w_1 = 0.33 \pm 0.09$, with posteriors of $a_1 = 1.56\pm0.93$ and $b_1 = 1.64\pm1.58$.
-
-\begin{figure}[ht!]
-    \script{psi_plot.py}
-    \begin{centering}
-        \includegraphics{figures/psi_dist.pdf}
-        \caption{Inferred stellar obliquity distribution for all exoplanetary systems with sky-projected stellar obliquity measurements.
-This inference is based purely on the observed sky-projected obliquities and assumes an isotropic distribution for the stellar inclinations.
-        \dfm{Might be worth plotting the two mixture components too.}}
-        \label{fig:psi_dist}
-    \end{centering}
-\end{figure}
-
-The discrepancy between this inference for the full sample of exoplanet systems and the previous analysis of the subsample with $i_\star$ measurements \citep{Albrecht21} warrants further investigation.
-This earlier study identified a preponderance (\dfm{maybe different word.}) of perpendicular planets and disfavored an isotropic stellar obliquity distribution \citep{Albrecht21}.
-There are at least two potential explanations for this difference: (1) the subsample with $i_\star$ measurements is small and only includes about 20 misaligned systems, and (2) the requirement for $i_\star$ measurements could introduce selection biases in the sample.
-\dfm{Expand on this a little bit, I think.
-Maybe add a short paragraph for each.}
-
-The inferred stellar obliquity distribution indicates that approximately $67\pm9$\% of the systems have a stellar obliquity of less than $40\degr$, and approximately $33\pm9$\% of the systems follow a nearly isotropic stellar obliquity distribution ranging from $\sim$$40\degr$ to $\sim$$140\degr$.
-These findings could have significant implications for the formation and evolution of close-in planetary systems.
-The broad distribution of misaligned systems is in good agreement with the predicted outcome of multiple giant planets scattering after a convergent disk migration, as proposed by various studies, such as \cite{Nagasawa11} and \cite{Beague12}.
-The intriguing result should be further examined with a more carefully selected sample of Hot Jupiters and provides opportunities to place constraints on their origin channels of Hot Jupiters.
-
-\section{Relations between the $\psi$, $\lambda$, and \lowercase{$i_\star$} distributions}\label{sec:jacobian}
-
-In this section, we aim to gain insight into the reasons behind the predominant role of the sky-projected stellar obliquity distribution and the less significant impact of the stellar inclination distribution in the inference of the stellar obliquity distribution.
+In this subsection, we aim to gain insight into the reasons behind the predominant role of the sky-projected stellar obliquity distribution and the less significant impact of the stellar inclination distribution in the inference of the stellar obliquity distribution.
 
@@
 
+\begin{figure*}[ht!]
+    \script{transform.py}
+    \includegraphics[width=\linewidth]{figures/transform.pdf}
+    \caption{Simulated $\cos{\psi}$ distributions (\nth{1} column) and the corresponding distributions of sky-projected stellar obliquity $\lambda$ (\nth{2} column) and stellar inclination $i_\star$ (\nth{3} column).
+The inferred $\cos{\psi}$ distributions assuming isotropic stellar inclinations are shown in the \nth{4} column.
+The grey histograms present the random samplings of $\lambda$ and $i_\star$ from the $\cos{\psi}$ distributions, and the blue curves present the numerical solutions.}
+    \label{fig:transform}
+\end{figure*}
+
 Using Equation~(\ref{eqn:jac_lam}) and (\ref{eqn:jac_istar}), we can now derive the $\lambda$ and $i_\star$ distributions for any given $\psi$ distributions, assuming the azimuthal angle of the stellar spin axis $\theta$ is random.
@@
 
-\begin{figure*}[ht!]
-    \script{transform.py}
-    \includegraphics[width=\linewidth]{figures/transform.pdf}
-    \caption{Simulated $\cos{\psi}$ distributions (\nth{1} column) and the corresponding distributions of sky-projected stellar obliquity $\lambda$ (\nth{2} column) and stellar inclination $i_\star$ (\nth{3} column).
-The inferred $\cos{\psi}$ distributions assuming isotropic stellar inclinations are shown in the \nth{4} column.
-The grey histograms present the random samplings of $\lambda$ and $i_\star$ from the $\cos{\psi}$ distributions, and the blue curves present the numerical solutions.}
-    \label{fig:transform}
-\end{figure*}
+\section{Application to Observations}\label{sec:observation}
+
+We next apply our hierarchical Bayesian framework to a sample of 161 exoplanetary systems with sky-projected stellar obliquity measurements, primarily consisting of Hot Jupiter systems, as summarized in \cite{Albrecht22} Table A1.
+The inferred $\cos{\psi}$ distribution is shown in Figure~\ref{fig:psi_dist}.
+The $\cos{\psi}$ distribution is peaked at 1, with nearly flat behavior between $-0.75$ and $0.75$ and no significant clustering.
+The distribution suggests that there is a pile-up of planetary systems with stellar obliquities less than $40^\circ$ and an isotropic distribution for obliquities between $40^\circ$ and $140^\circ$.
+The fraction of aligned systems $w_0$ dominates the distribution, with $w_0 = 0.67 \pm 0.09$.
+The corresponding posteriors for the parameters of the population Beta distribution are $a_0 = 31.0\pm12.8$ and $b_0 = 0.40\pm0.12$.
+On the other hand, the fraction of misaligned systems is estimated to be $w_1 = 0.33 \pm 0.09$, with posteriors of $a_1 = 1.56\pm0.93$ and $b_1 = 1.64\pm1.58$.
+
+\begin{figure}[ht!]
+    \script{psi_plot.py}
+    \begin{centering}
+        \includegraphics{figures/psi_dist.pdf}
+        \caption{Inferred stellar obliquity distribution for all exoplanetary systems with sky-projected stellar obliquity measurements.
+This inference is based purely on the observed sky-projected obliquities.}
+        \label{fig:psi_dist}
+    \end{centering}
+\end{figure}
+
+The discrepancy between this inference for the full sample of exoplanet systems and the previous analysis of the subsample with $i_\star$ measurements \citep{Albrecht21} warrants further investigation.
+This earlier study identified a concentration of perpendicular planets and disfavored an isotropic stellar obliquity distribution \citep{Albrecht21}.
+There are at least two potential explanations for this difference: (1) the subsample with $i_\star$ measurements is small and only includes about 20 misaligned systems, and (2) the requirement for $i_\star$ measurements could introduce selection biases in the sample.
+\dfm{Expand on this a little bit, I think.
+Maybe add a short paragraph for each.}
+
+The inferred stellar obliquity distribution indicates that approximately $67\pm9$\% of the systems have a stellar obliquity of less than $40\degr$, and approximately $33\pm9$\% of the systems follow a nearly isotropic stellar obliquity distribution ranging from $\sim$$40\degr$ to $\sim$$140\degr$.
+These findings could have significant implications for the formation and evolution of close-in planetary systems.
+The broad distribution of misaligned systems is in good agreement with the predicted outcome of multiple giant planets scattering after a convergent disk migration, as proposed by various studies, such as \cite{Nagasawa11} and \cite{Beague12}.
+The intriguing result should be further examined with a more carefully selected sample of Hot Jupiters and provides opportunities to place constraints on their origin channels of Hot Jupiters.
 
@@
 Stellar inclination measurements are optional input in the model, and if not available, they are assumed to be isotropically distributed.
-Our open-source hierarchical Bayesian model, available on \href{https://github.com/jiayindong/obliquity}{GitHub\,\faGithub\,(https://github.com/jiayindong/obliquity)}, can be customized to different stellar obliquity distributions and priors for specific target samples.
+Our open-source hierarchical Bayesian model, available on GitHub \href{https://github.com/jiayindong/obliquity}{(https://github.com/jiayindong/obliquity\,\faGithub)}, can be customized to different stellar obliquity distributions and priors for specific target samples.
 

### 2023-04-10 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 We have $\bm{w} = \{w_0, w_1\}$, $\bm{\mu} = \{\mu_0, \mu_1\}$, and $\bm{\kappa} = \{\kappa_0, \kappa_1\}$.
-The hyperparameter $\bm{w}$ describes the weight of each component, and $\bm{\mu}$ and 1/$\bm{\kappa}$ correspond to the mean and variance of each Beta distribution component, respectively.
+The hyperparameter $\bm{w}$ describes the weight of each component.
+The $\bm{\mu}$ and 1/$\bm{\kappa}$ correspond to the mean and variance of each Beta distribution component, respectively, a reparametrization of parameters suggested by \cite{Gelman14}.
 The greater value of the $\kappa$, the smaller the variance (i.e., the distribution is more concentrated).
@@
 Since the Beta distribution is defined on the interval $[0, 1]$ whereas the support of $\cos{\psi}$ is from -1 to 1, we extend the mixture distribution's support $[0,1]$ to $[-1,1]$ using a linear transformation (i.e., $2\,( w_0 u_0 + w_1 u_1)-1$).
-Lastly, to avoid label switching in the mixture model, we apply a transformation to the distribution means $\vb*{\mu}$ to force the vector to be ordered.
 For $\mu$ and $\kappa$, we adopt the following priors:
@@
 where $\mu_{0,1}$ are uniformly distributed between 0 and 1 and $\log{\kappa}_{0,1}$ are normally distributed with a mean of 0 and standard deviation of 3.
+To deal with label switching in the mixture model, we remove the symmetry by forcing the vector $\vb*{\mu}$ to be ordered.
 Notably, when applying the framework to a small sample size with $N \lesssim 50$, the choice of hyperpriors for the Beta distribution could impact the inferred distribution \citep{Nagpal22}.
@@
 Besides, since $\lambda$ and $-\lambda$ correspond to the same $\psi$ solution, we limit $\lambda$ to $\left[0\degr, 180\degr\right]$ and thus $\theta$ to $\left[0\degr, 180\degr\right]$.
-We find doing so, while not compromising the inference of the $\psi$ distribution, greatly improves the sampling performance as it avoids chains stuck at $\theta = 0$ or $2\pi$ due to discontinuity.}
+We find doing so greatly improves the sampling performance while not compromising the inference of the $\psi$ distribution.
+The possible explanation for the improvement is that the $\left[0\degr, 180\degr\right]$ support prevents chains from getting stuck at $\theta = 0\degr$ and $\theta = 360\degr$ discontinuity.}
 

### 2023-04-10 src/scripts/transform.py
figures

@@ -148,40 +148,34 @@ if __name__ == '__main__':
 
     ### Make the plot ###
 
-    plt.figure(figsize=(7,5),dpi=110)
+    plt.figure(figsize=(5,5),dpi=110)
+
 
     ### cosψ ~ Uniform(-1,1) ###
 
     cosψ = uni.posterior.cosψ.values.ravel()
     λ = uni.posterior.λ.values.ravel()
     i = uni.posterior.i.values.ravel()
-    iso_cosψ = uni.posterior.iso_cosψ.values.ravel()
 
-    plt.subplot(4,4,1)
+    plt.subplot(4,3,1)
     plt.hist(cosψ, color='#e5e1e0', density=True, bins=40)
     plt.title(r'$\cos{\psi} \sim \mathcal{U}(-1,1)$')
     plt.ylim([0,0.6])
     plt.xlim([-1,1])
 
-    plt.subplot(4,4,2)
+    plt.subplot(4,3,2)
     plt.hist(λ, color='#e5e1e0', density=True, bins=40)
     plt.title(r'$\lambda$')
     plt.ylim([0,0.5])
     plt.xlim([0,np.pi])
-    plt.xticks([0,np.pi/4,np.pi/2,3*np.pi/4,np.pi], ['0', '', r'$\pi/2$', '', r'$\pi$'])
+    plt.xticks([0,np.pi/2,np.pi], ['0', r'$\pi/2$', r'$\pi$'])
 
-    plt.subplot(4,4,3)
+    plt.subplot(4,3,3)
     plt.hist(i, color='#e5e1e0', density=True, bins=40)
     plt.title(r'$i_\star$')
     plt.xlim([0,np.pi])
-    plt.xticks([0,np.pi/4,np.pi/2,3*np.pi/4,np.pi], ['0', '', r'$\pi/2$', '', r'$\pi$'])
+    plt.xticks([0,np.pi/4,np.pi/2,3*np.pi/4,np.pi], ['0', r'$\pi/4$', r'$\pi/2$', r'$3\pi/4$', r'$\pi$'])
     plt.ylim([0,0.55])
-
-    plt.subplot(4,4,4)
-    plt.hist(iso_cosψ, color='#e5e1e0', density=True, bins=40)
-    plt.title(r'Inferred $\cos{\psi}$')
-    plt.ylim([0,0.6])
-    plt.xlim([-1,1])
 
 
     cosψ = np.linspace(-1,1,200)
[…]

### 2023-04-13 src/scripts/sim_plot.py
figures

@@ -55,22 +55,22 @@ def psi_dist_draws(model_name):
         for b in range(1000):
             noistar_draws[:, a*1000+b] = beta.pdf(x, post.a[a,b], post.b[a,b])
 
-    post = nolam_idata.posterior
-    nolam_draws = np.zeros(shape=(len(x),16000))
-    for a in range(16):
-        for b in range(1000):
-            nolam_draws[:, a*1000+b] = beta.pdf(x, post.a[a,b], post.b[a,b])
+    # post = nolam_idata.posterior
+    # nolam_draws = np.zeros(shape=(len(x),16000))
+    # for a in range(16):
+    #     for b in range(1000):
+    #         nolam_draws[:, a*1000+b] = beta.pdf(x, post.a[a,b], post.b[a,b])
 
-    return draws, noistar_draws, nolam_draws
+    return draws, noistar_draws
 
-uni_draws, uni_noistar_draws, uni_nolam_draws = psi_dist_draws("uni")
-norm1_draws, norm1_noistar_draws, norm1_nolam_draws = psi_dist_draws("norm1")
-norm2_draws, norm2_noistar_draws, norm2_nolam_draws = psi_dist_draws("norm2")
-norm3_draws, norm3_noistar_draws, norm3_nolam_draws = psi_dist_draws("norm3")
+uni_draws, uni_noistar_draws = psi_dist_draws("uni")
+norm1_draws, norm1_noistar_draws = psi_dist_draws("norm1")
+norm2_draws, norm2_noistar_draws = psi_dist_draws("norm2")
+norm3_draws, norm3_noistar_draws = psi_dist_draws("norm3")
 
 
 ### Make the plot ###
-fig, big_axes = plt.subplots(figsize=(6.5,7),dpi=150,nrows=4,ncols=1,sharey=True) 
+fig, big_axes = plt.subplots(figsize=(3.5,7),dpi=150,nrows=4,ncols=1,sharey=True) 
 
[…]

### 2023-04-13 src/scripts/transform.py

@@ -168,13 +168,13 @@ if __name__ == '__main__':
     plt.title(r'$\lambda$')
     plt.ylim([0,0.5])
     plt.xlim([0,np.pi])
-    plt.xticks([0,np.pi/2,np.pi], ['0', r'$\pi/2$', r'$\pi$'])
+    plt.xticks([0,np.pi/4,np.pi/2,3*np.pi/4,np.pi], ['0', '', r'$\pi/2$', '', r'$\pi$'])
 
     plt.subplot(4,3,3)
     plt.hist(i, color='#e5e1e0', density=True, bins=40)
     plt.title(r'$i_\star$')
     plt.xlim([0,np.pi])
-    plt.xticks([0,np.pi/4,np.pi/2,3*np.pi/4,np.pi], ['0', r'$\pi/4$', r'$\pi/2$', r'$3\pi/4$', r'$\pi$'])
+    plt.xticks([0,np.pi/4,np.pi/2,3*np.pi/4,np.pi], ['0', '', r'$\pi/2$', '', r'$\pi$'])
     plt.ylim([0,0.55])
 
 
@@ -228,13 +228,13 @@ if __name__ == '__main__':
     plt.title(r'$\lambda$')
     plt.ylim([0,1.5])
     plt.xlim([0,np.pi])
-    plt.xticks([0,np.pi/2,np.pi], ['0', r'$\pi/2$', r'$\pi$'])
+    plt.xticks([0,np.pi/4,np.pi/2,3*np.pi/4,np.pi], ['0', '', r'$\pi/2$', '', r'$\pi$'])
 
     plt.subplot(4,3,6)
     plt.hist(i, color='#e5e1e0', density=True, bins=40)
     plt.title(r'$i_\star$')
     plt.xlim([0,np.pi])
-    plt.xticks([0,np.pi/4,np.pi/2,3*np.pi/4,np.pi], ['0', r'$\pi/4$', r'$\pi/2$', r'$3\pi/4$', r'$\pi$'])
+    plt.xticks([0,np.pi/4,np.pi/2,3*np.pi/4,np.pi], ['0', '', r'$\pi/2$', '', r'$\pi$'])
     plt.ylim([0,0.55])
 
 
@@ -289,7 +289,7 @@ if __name__ == '__main__':
     plt.title(r'$\lambda$')
     plt.ylim([0,1.5])
     plt.xlim([0,np.pi])
[…]

### 2023-04-13 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 Misaligned systems have nearly isotropic stellar obliquities \edits{with no strong clustering near 90$\degr$.
-Furthermore, misaligned Hot Jupiters have a diverse range of stellar obliquities, suggesting that dynamic mechanisms, such as planet-planet scattering after a convergent disk migration or planetary Kozai, are likely responsible for producing the population.}
+Furthermore, misaligned Hot Jupiters have a diverse range of stellar obliquities from $0\degr$ to $180\degr$, preferring dynamic mechanisms, such as planet-planet scattering after a convergent disk migration or planetary Kozai, which could produce broad distributions.}
 
@@
 \begin{figure*}[ht!]
+    \centering
     \script{transform.py}
-    \includegraphics[width=\linewidth]{figures/transform.pdf}
+    \includegraphics[width=0.8\linewidth]{figures/transform.pdf}
     \caption{Simulated $\cos{\psi}$ distributions (\nth{1} column) and the corresponding distributions of sky-projected stellar obliquity $\lambda$ (\nth{2} column) and stellar inclination $i_\star$ (\nth{3} column).

### 2023-04-13 src/scripts/sim_plot.py
figures

@@ -145,7 +145,7 @@ pcosψ = 1/np.sqrt(2*np.pi*σ**2)*np.exp(-(cosψ)**2/2/σ**2)
 plt.plot(cosψ, pcosψ, c='slategrey', ls='--', lw=1, zorder=0)
 
 # cosψ ~ N(0,0.2), noistar
-ax = fig.add_subplot(4,3,4)
+ax = fig.add_subplot(4,2,4)
 
 q025, q16, q50, q84, q975 = np.percentile(norm1_noistar_draws, [2.5, 16, 50, 84, 97.5], axis=1)/2
 plt.plot(2*x-1, q50, color='#f56e4a', lw=1.2)

### 2023-04-13 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 To investigate the performance of our hierarchical Bayesian framework, we apply it to simulated data, generated with a known ground-truth stellar obliquity distribution.
-\edits{We test the following four $\cos{\psi}$ distributions: a uniform distribution such that $\cos{\psi} \sim \mathcal{U}(-1,1)$, and three truncated Normal distributions bounded between -1 and 1, such that $\cos{\psi} \sim \mathcal{N}(0,0.2)$, $\cos{\psi} \sim \mathcal{N}(-0.4,0.2)$, and $\cos{\psi} \sim \mathcal{N}(0.4,0.2)$.}
+\edits{We test the following four $\cos{\psi}$ distributions: a uniform distribution bounded between $-1$ and $1$ such that $\cos{\psi} \sim \mathcal{U}(-1,1)$, and three truncated Normal distributions bounded between $-1$ and $1$, such that $\cos{\psi} \sim \textit{truncnorm}(0, 0.2)$, $\cos{\psi} \sim \textit{truncnorm}(-0.4, 0.2)$, and $\cos{\psi} \sim \textit{truncnorm}(0.4, 0.2)$, where the first and second number are the mean and standard deviation of the normal distribution, respectively.}
 For each $\cos{\psi}$ distribution, we randomly generate 200 samples of sky-projected stellar obliquity $\lambda$ and stellar inclination $i_\star$.
@@
 To simulate the observation process, we add Gaussian noise to the \emph{true} $\lambda$ and $i_\star$, using uncertainties of $\sigma_{\lambda} = 8\degr$ and $\sigma_{i_\star} = 10\degr$, which are typical of the literature sample \cite{Albrecht22}.
-Using these simulated $\lambda$ measurements and their uncertainties, we infer the $\cos{\psi}$ distribution of the sample with \edits{both $\lambda$ and $i_\star$ likelihoods, $\lambda$ likelihood only, and $i_\star$ likelihood only.}
+Using these simulated $\lambda$ measurements and their uncertainties, we infer the $\cos{\psi}$ distribution of the sample with both $\lambda$ and $i_\star$ likelihoods or $\lambda$ likelihood only.
 
@@
 
-\begin{figure*}[ht!]
+\begin{figure}[ht!]
     \centering
     \script{sim_plot.py}
-    \includegraphics[width=0.8\textwidth]{figures/simulation.pdf}
+    \includegraphics{figures/simulation.pdf}
     \caption{Inferred stellar obliquity distributions from sky-projected stellar obliquities with and without information on the stellar inclination, depicted in blue and orange curves, respectively.
@@
     \label{fig:simulation}
-\end{figure*}
+\end{figure}
 
@@
     \script{transform.py}
-    \includegraphics[width=0.8\linewidth]{figures/transform.pdf}
+    \includegraphics[width=0.7\linewidth]{figures/transform.pdf}
     \caption{Simulated $\cos{\psi}$ distributions (\nth{1} column) and the corresponding distributions of sky-projected stellar obliquity $\lambda$ (\nth{2} column) and stellar inclination $i_\star$ (\nth{3} column).

### 2023-04-17 src/scripts/sim_plot.py
update plots

@@ -70,7 +70,7 @@ norm3_draws, norm3_noistar_draws = psi_dist_draws("norm3")
 
 
 ### Make the plot ###
-fig, big_axes = plt.subplots(figsize=(3.5,7),dpi=150,nrows=4,ncols=1,sharey=True) 
+fig, big_axes = plt.subplots(figsize=(3.5,6.5),dpi=150,nrows=4,ncols=1,sharey=True) 
 
 for row, big_ax in enumerate(big_axes, start=1):
     

### 2023-04-17 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 Misaligned systems have nearly isotropic stellar obliquities \edits{with no strong clustering near 90$\degr$.
-Furthermore, misaligned Hot Jupiters have a diverse range of stellar obliquities from $0\degr$ to $180\degr$, preferring dynamic mechanisms, such as planet-planet scattering after a convergent disk migration or planetary Kozai, which could produce broad distributions.}
+Furthermore, misaligned Hot Jupiters present a diverse range of stellar obliquities, preferring dynamic mechanisms, such as planet-planet scattering after a convergent disk migration, which could produce both prograde and retrograde orbits with no particular inclination concentrations.}
 

### 2023-04-17 src/tex/ms.tex
[showyourwork] overleaf sync

--- 2023-04-13 src/tex/ms.tex
@@
 We introduce a flexible, hierarchical Bayesian framework that can be used to infer the stellar obliquity distribution solely from sky-projected stellar obliquities, including stellar inclination measurements when available.
-When applying the framework to all exoplanetary systems with measured sky-projected stellar obliquity, we find that the inferred population-level obliquity distribution is unimodal and peaked at zero degrees.
-Misaligned systems have nearly isotropic stellar obliquities \edits{with no strong clustering near 90$\degr$.
-Furthermore, misaligned Hot Jupiters have a diverse range of stellar obliquities from $0\degr$ to $180\degr$, preferring dynamic mechanisms, such as planet-planet scattering after a convergent disk migration or planetary Kozai, which could produce broad distributions.}
+When applying the framework to all exoplanetary systems with measured sky-projected stellar obliquity, we find that the inferred population-level obliquity distribution is unimodal and peaked at zero degrees \edits{with a fraction of $72\pm9$\,\%.
+The small fraction of misaligned systems ($28\pm9$\,\%) has nearly isotropic stellar obliquities with no strong clustering near 90$\degr$.
+The diverse range of stellar obliquities prefers dynamic mechanisms, such as planet-planet scattering after a convergent disk migration, which could produce both prograde and retrograde orbits with no strong inclination concentrations.}
 
@@
 It is as yet unclear if, and to what extent, all of these physical and dynamic processes apply to exoplanetary systems.
-These proposed mechanisms all make different predictions on stellar obliquity distributions \citep[see][and references therein]{Albrecht22}.
+These proposed mechanisms all make different predictions on stellar obliquity distributions \citep[see][and references therein]{Albrecht22, Dawson18}.
 For example, the secular chaos mechanism tends to produce a stellar obliquity distribution with $\psi < 90\degr$ \citep[e.g.,][]{Teyssandier19}.
-The stellar Kozai mechanism predicts a bimodal stellar obliquity distribution, concentrated at $40\degr$ and $140\degr$ \citep[e.g.,][]{Fabrycky07, Anderson16, Vick19}.
-The multiple-planet scattering mechanism results in a majority of aligned systems, with a small fraction of systems at a diverse range of stellar obliquities \citep[e.g.,][]{Beague12}.
+\edits{The stellar von Zeipel-Lidov-Kozai (ZLK) mechanism predicts a bimodal stellar obliquity distribution, concentrated at $40\degr$ and $140\degr$ \citep[e.g.,][]{Fabrycky07, Anderson16, Vick19}, assuming a zero stellar obliquity when the ZLK oscillation begins, or a broad stellar obliquity peaked near $90\degr$ if we drop the assumption \citep{Vick23}.} The multiple-planet scattering mechanism results in a majority of aligned systems, with a small fraction of systems at a diverse range of stellar obliquities \citep[e.g.,][]{Beague12}.
 With these predictions in mind, we aim to determine the dominant mechanisms responsible for shaping close-in planetary systems by characterizing the stellar obliquity distribution of exoplanetary systems through a Bayesian approach.
@@
 In Figure~\ref{fig:coord}, we introduce two coordinate systems that describe the stellar spin axis $\vb{\hat{n}_\star}$ and its planet's orbital axis $\vb{\hat{n}_{\rm orb}}$.
-The setup is similar to the coordinate system setup in \cite{Fabrycky09} but has a different definition of the azimuthal angle of the stellar spin and coordinate orientations.
+The setup is similar to the coordinate system setup in \cite{Fabrycky09} but has a different definition of the azimuthal angle of the stellar spin and coordinate orientation.
 
@@
     w_{0,1} &\sim {\rm Dirichlet}(1, 1) \nonumber\\
-    u_0 &\sim {\rm Beta}(\mu_0\kappa_0, (1-\mu_0)\kappa_0) \nonumber\\
-    u_1 &\sim {\rm Beta}(\mu_1\kappa_1, (1-\mu_1)\kappa_1) \nonumber\\
-    \cos{\psi} &\sim 2\,( w_0 u_0 + w_1 u_1)-1.
+    u_{0,1} &\sim {\rm Beta}(\mu_{0,1}\kappa_{0,1}, (1-\mu_{0,1})\kappa_{0,1}) \nonumber\\
+    \cos{\psi} &\sim 2\,(w_0 u_0 + w_1 u_1) - 1.
 \end{align}
-Since the Beta distribution is defined on the interval $[0, 1]$ whereas the support of $\cos{\psi}$ is from -1 to 1, we extend the mixture distribution's support $[0,1]$ to $[-1,1]$ using a linear transformation (i.e., $2\,( w_0 u_0 + w_1 u_1)-1$).
-For $\mu$ and $\kappa$, we adopt the following priors:
+Since the Beta distribution is defined on the interval $[0, 1]$ whereas the support of $\cos{\psi}$ is from $-1$ to $1$, we extend the mixture distribution's support $[0,1]$ to $[-1,1]$ using a linear transformation (i.e., $2\,( w_0 u_0 + w_1 u_1)-1$).
+For hyperparmeters $\mu$ and $\kappa$, we adopt the following priors:
 \begin{align}
@@
 \end{align}
-where $\mu_{0,1}$ are uniformly distributed between 0 and 1 and $\log{\kappa}_{0,1}$ are normally distributed with a mean of 0 and standard deviation of 3.
+where $\mu_{0,1}$ is uniformly distributed between $0$ and $1$ and $\log{\kappa}_{0,1}$ is normally distributed with a mean of $0$ and standard deviation of $3$.
 To deal with label switching in the mixture model, we remove the symmetry by forcing the vector $\vb*{\mu}$ to be ordered.
-Notably, when applying the framework to a small sample size with $N \lesssim 50$, the choice of hyperpriors for the Beta distribution could impact the inferred distribution \citep{Nagpal22}.
+Notably, when applying the framework to a small sample size with $N \lesssim 50$, the choice of hyperpriors for the Beta distribution could impact the inferred distribution \citep[e.g.,][]{Nagpal22, Gelman14}.
 To ensure the robustness of the inferred distributions in such cases, it is crucial to test their sensitivity to different hyperpriors.
-Here we present a two-component mixture model for demonstration. In practice, the distribution can be modified to a single component or higher components.
 }
 
-Next, the parameter priors are the following:
+Next, the model parameter priors are the following:
 \begin{align}
@@
 If $\gamma_{\star, n}$ is available, we construct Normal distributions with means and standard deviations from measurements.
-\edits{For the orbital inclination $i_{\rm orb}$, following the convention, we limit it to $\left[0\degr, 90\degr\right]$.
+\edits{For the orbital inclination $i_{\rm orb}$, following the convention, we limit it to $\left[0, \pi/2\right]$.
 To not underestimate the $\psi$, we then set $i_\star$ to be varied from $0\degr$ to $180\degr$.
-Besides, since $\lambda$ and $-\lambda$ correspond to the same $\psi$ solution, we limit $\lambda$ to $\left[0\degr, 180\degr\right]$ and thus $\theta$ to $\left[0\degr, 180\degr\right]$.
-We find doing so greatly improves the sampling performance while not compromising the inference of the $\psi$ distribution.
-The possible explanation for the improvement is that the $\left[0\degr, 180\degr\right]$ support prevents chains from getting stuck at $\theta = 0\degr$ and $\theta = 360\degr$ discontinuity.}
+Besides, since $\lambda$ and $-\lambda$ correspond to the same $\psi$ solution, we limit $\lambda$ to $\left[0, \pi\right]$ and thus $\theta$ to $\left[0, \pi\right]$.
+We find doing so greatly improves the sampling performance while not compromising the inference of the $\psi$ distribution due to the symmetry.
+The $\left[0, \pi\right]$ support avoids the otherwise discontinuity at $\theta = 0$ and $\theta = \pi$.}
 
@@
 \begin{align}
-    \mathcal{L}(\lambda) &\sim \prod_{i=1}^N\mathcal{N}(\hat{\lambda}_n, \sigma_{\hat{\lambda}_n}) \nonumber\\
-    \mathcal{L}(i_{\rm orb}) &\sim \prod_{i=1}^N\mathcal{N}(\hat{i}_{{\rm orb},n}, \sigma_{\hat{i}_{{\rm orb},n}}) \nonumber\\
-    \mathcal{L}(i_\star) &\sim \prod_{i=1}^N\mathcal{N}({\rm Obs}_{\star,n}, \sigma_{{\rm Obs}_{\star,n}}).
+    \mathcal{L}(\lambda) &\sim \prod_{n=1}^N\mathcal{N}(\hat{\lambda}_n, \sigma_{\hat{\lambda}_n}) \nonumber\\
+    \mathcal{L}(i_\star) &\sim \prod_{n=1}^N\mathcal{N}({\rm Obs}_{\star,n}, \sigma_{{\rm Obs}_{\star,n}}) \,(\textit{optional})\nonumber\\
+    \mathcal{L}(i_{\rm orb}) &\sim \prod_{n=1}^N\mathcal{N}(\hat{i}_{{\rm orb},n}, \sigma_{\hat{i}_{{\rm orb},n}}).
 \end{align}
@@
 
+\subsection{Simulated Data}\label{subsec:sim}
 To investigate the performance of our hierarchical Bayesian framework, we apply it to simulated data, generated with a known ground-truth stellar obliquity distribution.
-\edits{We test the following four $\cos{\psi}$ distributions: a uniform distribution bounded between $-1$ and $1$ such that $\cos{\psi} \sim \mathcal{U}(-1,1)$, and three truncated Normal distributions bounded between $-1$ and $1$, such that $\cos{\psi} \sim \textit{truncnorm}(0, 0.2)$, $\cos{\psi} \sim \textit{truncnorm}(-0.4, 0.2)$, and $\cos{\psi} \sim \textit{truncnorm}(0.4, 0.2)$, where the first and second number are the mean and standard deviation of the normal distribution, respectively.}
+\edits{We test the following four $\cos{\psi}$ distributions: a uniform distribution bounded between $-1$ and $1$ such that $\cos{\psi} \sim \mathcal{U}(-1,1)$, and three truncated Normal distributions bounded between $-1$ and $1$, such that $\cos{\psi} \sim \mathcal{N}(0, 0.2)$, $\cos{\psi} \sim \mathcal{N}(-0.4, 0.2)$, and $\cos{\psi} \sim \mathcal{N}(0.4, 0.2)$, where the first number is the mean of the distribution and the second number is the standard deviation.}
 For each $\cos{\psi}$ distribution, we randomly generate 200 samples of sky-projected stellar obliquity $\lambda$ and stellar inclination $i_\star$.
@@
 Despite the fact that the inferred distributions without $i_\star$ measurements have marginally wider uncertainties, as indicated by shallow color contours, the modes and widths of the inferred stellar obliquity distributions are consistent with or without $i_\star$ likelihood.
-Since the injected distributions in rows 2-4 are Normal distributions, it should not be surprising that the inferred distributions, which are Beta distributions, may not exactly match the injected distributions.
-\dfm{Add some more here about how the true distribution is not included within the support of the distribution we're using to fit.
-One thought: perhaps we would actually do better if we used the two betas here too?}
+\edits{Since the injected distributions in rows 2-4 are Normal distributions, it should not be surprising that the inferred distributions, which are Beta distributions, may not exactly match the injected distributions.
+Given that the true distributions are not included within the support of the distributions we are using to fit, it is impressive to see how well the underlying distributions could be recovered.}
 
@@
 
-\subsection{Jacobian transformations between the $\psi$, $\lambda$, and \lowercase{$i_\star$} distributions}\label{sec:jacobian}
+\subsection{Jacobian Transformations Between the $\psi$, $\lambda$, and \lowercase{$i_\star$} Distributions}\label{subsec:jacobian}
 
-In this subsection, we aim to gain insight into the reasons behind the predominant role of the sky-projected stellar obliquity distribution and the less significant impact of the stellar inclination distribution in the inference of the stellar obliquity distribution.
-
-In Figure~\ref{fig:coord}, we present two coordinate systems that describe the stellar spin axis $\bf{n}_{\star}$ for a given orbital axis $\bf{n}_{\rm orb}$.
-In both coordinates, we set the observer's line of sight to be one of the horizontal axes (represented as the conventional $x$-axis in Cartesian coordinates), and the orbital axis of the planet to be the vertical axis (represented as the conventional $z$-axis in Cartesian coordinates).
+Next, we aim to gain insight into the reasons behind the predominant role of the sky-projected stellar obliquity distribution and the less significant impact of the stellar inclination distribution in the inference of the stellar obliquity distribution.
 To simplify the problem, the orbital inclination of the transiting planet is assumed to be $90\degr$ in this illustration.
-As discussed in Section~\ref{subsec:sims}, the assumption will not compromise the stellar obliquity distribution inference for transiting planets.
-
-The $\{\psi, \theta\}$ coordinate system shown in panel (a) relates to the physical properties of a planetary system.
-$\psi$ is the angle between the stellar spin axis and the planetary orbital axis, and $\theta$ is the azimuthal angle of the stellar spin axis.
-For the given $\bf{n}_{\rm orb}$ axis, if $\bf{n}_{\star}$ is a random vector with uniform distribution on a three-dimensional sphere, i.e., $\bf{n}_{\star}$ is isotropically distributed, the probability density function of $\psi$ follows $p(\psi) = \sin{\psi}$ and $p(\theta) = 1/2\pi$.
-\edits{Here $p(x)$ denote the probability density distribution of $x$.}
-Since $p(\cos{\psi}) = p(\psi) \abs{\dv*{\psi}{\cos{\psi}}}$, $p(\cos{\psi})$ is uniformly distributed between $-1$ and $1$ for isotropic stellar obliquity.
-The $\{\lambda, i_\star\}$ coordinate system shown in panel (b) relates to observed properties.
-$\lambda$ is the sky-projected stellar obliquity and $i_\star$ is the stellar inclination.
-If $\bf{n}_{\star}$ is isotropically distributed, $p(i_\star) = \sin{i_\star}$ and $p(\lambda) = 1/2\pi$.
-Again, since $p(\cos{i_\star}) = p(i_\star) \abs{\dv*{i_\star}{\cos{i_\star}}}$, $p(\cos{i_\star})$ is uniformly distributed between $-1$ and $1$.
-Because of the observational degeneracy between $i_\star$ and $\pi - i_\star$, the convention is to have $0 \leqslant i_\star \leqslant \pi/2$, and thus $\cos{i_\star} \sim \mathcal{U}(0, 1)$ (i.e., $p(\cos{i_\star}) = 1$).
-This also suggests $\theta \sim \mathcal{U}(-\pi/2, \pi/2)$.
-
-We could find the mathematical relations between $\{\psi, \theta\}$ and $\{\lambda, i_\star\}$ by pairing the Cartesian components of $\bf{n}_{\star}$ in two coordinate systems:
+As discussed earlier, the assumption will not compromise the stellar obliquity distribution inference for transiting planets.
+\edits{We pair the Cartesian components in Equation~(\ref{eqn:nstar1}) and Equation~(\ref{eqn:nstar2}), assuming $i_{\rm orb} = 90\degr$, and get:
 \begin{align}
@@
 \end{align}
+}
 
@@
 We could find the distribution of $\cos{\lambda}$ using the Jacobian transformation from $\cos{\psi}$ and $\cos{\theta}$.
-Since $\psi$ and $\theta$ are independent variables, we could marginalize over $\theta$ to find the relation between the probability density functions between $\lambda$ and $\psi$.
-
+Since $\psi$ and $\theta$ are assumed to be independent variables, we could marginalize over $\theta$ to find the relation between the probability density functions between $\lambda$ and $\psi$.
 The Jacobian transformation follows
@@
 \end{equation}
-Replacing $\sin{i_\star}$ in Equation~(\ref{eq:coord3}) using Equation~(\ref{eq:coord1}), we find 
+\edits{Here $p(x)$ denotes the probability density distribution of $x$.
+To find $\psi$ in terms of $\lambda$ and $\theta$, we replace $\sin{i_\star}$ in Equation~(\ref{eq:coord3}) using Equation~(\ref{eq:coord1}) and find}
 \begin{equation}
@@
 \end{equation}
-Assume $\theta$ is uniformly distributed between $-\pi/2$ and $\pi/2$, $p(\cos{\theta}) = p(\theta) \abs{\dv*{\theta}{\cos{\theta}}} = 2/\pi/(1-\cos^2{\theta)^{1/2}}$, where the factor of 2 is due to two solutions of $\theta$ to the $\cos{\theta}$.
-Putting all the parts together, we get
+Since $\theta$ is uniformly distributed between $0$ and $\pi$, $p(\cos{\theta}) = p(\theta) \abs{\dv*{\theta}{\cos{\theta}}} = 1/\pi/(1-\cos^2{\theta)^{1/2}}$. Putting all the parts together, we get
 \begin{equation}\label{eqn:jac_lam}
-    p(\cos{\lambda}) = \frac{2}{\pi} \int_{0}^{1} (1-\cos^2{\theta}\cos^2{\lambda})^{-3/2} p(\cos{\psi}) d\cos{\theta}.
+    p(\cos{\lambda}) = \frac{1}{\pi} \int_{-1}^{1} (1-\cos^2{\theta}\cos^2{\lambda})^{-3/2} p(\cos{\psi}) d\cos{\theta}.
 \end{equation}
-If $\cos{\psi}$ is uniformly distribution, i.e., $p(\cos{\psi}) = 1/2$, Equation~(\ref{eqn:jac_lam}) becomes $p(\cos{\lambda}) = 1/\pi/\sqrt{1-\cos^2{\lambda}}$, which is equivalent to $\lambda \sim \mathcal{U}(-\pi, \pi)$.
-This suggests $\lambda$ is uniformly distributed for an isotropic $\psi$ distribution.
-For different $\cos{\psi}$ distributions, $p(\lambda)$ can be found by evaluating Equation~(\ref{eqn:jac_lam}).
+In the special case that $\cos{\psi}$ is uniformly distribution, i.e., $p(\cos{\psi}) = 1/2$, Equation~(\ref{eqn:jac_lam}) becomes $p(\cos{\lambda}) = 1/\pi/\sqrt{1-\cos^2{\lambda}}$, which is equivalent to $\lambda \sim \mathcal{U}(0, \pi)$.
+This suggests $\lambda$ is uniformly distributed for an isotropic $\psi$ distribution, as expected.
 
@@
 \end{equation} 
-Again, we assume $\theta$ is uniformly distributed, and this gives $p(\cos{\theta}) = 2/\pi/(1-\cos^2{\theta)^{1/2}}$. Lastly, we transform the $p(\sin{\psi})$ to $p(\cos{\psi})$,
+Again, we assume $\theta$ is uniformly distributed, and this gives $p(\cos{\theta}) = 1/\pi/(1-\cos^2{\theta)^{1/2}}$. Lastly, we transform the $p(\sin{\psi})$ to $p(\cos{\psi})$,
 \begin{equation}
@@
 \begin{align}\label{eqn:jac_istar}
-    p(\cos{i_\star}) = \frac{4}{\pi} \int_{\cos{i_\star}}^{1}& \frac{\cos{i_\star}/\cos{\theta}}{\sqrt{\cos^2{\theta}-\cos^2{i_\star}}} \nonumber\\ 
-    &\frac{1}{\sqrt{1-\cos^2{\theta}}} p(\cos{\psi}) d\cos{\theta}.
+    p(\cos{i_\star}) = \nonumber\\ 
+    \frac{2}{\pi} \int_{\cos{i_\star}}^{1,-1}& \frac{\cos{i_\star}/\cos{\theta}}{\sqrt{\cos^2{\theta}-\cos^2{i_\star}}}
+    \frac{1}{\sqrt{1-\cos^2{\theta}}} p(\cos{\psi}) d\cos{\theta},
 \end{align}
+\edits{where the integral is from $\cos{i_\star}$ to $1$ for $\cos{\theta} > 0$, or from $\cos{i_\star}$ to $-1$ for $\cos{\theta} \leqslant 0$.}
 Note that the lower limit of the integral is $\cos{i_\star}$ instead of 0 since $\lvert \cos{\theta}/\cos{i_\star} \rvert \geqslant 1$.
@@
     \caption{Simulated $\cos{\psi}$ distributions (\nth{1} column) and the corresponding distributions of sky-projected stellar obliquity $\lambda$ (\nth{2} column) and stellar inclination $i_\star$ (\nth{3} column).
-The inferred $\cos{\psi}$ distributions assuming isotropic stellar inclinations are shown in the \nth{4} column.
 The grey histograms present the random samplings of $\lambda$ and $i_\star$ from the $\cos{\psi}$ distributions, and the blue curves present the numerical solutions.}
@@
 Using Equation~(\ref{eqn:jac_lam}) and (\ref{eqn:jac_istar}), we can now derive the $\lambda$ and $i_\star$ distributions for any given $\psi$ distributions, assuming the azimuthal angle of the stellar spin axis $\theta$ is random.
-In Figure~\ref{fig:transform}, we present numerical solutions of the $\lambda$ and $i_\star$ distributions for four different $\cos{\psi}$ distributions.
+In Figure~\ref{fig:transform}, we present numerical solutions of the $\lambda$ and $i_\star$ distributions for four different $\cos{\psi}$ distributions used in Section~\ref{subsec:sim}.
 The top row of Figure~\ref{fig:transform} shows an isotropic $\psi$ distribution, where $\cos{\psi} \sim \mathcal{U}(-1,1)$.
-The second, third, and fourth rows of Figure~\ref{fig:transform} present Normal distributions of $\cos{\psi}$ following $\mathcal{N}(0,0.2)$, $\mathcal{N}(-0.4,0.2)$, and $\mathcal{N}(0.4,0.2)$, respectively.
+The second, third, and fourth rows of Figure~\ref{fig:transform} present truncated Normal distributions of $\cos{\psi}$ following $\mathcal{N}(0,0.2)$, $\mathcal{N}(-0.4,0.2)$, and $\mathcal{N}(0.4,0.2)$, respectively.
 The blue curves in each row show the numerical solutions of the $\lambda$ and $i_\star$ distributions, while the grey histograms show the sampling of $\lambda$ and $i_\star$ from the $\cos{\psi}$ and $\theta$ distributions.
@@
 Additionally, the degeneracy of the solution could be a significant issue when attempting to infer the $\psi$ distribution from the $i_\star$ distribution.
-For example, when $\cos{\psi} \sim \mathcal{N}(-0.4,0.2)$ or $\cos{\psi} \sim \mathcal{N}(0.4,0.2)$, two $i_\star$ distributions are exactly the same.
+For example, when $\cos{\psi} \sim \mathcal{N}(-0.4,0.2)$ or $\cos{\psi} \sim \mathcal{N}(0.4,0.2)$, two $i_\star$ distributions are exactly the same, \edits{since the corresponding $i_\star$ distributions are identical for $\psi$ distributions symmetric around $\psi = \pi/2$.}
 
@@
 
-In the \nth{4} column in Figure~\ref{fig:transform}, the $\psi$ distribution is found from the $\lambda$ distribution by assuming an isotropic $i_\star$ distribution.
-The method could accurately recover the peaks and widths of the underlying $\psi$ distributions (represented by blue dashed lines) shown in grey histograms.
-The method over-predicts $\cos{\psi}$ at $\pm 1$ because of the over-prediction of $i_\star$ near $90\degr$.
-However, this small deviation from the true distribution has a limited impact on population inference because of the hyperparameters have limited degrees of freedom.
-By comparing the blue curves with the orange curves in Figure~\ref{fig:simulation}, it can be seen that the overestimation of $\cos{\psi}$ near $\pm1$ under the assumption of isotropic $i_\star$ has little effect on the overall inference.
-The distribution is primarily determined by the peak and width of the stellar obliquity distribution, both of which are correctly predicted by the isotropic $i_\star$ model.
+\section{Application to Exoplanetary Systems}\label{sec:observation}
 
-\section{Application to Observations}\label{sec:observation}
-
+\subsection{Stellar Obliquity Distribution of All Exoplanetary Systems}
 We next apply our hierarchical Bayesian framework to a sample of 161 exoplanetary systems with sky-projected stellar obliquity measurements, primarily consisting of Hot Jupiter systems, as summarized in \cite{Albrecht22} Table A1.
@@
 The distribution suggests that there is a pile-up of planetary systems with stellar obliquities less than $40^\circ$ and an isotropic distribution for obliquities between $40^\circ$ and $140^\circ$.
-The fraction of aligned systems $w_0$ dominates the distribution, with $w_0 = 0.67 \pm 0.09$.
-The corresponding posteriors for the parameters of the population Beta distribution are $a_0 = 31.0\pm12.8$ and $b_0 = 0.40\pm0.12$.
-On the other hand, the fraction of misaligned systems is estimated to be $w_1 = 0.33 \pm 0.09$, with posteriors of $a_1 = 1.56\pm0.93$ and $b_1 = 1.64\pm1.58$.
+The fraction of aligned systems $w_1$ dominates the distribution, with $w_1 = 0.719 \pm 0.085$.
+The corresponding posteriors for the parameters of the population Beta distribution are $\mu_1 = 0.976 \pm 0.022$ and $\kappa_1 = 14.1 \pm 26.6$.
+On the other hand, the fraction of misaligned systems is estimated to be $w_0 = 0.281 \pm 0.085$, with posteriors of $\mu_1 = 0.434 \pm 0.088$ and $\kappa_1 = 4.2 \pm 5.6$.
+\edits{Note that the $\mu_{0,1}$ need to be transformed by $2\mu-1$ to represent to true means of the $\cos{\psi}$ distributions.}
 
@@
         \caption{Inferred stellar obliquity distribution for all exoplanetary systems with sky-projected stellar obliquity measurements.
-This inference is based purely on the observed sky-projected obliquities.}
+This inference is based purely on the observed sky-projected obliquities using a two-component model.
+\edits{The blue line and shallow contours represent the median, 1-$\sigma$ and 2-$\sigma$ uncertainties of the inferred distributions.
+The medians of the two components are labeled in dashed and dotted-dashed lines, respectively.}}
         \label{fig:psi_dist}
@@
 There are at least two potential explanations for this difference: (1) the subsample with $i_\star$ measurements is small and only includes about 20 misaligned systems, and (2) the requirement for $i_\star$ measurements could introduce selection biases in the sample.
-\dfm{Expand on this a little bit, I think.
-Maybe add a short paragraph for each.}
+\edits{First, since the sample size of the misaligned sample with $i_\star$ measurements is relatively small ($N < 30$), the observed sample, even if it is unbiased, may not be able to represent the underlying distribution.
+Second, the requirement of $i_\star$ measurements could introduce selection biases.
+The rotation modulation technique is most applicable to cool stars with spots, whereas misaligned Hot Jupiters are mostly found around hot stars.
+To measure the stellar inclinations of the hot-star hosts of misaligned Hot Jupiters, the gravity darkening technique is commonly used, but the technique is biased at detecting polar-orbit planets.}
 
-The inferred stellar obliquity distribution indicates that approximately $67\pm9$\% of the systems have a stellar obliquity of less than $40\degr$, and approximately $33\pm9$\% of the systems follow a nearly isotropic stellar obliquity distribution ranging from $\sim$$40\degr$ to $\sim$$140\degr$.
-These findings could have significant implications for the formation and evolution of close-in planetary systems.
+The inferred stellar obliquity distribution indicates that approximately $72\pm9$\% of the systems have a stellar obliquity of less than $40\degr$, and approximately $28\pm9$\% of the systems follow a nearly isotropic stellar obliquity distribution ranging from $\sim$$40\degr$ to $\sim$$140\degr$.
+These findings could have significant implications for the formation and evolution of close-in planetary systems, \edits{particularly on Hot Jupiters.
+The diverse distribution disfavors dynamical mechanisms, such as secular chaos which tends to produce stellar obliquities less than $90\degr$, or stellar Kozai which tends to produce stellar obliquities clustered at certain angles.}
 The broad distribution of misaligned systems is in good agreement with the predicted outcome of multiple giant planets scattering after a convergent disk migration, as proposed by various studies, such as \cite{Nagasawa11} and \cite{Beague12}.

### 2023-04-18 src/scripts/sim.py
update jn

@@ -139,52 +139,52 @@ def posteriors(this_model):
                                   chains=4, random_seed=int(datetime.now().strftime("%Y%m%d")))
 
 
-    with pm.Model() as model_nolam:
+    # with pm.Model() as model_nolam:
 
-        ncomps = 1
+    #     ncomps = 1
         
-        # hyperprior
-        w = pm.Dirichlet('w', np.ones(ncomps))
+    #     # hyperprior
+    #     w = pm.Dirichlet('w', np.ones(ncomps))
         
-        if ncomps > 1:
-            μ = pm.Uniform('μ', lower=0., upper=1., shape=ncomps, 
-                           transform=pm.distributions.transforms.Ordered(), 
-                           initval=np.sort(np.random.rand(ncomps)))
-        else:
-            μ = pm.Uniform('μ', lower=0., upper=1., shape=ncomps)
+    #     if ncomps > 1:
+    #         μ = pm.Uniform('μ', lower=0., upper=1., shape=ncomps, 
+    #                        transform=pm.distributions.transforms.Ordered(), 
+    #                        initval=np.sort(np.random.rand(ncomps)))
+    #     else:
+    #         μ = pm.Uniform('μ', lower=0., upper=1., shape=ncomps)
             
-        logκ = pm.Normal('logκ', 3., shape=ncomps)
-        κ = pm.Deterministic('κ', pm.math.exp(logκ))
+    #     logκ = pm.Normal('logκ', 3., shape=ncomps)
+    #     κ = pm.Deterministic('κ', pm.math.exp(logκ))
       
-        a = pm.Deterministic('a', μ*κ)
-        b = pm.Deterministic('b', (1.-μ)*κ)
+    #     a = pm.Deterministic('a', μ*κ)
+    #     b = pm.Deterministic('b', (1.-μ)*κ)
         
-        # mixture cosψ distribution
-        u = pm.Mixture('u', w=w, comp_dists=pm.Beta.dist(a, b, shape=(ncomps,)), shape=nsample)
+    #     # mixture cosψ distribution
+    #     u = pm.Mixture('u', w=w, comp_dists=pm.Beta.dist(a, b, shape=(ncomps,)), shape=nsample)
         
[…]

### 2023-04-18 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 When applying the framework to all exoplanetary systems with measured sky-projected stellar obliquity, we find that the inferred population-level obliquity distribution is unimodal and peaked at zero degrees \edits{with a fraction of $72\pm9$\,\%.
-The small fraction of misaligned systems ($28\pm9$\,\%) has nearly isotropic stellar obliquities with no strong clustering near 90$\degr$.
+The $28\pm9$\,\% fraction of misaligned systems has nearly isotropic stellar obliquities with no strong clustering near 90$\degr$.
 The diverse range of stellar obliquities prefers dynamic mechanisms, such as planet-planet scattering after a convergent disk migration, which could produce both prograde and retrograde orbits with no strong inclination concentrations.}
@@
 
-\subsection{Stellar Obliquity Distribution of All Exoplanetary Systems}
 We next apply our hierarchical Bayesian framework to a sample of 161 exoplanetary systems with sky-projected stellar obliquity measurements, primarily consisting of Hot Jupiter systems, as summarized in \cite{Albrecht22} Table A1.
@@
 
-Finally, we applied the framework to all exoplanetary systems to all exoplanetary systems with available sky-projected stellar obliquities and found that approximately $67\pm9$\% of the systems have a stellar obliquity less than $40\degr$, and approximately $33\pm9$\% of the systems follow a nearly isotropic stellar obliquity distribution between $\sim$$40\degr$ and $\sim$$140\degr$.
+Finally, we applied the framework to all exoplanetary systems to all exoplanetary systems with available sky-projected stellar obliquities and found that approximately $72\pm9$\% of the systems have a stellar obliquity less than $40\degr$, and approximately $28\pm9$\% of the systems follow a nearly isotropic stellar obliquity distribution between $\sim$$40\degr$ and $\sim$$140\degr$.
 The distribution could have important implications for the formation and evolution of close-in planetary systems and is worth further investigation.

### 2023-05-01 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-Stellar obliquity $\psi$, the angle between a planet's orbital axis and its host star's spin axis, traces the formation and evolution of a planetary system.
-In transiting exoplanet observations, only the sky-projected stellar obliquity $\lambda$ can be measured.
-To determine the true stellar obliquity $\psi$, information about the stellar inclination $i_\star$ is also needed.
-In this paper, we demonstrate that while a constraint on the stellar inclination is crucial for measuring the obliquity of an individual system, it is not required for robust determination of the population-level stellar obliquity distribution.
+Stellar obliquity, the angle between a planet's orbital axis and its host star's spin axis, traces the formation and evolution of a planetary system.
+In transiting exoplanet observations, only the sky-projected stellar obliquity can be measured, but this can be de-projected using an estimate of the stellar obliquity.
+In this paper, we introduce a flexible, hierarchical Bayesian framework that can be used to infer the stellar obliquity distribution solely from sky-projected stellar obliquities, including stellar inclination measurements when available.
+We demonstrate that while a constraint on the stellar inclination is crucial for measuring the obliquity of an individual system, it is not required for robust determination of the population-level stellar obliquity distribution.
 In practice, the constraints on the stellar obliquity distribution are mainly driven by the sky-projected stellar obliquities.
-We introduce a flexible, hierarchical Bayesian framework that can be used to infer the stellar obliquity distribution solely from sky-projected stellar obliquities, including stellar inclination measurements when available.
-When applying the framework to all exoplanetary systems with measured sky-projected stellar obliquity, we find that the inferred population-level obliquity distribution is unimodal and peaked at zero degrees \edits{with a fraction of $72\pm9$\,\%.
-The $28\pm9$\,\% fraction of misaligned systems has nearly isotropic stellar obliquities with no strong clustering near 90$\degr$.
-The diverse range of stellar obliquities prefers dynamic mechanisms, such as planet-planet scattering after a convergent disk migration, which could produce both prograde and retrograde orbits with no strong inclination concentrations.}
+
+When applying the framework to all exoplanetary systems with measured sky-projected stellar obliquity, which are mostly Hot Jupiter systems, we find that the inferred population-level obliquity distribution is unimodal and peaked at zero degrees.
+The misaligned systems have nearly isotropic stellar obliquities with no strong clustering near 90$\degr$.
+The diverse range of stellar obliquities prefers dynamic mechanisms, such as planet-planet scattering after a convergent disk migration, which could produce both prograde and retrograde orbits with no strong inclination concentrations.
 
@@
 It is as yet unclear if, and to what extent, all of these physical and dynamic processes apply to exoplanetary systems.
-These proposed mechanisms all make different predictions on stellar obliquity distributions \citep[see][and references therein]{Albrecht22, Dawson18}.
+These proposed mechanisms all make different predictions on stellar obliquity distributions with a focus on Hot Jupiter systems \citep[see][and references therein]{Albrecht22, Dawson18}.
 For example, the secular chaos mechanism tends to produce a stellar obliquity distribution with $\psi < 90\degr$ \citep[e.g.,][]{Teyssandier19}.
-\edits{The stellar von Zeipel-Lidov-Kozai (ZLK) mechanism predicts a bimodal stellar obliquity distribution, concentrated at $40\degr$ and $140\degr$ \citep[e.g.,][]{Fabrycky07, Anderson16, Vick19}, assuming a zero stellar obliquity when the ZLK oscillation begins, or a broad stellar obliquity peaked near $90\degr$ if we drop the assumption \citep{Vick23}.} The multiple-planet scattering mechanism results in a majority of aligned systems, with a small fraction of systems at a diverse range of stellar obliquities \citep[e.g.,][]{Beague12}.
+The stellar von Zeipel-Lidov-Kozai (ZLK) mechanism predicts a bimodal stellar obliquity distribution, concentrated at $40\degr$ and $140\degr$ \citep[e.g.,][]{Fabrycky07, Anderson16, Vick19}, assuming a zero stellar obliquity when the ZLK oscillation begins, or a broad stellar obliquity peaked near $90\degr$ if we drop the assumption \citep{Vick23}.
+The multiple-planet scattering mechanism results in a majority of aligned systems, with a small fraction of systems at a diverse range of stellar obliquities \citep[e.g.,][]{Beague12}.
 With these predictions in mind, we aim to determine the dominant mechanisms responsible for shaping close-in planetary systems by characterizing the stellar obliquity distribution of exoplanetary systems through a Bayesian approach.
@@
 In this study, we aim to gain a deeper understanding of this relationship and develop a statistical approach to infer the stellar obliquity distribution.
-\edits{In Section~\ref{sec:coords}, we find the expression of sky-projected stellar olibquity $\lambda$ and stellar inclination $i_\star$ in terms of the orbital inclination $i_{\rm orb}$, stellar obliquity $\psi$, and the azimuthal angle of the stellar spin axis $\theta$.
+In Section~\ref{sec:coords}, we find the expression of sky-projected stellar obliquity $\lambda$ and stellar inclination $i_\star$ in terms of the orbital inclination $i_{\rm orb}$, stellar obliquity $\psi$, and the azimuthal angle of the stellar spin axis $\theta$.
 In Section~\ref{sec:hbm}, we introduce a flexible, hierarchical Bayesian framework that allows us to infer the stellar obliquity distribution of a sample.
@@
 We discuss the implication of the stellar obliquity distribution for Hot Jupiter origins.
-}
-
-{\color{purple}
 
@@
         \fig{figures/coord_psi.pdf}{0.45\textwidth}{\vspace*{-1.8cm}(a) The $\{\psi, \theta\}$ coordinate system.
-\edits{The planetary orbital axis is set to be aligned with $\vb*{z}$-axis.} The grey circle corresponds to a constant $\psi$ value and its circumference is proportional to $\sin{\psi}$.}
+The planetary orbital axis is set to be aligned with $\vb*{z}$-axis.
+The grey circle corresponds to a constant $\psi$ value and its circumference is proportional to $\sin{\psi}$.}
         \fig{figures/coord_lam.pdf}{0.45\textwidth}{\vspace*{-1.8cm}(b) The $\{\lambda, i_\star\}$ coordinate system.
-\edits{The observer's light of sight is set to be aligned with $\vb*{x'}$-axis.} The grey circle corresponds to a constant $i_\star$ value and its circumference is proportional to $\sin{i_\star}$.}
+The observer's light of sight is set to be aligned with $\vb*{x'}$-axis.
+The grey circle corresponds to a constant $i_\star$ value and its circumference is proportional to $\sin{i_\star}$.}
     }
@@
     \caption{Two coordinate systems that describe the stellar spin axis $\vb{\hat{n}_\star}$ and the planet's orbital axis $\vb{\hat{n}_{\rm orb}}$.
-\edits{The $\{\psi, \theta\}$ coordinate system setup is motivated by the physical properties of a planetary system.
-The $\{\lambda, i_\star\}$ coordinate system setup is motivated by observables.} }
+The $\{\psi, \theta\}$ coordinate system setup is motivated by the physical properties of a planetary system.
+The $\{\lambda, i_\star\}$ coordinate system setup is motivated by observables.}
     \label{fig:coord}
@@
 \end{equation}
-Particularly, we are interested in finding the distribution of the angle $\psi$.
 
@@
 
-}
-
 \section{Hierarchical Bayesian Framework}\label{sec:hbm}
@@
     \includegraphics[width=0.8\linewidth]{figures/graph.pdf}
-    \caption{\edits{A probabilistic graphical model showing the conditional structure of our hierarchical Bayesian framework for inferring the stellar obliquity distribution of exoplanetary systems.
+    \caption{A probabilistic graphical model showing the conditional structure of our hierarchical Bayesian framework for inferring the stellar obliquity distribution of exoplanetary systems.
 The population model is parameterized by hyperparameters $\vb*{\beta}$, and it is constrained by the stellar obliquity $\psi_n$ of each $n_{\rm th}$ individual system.
@@
     Properties of the star other than its inclination, such as its radius and rotation period, are represented by $\gamma_{\star, n}$, and these can inform the constraint on $i_{\star, n}$, if available.
-    $\hat{\lambda}_n$ has the measured sky-projected stellar obliquity and its uncertainty, Obs$_{\star,n}$ contains all observed properties of the star with uncertainties, and $\hat{i}_{{\rm orb}, n}$ has measured orbital inclination and its uncertainty.}}
+    $\hat{\lambda}_n$ has the measured sky-projected stellar obliquity and its uncertainty, Obs$_{\star,n}$ contains all observed properties of the star with uncertainties, and $\hat{i}_{{\rm orb}, n}$ has measured orbital inclination and its uncertainty.}
     \label{fig:graph}
@@
 If stellar obliquity distribution is isotropic, $\cos{\psi}$ is uniformly distributed between -1 and 1.
-To flexibly model the stellar obliquity distribution, \edits{we select a multi-component mixture of Beta distributions with hyperparameters $\bm{\beta} = \{\bm{w},\bm{\mu},\bm{\kappa}\}$, where each hyperparameter has a dimension of the number of components.
+To flexibly model the stellar obliquity distribution, we select a multi-component mixture of Beta distributions with hyperparameters $\bm{\beta} = \{\bm{w},\bm{\mu},\bm{\kappa}\}$, where each hyperparameter has a dimension of the number of components.
 For example, for a two-component mixture model, each hyperparameter has a dimension of 2.
@@
 To ensure the robustness of the inferred distributions in such cases, it is crucial to test their sensitivity to different hyperpriors.
-}
 
@@
 If $\gamma_{\star, n}$ is available, we construct Normal distributions with means and standard deviations from measurements.
-\edits{For the orbital inclination $i_{\rm orb}$, following the convention, we limit it to $\left[0, \pi/2\right]$.
+For the orbital inclination $i_{\rm orb}$, following the convention, we limit it to $\left[0, \pi/2\right]$.
 To not underestimate the $\psi$, we then set $i_\star$ to be varied from $0\degr$ to $180\degr$.
@@
 We find doing so greatly improves the sampling performance while not compromising the inference of the $\psi$ distribution due to the symmetry.
-The $\left[0, \pi\right]$ support avoids the otherwise discontinuity at $\theta = 0$ and $\theta = \pi$.}
+The $\left[0, \pi\right]$ support avoids the otherwise discontinuity at $\theta = 0$ and $\theta = \pi$.
 
@@
 To investigate the performance of our hierarchical Bayesian framework, we apply it to simulated data, generated with a known ground-truth stellar obliquity distribution.
-\edits{We test the following four $\cos{\psi}$ distributions: a uniform distribution bounded between $-1$ and $1$ such that $\cos{\psi} \sim \mathcal{U}(-1,1)$, and three truncated Normal distributions bounded between $-1$ and $1$, such that $\cos{\psi} \sim \mathcal{N}(0, 0.2)$, $\cos{\psi} \sim \mathcal{N}(-0.4, 0.2)$, and $\cos{\psi} \sim \mathcal{N}(0.4, 0.2)$, where the first number is the mean of the distribution and the second number is the standard deviation.}
+We test the following four $\cos{\psi}$ distributions: a uniform distribution bounded between $-1$ and $1$ such that $\cos{\psi} \sim \mathcal{U}(-1,1)$, and three truncated Normal distributions bounded between $-1$ and $1$, such that $\cos{\psi} \sim \mathcal{N}(0, 0.2)$, $\cos{\psi} \sim \mathcal{N}(-0.4, 0.2)$, and $\cos{\psi} \sim \mathcal{N}(0.4, 0.2)$, where the first number is the mean of the distribution and the second number is the standard deviation.
 For each $\cos{\psi}$ distribution, we randomly generate 200 samples of sky-projected stellar obliquity $\lambda$ and stellar inclination $i_\star$.
@@
 Despite the fact that the inferred distributions without $i_\star$ measurements have marginally wider uncertainties, as indicated by shallow color contours, the modes and widths of the inferred stellar obliquity distributions are consistent with or without $i_\star$ likelihood.
-\edits{Since the injected distributions in rows 2-4 are Normal distributions, it should not be surprising that the inferred distributions, which are Beta distributions, may not exactly match the injected distributions.
-Given that the true distributions are not included within the support of the distributions we are using to fit, it is impressive to see how well the underlying distributions could be recovered.}
+Since the injected distributions in rows 2-4 are Normal distributions, it should not be surprising that the inferred distributions, which are Beta distributions, may not exactly match the injected distributions.
+Given that the true distributions are not included within the support of the distributions we are using to fit, it is impressive to see how well the underlying distributions could be recovered.
 
@@
 As discussed earlier, the assumption will not compromise the stellar obliquity distribution inference for transiting planets.
-\edits{We pair the Cartesian components in Equation~(\ref{eqn:nstar1}) and Equation~(\ref{eqn:nstar2}), assuming $i_{\rm orb} = 90\degr$, and get:
+We pair the Cartesian components in Equation~(\ref{eqn:nstar1}) and Equation~(\ref{eqn:nstar2}), assuming $i_{\rm orb} = 90\degr$, and get:
 \begin{align}
@@
 \end{align}
-}
 
@@
 \end{equation}
-\edits{Here $p(x)$ denotes the probability density distribution of $x$.
-To find $\psi$ in terms of $\lambda$ and $\theta$, we replace $\sin{i_\star}$ in Equation~(\ref{eq:coord3}) using Equation~(\ref{eq:coord1}) and find}
+Here $p(x)$ denotes the probability density distribution of $x$.
+To find $\psi$ in terms of $\lambda$ and $\theta$, we replace $\sin{i_\star}$ in Equation~(\ref{eq:coord3}) using Equation~(\ref{eq:coord1}) and find
 \begin{equation}
-    \cos{\lambda} = \cos{\psi}/\sqrt{1-(1-\cos^2{\psi})\cos^2{\theta}}.
+    \cos{\lambda} = \frac{\cos{\psi}}{\sqrt{1-(1-\cos^2{\psi})\cos^2{\theta}}}.
 \end{equation} 
@@
 \begin{equation}
-    \sin{\psi} = \cos{i_\star}/\cos{\theta}
+    \sin{\psi} = \frac{\cos{i_\star}}{\cos{\theta}}
 \end{equation} and the partial derivative
 \begin{equation}
-    \abs{\pdv{\sin{\psi}}{\cos{i_\star}}} = 1/\cos{\theta}.
+    \abs{\pdv{\sin{\psi}}{\cos{i_\star}}} = \frac{1}{\cos{\theta}}.
 \end{equation} 
@@
 \begin{equation}
-    p(\sin{\psi}) = 2p(\cos{\psi})\sin{\psi}/\sqrt{1-\sin^2{\psi}}.
+    p(\sin{\psi}) = \frac{2p(\cos{\psi})\sin{\psi}}{\sqrt{1-\sin^2{\psi}}}.
 \end{equation}
@@
 \end{align}
-\edits{where the integral is from $\cos{i_\star}$ to $1$ for $\cos{\theta} > 0$, or from $\cos{i_\star}$ to $-1$ for $\cos{\theta} \leqslant 0$.}
+where the integral is from $\cos{i_\star}$ to $1$ for $\cos{\theta} > 0$, or from $\cos{i_\star}$ to $-1$ for $\cos{\theta} \leqslant 0$.
 Note that the lower limit of the integral is $\cos{i_\star}$ instead of 0 since $\lvert \cos{\theta}/\cos{i_\star} \rvert \geqslant 1$.
@@
 Additionally, the degeneracy of the solution could be a significant issue when attempting to infer the $\psi$ distribution from the $i_\star$ distribution.
-For example, when $\cos{\psi} \sim \mathcal{N}(-0.4,0.2)$ or $\cos{\psi} \sim \mathcal{N}(0.4,0.2)$, two $i_\star$ distributions are exactly the same, \edits{since the corresponding $i_\star$ distributions are identical for $\psi$ distributions symmetric around $\psi = \pi/2$.}
+For example, when $\cos{\psi} \sim \mathcal{N}(-0.4,0.2)$ or $\cos{\psi} \sim \mathcal{N}(0.4,0.2)$, two $i_\star$ distributions are exactly the same, since the corresponding $i_\star$ distributions are identical for $\psi$ distributions symmetric around $\psi = \pi/2$.
 
@@
 On the other hand, the fraction of misaligned systems is estimated to be $w_0 = 0.281 \pm 0.085$, with posteriors of $\mu_1 = 0.434 \pm 0.088$ and $\kappa_1 = 4.2 \pm 5.6$.
-\edits{Note that the $\mu_{0,1}$ need to be transformed by $2\mu-1$ to represent to true means of the $\cos{\psi}$ distributions.}
+Note that the $\mu_{0,1}$ needs to be transformed by $2\mu-1$ to represent to true means of the $\cos{\psi}$ distributions.
 
@@
 This inference is based purely on the observed sky-projected obliquities using a two-component model.
-\edits{The blue line and shallow contours represent the median, 1-$\sigma$ and 2-$\sigma$ uncertainties of the inferred distributions.
-The medians of the two components are labeled in dashed and dotted-dashed lines, respectively.}}
+The blue line and shallow contours represent the median, 1-$\sigma$ and 2-$\sigma$ uncertainties of the inferred distributions.
+The medians of the two components are labeled in dashed and dotted-dashed lines, respectively.}
         \label{fig:psi_dist}
@@
 There are at least two potential explanations for this difference: (1) the subsample with $i_\star$ measurements is small and only includes about 20 misaligned systems, and (2) the requirement for $i_\star$ measurements could introduce selection biases in the sample.
-\edits{First, since the sample size of the misaligned sample with $i_\star$ measurements is relatively small ($N < 30$), the observed sample, even if it is unbiased, may not be able to represent the underlying distribution.
+First, since the sample size of the misaligned sample with $i_\star$ measurements is relatively small ($N < 30$), the observed sample, even if it is unbiased, may not be able to represent the underlying distribution.
 Second, the requirement of $i_\star$ measurements could introduce selection biases.
 The rotation modulation technique is most applicable to cool stars with spots, whereas misaligned Hot Jupiters are mostly found around hot stars.
-To measure the stellar inclinations of the hot-star hosts of misaligned Hot Jupiters, the gravity darkening technique is commonly used, but the technique is biased at detecting polar-orbit planets.}
+To measure the stellar inclinations of the hot-star hosts of misaligned Hot Jupiters, the gravity darkening technique is commonly used, but the technique is biased at detecting polar-orbit planets.
 
 The inferred stellar obliquity distribution indicates that approximately $72\pm9$\% of the systems have a stellar obliquity of less than $40\degr$, and approximately $28\pm9$\% of the systems follow a nearly isotropic stellar obliquity distribution ranging from $\sim$$40\degr$ to $\sim$$140\degr$.
-These findings could have significant implications for the formation and evolution of close-in planetary systems, \edits{particularly on Hot Jupiters.
-The diverse distribution disfavors dynamical mechanisms, such as secular chaos which tends to produce stellar obliquities less than $90\degr$, or stellar Kozai which tends to produce stellar obliquities clustered at certain angles.}
+These findings could have significant implications for the formation and evolution of close-in planetary systems, particularly on Hot Jupiters.
+The diverse distribution disfavors dynamical mechanisms, such as secular chaos which tends to produce stellar obliquities less than $90\degr$, or stellar Kozai which tends to produce stellar obliquities clustered at certain angles.
 The broad distribution of misaligned systems is in good agreement with the predicted outcome of multiple giant planets scattering after a convergent disk migration, as proposed by various studies, such as \cite{Nagasawa11} and \cite{Beague12}.

### 2023-05-01 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 % Define document class
-\documentclass[twocolumn,times]{aastex631}
+\documentclass[twocolumn,times,linenumbers]{aastex631}
 \usepackage{showyourwork}
@@
 
-When applying the framework to all exoplanetary systems with measured sky-projected stellar obliquity, which are mostly Hot Jupiter systems, we find that the inferred population-level obliquity distribution is unimodal and peaked at zero degrees.
+When applying the framework to all systems with measured sky-projected stellar obliquity, which are mostly Hot Jupiter systems, we find that the inferred population-level obliquity distribution is unimodal and peaked at zero degrees.
 The misaligned systems have nearly isotropic stellar obliquities with no strong clustering near 90$\degr$.
-The diverse range of stellar obliquities prefers dynamic mechanisms, such as planet-planet scattering after a convergent disk migration, which could produce both prograde and retrograde orbits with no strong inclination concentrations.
+The diverse range of stellar obliquities prefers dynamic mechanisms, such as planet-planet scattering after a convergent disk migration, which could produce both prograde and retrograde orbits with no strong inclination concentrations of close-in planets.
 

### 2023-05-01 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 The misaligned systems have nearly isotropic stellar obliquities with no strong clustering near 90$\degr$.
-The diverse range of stellar obliquities prefers dynamic mechanisms, such as planet-planet scattering after a convergent disk migration, which could produce both prograde and retrograde orbits with no strong inclination concentrations of close-in planets.
+The diverse range of stellar obliquities prefers dynamic mechanisms, such as planet-planet scattering after a convergent disk migration, which could produce both prograde and retrograde orbits of close-in planets with no strong inclination concentrations other than 0$\degr$.
 

### 2023-05-19 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-\received{}
+\received{May 1, 2023}
 \revised{}
@@
 \end{abstract}
+
+\keywords{Exoplanets (498) --- Exoplanet dynamics (490) --- Bayesian statistics (1900)}
 
@@
 \begin{align}
-    w_{0,1} &\sim {\rm Dirichlet}(1, 1) \nonumber\\
+    \edits{w_{0,1}} &\edits{\sim {\rm Bernoulli}(1/2)} \nonumber\\
     u_{0,1} &\sim {\rm Beta}(\mu_{0,1}\kappa_{0,1}, (1-\mu_{0,1})\kappa_{0,1}) \nonumber\\
-    \cos{\psi} &\sim 2\,(w_0 u_0 + w_1 u_1) - 1.
+    \cos{\psi} &= 2\,(w_0 u_0 + w_1 u_1) - 1.
 \end{align}
@@
 The probabilistic model is constructed using the $\mathtt{PyMC}$ package version $\mathtt{v5.1.2}$ \citep{pymc}, and the posteriors are sampled with the No-U-Turn Sampler \citep[NUTS;][]{Hoffman11}, which is a gradient-based Markov chain Monte Carlo (MCMC) sampling algorithm.
-This paper's figures and simulations are completely reproducible and were created using the \showyourwork package.
+This paper's figures and simulations are completely reproducible and were created using the $\mathtt{showyourwork}$ package.
 The open-source code is available on GitHub \href{https://github.com/jiayindong/obliquity}{(https://github.com/jiayindong/obliquity\,\faGithub)}.
@@
 
-This study was conducted using the \href{https://github.com/showyourwork/showyourwork}{\showyourwork} reproducible workflow \citep{Luger2021}, which leverages continuous integration to automate the data retrieval from \href{https://zenodo.org/}{zenodo.org}, figure generation, and manuscript compilation.
+This study was conducted using the \href{https://github.com/showyourwork/showyourwork}{$\mathtt{showyourwork}$} reproducible workflow \citep{Luger2021}, which leverages continuous integration to automate the data retrieval from \href{https://zenodo.org/}{zenodo.org}, figure generation, and manuscript compilation.
 The script used to produce each figure can be accessed via a link in the corresponding figure caption, as it corresponds to the latest build of the manuscript.
@@
 
-\software{$\mathtt{ArviZ}$ \citep{arviz_2019}, $\mathtt{Jupyter}$ \citep{Jupyter}, $\mathtt{Matplotlib}$ \citep{Matplotlib07, Matplotlib16}, $\mathtt{NumPy}$ \citep{NumPy11, NumPy20}, $\mathtt{pandas}$ \citep{mckinney-proc-scipy-2010, reback2020pandas}, $\mathtt{PyMC}$ \citep{pymc}, $\mathtt{SciPy}$ \citep{2020SciPy-NMeth}, \showyourwork \citep{Luger2021}}
+\software{$\mathtt{ArviZ}$ \citep{arviz_2019}, $\mathtt{Jupyter}$ \citep{Jupyter}, $\mathtt{Matplotlib}$ \citep{Matplotlib07, Matplotlib16}, $\mathtt{NumPy}$ \citep{NumPy11, NumPy20}, $\mathtt{pandas}$ \citep{mckinney-proc-scipy-2010, reback2020pandas}, $\mathtt{PyMC}$ \citep{pymc}, $\mathtt{SciPy}$ \citep{2020SciPy-NMeth}, $\mathtt{showyourwork}$ \citep{Luger2021}}
 

### 2023-05-21 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 Each row of Figure~\ref{fig:simulation} corresponds to a different simulation distribution.
-The orange curve and contours in the left column are the median, 1-$\sigma$ and 2-$\sigma$ uncertainties of the inferred $\cos{\psi}$ distribution with stellar inclination information.
-The right column shows the same inferences (in blue) when constraints on stellar inclination are not included.
+The \edits{blue} curve and contours in the left column are the median, 1-$\sigma$ and 2-$\sigma$ uncertainties of the inferred $\cos{\psi}$ distribution with stellar inclination information.
+The right column shows the same inferences (in \edits{orange}) when constraints on stellar inclination are not included.
 Surprisingly, Figure~\ref{fig:simulation} demonstrates that our inference procedure recovers the true distribution for $\cos{\psi}$ equally well, regardless of the inclusion of $i_\star$ measurements.
 Despite the fact that the inferred distributions without $i_\star$ measurements have marginally wider uncertainties, as indicated by shallow color contours, the modes and widths of the inferred stellar obliquity distributions are consistent with or without $i_\star$ likelihood.
-Since the injected distributions in rows 2-4 are Normal distributions, it should not be surprising that the inferred distributions, which are Beta distributions, may not exactly match the injected distributions.
+Since the injected distributions in rows 2--4 are Normal distributions, it should not be surprising that the inferred distributions, which are Beta distributions, may not exactly match the injected distributions.
 Given that the true distributions are not included within the support of the distributions we are using to fit, it is impressive to see how well the underlying distributions could be recovered.
@@
 
+\begin{figure*}[ht!]
+    \script{polar_plot.py}
+    \begin{centering}
+        \includegraphics{figures/polar.pdf}
+        \caption{Inferred stellar obliquity distribution for the subsample of exoplanetary systems with both sky-projected stellar obliquity measurements and stellar inclination measurements.}
+        \label{fig:polar}
+    \end{centering}
+\end{figure*}
+
 The discrepancy between this inference for the full sample of exoplanet systems and the previous analysis of the subsample with $i_\star$ measurements \citep{Albrecht21} warrants further investigation.

### 2023-05-21 src/scripts/polar.py
add polar plots

@@ -1,0 +1,245 @@
+import os
+from pathlib import Path
+os.environ["PATH"] += os.pathsep + str(Path.home() / "bin")
+
+import sys
+import subprocess
+from datetime import datetime
+
+import paths
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+import pymc as pm
+import arviz as az
+import pytensor.tensor as at
+import pytensor
+import scipy.stats as stats
+
+from matplotlib import rc
+rc('font', **{'family':'sans-serif'})
+rc('text', usetex=False)
+
+plt.rcParams['xtick.top'] =  True
+plt.rcParams['xtick.direction'] =  'in'
+plt.rcParams['xtick.major.width'] =  1.0
+plt.rcParams['xtick.minor.width'] =  1.0
+plt.rcParams['ytick.right'] =  True
+plt.rcParams['ytick.direction'] =  'in'
+plt.rcParams['ytick.major.width'] = 1.0
+plt.rcParams['ytick.minor.width'] =  1.0
+plt.rcParams['lines.markeredgewidth'] =  1.0
+
+def read_table2(df, var_name):
+    size = len(df)
+    
+    var = []
+    err_var = []
+    
+    for i in range(size):
+    
+        this_var = df['%s'%var_name][i]
+        
+        if this_var == "cdots":
+            var.append(np.nan)
+            err_var.append(np.nan)
+        elif '+or-' in this_var:
+            for j in range(len(this_var)):
+                if this_var[j] == '+':
+                    var.append(float(this_var[:j-1]))
+                if this_var[j] == '-':
+                    err_var.append(float(this_var[j+2:]))
+        else:
+            for j in range(len(this_var)-1):
+                if this_var[j:j+2] == '${':
+                    for k in range(6):
+                        if this_var[j+2+k] == '}':
+                            var.append(float(this_var[j+2:j+2+k]))
+                elif this_var[j] == '_':
+                    for k in range(10):
[…]

### 2023-05-21 src/scripts/polar_plot.py

@@ -1,0 +1,94 @@
+import paths
+import os
+from pathlib import Path
+os.environ["PATH"] += os.pathsep + str(Path.home() / "bin")
+
+import subprocess
+import sys
+
+import pymc as pm
+import arviz as az
+import pytensor.tensor as at
+
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+
+import scipy.integrate as integrate
+from scipy import stats
+
+from matplotlib import rc
+rc('font', **{'family':'sans-serif'})
+rc('text', usetex=True)
+rc('text.latex', preamble=r'\usepackage{physics}')
+
+plt.rcParams['xtick.top'] =  True
+plt.rcParams['xtick.direction'] =  'in'
+plt.rcParams['xtick.major.width'] =  1.0
+plt.rcParams['xtick.minor.width'] =  1.0
+plt.rcParams['ytick.right'] =  True
+plt.rcParams['ytick.direction'] =  'in'
+plt.rcParams['ytick.major.width'] = 1.0
+plt.rcParams['ytick.minor.width'] =  1.0
+plt.rcParams['lines.markeredgewidth'] =  1.0
+
+from scipy.stats import beta
+x = np.linspace(1e-5,1-1e-5,1000)
+
+# load the posteriors from the MCMC and calculate psi dists
+def polar_dist_draws(model_name):
+
+	istar_idata = az.from_netcdf(paths.data / "polar" / (model_name + "_istar.nc"))
+	noistar_idata = az.from_netcdf(paths.data / "polar" / (model_name + "_noistar.nc"))
+
+	post = istar_idata.posterior
+	istar_draws = np.zeros(shape=(len(x),4000))
+	for a in range(4):
+		for b in range(1000):
+			istar_draws[:, a*1000+b] = (post.w[a,b,0].values*beta.pdf(x, post.a[a,b,0], post.b[a,b,0])
+										+post.w[a,b,1].values*beta.pdf(x, post.a[a,b,1], post.b[a,b,1]))
+	
+	post = noistar_idata.posterior
+	noistar_draws = np.zeros(shape=(len(x),4000))
+	for a in range(4):
+		for b in range(1000):
[…]

### 2023-05-21 src/scripts/polar.py
add polar plots

@@ -1,3 +1,4 @@
+# dummy
 import os
 from pathlib import Path
 os.environ["PATH"] += os.pathsep + str(Path.home() / "bin")

### 2023-05-21 src/scripts/polar.py
add polar plots

@@ -1,4 +1,3 @@
-# dummy
 import os
 from pathlib import Path
 os.environ["PATH"] += os.pathsep + str(Path.home() / "bin")

### 2023-05-21 src/scripts/polar.py
add polar plots

@@ -98,6 +98,8 @@ if __name__ == '__main__':
 	# Subsample with Prot measurements
 	nsample2 = int(np.sum(mask2))
 
+	print('Sample size': nsample)
+
 	sim_dir = paths.data / "polar"
 	sim_dir.mkdir(exist_ok=True, parents=True)
 

### 2023-05-21 src/scripts/polar.py
add polar plots

@@ -98,8 +98,6 @@ if __name__ == '__main__':
 	# Subsample with Prot measurements
 	nsample2 = int(np.sum(mask2))
 
-	print('Sample size': nsample)
-
 	sim_dir = paths.data / "polar"
 	sim_dir.mkdir(exist_ok=True, parents=True)
 

### 2023-05-22 src/scripts/sim.py
add beta to sims

@@ -39,9 +39,6 @@ def posteriors(this_model):
     true_lam = idata.posterior.λ.values.ravel()
     obs_lam = true_lam + err_lam*np.random.normal(size=nsample)
 
-    # Not neccessary to limit obs_istar to [0, pi/2]
-    # obs_istar[obs_istar>np.pi/2] = np.pi-obs_istar[obs_istar>np.pi/2]
-
     # Limit obs_lam to [0, pi]
     obs_lam[obs_lam<0] = -obs_lam[obs_lam<0]
 
@@ -294,6 +291,33 @@ if __name__ == '__main__':
         
         iso_cosi = pm.Uniform('iso_cosi', lower=0., upper=1.)
         iso_cosψ = pm.Deterministic('iso_cosψ', at.sqrt(1.-iso_cosi**2)*cosλ)
+
+    # cosψ ~ 2*Beta(3,6)-1
+    with pm.Model() as model_beta:
+
+        cosψ_ = pm.Beta('cosψ_', alpha=3, beta=6)
+        cosψ = pm.Deterministic('cosψ', 2*cosψ_-1)
+        
+        ψ = pm.Deterministic('ψ', at.arccos(cosψ))
+        sinψ = pm.Deterministic('sinψ', at.sqrt(1.-cosψ**2))
+        
+        θ = pm.Uniform('θ', lower=0., upper=np.pi)
+        cosθ = pm.Deterministic('cosθ', at.cos(θ))
+        sinθ = pm.Deterministic('sinθ', at.sin(θ))
+
+        # iorb
+        iorb = np.pi/2
+
+        # find λ in terms of ψ, θ, and iorb
+        λ = pm.Deterministic('λ', at.arctan2(sinψ*sinθ, cosψ*at.sin(iorb)-sinψ*cosθ*at.cos(iorb)))
+        cosλ = pm.Deterministic('cosλ', at.cos(λ))
+
+        # find i in terms of ψ, θ, and iorb
+        cosi = pm.Deterministic('cosi', sinψ*cosθ*at.sin(iorb)+cosψ*at.cos(iorb))
+        i = pm.Deterministic('i', at.arccos(cosi))
+        
+        iso_cosi = pm.Uniform('iso_cosi', lower=0., upper=1.)
+        iso_cosψ = pm.Deterministic('iso_cosψ', at.sqrt(1.-iso_cosi**2)*cosλ)
         
         
     sim_dir = paths.data / "simulation"
[…]

### 2023-05-23 src/scripts/polar.py
add polar data

@@ -30,49 +30,89 @@ plt.rcParams['ytick.major.width'] = 1.0
 plt.rcParams['ytick.minor.width'] =  1.0
 plt.rcParams['lines.markeredgewidth'] =  1.0
 
-def read_table2(df, var_name):
-    size = len(df)
-    
-    var = []
-    err_var = []
-    
-    for i in range(size):
-    
-        this_var = df['%s'%var_name][i]
-        
-        if this_var == "cdots":
-            var.append(np.nan)
-            err_var.append(np.nan)
-        elif '+or-' in this_var:
-            for j in range(len(this_var)):
-                if this_var[j] == '+':
-                    var.append(float(this_var[:j-1]))
-                if this_var[j] == '-':
-                    err_var.append(float(this_var[j+2:]))
-        else:
-            for j in range(len(this_var)-1):
-                if this_var[j:j+2] == '${':
-                    for k in range(6):
-                        if this_var[j+2+k] == '}':
-                            var.append(float(this_var[j+2:j+2+k]))
-                elif this_var[j] == '_':
-                    for k in range(10):
-                        if this_var[j+k] == '^':
-                            low = float(this_var[j+3:j+k-1])
-                            high = float(this_var[j+k+3:len(this_var)-2])
-                    err_var.append((low+high)/2)
-                    
-    return np.array(var), np.array(err_var)
+# Data from Table 2 in Albrecht+21
 
-# Load Table 2 of Albrecht+21
-df = pd.read_csv(paths.data / "Albrecht21_Table2.csv")
+# Radii, err_Radii in Solar unit
[…]

### 2023-05-23 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \received{May 1, 2023}
-\revised{}
+\revised{\today}
 \accepted{}
@@
 To investigate the performance of our hierarchical Bayesian framework, we apply it to simulated data, generated with a known ground-truth stellar obliquity distribution.
-We test the following four $\cos{\psi}$ distributions: a uniform distribution bounded between $-1$ and $1$ such that $\cos{\psi} \sim \mathcal{U}(-1,1)$, and three truncated Normal distributions bounded between $-1$ and $1$, such that $\cos{\psi} \sim \mathcal{N}(0, 0.2)$, $\cos{\psi} \sim \mathcal{N}(-0.4, 0.2)$, and $\cos{\psi} \sim \mathcal{N}(0.4, 0.2)$, where the first number is the mean of the distribution and the second number is the standard deviation.
+We test the following \edits{five} $\cos{\psi}$ distributions: a uniform distribution bounded between $-1$ and $1$ such that $\cos{\psi} \sim \mathcal{U}(-1,1)$, three truncated Normal distributions bounded between $-1$ and $1$, such that $\cos{\psi} \sim \mathcal{N}(0, 0.2)$, $\cos{\psi} \sim \mathcal{N}(-0.4, 0.2)$, and $\cos{\psi} \sim \mathcal{N}(0.4, 0.2)$, where the first number is the mean of the distribution and the second number is the standard deviation, \edits{and lastly, a Beta distribution such that $\cos{\psi} \sim \mathrm{Beta}(3, 6)$ with the support being extended linearly from $[0, 1]$ to $[-1, 1]$}.
 For each $\cos{\psi}$ distribution, we randomly generate 200 samples of sky-projected stellar obliquity $\lambda$ and stellar inclination $i_\star$.
@@
 Each row of Figure~\ref{fig:simulation} corresponds to a different simulation distribution.
-The \edits{blue} curve and contours in the left column are the median, 1-$\sigma$ and 2-$\sigma$ uncertainties of the inferred $\cos{\psi}$ distribution with stellar inclination information.
-The right column shows the same inferences (in \edits{orange}) when constraints on stellar inclination are not included.
+The \edits{blue} curve and contours in the left column are the median, 1-$\sigma$ and 2-$\sigma$ uncertainties of the inferred $\cos{\psi}$ distribution \edits{when constraints on stellar inclination are not included}.
+The right column shows the same inferences (in \edits{orange}) \edits{with stellar inclination information}.
 Surprisingly, Figure~\ref{fig:simulation} demonstrates that our inference procedure recovers the true distribution for $\cos{\psi}$ equally well, regardless of the inclusion of $i_\star$ measurements.
@@
     \includegraphics{figures/simulation.pdf}
-    \caption{Inferred stellar obliquity distributions from sky-projected stellar obliquities with and without information on the stellar inclination, depicted in blue and orange curves, respectively.
+    \caption{Inferred stellar obliquity distributions from sky-projected stellar obliquities \edits{without and with} information on the stellar inclination, depicted in blue and orange curves, respectively.
 Each row presents a set of simulated data with the true distribution of $\cos{\psi}$ indicated by grey dashed curves.

### 2023-05-23 src/scripts/polar.py
add polar data

@@ -15,6 +15,7 @@ import arviz as az
 import pytensor.tensor as at
 import pytensor
 import scipy.stats as stats
+from numpy import nan
 
 from matplotlib import rc
 rc('font', **{'family':'sans-serif'})

### 2023-05-23 src/scripts/polar_plot.py
add polar data

@@ -59,10 +59,10 @@ def polar_dist_draws(model_name):
 
 istar_draws, noistar_draws= polar_dist_draws("polar")
 
-fig = plt.figure(figsize=(5,2.5),dpi=110)
+fig = plt.figure(figsize=(5,2.4),dpi=110)
 
 plt.subplot(1,2,1)
-q025, q16, q50, q84, q975 = np.percentile(istar_draws, [2.5, 16, 50, 84, 97.5], axis=1)/2
+q025, q16, q50, q84, q975 = np.percentile(noistar_draws, [2.5, 16, 50, 84, 97.5], axis=1)/2
 plt.plot(2*x-1, q50, color='C0', lw=1.2)
 plt.fill_between(2*x-1, q16, q84, alpha=0.3, label="posterior", color='#7dabd0')
 plt.fill_between(2*x-1, q025, q975, alpha=0.3, color='#7dabd0')
@@ -73,10 +73,10 @@ plt.xlim([-1,1])
 plt.xlabel(r'$\cos{\psi}$',fontsize=11)
 plt.ylabel('Probablity density function')
 
-plt.title(r'measured $\vb*{\lambda}$ and $\vb*{i_\star}$', fontsize=11)
+plt.title(r'measured $\vb*{\lambda}$ only', fontsize=11)
 
 plt.subplot(1,2,2)
-q025, q16, q50, q84, q975 = np.percentile(noistar_draws, [2.5, 16, 50, 84, 97.5], axis=1)/2
+q025, q16, q50, q84, q975 = np.percentile(istar_draws, [2.5, 16, 50, 84, 97.5], axis=1)/2
 plt.plot(2*x-1, q50, color='#f56e4a')
 plt.fill_between(2*x-1, q16, q84, alpha=0.3, label="posterior", color='#fbc1ad')
 plt.fill_between(2*x-1, q025, q975, alpha=0.3, color='#fbc1ad')
@@ -84,7 +84,7 @@ plt.fill_between(2*x-1, q025, q975, alpha=0.3, color='#fbc1ad')
 plt.ylim([0,1])
 plt.xlim([-1,1])
 
-plt.title(r'measured $\vb*{\lambda}$ only', fontsize=11)
+plt.title(r'measured $\vb*{\lambda}$ and $\vb*{i_\star}$', fontsize=11)
 
 plt.tight_layout()
 

### 2023-05-23 src/scripts/polar_plot.py
update polar figure

@@ -42,28 +42,37 @@ def polar_dist_draws(model_name):
 	noistar_idata = az.from_netcdf(paths.data / "polar" / (model_name + "_noistar.nc"))
 
 	post = istar_idata.posterior
-	istar_draws = np.zeros(shape=(len(x),4000))
+	istar_draws0 = np.zeros(shape=(len(x),4000))
+	istar_draws1 = np.zeros(shape=(len(x),4000))
 	for a in range(4):
 		for b in range(1000):
-			istar_draws[:, a*1000+b] = (post.w[a,b,0].values*beta.pdf(x, post.a[a,b,0], post.b[a,b,0])
-										+post.w[a,b,1].values*beta.pdf(x, post.a[a,b,1], post.b[a,b,1]))
+			istar_draws0[:, a*1000+b] = post.w[a,b,0].values*beta.pdf(x, post.a[a,b,0], post.b[a,b,0])
+			istar_draws1[:, a*1000+b] = post.w[a,b,1].values*beta.pdf(x, post.a[a,b,1], post.b[a,b,1])
 	
 	post = noistar_idata.posterior
-	noistar_draws = np.zeros(shape=(len(x),4000))
+	noistar_draws0 = np.zeros(shape=(len(x),4000))
+	noistar_draws1 = np.zeros(shape=(len(x),4000))
 	for a in range(4):
 		for b in range(1000):
-			noistar_draws[:, a*1000+b] = (post.w[a,b,0].values*beta.pdf(x, post.a[a,b,0], post.b[a,b,0])
-										  +post.w[a,b,1].values*beta.pdf(x, post.a[a,b,1], post.b[a,b,1]))
+			noistar_draws0[:, a*1000+b] = post.w[a,b,0].values*beta.pdf(x, post.a[a,b,0], post.b[a,b,0])
+			noistar_draws1[:, a*1000+b] = post.w[a,b,1].values*beta.pdf(x, post.a[a,b,1], post.b[a,b,1])
 
-	return istar_draws, noistar_draws
+	return istar_draws0, istar_draws1, noistar_draws0, noistar_draws1
 
-istar_draws, noistar_draws= polar_dist_draws("polar")
+istar_draws0, istar_draws1, noistar_draws0, noistar_draws1= polar_dist_draws("polar")
 
[…]

### 2023-05-23 src/scripts/psi_plot.py

@@ -41,7 +41,7 @@ for a in range(4):
         comp1[:, a*1000+b] = post.w[a,b,1].values*beta.pdf(x, post.μ[a,b,1].values*post.κ[a,b,1].values, (1-post.μ[a,b,1]).values*post.κ[a,b,1].values)
 
 # make the figure
-plt.figure(figsize=(3.5,2.7),dpi=150)
+fig = plt.figure(figsize=(3.5,2.7),dpi=150)
 
 q025, q16, q50, q84, q975 = np.percentile(comp0, [2.5, 16, 50, 84, 97.5], axis=1)/2
 plt.plot(2*x-1, q50, '-.', c='k', lw=1., zorder=10)
@@ -59,6 +59,8 @@ plt.xlim([-1,1])
 
 plt.xlabel(r'$\cos{\psi}$',fontsize=11)
 plt.ylabel('Probablity density function')
+
+fig.set_facecolor('w')
 
 plt.tight_layout()
 plt.savefig(paths.figures / "psi_dist.pdf", bbox_inches="tight", dpi=600)

### 2023-05-23 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 % Define document class
-\documentclass[twocolumn,times,linenumbers]{aastex631}
+\documentclass[twocolumn,times]{aastex631}
 \usepackage{showyourwork}
@@
 \newcommand{\dfm}[1]{\textcolor{blue}{\textbf{From DFM:} #1}}
-\newcommand{\edits}[1]{\textcolor{purple}{#1}}
-
-\received{May 1, 2023}
-\revised{}
-\accepted{}
-
-\submitjournal{the AAS Journals}
+\newcommand{\edits}[1]{\textcolor{black}{#1}}
 
@@
 \subsection{Simulated Data}\label{subsec:sim}
+
+\begin{figure}[ht!]
+    \centering
+    \script{sim_plot.py}
+    \includegraphics{figures/simulation.pdf}
+    \caption{Inferred stellar obliquity distributions from sky-projected stellar obliquities \edits{without and with} information on the stellar inclination, depicted in blue and orange curves, respectively.
+Each row presents a set of simulated data with the true distribution of $\cos{\psi}$ indicated by grey dashed curves.
+The shallow contours represent the 1-$\sigma$ and 2-$\sigma$ uncertainties of the inferred distributions.}
+    \label{fig:simulation}
+\end{figure}
+
 To investigate the performance of our hierarchical Bayesian framework, we apply it to simulated data, generated with a known ground-truth stellar obliquity distribution.
@@
 
-\begin{figure}[ht!]
+\subsection{Jacobian Transformations Between the $\psi$, $\lambda$, and \lowercase{$i_\star$} Distributions}\label{subsec:jacobian}
+
+\begin{figure*}[ht!]
     \centering
-    \script{sim_plot.py}
-    \includegraphics{figures/simulation.pdf}
-    \caption{Inferred stellar obliquity distributions from sky-projected stellar obliquities \edits{without and with} information on the stellar inclination, depicted in blue and orange curves, respectively.
-Each row presents a set of simulated data with the true distribution of $\cos{\psi}$ indicated by grey dashed curves.
-The shallow contours represent the 1-$\sigma$ and 2-$\sigma$ uncertainties of the inferred distributions.}
-    \label{fig:simulation}
-\end{figure}
-
-\subsection{Jacobian Transformations Between the $\psi$, $\lambda$, and \lowercase{$i_\star$} Distributions}\label{subsec:jacobian}
+    \script{transform.py}
+    \includegraphics[width=0.7\linewidth]{figures/transform.pdf}
+    \caption{Simulated $\cos{\psi}$ distributions (\nth{1} column) and the corresponding distributions of sky-projected stellar obliquity $\lambda$ (\nth{2} column) and stellar inclination $i_\star$ (\nth{3} column).
+The grey histograms present the random samplings of $\lambda$ and $i_\star$ from the $\cos{\psi}$ distributions, and the blue curves present the numerical solutions.}
+    \label{fig:transform}
+\end{figure*}
 
@@
 
-\begin{figure*}[ht!]
-    \centering
-    \script{transform.py}
-    \includegraphics[width=0.7\linewidth]{figures/transform.pdf}
-    \caption{Simulated $\cos{\psi}$ distributions (\nth{1} column) and the corresponding distributions of sky-projected stellar obliquity $\lambda$ (\nth{2} column) and stellar inclination $i_\star$ (\nth{3} column).
-The grey histograms present the random samplings of $\lambda$ and $i_\star$ from the $\cos{\psi}$ distributions, and the blue curves present the numerical solutions.}
-    \label{fig:transform}
-\end{figure*}
-
 Using Equation~(\ref{eqn:jac_lam}) and (\ref{eqn:jac_istar}), we can now derive the $\lambda$ and $i_\star$ distributions for any given $\psi$ distributions, assuming the azimuthal angle of the stellar spin axis $\theta$ is random.
@@
 \section{Application to Exoplanetary Systems}\label{sec:observation}
-
-We next apply our hierarchical Bayesian framework to a sample of 161 exoplanetary systems with sky-projected stellar obliquity measurements, primarily consisting of Hot Jupiter systems, as summarized in \cite{Albrecht22} Table A1.
-The inferred $\cos{\psi}$ distribution is shown in Figure~\ref{fig:psi_dist}.
-The $\cos{\psi}$ distribution is peaked at 1, with nearly flat behavior between $-0.75$ and $0.75$ and no significant clustering.
-The distribution suggests that there is a pile-up of planetary systems with stellar obliquities less than $40^\circ$ and an isotropic distribution for obliquities between $40^\circ$ and $140^\circ$.
-The fraction of aligned systems $w_1$ dominates the distribution, with $w_1 = 0.719 \pm 0.085$.
-The corresponding posteriors for the parameters of the population Beta distribution are $\mu_1 = 0.976 \pm 0.022$ and $\kappa_1 = 14.1 \pm 26.6$.
-On the other hand, the fraction of misaligned systems is estimated to be $w_0 = 0.281 \pm 0.085$, with posteriors of $\mu_1 = 0.434 \pm 0.088$ and $\kappa_1 = 4.2 \pm 5.6$.
-Note that the $\mu_{0,1}$ needs to be transformed by $2\mu-1$ to represent to true means of the $\cos{\psi}$ distributions.
 
@@
 \end{figure*}
+
+We next apply our hierarchical Bayesian framework to a sample of 161 exoplanetary systems with sky-projected stellar obliquity measurements, primarily consisting of Hot Jupiter systems, as summarized in \cite{Albrecht22} Table A1.
+The inferred $\cos{\psi}$ distribution is shown in Figure~\ref{fig:psi_dist}.
+The $\cos{\psi}$ distribution is peaked at 1, with nearly flat behavior between $-0.75$ and $0.75$ and no significant clustering.
+The distribution suggests that there is a pile-up of planetary systems with stellar obliquities less than $40^\circ$ and an isotropic distribution for obliquities between $40^\circ$ and $140^\circ$.
+The fraction of aligned systems $w_1$ dominates the distribution, with $w_1 = 0.719 \pm 0.085$.
+The corresponding posteriors for the parameters of the population Beta distribution are $\mu_1 = 0.976 \pm 0.022$ and $\kappa_1 = 14.1 \pm 26.6$.
+On the other hand, the fraction of misaligned systems is estimated to be $w_0 = 0.281 \pm 0.085$, with posteriors of $\mu_1 = 0.434 \pm 0.088$ and $\kappa_1 = 4.2 \pm 5.6$.
+Note that the $\mu_{0,1}$ needs to be transformed by $2\mu-1$ to represent to true means of the $\cos{\psi}$ distributions.
 

### 2023-05-23 src/scripts/polar_plot.py
update polar figure

@@ -65,11 +65,11 @@ fig = plt.figure(figsize=(5,2.4),dpi=110)
 
 plt.subplot(1,2,1)
 
-q025, q16, q50, q84, q975 = np.percentile(noistar_draws0, [2.5, 16, 50, 84, 97.5], axis=1)/2
-plt.plot(2*x-1, q50, '-.', c='k', lw=1., zorder=10)
+# q025, q16, q50, q84, q975 = np.percentile(noistar_draws0, [2.5, 16, 50, 84, 97.5], axis=1)/2
+# plt.plot(2*x-1, q50, '-.', c='k', lw=1., zorder=10)
 
-q025, q16, q50, q84, q975 = np.percentile(noistar_draws1, [2.5, 16, 50, 84, 97.5], axis=1)/2
-plt.plot(2*x-1, q50, '--', c='k', lw=1., zorder=10)
+# q025, q16, q50, q84, q975 = np.percentile(noistar_draws1, [2.5, 16, 50, 84, 97.5], axis=1)/2
+# plt.plot(2*x-1, q50, '--', c='k', lw=1., zorder=10)
 
 q025, q16, q50, q84, q975 = np.percentile(noistar_draws0+noistar_draws1, [2.5, 16, 50, 84, 97.5], axis=1)/2
 plt.plot(2*x-1, q50, color='C0', lw=1.5)
@@ -86,11 +86,11 @@ plt.title(r'measured $\vb*{\lambda}$ only', fontsize=11)
 
 plt.subplot(1,2,2)
 
-q025, q16, q50, q84, q975 = np.percentile(istar_draws0, [2.5, 16, 50, 84, 97.5], axis=1)/2
-plt.plot(2*x-1, q50, '-.', c='k', lw=1., zorder=10)
+# q025, q16, q50, q84, q975 = np.percentile(istar_draws0, [2.5, 16, 50, 84, 97.5], axis=1)/2
+# plt.plot(2*x-1, q50, '-.', c='k', lw=1., zorder=10)
 
-q025, q16, q50, q84, q975 = np.percentile(istar_draws1, [2.5, 16, 50, 84, 97.5], axis=1)/2
-plt.plot(2*x-1, q50, '--', c='k', lw=1., zorder=10)
+# q025, q16, q50, q84, q975 = np.percentile(istar_draws1, [2.5, 16, 50, 84, 97.5], axis=1)/2
[…]

### 2023-05-23 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 % Define document class
-\documentclass[twocolumn,times]{aastex631}
+\documentclass[twocolumn]{aastex631}
 \usepackage{showyourwork}
@@
 For the stellar obliquity distribution, we model the $\cos{\psi}$ distribution instead of $\psi$ distribution to understand if the stellar obliquity is isotropically distributed.
-If stellar obliquity distribution is isotropic, $\cos{\psi}$ is uniformly distributed between -1 and 1.
+If stellar obliquity distribution is isotropic, $\cos{\psi}$ is uniformly distributed between $-1$ and $1$.
 To flexibly model the stellar obliquity distribution, we select a multi-component mixture of Beta distributions with hyperparameters $\bm{\beta} = \{\bm{w},\bm{\mu},\bm{\kappa}\}$, where each hyperparameter has a dimension of the number of components.
@@
 To ensure the robustness of the inferred distributions in such cases, it is crucial to test their sensitivity to different hyperpriors.
+\edits{We also note that our framework is flexible to customize to distributions other than the Beta distribution for the population-level stellar obliquity distribution inference.
+For example, if instead studying the stellar obliquity distribution in the $\psi$-angle space, a von Mises distribution could be used to evaluate the mean value and the dispersion of the angle.}
 
@@
 If $\gamma_{\star, n}$ is available, we construct Normal distributions with means and standard deviations from measurements.
-For the orbital inclination $i_{\rm orb}$, following the convention, we limit it to $\left[0, \pi/2\right]$.
-To not underestimate the $\psi$, we then set $i_\star$ to be varied from $0\degr$ to $180\degr$.
-Besides, since $\lambda$ and $-\lambda$ correspond to the same $\psi$ solution, we limit $\lambda$ to $\left[0, \pi\right]$ and thus $\theta$ to $\left[0, \pi\right]$.
+For the orbital inclination $i_{\rm orb}$, following the convention, we limit it to $\left[0, \pi/2\right]$, \edits{i.e., the orbital axis always pointing to us}.
+To not underestimate the $\psi$, we then set $i_\star$ to be varied from $0\degr$ to $180\degr$ \edits{such that the stellar spin axis could either point to us or point away from us}.
+Besides, since $\lambda$ and $-\lambda$ correspond to the same $\psi$ solution, we limit $\lambda$ to $\left[0, \pi\right]$ and thus $\theta$ to $\left[0, \pi\right]$ \edits{to avoid bimodal distributions of $\lambda$ and $\theta$.}
 We find doing so greatly improves the sampling performance while not compromising the inference of the $\psi$ distribution due to the symmetry.
@@
 
-\section{Model Performance}\label{sec:simulation}
-
-\subsection{Simulated Data}\label{subsec:sim}
-
-\begin{figure}[ht!]
+\begin{figure}
     \centering
@@
 \end{figure}
+
+\begin{figure*}
+    \centering
+    \script{transform.py}
+    \includegraphics[width=0.7\linewidth]{figures/transform.pdf}
+    \caption{Simulated $\cos{\psi}$ distributions (\nth{1} column) and the corresponding distributions of sky-projected stellar obliquity $\lambda$ (\nth{2} column) and stellar inclination $i_\star$ (\nth{3} column).
+The grey histograms present the random samplings of $\lambda$ and $i_\star$ from the $\cos{\psi}$ distributions, and the blue curves present the numerical solutions.}
+    \label{fig:transform}
+\end{figure*}
+
+\section{Model Performance}\label{sec:simulation}
+
+\subsection{Simulated Data}\label{subsec:sim}
 
@@
 \subsection{Jacobian Transformations Between the $\psi$, $\lambda$, and \lowercase{$i_\star$} Distributions}\label{subsec:jacobian}
-
-\begin{figure*}[ht!]
-    \centering
-    \script{transform.py}
-    \includegraphics[width=0.7\linewidth]{figures/transform.pdf}
-    \caption{Simulated $\cos{\psi}$ distributions (\nth{1} column) and the corresponding distributions of sky-projected stellar obliquity $\lambda$ (\nth{2} column) and stellar inclination $i_\star$ (\nth{3} column).
-The grey histograms present the random samplings of $\lambda$ and $i_\star$ from the $\cos{\psi}$ distributions, and the blue curves present the numerical solutions.}
-    \label{fig:transform}
-\end{figure*}
 
@@
 
-\begin{figure}[ht!]
+\begin{figure}
     \script{psi_plot.py}
@@
 
-\begin{figure*}[ht!]
+\begin{figure*}
     \script{polar_plot.py}
@@
         \includegraphics{figures/polar.pdf}
-        \caption{Inferred stellar obliquity distribution for the subsample of exoplanetary systems with both sky-projected stellar obliquity measurements and stellar inclination measurements.}
+        \caption{Inferred stellar obliquity distribution for the subsample of exoplanetary systems with both sky-projected stellar obliquity measurements and stellar inclination measurements.
+Inference made purely from measured projected stellar obliquity and inference made from both measured projected stellar obliquity and stellar inclination are shown in blue and orange curves, respectively.
+The line and shallow contours represent the median, 1-$\sigma$ and 2-$\sigma$ uncertainties of the inferred distributions.}
         \label{fig:polar}
@@
 We next apply our hierarchical Bayesian framework to a sample of 161 exoplanetary systems with sky-projected stellar obliquity measurements, primarily consisting of Hot Jupiter systems, as summarized in \cite{Albrecht22} Table A1.
+\edits{We use a two-component model here for its flexibility in describing both an aligned-system population and a misaligned-system population that may or may not be concentrated at a certain angle.}
 The inferred $\cos{\psi}$ distribution is shown in Figure~\ref{fig:psi_dist}.
@@
 The discrepancy between this inference for the full sample of exoplanet systems and the previous analysis of the subsample with $i_\star$ measurements \citep{Albrecht21} warrants further investigation.
-This earlier study identified a concentration of perpendicular planets and disfavored an isotropic stellar obliquity distribution \citep{Albrecht21}.
+This earlier study identified a concentration of perpendicular planets and disfavored an isotropic stellar obliquity distribution.
+\edits{Indeed, when we applied our framework to this subsample, a concentration near $\cos{\psi} = -0.2$ was found in models including or without including the stellar inclination information, as shown in Figure~\ref{fig:polar}.}
 There are at least two potential explanations for this difference: (1) the subsample with $i_\star$ measurements is small and only includes about 20 misaligned systems, and (2) the requirement for $i_\star$ measurements could introduce selection biases in the sample.
 First, since the sample size of the misaligned sample with $i_\star$ measurements is relatively small ($N < 30$), the observed sample, even if it is unbiased, may not be able to represent the underlying distribution.
+\edits{The small sample size leads to a large uncertainty on the inferred stellar obliquity distribution, shown by the 1-$\sigma$ and 2-$\sigma$ contours in Figure~\ref{fig:polar}.}
 Second, the requirement of $i_\star$ measurements could introduce selection biases.

### 2023-06-05 src/scripts/polar.py
update polar plot and randseed

@@ -229,7 +229,7 @@ if __name__ == '__main__':
 	# Sampling
 	with model_istar:
 		idata_istar = pm.sample(nuts={'target_accept':0.99, 'max_treedepth':13}, # 'step_scale':0.01
-        						chains=4, random_seed=int(datetime.now().strftime("%Y%m%d")))
+        						chains=4, random_seed=123)
 
 	# Save the traces
 	az.to_netcdf(idata_istar.posterior, sim_dir / "polar_istar.nc")
@@ -281,6 +281,6 @@ if __name__ == '__main__':
 	# Sampling
 	with model_noistar:
 		idata_noistar = pm.sample(nuts={'target_accept':0.99, 'max_treedepth':13}, # 'step_scale':0.01
-									chains=4, random_seed=int(datetime.now().strftime("%Y%m%d")))
+									chains=4, random_seed=123)
 	# Save the traces
 	az.to_netcdf(idata_noistar.posterior, sim_dir / "polar_noistar.nc")

### 2023-06-05 src/scripts/polar_plot.py

@@ -35,11 +35,24 @@ plt.rcParams['lines.markeredgewidth'] =  1.0
 from scipy.stats import beta
 x = np.linspace(1e-5,1-1e-5,1000)
 
+numall = 151
+numistar = 62
+
 # load the posteriors from the MCMC and calculate psi dists
 def polar_dist_draws(model_name):
 
+	all_noistar = az.from_netcdf(paths.data / "all_noistar.nc")
+
 	istar_idata = az.from_netcdf(paths.data / "polar" / (model_name + "_istar.nc"))
 	noistar_idata = az.from_netcdf(paths.data / "polar" / (model_name + "_noistar.nc"))
+
+	post = all_noistar.posterior
+	comp0 = np.zeros(shape=(len(x),4000))
+	comp1 = np.zeros(shape=(len(x),4000))
+	for a in range(4):
+	    for b in range(1000):
+	        comp0[:, a*1000+b] = post.w[a,b,0].values*beta.pdf(x, post.μ[a,b,0].values*post.κ[a,b,0].values, (1-post.μ[a,b,0]).values*post.κ[a,b,0].values)
+	        comp1[:, a*1000+b] = post.w[a,b,1].values*beta.pdf(x, post.μ[a,b,1].values*post.κ[a,b,1].values, (1-post.μ[a,b,1]).values*post.κ[a,b,1].values)
 
 	post = istar_idata.posterior
 	istar_draws0 = np.zeros(shape=(len(x),4000))
@@ -57,50 +70,53 @@ def polar_dist_draws(model_name):
 			noistar_draws0[:, a*1000+b] = post.w[a,b,0].values*beta.pdf(x, post.a[a,b,0], post.b[a,b,0])
 			noistar_draws1[:, a*1000+b] = post.w[a,b,1].values*beta.pdf(x, post.a[a,b,1], post.b[a,b,1])
 
-	return istar_draws0, istar_draws1, noistar_draws0, noistar_draws1
+	return comp0, comp1, istar_draws0, istar_draws1, noistar_draws0, noistar_draws1
 
-istar_draws0, istar_draws1, noistar_draws0, noistar_draws1= polar_dist_draws("polar")
+comp0, comp1, istar_draws0, istar_draws1, noistar_draws0, noistar_draws1= polar_dist_draws("polar")
[…]

### 2023-06-05 src/scripts/psi.py

@@ -144,6 +144,6 @@ if __name__ == '__main__':
         logl_λ = pm.Normal('logl_λ', mu=λ, sigma=err_Lam, observed=Lam)
         
         all_noistar = pm.sample(nuts={'target_accept':0.99, 'max_treedepth':13}, # 'step_scale':0.01
-                                chains=4, random_seed=int(datetime.now().strftime("%Y%m")))
+                                chains=4, random_seed=123)
             
     az.to_netcdf(all_noistar.posterior, paths.data / "all_noistar.nc")

### 2023-06-05 src/scripts/sim.py

@@ -5,7 +5,6 @@ os.environ["PATH"] += os.pathsep + str(Path.home() / "bin")
 
 import subprocess
 import sys
-from datetime import datetime
 
 import pymc as pm
 import arviz as az
@@ -74,7 +73,7 @@ def posteriors(this_model):
         cosθ = pm.Deterministic('cosθ', at.cos(θ))
         
         # iorb
-        iorb = np.pi/2
+        iorb = np.pi/2*np.ones(nsample)
         
         # find λ in terms of ψ, θ, and iorb
         λ = pm.Deterministic('λ', at.arctan2(sinψ*sinθ, cosψ*at.sin(iorb)-sinψ*cosθ*at.cos(iorb)))
@@ -90,7 +89,7 @@ def posteriors(this_model):
         logl_i = pm.Normal('logl_i', mu=istar, sigma=err_istar, observed=obs_istar)
 
         idata = pm.sample(nuts={'target_accept':0.99, 'max_treedepth':13}, 
-                          chains=4, random_seed=int(datetime.now().strftime("%Y%m%d")))
+                          chains=4, random_seed=123)
 
     with pm.Model() as model_noistar:
 

### 2023-06-05 src/scripts/sim_plot.py

@@ -126,9 +126,46 @@ cosψ = np.linspace(-1,1,200)
 pcosψ = np.ones_like(cosψ)*0.5
 plt.plot(cosψ, pcosψ, c='slategrey', ls='--', lw=1)
 
+# cosψ ~ Beta(3,6), no istar
+ax = fig.add_subplot(5,2,3)
+        
+q025, q16, q50, q84, q975 = np.percentile(beta_noistar_draws, [2.5, 16, 50, 84, 97.5], axis=1)/2
+plt.plot(2*x-1, q50, color='C0', lw=1.2)
+plt.fill_between(2*x-1, q16, q84, alpha=0.3, label="posterior", color='#7dabd0')
+plt.fill_between(2*x-1, q025, q975, alpha=0.3, color='#7dabd0')
+
+plt.ylim([0,1.5])
+plt.xlim([-1,1])
+
+plt.xlabel(r'$\cos{\psi}$',fontsize=11)
+plt.title(r'measured $\vb*{\lambda}$ only', fontsize=10)
+
+cosψ = np.linspace(0,1,200)
+pcosψ = stats.beta.pdf(cosψ, 3, 6)/2
+plt.plot(2*cosψ-1, pcosψ, c='slategrey', ls='--', lw=1, zorder=0)
+
+
+# cosψ ~ Beta(3,6), with istar
+ax = fig.add_subplot(5,2,4)
+
+q025, q16, q50, q84, q975 = np.percentile(beta_draws, [2.5, 16, 50, 84, 97.5], axis=1)/2
+plt.plot(2*x-1, q50, color='#f56e4a', lw=1.2)
+plt.fill_between(2*x-1, q16, q84, alpha=0.3, label="posterior", color='#fbc1ad')
+plt.fill_between(2*x-1, q025, q975, alpha=0.3, color='#fbc1ad')
+
+plt.ylim([0,1.5])
+plt.xlim([-1,1])
+
+plt.xlabel(r'$\cos{\psi}$',fontsize=11)
+plt.title(r'measured $\vb*{\lambda}$ and $\vb*{i_\star}$', fontsize=10)
+
+cosψ = np.linspace(0,1,200)
+pcosψ = stats.beta.pdf(cosψ, 3, 6)/2
+plt.plot(2*cosψ-1, pcosψ, c='slategrey', ls='--', lw=1, zorder=0)
+
 
[…]

### 2023-06-06 src/scripts/polar_plot.py
polar plot normalization

@@ -78,13 +78,13 @@ comp0, comp1, istar_draws0, istar_draws1, noistar_draws0, noistar_draws1= polar_
 fig = plt.figure(figsize=(7,2.6),dpi=110)
 
 plt.subplot(1,3,1)
-q025, q16, q50, q84, q975 = np.percentile(comp0+comp1, [2.5, 16, 50, 84, 97.5], axis=1)
+q025, q16, q50, q84, q975 = np.percentile(comp0+comp1, [2.5, 16, 50, 84, 97.5], axis=1)/2
 plt.plot(2*x-1, q50, color='C0')
 plt.fill_between(2*x-1, q16, q84, alpha=0.3, label="posterior", color='#7dabd0')
 plt.fill_between(2*x-1, q025, q975, alpha=0.3, color='#7dabd0')
 
 
-plt.ylim([0,2])
+plt.ylim([0,1])
 plt.xlim([-1,1])
 
 plt.xlabel(r'$\cos{\psi}$',fontsize=11)
@@ -94,12 +94,12 @@ plt.title(r'\begin{center} all %i systems \\ \textbf{no} $i_{\star}$ likelihood 
           fontsize=11, y=1.1)
 
 plt.subplot(1,3,2)
-q025, q16, q50, q84, q975 = np.percentile(noistar_draws0+noistar_draws1, [2.5, 16, 50, 84, 97.5], axis=1)
+q025, q16, q50, q84, q975 = np.percentile(noistar_draws0+noistar_draws1, [2.5, 16, 50, 84, 97.5], axis=1)/2
 plt.plot(2*x-1, q50, color='#69a9a3')
 plt.fill_between(2*x-1, q16, q84, alpha=0.5, label="posterior", color='#cfe7ea')
 plt.fill_between(2*x-1, q025, q975, alpha=0.5, color='#cfe7ea')
 
-plt.ylim([0,2])
+plt.ylim([0,1])
 plt.xlim([-1,1])
 
 plt.title(r'\begin{center} %i systems with observed $i_{\star}$ \\ \textbf{no} $i_{\star}$ likelihood \end{center}'%numistar, 
@@ -107,12 +107,12 @@ plt.title(r'\begin{center} %i systems with observed $i_{\star}$ \\ \textbf{no} $
 
 
 plt.subplot(1,3,3)
[…]

### 2023-06-06 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 % Define document class
-\documentclass[twocolumn]{aastex631}
+\documentclass[twocolumn,times]{aastex631}
 \usepackage{showyourwork}
@@
 \usepackage[super]{nth}
-\usepackage{fontawesome}
+\usepackage{fontawesome5}
 \usepackage{hyperref}
 
-\newcommand{\dfm}[1]{\textcolor{blue}{\textbf{From DFM:} #1}}
-\newcommand{\edits}[1]{\textcolor{black}{#1}}
+\newcommand{\edits}[1]{\textcolor{purple}{#1}}
 
@@
 To investigate the performance of our hierarchical Bayesian framework, we apply it to simulated data, generated with a known ground-truth stellar obliquity distribution.
-We test the following \edits{five} $\cos{\psi}$ distributions: a uniform distribution bounded between $-1$ and $1$ such that $\cos{\psi} \sim \mathcal{U}(-1,1)$, three truncated Normal distributions bounded between $-1$ and $1$, such that $\cos{\psi} \sim \mathcal{N}(0, 0.2)$, $\cos{\psi} \sim \mathcal{N}(-0.4, 0.2)$, and $\cos{\psi} \sim \mathcal{N}(0.4, 0.2)$, where the first number is the mean of the distribution and the second number is the standard deviation, \edits{and lastly, a Beta distribution such that $\cos{\psi} \sim \mathrm{Beta}(3, 6)$ with the support being extended linearly from $[0, 1]$ to $[-1, 1]$}.
-For each $\cos{\psi}$ distribution, we randomly generate 200 samples of sky-projected stellar obliquity $\lambda$ and stellar inclination $i_\star$.
+We test the following \edits{five} $\cos{\psi}$ distributions: a uniform distribution bounded between $-1$ and $1$ such that $\cos{\psi} \sim \mathcal{U}(-1,1)$, \edits{a Beta distribution such that $\cos{\psi} \sim \mathrm{Beta}(3, 6)$ with the support being extended linearly from $[0, 1]$ to $[-1, 1]$}, and three truncated Normal distributions bounded between $-1$ and $1$, such that $\cos{\psi} \sim \mathcal{N}(0, 0.2)$, $\cos{\psi} \sim \mathcal{N}(-0.4, 0.2)$, and $\cos{\psi} \sim \mathcal{N}(0.4, 0.2)$, where the first number is the mean of the distribution and the second number is the standard deviation.
+For each $\cos{\psi}$ distribution, we randomly generate \edits{400} samples of sky-projected stellar obliquity $\lambda$ and stellar inclination $i_\star$.
 We assume the stellar spin axis is uniformly distributed around the planetary orbital axis in the azimuthal direction and the orbital inclination is $90\degr$.
@@
         \includegraphics{figures/polar.pdf}
-        \caption{Inferred stellar obliquity distribution for the subsample of exoplanetary systems with both sky-projected stellar obliquity measurements and stellar inclination measurements.
-Inference made purely from measured projected stellar obliquity and inference made from both measured projected stellar obliquity and stellar inclination are shown in blue and orange curves, respectively.
-The line and shallow contours represent the median, 1-$\sigma$ and 2-$\sigma$ uncertainties of the inferred distributions.}
+        \caption{Inferred stellar obliquity distribution for the full sample of exoplanetary systems with sky-projected stellar obliquity measurements (left panel) and the subsample of systems with both sky-projected stellar obliquity measurements and stellar inclination measurements (middle and right panels).
+The subsample inference made purely from measured projected stellar obliquity is shown in the middle panel, and the subsample inference made from both measured projected stellar obliquity and stellar inclination is shown in the right panel.
+The line and shallow contours represent the median, 1-$\sigma$ and 2-$\sigma$ uncertainties of the inferred distributions.
+The samples are extracted from \cite{Albrecht21} and \cite{Albrecht22}.}
         \label{fig:polar}
